### 2012-05-20 paper.tex
Generate pdf document from README.markdown

%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
%               sigplanconf-guide.pdf
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------

\documentclass[preprint]{sigplanconf}

% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\usepackage{amsmath}
\usepackage{enumerate}

\begin{document}

\conferenceinfo{WXYZ '05}{date, City.} 
\copyrightyear{2005} 
\copyrightdata{[to be supplied]}

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Title Text}
\subtitle{Subtitle Text, if any}

\authorinfo{Name1}
           {Affiliation1}
           {Email1}
\authorinfo{Name2\and Name3}
           {Affiliation2/3}
           {Email2/3}

\maketitle

\begin{abstract}
This is the text of the abstract.
\end{abstract}

\category{CR-number}{subcategory}{third-level}

\terms
term1, term2

\keywords
keyword1, keyword2

\section{Introduction}

\input{main.tex}

\appendix
\section{Appendix Title}

This is the text of the appendix, if you need one.

\acks

Acknowledgments, if needed.

% We recommend abbrvnat bibliography style.

\bibliographystyle{abbrvnat}

% The bibliography should be embedded for final submission.

\begin{thebibliography}{}
\softraggedright

\bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
P. Q. Smith, and X. Y. Jones. ...reference text...

\end{thebibliography}

\end{document}

### 2012-05-22 paper.tex
structuring a paper.

@@
 \usepackage{amsmath}
-\usepackage{enumerate}
 
@@
 
-\titlebanner{banner above paper title}        % These are ignored unless
+\titlebanner{A draft paper to be submitted to Haskell Symposium 2012}
 \preprintfooter{short description of paper}   % 'preprint' option specified.
 
-\title{Title Text}
-\subtitle{Subtitle Text, if any}
+\title{Behavior Driven Development in Haskell}
+%\subtitle{Subtitle Text, if any}
 
-\authorinfo{Name1}
-           {Affiliation1}
-           {Email1}
-\authorinfo{Name2\and Name3}
-           {Affiliation2/3}
-           {Email2/3}
+\authorinfo{Simon HENGEL}
+           {TBD: Affiliation}
+           {TBD: Email}
+\authorinfo{Trystan SPANGLER}
+           {TBD: Affiliation}
+           {TBD: Email}
+\authorinfo{Kazuhiko YAMAMOTO}
+           {IIJ Innovation Institute Inc.}
+           {kazu@iij.ad.jp}
 
@@
 \begin{abstract}
-This is the text of the abstract.
+TBD: This is the text of the abstract.
 \end{abstract}
@@
 \keywords
-keyword1, keyword2
+software design, documentation, automatic test
 
@@
 
-\input{main.tex}
+The text of the paper begins here.
 
-\appendix
-\section{Appendix Title}
+The contribution of this paper is as follows:
 
-This is the text of the appendix, if you need one.
+\begin{itemize}
+\item Introduced doctest from Python community to Haskell community
+\item Introduced *spec from Ruby community to Haskell community
+\item Integrated doctest and QuickCheck
+\item Show best current practice of design/documentation/tests
+\end{itemize}
+
+\section{Observation}
+
+Should talk about equation property.
+
+\section{Solution}
+
+Note: Kazu defines functions without signatures if they are simple enough.
+And their signatures are automatically inserted by IDE (like ghc-mod) later.
+
+\section{Implementation}
+
+\subsection{doctest}
+
+\subsection{doctest and QuickCheck}
+
+\subsection{hspec}
+
+\section{Conclusion}
+
+%% \appendix
+%% \section{Appendix Title}
+
+%% This is the text of the appendix, if you need one.
 
@@
 
-Acknowledgments, if needed.
+TBD: Acknowledgments, if needed.
 

### 2012-05-22 paper.tex
adding terminology.

@@
 
+\section{Terminology}
+
+\begin{description}
+\item[unit test] TBD
+\item[example] TBD
+\item[property] TBD
+\item[behavior] TBD
+\item[documentation] Text written in Haskell comments and parsed by Haddock
+\end{description}
+
 \section{Observation}
@@
 
-Note: Kazu defines functions without signatures if they are simple enough.
-And their signatures are automatically inserted by IDE (like ghc-mod) later.
+The authors are not fundamentalist.
+If functions to be defined are simple enough,
+we might implement them first withtout signatures.
+And their signatures can be automatically
+inserted by IDE (like ghc-mod) later
+if Haskell compilers can infer correctly.
+Also, we can cut and paste typical examples and their results
+from Haskell interpreters to their document.
 

### 2012-05-22 paper.tex
adding TBD.
writing ack.

@@
 
-\conferenceinfo{WXYZ '05}{date, City.} 
-\copyrightyear{2005} 
+\conferenceinfo{Haskell Symposium 2012}{13th September, Copenhagen.}
+\copyrightyear{2012}
 \copyrightdata{[to be supplied]}
@@
 \titlebanner{A draft paper to be submitted to Haskell Symposium 2012}
-\preprintfooter{short description of paper}   % 'preprint' option specified.
+\preprintfooter{TBD: short description of paper}   % 'preprint' option specified.
 
@@
 
-\category{CR-number}{subcategory}{third-level}
+\category{TBD:CR-number}{subcategory}{third-level}
 
 \terms
-term1, term2
+TBD
 
@@
 \section{Introduction}
-
-The text of the paper begins here.
 
@@
 
-TBD: Acknowledgments, if needed.
+We would like to thank
+Nick Smallbone for having improved QuickCheck so that boolean property is tested
+just once.
+We are grateful to David Waern and Simon Marlow
+for introducing the property syntax to haddock.
 

### 2012-05-23 paper.tex
adding memo and advice from "How to write a great research paper".

@@
 \begin{abstract}
-TBD: This is the text of the abstract.
+
+State the problem.
+
+Say why it’s an interesting problem.
+
+Say what your solution achieves.
+
+Say what follows from your solution.
+
 \end{abstract}
@@
 
+%%%% Describe the problem.
+
+TBD: Strong type system.
+
+TBD: Tend not to write tests
+
+The purpose of this paper is to
+encourage Haskllers to write documentation, which can be used for testing, in their libraries.
+
+%%%% State your contributions.
+
 The contribution of this paper is as follows:
@@
 \begin{itemize}
-\item Introduced doctest from Python community to Haskell community
-\item Introduced *spec from Ruby community to Haskell community
-\item Integrated doctest and QuickCheck
-\item Show best current practice of design/documentation/tests
+\item Introduced {\tt doctest} from Python community to Haskell community
+and integrated it with QuickCheck (Section XXX)
+\item Introduced {\tt hspec} from Ruby community to Haskell community and
+integrated it with QuickCheck (Section XXX)
+\item Show best current practice of design/documentation/tests (Section XXX)
 \end{itemize}
@@
 
-Should talk about equation property.
+Haddock example and doctest.
+
+Should talk about equation property. (e.g. container)
+Quickcheck.
+
+pure vs IO.
 
 \section{Solution}
+
+FIRST convey the idea.
+
+\subsection{Tests suitable with doctest}
+
+\subsection{Tests suitable with hspec}
+
+\subsection{Integrating multiple test suites}
+
+\section{Best current practice}
+
+undefined.
 
@@
 \subsection{hspec}
+
+\section{Related work}
+
+literate programming.
+
+hunit.
+
+test-framework.
+
+smallCheck.
+
+And others?
 

### 2012-05-23 paper.tex
writing intro.

@@
 
+TBD: Asbtract will be written at the final moment.
+
 State the problem.
@@
 
-TBD: Strong type system.
+One of the best features of Haskell is its strong type system, 
+which ensures no type errors in programs if they can be complied.
+Thanks to this, programs written in Haskell have less bugs
+than other programming languages which do not provide such strong type systems.
 
-TBD: Tend not to write tests
+Unfortunately, the strong type system of Haskell does not guarantee
+no value errors.
+Haskellers is, however, apt to rely on the strong type system too much.
+They tend to write less test programs than programmers of other
+programming languages.
 
-The purpose of this paper is to
-encourage Haskllers to write documentation, which can be used for testing, in their libraries.
+This also means that typical usage examples cannot be found from test programs.
+We agree that type signatures are good abstract document but
+there are many functions whose behaviors cannot be guessed from
+their names and type signatures.
+
+Why do we dislike to write test programs?
+It is probably because writing test programs is not fun.
+To encourage Haskellers to write more test programs
+to detect value errors,
+we need a scheme to make writing test programs fun.
+
+For this purpose,
+we brought design methods from other programming language communities and
+integrated them with a Haskell-original test method.
+In particular, we introduced {\tt doctest} and
+integrated with QuickCheck. 
+We also implemented {\tt hspec} with QuickCheck enabled.
+
+In our method, the text for design takes role of documentation and
+is used as test cases. 
+This method would motivate Haskellers
+bacause it has a kill-three-birds-with-one-stone aspect.
+
+%The purpose of this paper is to encourage Haskllers to write documentation, which can be used for testing, in their libraries.
 
@@
 
+how to handle error cases.
+
 \subsection{doctest and QuickCheck}

### 2012-05-23 paper.tex
writing observation.

@@
 
+TBD: we need evidence to support above.
+
 This also means that typical usage examples cannot be found from test programs.
@@
 to detect value errors,
-we need a scheme to make writing test programs fun.
+we need a scheme to make writing test programs more interesting.
 
@@
 
-In our method, the text for design takes role of documentation and
-is used as test cases. 
+In our method, design specification takes role of documentation for users and
+is used as test cases by its developer. 
 This method would motivate Haskellers
@@
 
+The definitions of words used in this paper is as follows:
+
 \begin{description}
-\item[unit test] TBD
-\item[example] TBD
-\item[property] TBD
-\item[behavior] TBD
-\item[documentation] Text written in Haskell comments and parsed by Haddock
+\item[unit test] tests for individual programs or libraries.
+\item[example] Typical usage of functions.
+\item[property] Invariants which functions have.
+\item[behavior] How do functions behave for users. Behavior can be shown by using both example and property.
+\item[design] TBD
+\item[documentation] Explanation of functions.
 \end{description}
@@
 
-Haddock example and doctest.
+We insist that examples and properties, which are useful for users, should be
+documented.
+So, let us first observe Haddock, the documentation system of Haskell. Haddock defines annotation syntax in Haskell's comment.
 
-Should talk about equation property. (e.g. container)
-Quickcheck.
+\subsection{Example annotation}
 
-pure vs IO.
+We can specify examples with the annotation syntax {\tt >>>}.
+The following is an example of how to use the {\tt length} function
+for lists.
+
+\begin{quote}
+\begin{verbatim}
+> length []
+0
+\end{verbatim}
+\end{quote}
+
+\noindent We can write this by copy-and-pasting from an interpreter.
+It's worth explaining that we can specify operation sequences:
+
+\begin{quote}
+\begin{verbatim}
+> writeFile "tmpfile" "Hello"
+> readFile "tmpfile"
+"Hello"
+\end{verbatim}
+\end{quote}
+
+\noindent But example annotations are useful for users but 
+it is not used as test case from the developer point of view (before we implemented {\tt doctest}).
+
+\subsection{Property annotation}
+
+The {\tt container} packages has equation properties:
+
+\begin{quote}
+\begin{verbatim}
+>>> Data.Map.null (empty) == True
+\end{verbatim}
+\end{quote}
+
+It is written in code blocks ({\tt >}).
+This cannot be used as test case.
+
+Equation property is tested 100 times.
+
+Quickcheck property should be able to be specified.
+
+To our experience, it is not difficult to find properties for pure functions.
+We believe that Haskellers should show properties of pure functions
+in documentation.
+
+\subsection{Internal documentation}
+
+not useful for users
 
@@
 We are grateful to David Waern and Simon Marlow
-for introducing the property syntax to haddock.
+for introducing the property syntax to Haddock.
 

### 2012-05-23 paper.tex
writing property annotation.

@@
 
-The {\tt container} packages has equation properties:
+Haddock also provides code block annotations indicated wwith ({\tt >}).
+For example, the {\tt container} packages has the following code block annotation:
 
@@
 
-It is written in code blocks ({\tt >}).
-This cannot be used as test case.
+\noindent Some Haskellers like this kind of equations and dislike to write it with the example annotation as follow:
+
+\begin{quote}
+\begin{verbatim}
+> Data.Map.null (empty)
+True
+\end{verbatim}
+\end{quote}
+
+We can see the equation as a boolean property for {\tt null} and {\tt empty} and
+can say that there are potentially demands to show properties with Haddock annotation.
+But there is no annotation syntax to describe properties in Haddock.
+Even if property annotation syntax is defined, 
+we have no agreement on how to describe non-boolean properties: e.g. how to specify types of funcation parameters.
 

### 2012-05-23 paper.tex
fixing mistakes.

@@
 \begin{verbatim}
-> length []
+>>> length []
 0
@@
 \begin{verbatim}
-> writeFile "tmpfile" "Hello"
-> readFile "tmpfile"
+>>> writeFile "tmpfile" "Hello"
+>>> readFile "tmpfile"
 "Hello"
@@
 
-Haddock also provides code block annotations indicated wwith ({\tt >}).
+Haddock also provides code block annotations indicated with ({\tt >}).
 For example, the {\tt container} packages has the following code block annotation:
@@
 \begin{verbatim}
->>> Data.Map.null (empty) == True
+> Data.Map.null (empty) == True
 \end{verbatim}
@@
 \begin{verbatim}
-> Data.Map.null (empty)
+>>> Data.Map.null (empty)
 True

### 2012-05-23 paper.tex
more fix.

@@
 
-\noindent We can write this by copy-and-pasting from an interpreter.
+\noindent The first line starts with {\tt >>>} and an expression follows. The second line contains its result. We can write this by copy-and-pasting from an interpreter and modifying the prompt.
 It's worth explaining that we can specify operation sequences:
@@
 
-\noindent But example annotations are useful for users but 
+\noindent Example annotations are useful for users but 
 it is not used as test case from the developer point of view (before we implemented {\tt doctest}).

### 2012-05-23 paper.tex
writing solution.

@@
 
-\category{TBD:CR-number}{subcategory}{third-level}
+\category{D.2.2}{Software Engineering}{Testing tools}
 
-\terms
-TBD
+\terms Languages, Design, Verification
 
-\keywords
-software design, documentation, automatic test
+\keywords Haddock, QuickCheck, doctest, hspec
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Terminology}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Observation}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Example annotation}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Property annotation}
@@
 can say that there are potentially demands to show properties with Haddock annotation.
-But there is no annotation syntax to describe properties in Haddock.
-Even if property annotation syntax is defined, 
-we have no agreement on how to describe non-boolean properties: e.g. how to specify types of funcation parameters.
-
-Equation property is tested 100 times.
-
-Quickcheck property should be able to be specified.
-
 To our experience, it is not difficult to find properties for pure functions.
@@
 
+But there is no annotation syntax to describe properties in Haddock.
+Even if a prefix keyword for properties is defined, 
+we have no agreement on how to describe non-boolean properties: e.g. how to specify types of funcation parameters. And we don't have automatic test mechanism to use properties in documentation.
+
+Note that QuickCheck 2.3 or earlier tests a boolean property 100 times though just one test is enough. (should got to appendix?)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Internal documentation}
 
-not useful for users
+There are some examples and properties which are not useful for users but useful for developers.
+One example is a corner case registered in a bug track system.
+We don't have a mechanism to integrate design, documentation and tests.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Solution}
 
-FIRST convey the idea.
+To specify visible behavior, we propose to use {\tt doctest} integrated with QuickCheck. To describe internal behavior, we propose to use {\tt hspec} aslo integrated with QuickCheck.
 
-\subsection{Tests suitable with doctest}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Visible behavior}
 
-\subsection{Tests suitable with hspec}
+talk about doctest.
 
+talk about the {\tt prop>} syntax.
+
+talk about how to specify non-boolean properties.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Internal behavior}
+
+TBD: talk about hspec.
+
+\begin{quote}
+\begin{verbatim}
+TBD: Example
+\end{verbatim}
+\end{quote}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Integrating multiple test suites}
 
+TBD: talk about Cabal.
+
+\begin{quote}
+\begin{verbatim}
+TBD: Example
+Test-Suite spec
+  Main-Is: Spec.hs
+  Type: exitcode-stdio-1.0
+  ...
+
+Test-Suite doctests
+  Main-Is: doctests.hs
+  Type: exitcode-stdio-1.0
+  ...
+\end{verbatim}
+\end{quote}
+
+\begin{quote}
+\begin{verbatim}
+TBD: Spec.hs comes here
+\end{verbatim}
+\end{quote}
+
+\begin{quote}
+\begin{verbatim}
+TBD: doctest.hs comes here
+\end{verbatim}
+\end{quote}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Best current practice}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Implementation}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{doctest}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{doctest and QuickCheck}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{hspec}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Related work}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusion}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \acks

### 2012-05-23 paper.tex
writing solution.

@@
 
-\noindent Some Haskellers like this kind of equations and dislike to write it with the example annotation as follow:
+\noindent Some Haskellers like this kind of equations and dislike to write it with the example annotation as follows:
 
@@
 
-talk about doctest.
+To execute examples in haddock annotations, we introduced {\tt doctest} from Python community and implemented it in Haskell.
 
-talk about the {\tt prop>} syntax.
+TBD: talk about exception.
 
-talk about how to specify non-boolean properties.
+To specify properties in haddock annotations, we propose a new syntax. In this paper, we use the {\tt prop>} prefix. We can write the boolean equation as follows:
+
+\begin{quote}
+\begin{verbatim}
+prop> Data.Map.null (empty) == True
+\end{verbatim}
+\end{quote}
+
+\noindent talk about how to specify non-boolean properties.
+
+\begin{quote}
+\begin{verbatim}
+prop> \xs -> sort xs == sort (sort (xs::[Int]))
+\end{verbatim}
+\end{quote}
+
+\noindent talk about open lambda term.
+
+\begin{quote}
+\begin{verbatim}
+prop> sort xs == sort (sort (xs::[Int]))
+\end{verbatim}
+\end{quote}
 

### 2012-05-23 paper.tex
change wording.

@@
 They tend to write less test programs than programmers of other
-programming languages.
-
-TBD: we need evidence to support above.
+programming languages. TBD: we need evidence to support this.
 
@@
 
-We can see the equation as a boolean property for {\tt null} and {\tt empty} and
+We can see the equation as a no-parameter property for {\tt null} and {\tt empty} and
 can say that there are potentially demands to show properties with Haddock annotation.
@@
 Even if a prefix keyword for properties is defined, 
-we have no agreement on how to describe non-boolean properties: e.g. how to specify types of funcation parameters. And we don't have automatic test mechanism to use properties in documentation.
-
-Note that QuickCheck 2.3 or earlier tests a boolean property 100 times though just one test is enough. (should got to appendix?)
+we have no agreement on how to describe parameterized properties: e.g. how to specify types of funcation parameters. And we don't have automatic test mechanism to use properties in documentation.
 
@@
 One example is a corner case registered in a bug track system.
+TBD: setup-teardown.
 We don't have a mechanism to integrate design, documentation and tests.
@@
 
+TBD: command and library. 
 TBD: talk about exception.
 
-To specify properties in haddock annotations, we propose a new syntax. In this paper, we use the {\tt prop>} prefix. We can write the boolean equation as follows:
+To specify properties in haddock annotations, we propose a new syntax. In this paper, we use the {\tt prop>} prefix. We can write the no-parameter equation as follows:
 
@@
 
-\noindent talk about how to specify non-boolean properties.
+\noindent Of course, we allowed parameterized properties by
+using closed lambda terms.
+Type signatures can be specified for each parameter:
 
@@
 
-\noindent talk about open lambda term.
+\noindent Since the right hand side of {\tt ->} is verbose as documentation,
+open lambda terms are also allowed:
 
@@
 \end{quote}
+
+\noindent To test properties in haddock annotations, we extended doctest to handle
+them with QuickCheck.
 
@@
 
-%% \appendix
-%% \section{Appendix Title}
+\appendix
 
-%% This is the text of the appendix, if you need one.
+\section{Package Survey}
+
+TBD
+
+\section{Improving QuickCheck}
+
+Since no-parameter properties has no parameters, it is enough to test them
+just one.
+However, QuickCheck 2.3 or earlier tests no-parameter properties 100 times.
+More precisely, the evaluation results are memorized and they are
+compared with {\tt True} 100 times.
+This will be fixed in the next release of QuickCheck.
 
@@
 We would like to thank
-Nick Smallbone for having improved QuickCheck so that boolean property is tested
-just once.
+Nick Smallbone for having improved QuickCheck so that no-parameter properties are tested just once.
 We are grateful to David Waern and Simon Marlow

### 2012-05-23 paper.tex
Add section on combining specs

@@
 
+\subsection{Combining multiple specs}
+
+\begin{quote}
+\emph{This is optional, so we can skip it, if necessary.}
+\end{quote}
+
+\begin{itemize}
+\item We create a spec file for each source file
+\item By convention, spec files end in Spec.hs
+\end{itemize}
+
+\noindent If we have a source tree like so:
+
+\begin{quote}
+\begin{verbatim}
+src/Foo.hs
+src/Foo/Bar.hs
+src/Baz.hs
+\end{verbatim}
+\end{quote}
+
+\noindent We would end up with the following spec files.
+
+\begin{quote}
+\begin{verbatim}
+test/FooSpec.hs
+test/Foo/BarSpec.hs
+test/BazSpec.hs
+\end{verbatim}
+\end{quote}
+
+\begin{itemize}
+\item We still want to combine those specs into a single spec.
+\item By convention we put the combined spec into test/Spec.hs
+\end{itemize}
+
+\begin{quote}
+\begin{verbatim}
+import Test.Hspec.Monadic
+
+import qualified FooSpec
+import qualified Foo.BarSpec
+import qualified BazSpec
+
+main :: IO ()
+main = hspecX $ do
+  describe "Foo"     FooSpec.spec
+  describe "Foo.Bar" Foo.BarSpec.spec
+  describe "Baz"     BazSpec.spec
+\end{verbatim}
+\end{quote}
+
+\begin{itemize}
+\item We use a preprocessor (http://hackage.haskell.org/package/hspec-discover) to automatically generate this boilerplate.
+\end{itemize}
+
+\begin{quote}
+\begin{verbatim}
+{-# OPTIONS_GHC -F -pgmF hspec-discover #-}
+\end{verbatim}
+\end{quote}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-05-24 paper.tex
Add two notes
re-structuring.

@@
 which ensures no type errors in programs if they can be complied.
-Thanks to this, programs written in Haskell have less bugs
+Thanks to this, it is said that programs written in Haskell have less bugs
 than other programming languages which do not provide such strong type systems.
+
+\begin{quote}
+    \emph{Would we need evidence to support the claim that "Haskell programs have less bugs"?}
+\end{quote}
 
@@
 Haskellers is, however, apt to rely on the strong type system too much.
-They tend to write less test programs than programmers of other
-programming languages. TBD: we need evidence to support this.
+They tend to write less test code than programmers of other
+programming languages.
 
-This also means that typical usage examples cannot be found from test programs.
+\begin{quote}
+    \emph{we need evidence to support this.}
+\end{quote}
+
+This also means that typical usage examples cannot be found from test code.
 We agree that type signatures are good abstract document but
@@
 
-Why do we dislike to write test programs?
-It is probably because writing test programs is not fun.
-To encourage Haskellers to write more test programs
+Why do we dislike to write test code?
+It is probably because writing test code is not fun.
+To encourage Haskellers to write more test code
 to detect value errors,
-we need a scheme to make writing test programs more interesting.
+we need a scheme to make writing test code more interesting.
+
+\begin{quote}
+    \emph{We should explain what is BDD around here}
+\end{quote}
 
@@
 \item[property] Invariants which functions have.
-\item[behavior] How do functions behave for users. Behavior can be shown by using both example and property.
+\item[behavior] How do functions behave for users. Behavior can be shown by using both examples and properties.
 \item[design] TBD
@@
 documented.
-So, let us first observe Haddock, the documentation system of Haskell. Haddock defines annotation syntax in Haskell's comment.
+So, let us first observe Haddock (version 2.7.2 or earlier), the documentation system of Haskell. 
+Haddock defines annotation syntax in Haskell's comment.
+
+Haddock provides code block annotations indicated with ({\tt >}).
+For example, the {\tt container} packages has the following code block annotation:
+
+\begin{quote}
+\begin{verbatim}
+-- | /O(1)/. Is the map empty?
+--
+-- > Data.Map.null (empty)           == True
+-- > Data.Map.null (singleton 1 'a') == False
+
+null :: Map k a -> Bool
+null Tip      = True
+null (Bin {}) = False
+\end{verbatim}
+\end{quote}
+
+TBD: code block cannot show results. properties should be distinguished from examples.
+
+TBD: Talk about test-framework. it integrates HUnit and QuickCheck but not 
+suitable for design.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Solution}
+
+To specify visible behavior, we propose to add two new annotation syntax to Haddock and to use {\tt doctest} integrated with QuickCheck.
+To describe internal behavior, we propose to use {\tt hspec} aslo integrated with QuickCheck.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Example annotation}
+\subsection{Visible behavior}
 
-We can specify examples with the annotation syntax {\tt >>>}.
-The following is an example of how to use the {\tt length} function
-for lists.
+Useful behavior should be documented in Haddock. 
+For this purpose, we propose the followings:
+
+\begin{itemize}
+\item Define a new Haddock annotation syntax for examples.
+\item Define a new Haddock annotation syntax for properties.
+\item Bring {\tt doctest} from the Python community to the Haskell
+community and integrate it with QuickCheck.
+\end{itemize}
+
+First, we introduce a new Haddock annotation syntax for the example.
+In this paper, we use {\tt >>>} which is the same as Python's one.
+Here is an example:
 
@@
 
-\noindent Example annotations are useful for users but 
-it is not used as test case from the developer point of view (before we implemented {\tt doctest}).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Property annotation}
-
-Haddock also provides code block annotations indicated with ({\tt >}).
-For example, the {\tt container} packages has the following code block annotation:
+\noindent Also, we can show exceptions.
 
@@
 \begin{verbatim}
-> Data.Map.null (empty) == True
+>>> head []
+*** Exception: Prelude.head: empty list
 \end{verbatim}
@@
 
-\noindent Some Haskellers like this kind of equations and dislike to write it with the example annotation as follows:
+To specify properties in Haddock annotations, we also propose a new syntax. In this paper, we use the {\tt prop>} prefix. We can write the no-parameter equation as follows:
 
@@
 \begin{verbatim}
->>> Data.Map.null (empty)
-True
-\end{verbatim}
-\end{quote}
-
-We can see the equation as a no-parameter property for {\tt null} and {\tt empty} and
-can say that there are potentially demands to show properties with Haddock annotation.
-To our experience, it is not difficult to find properties for pure functions.
-We believe that Haskellers should show properties of pure functions
-in documentation.
-
-But there is no annotation syntax to describe properties in Haddock.
-Even if a prefix keyword for properties is defined, 
-we have no agreement on how to describe parameterized properties: e.g. how to specify types of funcation parameters. And we don't have automatic test mechanism to use properties in documentation.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Internal documentation}
-
-There are some examples and properties which are not useful for users but useful for developers.
-One example is a corner case registered in a bug track system.
-TBD: setup-teardown.
-We don't have a mechanism to integrate design, documentation and tests.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Solution}
-
-To specify visible behavior, we propose to use {\tt doctest} integrated with QuickCheck. To describe internal behavior, we propose to use {\tt hspec} aslo integrated with QuickCheck.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Visible behavior}
-
-To execute examples in haddock annotations, we introduced {\tt doctest} from Python community and implemented it in Haskell.
-
-TBD: command and library. 
-TBD: talk about exception.
-
-To specify properties in haddock annotations, we propose a new syntax. In this paper, we use the {\tt prop>} prefix. We can write the no-parameter equation as follows:
-
-\begin{quote}
-\begin{verbatim}
-prop> Data.Map.null (empty) == True
+prop> Data.Map.null empty == True
 \end{verbatim}
@@
 
-\noindent To test properties in haddock annotations, we extended doctest to handle
-them with QuickCheck.
+Examples and properties in Haddock annotations are automatically
+tested by the {\tt doctest} command. We also provided a library
+to do this job. Its implementation will discussed in Section XXX.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Implementation}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Haddock}
+
+Haddock version 2.8 or later provides the {\tt >>>} syntax. 
+We extended Haddock to support the {\tt prop>} syntax by ourselves
+for proof of concept.
+The maintainers of Haddock agreed to introduce 
+the {\tt prop>} syntax and is now supporting it by their own.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{QuickCheck}
+
+Since no-parameter properties has no parameters, 
+it is enough to test them just one.
+However, QuickCheck 2.3 or earlier tests no-parameter properties 100 times.
+More precisely, the evaluation results are memorized and they are
+compared with {\tt True} 100 times.
+The current maintainer of QuickCheck modified that
+no-parameter properties are tested just once in its source repository.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{doctest}
+
+TBD(@sol): how to obtain examples and properties from Haddock annotations.
+Old version uses Haddock API but now uses GHC API directly. Why?
+
+TBD(@sol): escape from HUnit. Why?
+
+TBD(@sol): Old version executes one GHCi for each function examples.
+But new version executes one GHCI per module.
+Drastic speed up. Do We need a graph for execution time?
+
+TBD: how to handle error cases.
+
+TBD: how to integrate QuickCheck.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{hspec}
+
+TBD.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Best current practice}
 
-undefined.
+undefined. (The placeholders library provide better keyword such as {\tt notImplemented} and {\tt todo})
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Implementation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{doctest}
-
-how to handle error cases.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{doctest and QuickCheck}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{hspec}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Related work}
@@
 
-\section{Improving QuickCheck}
-
-Since no-parameter properties has no parameters, it is enough to test them
-just one.
-However, QuickCheck 2.3 or earlier tests no-parameter properties 100 times.
-More precisely, the evaluation results are memorized and they are
-compared with {\tt True} 100 times.
-This will be fixed in the next release of QuickCheck.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+We are grateful to David Waern and Simon Marlow
+for introducing two new syntax to Haddock.
 We would like to thank
 Nick Smallbone for having improved QuickCheck so that no-parameter properties are tested just once.
-We are grateful to David Waern and Simon Marlow
-for introducing the property syntax to Haddock.
 

### 2012-05-24 paper.tex
adding an example for BCP.

@@
 
-undefined. (The placeholders library provide better keyword such as {\tt notImplemented} and {\tt todo})
+First, we write a signature and leave the definition of the function undefined.
+
+\begin{quote}
+\begin{verbatim}
+rev :: [a] -> [a]
+rev = undefined
+\end{verbatim}
+\end{quote}
+
+\noindent Note that the placeholders library provide better keyword such as {\tt notImplemented} and {\tt todo}. Next we wrote documentation.
+
+\begin{quote}
+\begin{verbatim}
+-- |
+-- 'rev' @xs@ returns the elements of @xs@
+-- in rev order. @xs@ must be finite.
+-- 
+-- >>> rev [1,2,3]
+-- [3,2,1]
+--
+-- prop> rev [] == rev []
+-- prop> rev (xs ++ ys) == rev ys ++ rev (xs::[Int])
+-- prop> rev (rev xs) == (xs::[Int])
+
+rev :: [a] -> [a]
+rev = undefined
+\end{verbatim}
+\end{quote}
+
+\begin{figure*}
+\begin{verbatim}
+  describe "rev" $ do
+    it "return the first element in the last" $ property $
+        \xs -> not (null xs) ==> head (rev xs) == last (xs :: [Int])
+
+    it "return the last element in the first" $ property $
+        \xs -> not (null xs) ==> last (rev xs) == head (xs :: [Int])
+\end{verbatim}
+\end{figure*}
+
+\noindent We should run test in order to find errors in Haddock annotation.
+Then we implement the function.
+
+\begin{quote}
+\begin{verbatim}
+-- |
+-- 'rev' @xs@ returns the elements of @xs@
+-- in rev order. @xs@ must be finite.
+-- 
+-- >>> rev [1,2,3]
+-- [3,2,1]
+--
+-- prop> rev [] == rev []
+-- prop> rev (xs ++ ys) == rev ys ++ rev (xs::[Int])
+-- prop> rev (rev xs) == (xs::[Int])
+
+rev :: [a] -> [a]
+rev = foldl (flip (:)) []
+\end{verbatim}
+\end{quote}
+
+\noindent And test.
 

### 2012-05-24 paper.tex
writing implementation.

@@
 \begin{quote}
+\small
 \begin{verbatim}
@@
 tested by the {\tt doctest} command. We also provided a library
-to do this job. Its implementation will discussed in Section XXX.
+to do this job. Its implementation will discussed in Section~\ref{sec:doctest}.
 
@@
 \subsection{doctest}
+\label{sec:doctest}
 
@@
 
-TBD: how to handle error cases.
+Since test is done by comparing result string in Haddock notation
+and result string from GHCi, exceptions can be handled.
 
-TBD: how to integrate QuickCheck.
+To test properties, we also uses GHCi. 
+The `quickCheck'' string is prepended to each property
+and GHCi evaluates it.
+Success or failures is decided according to result string from GHCi.
+
+If properties contain free variables,
+we need to add lambda abstraction prefix before prepending
+the `quickCheck'' string.
+To check if properties contain free variables, 
+we also use GHCi for easy implementation at this moment.
+GHCi reports free variables if found:
+
+\begin{quote}
+\small
+\begin{verbatim}
+Prelude Data.List> sort xs == sort (sort (xs::[Int]))
+<interactive>:1:6: Not in scope: `xs'
+<interactive>:1:24: Not in scope: `xs'
+\end{verbatim}
+\end{quote}
+
+\noindent In this case, we should prepend the ``\verb|\xs -> |" string
+to the property.
 
@@
 \begin{figure*}
+\small
 \begin{verbatim}

### 2012-05-24 paper.tex
adding labels.

@@
 \item Introduced {\tt doctest} from Python community to Haskell community
-and integrated it with QuickCheck (Section XXX)
+and integrated it with QuickCheck
+(Section \ref{sec:Visible-behavior} and Section \ref{sec:doctest})
 \item Introduced {\tt hspec} from Ruby community to Haskell community and
-integrated it with QuickCheck (Section XXX)
-\item Show best current practice of design/documentation/tests (Section XXX)
+integrated it with QuickCheck
+(Section \ref{sec:Internal-behavior} and Section \ref{sec:hspec})
+\item Show best current practice of behavior driven development in Haskell (Section \ref{ref:Best-current-practice})
 \end{itemize}
@@
 \subsection{Visible behavior}
+\label{sec:Visible-behavior}
 
@@
 \subsection{Internal behavior}
+\label{sec:Internal-behavior}
 
@@
 \subsection{hspec}
+\label{sec:hspec}
 
@@
 \section{Best current practice}
+\label{ref:Best-current-practice}
 

### 2012-05-24 paper.tex
more precise explanation.
adding references.

@@
 integrated them with a Haskell-original test method.
-In particular, we introduced {\tt doctest} and
-integrated with QuickCheck. 
-We also implemented {\tt hspec} with QuickCheck enabled.
+In particular, we introduced {\tt doctest}~\cite{doctest} and
+integrated with QuickCheck~\cite{quickcheck}. 
+We also implemented {\tt hspec} inspired by rspec~\cite{rspec} 
+and integrated with QuickCheck.
 
@@
 
+TBD: behavior would not be a proper word since Haskell's pure
+functions are very static.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 documented.
-So, let us first observe Haddock (version 2.7.2 or earlier), the documentation system of Haskell. 
+So, let us first observe Haddock~\cite{haddock} (version 2.7.2 or earlier), the documentation system of Haskell. 
 Haddock defines annotation syntax in Haskell's comment.
@@
 To test properties, we also uses GHCi. 
-The `quickCheck'' string is prepended to each property
+Each property is enclosed with 
+the ``\verb|quickCheck (|'' and ``\verb|)|'' string
 and GHCi evaluates it.
@@
 
-\bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
-P. Q. Smith, and X. Y. Jones. ...reference text...
+\bibitem[Claessen et al. 2000]{quickcheck}
+Claessen K. and Hughes J., 
+``QuickCheck: a lightweight tool for random testing of Haskell programs'',
+ICFP '00 Proceedings of the fifth ACM SIGPLAN international conference on Functional programming,
+pp 268 -- 279,
+2000
+
+\bibitem[Marlow et al. 2010]{haddock}
+Marlow S. and Waern D.,
+``Haddock User Guide'',
+URL http://www.haskell.org/haddock/doc/html/,
+2010
+
+\bibitem[Python 2012]{doctest}
+Python Software Foundation, 
+``Python documentation'', 
+URL http://docs.python.org/library/doctest.html,
+2012
+
+\bibitem[Chelimsky et al. 2010]{rspec}
+Chelimsky D.,
+Astels D., Dennis Z., Hellesoy A., Helmkamp B., and North D.,
+``The RSpec Book: Behaviour-Driven 
+Development with RSpec, Cucumber, and Friends'',
+2010
+%% Hellesoy -- o/
 

### 2012-05-24 paper.tex
brushing up intro.

@@
 
-Why do we dislike to write test code?
-It is probably because writing test code is not fun.
-To encourage Haskellers to write more test code
-to detect value errors,
-we need a scheme to make writing test code more interesting.
+In other programming language communities, 
+behavior driven development (BDD) is getting much attention
+slowly replacing test driven development (TDD).
+In BDD, test code is written with specification terminology.
+In other words, specification can be used for automatic testing.
 
-\begin{quote}
-    \emph{We should explain what is BDD around here}
-\end{quote}
-
-For this purpose,
-we brought design methods from other programming language communities and
+In this paper, we consider how put BDD into practice in Haskell.
+We brought BDD methods from other programming language communities and
 integrated them with a Haskell-original test method.
@@
 and integrated with QuickCheck.
-
-In our method, design specification takes role of documentation for users and
-is used as test cases by its developer. 
-This method would motivate Haskellers
-bacause it has a kill-three-birds-with-one-stone aspect.
-
-%The purpose of this paper is to encourage Haskllers to write documentation, which can be used for testing, in their libraries.
-
-%%%% State your contributions.
 
@@
 
+\noindent Our {\tt doctest} implementation 
+can be used for design, documentation and automatic testing.
+We hope that this kill-three-birds-with-one-stone aspect 
+motives Haskellers to write test code.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{description}
+\item[developer] programmers who develop libraries
+\item[users] programmers who use libraries
 \item[unit test] tests for individual programs or libraries.
@@
 
+This section show best current practice of
+BDD in Haskell.
+
+\subsection{Procedure of BBD}
+
 First, we write a signature and leave the definition of the function undefined.
@@
 
-\noindent Note that the placeholders library provide better keyword such as {\tt notImplemented} and {\tt todo}. Next we wrote documentation.
+\noindent Note that the placeholders library provide better keyword such as {\tt notImplemented} and {\tt todo}. 
+Next we wrote visible behaviors in Haddock annotation:
 
@@
 
-\begin{figure*}
+\noindent We should execute {\tt doctest} to ensure no syntax errors.
+Also, we specifies internal behaviors with {\tt hspec}:
+
+\begin{quote}
 \small
@@
 describe "rev" $ do
-    it "return the first element in the last" $ property $
-        \xs -> not (null xs) ==> head (rev xs) == last (xs :: [Int])
+  it "returns the first element in the last" $ 
+    property $ \xs -> not (null xs) ==>
+      head (rev xs) == last (xs :: [Int])
 
-    it "return the last element in the first" $ property $
-        \xs -> not (null xs) ==> last (rev xs) == head (xs :: [Int])
+  it "returns the last element in the first" $ 
+    property $ \xs -> not (null xs) ==>
+      last (rev xs) == head (xs :: [Int])
 \end{verbatim}
-\end{figure*}
+\end{quote}
 
-\noindent We should run test in order to find errors in Haddock annotation.
+\noindent Also, we should execute this hspec module
+to ensure no syntax errors.
 Then we implement the function.
@@
 \begin{verbatim}
--- |
--- 'rev' @xs@ returns the elements of @xs@
--- in rev order. @xs@ must be finite.
--- 
--- >>> rev [1,2,3]
--- [3,2,1]
---
--- prop> rev [] == rev []
--- prop> rev (xs ++ ys) == rev ys ++ rev (xs::[Int])
--- prop> rev (rev xs) == (xs::[Int])
-
 rev :: [a] -> [a]
@@
 
-\noindent And test.
+\noindent Now we run test suites until
+all test failures disappear.
 
-The authors are not fundamentalist.
+Note that the authors are not fundamentalist.
 If functions to be defined are simple enough,
@@
 if Haskell compilers can infer correctly.
-Also, we can cut and paste typical examples and their results
-from Haskell interpreters to their document.
+Also, we can cut and paste examples and their results
+from Haskell interpreters to their Haddock annotations.
 

### 2012-05-24 paper.tex
writing problems.

@@
 
-In this paper, we consider how put BDD into practice in Haskell.
+In this paper, we consider how to put BDD into practice in Haskell.
 We brought BDD methods from other programming language communities and
@@
 \item[users] programmers who use libraries
-\item[unit test] tests for individual programs or libraries.
+%%\item[unit test] tests for individual programs or libraries.
 \item[example] Typical usage of functions.
@@
 \item[behavior] How do functions behave for users. Behavior can be shown by using both examples and properties.
-\item[design] TBD
 \item[documentation] Explanation of functions.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Observation}
+\section{Problems}
 
-We insist that examples and properties, which are useful for users, should be
-documented.
-So, let us first observe Haddock~\cite{haddock} (version 2.7.2 or earlier), the documentation system of Haskell. 
+In Haskell community, HUnit and QuickCheck are commonly used %% what about SmallCheck?
+and the {\tt test-framework} package integrates them. 
+We can categorize the {\tt test-framework} package into
+a tool of TDD.
+Before our work, there is no BDD tool commonly used in Haskell community.
+
+BDD tools used in other programming language communities are
+mainly based on example test. When we bring them to
+Haskell community, we should integrate them with properties test.
+
+We can categorize behaviors into two types: visible behaviors
+and internal behaviors.
+Visible behaviors are those which are useful for for both developers and users.
+By contrast, internal behaviors are those which are usefull only for developers.
+Behavior for a corner case registered in a bug track system is
+a good example of internal behaviors.
+
+BDD tools should be treat both of them and provide functionality
+to disclose visible behaviors as documentation for users.
+
+TBD: doctest and comments
+
+The standard documentation mechanism for Haskell is
+Haddock~\cite{haddock}.
 Haddock defines annotation syntax in Haskell's comment.
-
 Haddock provides code block annotations indicated with ({\tt >}).
@@
 
+Haddock version 2.7.2 or earlier does not provide annotation syntax 
+for examples and properties.
+
 TBD: code block cannot show results. properties should be distinguished from examples.
 
-TBD: Talk about test-framework. it integrates HUnit and QuickCheck but not 
-suitable for design.
+Here are summary of the problems:
+
+\begin{itemize}
+\item Haskell community does not have common syntax to describe
+behaviors in Haddock.
+\item Haskell community does not have a BBD tool for visible behaviors.
+\item Haskell community does not have a BBD tool for internal behaviors.
+\end{itemize}
 

### 2012-05-24 paper.tex
writing problems.

@@
 We hope that this kill-three-birds-with-one-stone aspect 
-motives Haskellers to write test code.
+motivates Haskellers to write test code.
 
@@
 BDD tools used in other programming language communities are
-mainly based on example test. When we bring them to
+mainly based on example test
+and do not support property test.
+To our experience, it is not difficult to find properties for 
+pure Haskell functions.
+So, when we bring BDD tools to
 Haskell community, we should integrate them with properties test.
@@
 and internal behaviors.
-Visible behaviors are those which are useful for for both developers and users.
+Visible behaviors are those which are useful for both developers and users.
 By contrast, internal behaviors are those which are usefull only for developers.
@@
 to disclose visible behaviors as documentation for users.
+If we can archive this, BDD tools can be used for
+triple purpose: design, documentation and automatic testing.
 
-TBD: doctest and comments
+A good candidate for this is {\tt doctest} in Python community.
+It extends Python's documentation strings so that they
+can contain examples. 
+Examples can be tested by the Python interpreter.
 
@@
 
-Haddock version 2.7.2 or earlier does not provide annotation syntax 
+TBD: code block cannot show results. properties should be distinguished from examples.
+
+TBD: Haddock version 2.7.2 or earlier does not provide annotation syntax 
 for examples and properties.
-
-TBD: code block cannot show results. properties should be distinguished from examples.
 

### 2012-05-24 paper.tex
adding annotation.
Use "correct semantic" instead of "no type error"
Introduction: wording

@@
 
-One of the best features of Haskell is its strong type system, 
-which ensures no type errors in programs if they can be complied.
-Thanks to this, it is said that programs written in Haskell have less bugs
-than other programming languages which do not provide such strong type systems.
+Haskell has a very strong type system, which ensures that most type errors are
+detected at compile time rather than runtime.  Thanks to this, we assume that
+programs written in Haskell have less bugs than programs written in programming
+languages that do not provide such a strong type system.
 
@@
 
-Unfortunately, the strong type system of Haskell does not guarantee
-no value errors.
+Unfortunately, the strong type system of Haskell does not guarantee correct
+semantic.
 Haskellers is, however, apt to rely on the strong type system too much.
@@
 
-TBD: talk about hspec.
+TBD(@sol): talk about hspec.
 
@@
 
-TBD: talk about Cabal.
+TBD(@sol): talk about Cabal.
 

### 2012-05-24 paper.tex
Introduction: wording
Remove trailing whitespace
Describe hspec

@@
 \usepackage{amsmath}
+
+\newcommand{\hspec}{Hspec}
 
@@
 
+Unfortunately, the strong type system of Haskell does not guarantee correct
+semantic.
+
 \begin{quote}
-    \emph{Would we need evidence to support the claim that "Haskell programs have less bugs"?}
+    \emph{TBD: What about dependent types and GHC's new features?}
 \end{quote}
 
-Unfortunately, the strong type system of Haskell does not guarantee correct
-semantic.
-Haskellers is, however, apt to rely on the strong type system too much.
-They tend to write less test code than programmers of other
-programming languages.
+Haskell programmers, however, tend to rely on the type system too much.  They
+tend to write less test code than programmers of other programming languages.
 
@@
 
-TBD(@sol): talk about hspec.
+If we have some behavior that does not fit into the user documentation, but
+would rather be part of a design document, we use \hspec{}.
+
+\hspec{} provides an EDSL for defining specs.  A spec organizes behavior in a
+tree structure, and it is defined with \texttt{it} and \texttt{describe}.
 
@@
 \begin{verbatim}
-TBD: Example
+spec :: Specs
+spec = do
+  describe "reverse" $ do
+    it "reverses a list" $
+      reverse [1 :: Int, 2, 3] @?= [3, 2, 1]
+
+    it "gives the original list, if applied twice" $ property $
+      \xs -> reverse (reverse xs) == (xs :: [Int])
 \end{verbatim}
 \end{quote}
+
+\noindent The {\tt it} function combines a text description for a behavior and
+a property or an example that gives some confidence in that behavior into a
+spec.
+Examples can be arbitrary HUnit assertions.  Properties must be of type {\tt
+Property}; QuickCheck's \texttt{property} function can be used to turn
+arbitrary properties inot a \texttt{Property}.
+
+The \texttt{describe} function combines a list of specs into a larger spec.  It
+can be used to organize a spec as a tree.
+Internally, a \texttt{Writer} is used to build up specs in a monadic way.
+
+A spec can be run with the \texttt{hspecX} function.  When a spec is run, all
+properties and examples are evaluated, and failing properties and examples are
+reported to the user (FIXME: user does not fit our definition of \emph{user}
+here!).
 

### 2012-05-24 paper.tex
Add code examples for "integrating multiple test suites"

@@
 \usepackage{amsmath}
+\usepackage{url}
 
@@
 \begin{verbatim}
+import Test.Hspec.Monadic
 spec :: Specs
@@
 
-TBD(@sol): talk about Cabal.
+\begin{itemize}
+
+\item @kazu: Can you add a reference to:
+\url{http://www.haskell.org/cabal/users-guide/developing-packages.html#test-suites}
+\item @kazu: should this subsection go into "best current practice"?
+
+\item It is important that testing is automated, including running of test
+    suites.  Where should we put this?
+\end{itemize}
+
+\noindent Cabals test feature can be used to automatically run tests.  A cabal
+file can contain multiple test suites, so it is possible to define test suites
+for Doctest and \hspec{} in a single cabal file.
+
+A driver for Hspec looks like so:
 
@@
 \begin{verbatim}
-TBD: Example
-Test-Suite spec
-  Main-Is: Spec.hs
-  Type: exitcode-stdio-1.0
-  ...
-
-Test-Suite doctests
-  Main-Is: doctests.hs
-  Type: exitcode-stdio-1.0
-  ...
+-- file Spec.hs
+import Test.Hspec.Monadic
+main = hspecX spec
 \end{verbatim}
 \end{quote}
+
+\noindent And a corresponding Cabal test suite section like so:
 
@@
 \begin{verbatim}
-TBD: Spec.hs comes here
+test-suite spec
+  type:          exitcode-stdio-1.0
+  main-is:       Spec.hs
+  build-depends: base, hspec
 \end{verbatim}
 \end{quote}
+
+\noindent A driver for Doctest looks like so:
 
@@
 \begin{verbatim}
-TBD: doctest.hs comes here
+-- file doctests.hs
+import Test.DocTest
+main = doctest ["src/Main.hs"]
+\end{verbatim}
+\end{quote}
+
+\noindent And a corresponding Cabal test suite section like so:
+
+\begin{quote}
+\small
+\begin{verbatim}
+test-suite doctests
+  type:          exitcode-stdio-1.0
+  main-is:       doctests.hs
+  build-depends: base, doctest
 \end{verbatim}

### 2012-05-24 paper.tex
Add a note

@@
 This also means that typical usage examples cannot be found from test code.
-We agree that type signatures are good abstract document but
-there are many functions whose behaviors cannot be guessed from
-their names and type signatures.
+
+We agree that type signatures are good abstract document
+
+\begin{quote}
+    \emph{Should we quote "Theorems for free!" by P. Wadler here?}
+\end{quote}
+
+but there are many
+functions whose behaviors cannot be guessed from their names and type
+signatures.
 

### 2012-05-24 paper.tex
Add an other commet

@@
 \label{sec:Visible-behavior}
+
+\begin{quote}
+    @kazu: I think all behavior is user visible, in the way that the user can
+    observe it when he uses the function.  Otherwise we would not be interested
+    in that behavior at all.  What do you think about using \emph{Behavior as
+    user documentation} and \emph{Behavior as code documentation} instead of
+    visible/internal behavior?  Or can we come up with something else?
+\end{quote}
 

### 2012-05-25 paper.tex
minor fix.

@@
 By contrast, internal behaviors are those which are usefull only for developers.
-Behavior for a corner case registered in a bug track system is
-a good example of internal behaviors.
+A good example of internal behaviors
+is behavior for a corner case registered in a bug track system.
 
@@
 
-Since no-parameter properties has no parameters,
+Since no-parameter properties have no parameters,
 it is enough to test them just one.

### 2012-05-25 paper.tex
minor fix.

@@
 \noindent Note that the placeholders library provide better keyword such as {\tt notImplemented} and {\tt todo}.
-Next we wrote visible behaviors in Haddock annotation:
+Next we write visible behaviors in Haddock annotation:
 
@@
 And their signatures can be automatically
-inserted by IDE (like ghc-mod) later
+inserted by IDE (such as ghc-mod) later
 if Haskell compilers can infer correctly.

### 2012-05-25 paper.tex
credit to adding new two syntax to Haddock goes to us.

@@
 
-Haddock version 2.8 or later provides the {\tt >>>} syntax.
-We extended Haddock to support the {\tt prop>} syntax by ourselves
-for proof of concept.
-The maintainers of Haddock agreed to introduce
-the {\tt prop>} syntax and is now supporting it by their own.
+We implemented the {\tt >>>} syntax and it
+is available in Haddock version 2.8 or later.
+We also extended Haddock to support the {\tt prop>} syntax
+which will be available in the future release.
 
@@
 
-We are grateful to David Waern and Simon Marlow
-for introducing two new syntax to Haddock.
+We are grateful to David Waern, 
+one of the Haddock maintainers, 
+for reviewing two new syntax to Haddock.
 We would like to thank

### 2012-05-25 paper.tex
adding open problem sections and moving cabal to BCP section.
add the "free!" paper and nub.

@@
 
-We agree that type signatures are good abstract document
-
-\begin{quote}
-    \emph{Should we quote "Theorems for free!" by P. Wadler here?}
-\end{quote}
-
+We agree that type signatures are good abstract document~\cite{free}
 but there are many
 functions whose behaviors cannot be guessed from their names and type
-signatures.
+signatures. A good example is \verb|nub :: [a] -> [a]| in 
+the Data.List module.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Integrating multiple test suites}
-
-\begin{itemize}
-
-\item @kazu: Can you add a reference to:
-\url{http://www.haskell.org/cabal/users-guide/developing-packages.html#test-suites}
-\item @kazu: should this subsection go into "best current practice"?
-
-\item It is important that testing is automated, including running of test
-    suites.  Where should we put this?
-\end{itemize}
-
-\noindent Cabals test feature can be used to automatically run tests.  A cabal
-file can contain multiple test suites, so it is possible to define test suites
-for Doctest and \hspec{} in a single cabal file.
-
-A driver for Hspec looks like so:
-
-\begin{quote}
-\small
-\begin{verbatim}
--- file Spec.hs
-import Test.Hspec.Monadic
-main = hspecX spec
-\end{verbatim}
-\end{quote}
-
-\noindent And a corresponding Cabal test suite section like so:
-
-\begin{quote}
-\small
-\begin{verbatim}
-test-suite spec
-  type:          exitcode-stdio-1.0
-  main-is:       Spec.hs
-  build-depends: base, hspec
-\end{verbatim}
-\end{quote}
-
-\noindent A driver for Doctest looks like so:
-
-\begin{quote}
-\small
-\begin{verbatim}
--- file doctests.hs
-import Test.DocTest
-main = doctest ["src/Main.hs"]
-\end{verbatim}
-\end{quote}
-
-\noindent And a corresponding Cabal test suite section like so:
-
-\begin{quote}
-\small
-\begin{verbatim}
-test-suite doctests
-  type:          exitcode-stdio-1.0
-  main-is:       doctests.hs
-  build-depends: base, doctest
-\end{verbatim}
-\end{quote}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Integrating multiple test suites}
+
+\begin{itemize}
+
+\item @kazu: Can you add a reference to:
+\url{http://www.haskell.org/cabal/users-guide/developing-packages.html#test-suites}
+\item @kazu: should this subsection go into "best current practice"?
+
+\item It is important that testing is automated, including running of test
+    suites.  Where should we put this?
+\end{itemize}
+
+\noindent Cabals test feature can be used to automatically run tests.  A cabal
+file can contain multiple test suites, so it is possible to define test suites
+for Doctest and \hspec{} in a single cabal file.
+
+A driver for Hspec looks like so:
+
+\begin{quote}
+\small
+\begin{verbatim}
+-- file Spec.hs
+import Test.Hspec.Monadic
+main = hspecX spec
+\end{verbatim}
+\end{quote}
+
+\noindent And a corresponding Cabal test suite section like so:
+
+\begin{quote}
+\small
+\begin{verbatim}
+test-suite spec
+  type:          exitcode-stdio-1.0
+  main-is:       Spec.hs
+  build-depends: base, hspec
+\end{verbatim}
+\end{quote}
+
+\noindent A driver for Doctest looks like so:
+
+\begin{quote}
+\small
+\begin{verbatim}
+-- file doctests.hs
+import Test.DocTest
+main = doctest ["src/Main.hs"]
+\end{verbatim}
+\end{quote}
+
+\noindent And a corresponding Cabal test suite section like so:
+
+\begin{quote}
+\small
+\begin{verbatim}
+test-suite doctests
+  type:          exitcode-stdio-1.0
+  main-is:       doctests.hs
+  build-depends: base, doctest
+\end{verbatim}
+\end{quote}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Related work}
+\section{Open problems and related work}
 
@@
 
+\bibitem[Wadler 1989]{free}
+Wadler P.,
+``Theorems for free!'',
+In proceeding of 4th International Conference on Functional Programming and Computer Architecture, 1989.
+
 \bibitem[Claessen et al. 2000]{quickcheck}

### 2012-05-25 paper.tex
adding excuse.

@@
 
-TBD: behavior would not be a proper word since Haskell's pure
-functions are very static.
+The word ``behavior'' is associated with actions.
+Since pure functions of Haskell are static,
+this word might be improper for Haskell.
+However, to avoid to define a new terminology,
+we use ``behavior'' in this paper.
 

### 2012-05-25 paper.tex
brushing up.

@@
 
-\keywords Haddock, QuickCheck, doctest, hspec
+\keywords Haddock, QuickCheck, doctest, Hspec
 
@@
 integrated with QuickCheck~\cite{quickcheck}.
-We also implemented {\tt hspec} inspired by rspec~\cite{rspec}
+We also implemented {\tt Hspec} inspired by rspec~\cite{rspec}
 and integrated with QuickCheck.
@@
 (Section \ref{sec:Visible-behavior} and Section \ref{sec:doctest})
-\item Introduced {\tt hspec} from Ruby community to Haskell community and
+\item Introduced {\tt Hspec} from Ruby community to Haskell community and
 integrated it with QuickCheck
@@
 To specify visible behavior, we propose to add two new annotation syntax to Haddock and to use {\tt doctest} integrated with QuickCheck.
-To describe internal behavior, we propose to use {\tt hspec} aslo integrated with QuickCheck.
+To describe internal behavior, we propose to use {\tt Hspec} aslo integrated with QuickCheck.
 
@@
 \noindent Since the right hand side of {\tt ->} is verbose as documentation,
-open lambda terms are also allowed:
+lambda terms without the lambda prefix are also allowed:
 
@@
 
+\begin{quote}
+    \emph{TBD: provide an example?}
+\end{quote}
+
 A spec can be run with the \texttt{hspecX} function.  When a spec is run, all
@@
 here!).
+
+\begin{quote}
+    \emph{TBD: should user be replaced with developer?}
+\end{quote}
 
@@
 which will be available in the future release.
+Both examples and properties in Haddock annotation are pretty printed when producing document
+such as in HTML.
 
@@
 compared with {\tt True} 100 times.
-The current maintainer of QuickCheck modified that
+The current maintainer of QuickCheck modified so that
 no-parameter properties are tested just once in its source repository.
+This feature will be available in the future release.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{hspec}
+\subsection{Hspec}
 \label{sec:hspec}
@@
 BDD in Haskell.
+
+\begin{quote}
+    \emph{TBD: one more example relating to IO is desired.}
+\end{quote}
 
@@
 \noindent We should execute {\tt doctest} to ensure no syntax errors.
-Also, we specifies internal behaviors with {\tt hspec}:
+Also, we specifies internal behaviors with {\tt Hspec}:
 
@@
 
-\noindent Also, we should execute this hspec module
+\noindent Also, we should execute this Hspec module
 to ensure no syntax errors.
@@
 And their signatures can be automatically
-inserted by IDE (such as ghc-mod) later
+inserted by IDE (such as Emacs/vim with {\tt ghc-mod}) later
 if Haskell compilers can infer correctly.
@@
 \emph{This is optional, so we can skip it, if necessary.}
+\end{quote}
+
+\begin{quote}
+\emph{How about moving this to appendix?}
 \end{quote}
@@
 \url{http://www.haskell.org/cabal/users-guide/developing-packages.html#test-suites}
-\item @kazu: should this subsection go into "best current practice"?
 

### 2012-05-25 paper.tex
brushing up.

@@
 
-TBD: code block cannot show results. properties should be distinguished from examples.
-
-TBD: Haddock version 2.7.2 or earlier does not provide annotation syntax
-for examples and properties.
+\noindent Though the code block itslef is a necessary feature,
+we think this kind usage has problems.
+If the equations above are examples, we cannot distinguish
+their results from their expression to be evaluated.
+If they are properties, they should be distinguished
+from example code.
 
@@
 
-We implemented the {\tt >>>} syntax and it
-is available in Haddock version 2.8 or later.
-We also extended Haddock to support the {\tt prop>} syntax
+As we described, Haddock version 2.7.2 or earlier has
+anotation syntax for code block but does not provide
+annotation syntax for examples and properties.
+We implemented annotation syntax for examples ({\tt >>>}) and 
+it is available in Haddock version 2.8 or later.
+We also extended Haddock to support annotation syntax for
+properties ({\tt prop>})
 which will be available in the future release.

### 2012-05-25 paper.tex
typo fix.
Wording

@@
 In other programming language communities,
-behavior driven development (BDD) is getting much attention
+behavior driven development (BDD) is getting much attention, and is
 slowly replacing test driven development (TDD).
@@
 
-In this paper, we consider how to put BDD into practice in Haskell.
+In this paper, we explore how to apply BDD to Haskell.
 We brought BDD methods from other programming language communities and
@@
 -- 'rev' @xs@ returns the elements of @xs@
--- in rev order. @xs@ must be finite.
+-- in reverse order. @xs@ must be finite.
 --

### 2012-05-25 paper.tex
going without Hspec credit.

@@
            {TBD: Email}
-\authorinfo{Trystan SPANGLER}
-           {TBD: Affiliation}
-           {TBD: Email}
+%% \authorinfo{Trystan SPANGLER}
+%%            {TBD: Affiliation}
+%%            {TBD: Email}
 \authorinfo{Kazuhiko YAMAMOTO}
@@
 In this paper, we explore how to apply BDD to Haskell.
-We brought BDD methods from other programming language communities and
-integrated them with a Haskell-original test method.
+We brought a test method from other programming language community and
+integrated it with a Haskell-original test method.
 In particular, we introduced {\tt doctest}~\cite{doctest} and
-integrated with QuickCheck~\cite{quickcheck}.
-We also implemented {\tt Hspec} inspired by rspec~\cite{rspec}
-and integrated with QuickCheck.
+integrated with QuickCheck~\cite{quickcheck} as a BDD tool.
+We also make use of \hspec{} inspired by rspec~\cite{rspec}.
 
@@
 (Section \ref{sec:Visible-behavior} and Section \ref{sec:doctest})
-\item Introduced {\tt Hspec} from Ruby community to Haskell community and
-integrated it with QuickCheck
-(Section \ref{sec:Internal-behavior} and Section \ref{sec:hspec})
-\item Show best current practice of behavior driven development in Haskell (Section \ref{ref:Best-current-practice})
+%% \item Introduced {\tt Hspec} from Ruby community to Haskell community and
+%% integrated it with QuickCheck
+%% (Section \ref{sec:Internal-behavior} and Section \ref{sec:hspec})
+\item Show best current practice of behavior driven development in Haskell using {\tt doctest} and \hspec{} (Section \ref{ref:Best-current-practice})
 \end{itemize}
@@
 a tool of TDD.
-Before our work, there is no BDD tool commonly used in Haskell community.
+Before our work and \hspec{}, there is no BDD tool commonly used in Haskell community.
 
@@
 To specify visible behavior, we propose to add two new annotation syntax to Haddock and to use {\tt doctest} integrated with QuickCheck.
-To describe internal behavior, we propose to use {\tt Hspec} aslo integrated with QuickCheck.
+To describe internal behavior, we propose to use \hspec{} aslo integrated with QuickCheck.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Hspec}
-\label{sec:hspec}
+%% \subsection{Hspec}
+%% \label{sec:hspec}
 
-TBD.
+%% TBD.
 
@@
 \noindent We should execute {\tt doctest} to ensure no syntax errors.
-Also, we specifies internal behaviors with {\tt Hspec}:
+Also, we specifies internal behaviors with \hspec{}:
 
@@
 
-\noindent Also, we should execute this Hspec module
+\noindent Also, we should execute this \hspec{} module
 to ensure no syntax errors.
@@
 
-A driver for Hspec looks like so:
+A driver for \hspec{} looks like so:
 

### 2012-05-25 paper.tex
s/annotation/markup/

@@
 Haddock~\cite{haddock}.
-Haddock defines annotation syntax in Haskell's comment.
-Haddock provides code block annotations indicated with ({\tt >}).
-For example, the {\tt container} packages has the following code block annotation:
+Haddock defines markup syntax in Haskell's comment.
+Haddock provides code block markup indicated with ({\tt >}).
+For example, the {\tt container} packages has the following code block markup:
 
@@
 
-To specify visible behavior, we propose to add two new annotation syntax to Haddock and to use {\tt doctest} integrated with QuickCheck.
+To specify visible behavior, we propose to add two new markups to Haddock and to use {\tt doctest} integrated with QuickCheck.
 To describe internal behavior, we propose to use \hspec{} aslo integrated with QuickCheck.
@@
 \begin{itemize}
-\item Define a new Haddock annotation syntax for examples.
-\item Define a new Haddock annotation syntax for properties.
+\item Define a new Haddock markup for examples.
+\item Define a new Haddock markup for properties.
 \item Bring {\tt doctest} from the Python community to the Haskell
@@
 
-First, we introduce a new Haddock annotation syntax for the example.
+First, we introduce a new Haddock markup for the example.
 In this paper, we use {\tt >>>} which is the same as Python's one.
@@
 
-To specify properties in Haddock annotations, we also propose a new syntax. In this paper, we use the {\tt prop>} prefix. We can write the no-parameter equation as follows:
+To specify properties, we also propose a new markup. In this paper, we use the {\tt prop>} prefix. We can write the no-parameter equation as follows:
 
@@
 
-Examples and properties in Haddock annotations are automatically
+Examples and properties in Haddock markup are automatically
 tested by the {\tt doctest} command. We also provided a library
@@
 As we described, Haddock version 2.7.2 or earlier has
-anotation syntax for code block but does not provide
-annotation syntax for examples and properties.
-We implemented annotation syntax for examples ({\tt >>>}) and 
+the markup for code block but does not provide
+markups for examples and properties.
+We implemented a markup for examples ({\tt >>>}) and 
 it is available in Haddock version 2.8 or later.
-We also extended Haddock to support annotation syntax for
+We also extended Haddock to support a markup for
 properties ({\tt prop>})
 which will be available in the future release.
-Both examples and properties in Haddock annotation are pretty printed when producing document
+Both examples and properties in Haddock markup are pretty printed when producing document
 such as in HTML.
@@
 
-TBD(@sol): how to obtain examples and properties from Haddock annotations.
+TBD(@sol): how to obtain examples and properties from Haddock markup.
 Old version uses Haddock API but now uses GHC API directly. Why?
@@
 \noindent Note that the placeholders library provide better keyword such as {\tt notImplemented} and {\tt todo}.
-Next we write visible behaviors in Haddock annotation:
+Next we write visible behaviors in Haddock markup:
 
@@
 Also, we can cut and paste examples and their results
-from Haskell interpreters to their Haddock annotations.
+from Haskell interpreters to their Haddock markup.
 
@@
 one of the Haddock maintainers, 
-for reviewing two new syntax to Haddock.
+for reviewing two new markups to Haddock.
 We would like to thank

### 2012-05-25 paper.tex
s/visible/for users/ s/internal/for developers/

@@
 and integrated it with QuickCheck
-(Section \ref{sec:Visible-behavior} and Section \ref{sec:doctest})
+(Section \ref{sec:for-users} and Section \ref{sec:doctest})
 %% \item Introduced {\tt Hspec} from Ruby community to Haskell community and
 %% integrated it with QuickCheck
-%% (Section \ref{sec:Internal-behavior} and Section \ref{sec:hspec})
+%% (Section \ref{sec:for-developers} and Section \ref{sec:hspec})
 \item Show best current practice of behavior driven development in Haskell using {\tt doctest} and \hspec{} (Section \ref{ref:Best-current-practice})
@@
 
-We can categorize behaviors into two types: visible behaviors
-and internal behaviors.
-Visible behaviors are those which are useful for both developers and users.
-By contrast, internal behaviors are those which are usefull only for developers.
-A good example of internal behaviors
-is behavior for a corner case registered in a bug track system.
+We can categorize examples/properties into two types: for users and
+and for developers.
+Examples/properties for users are those which are useful as user documentation.
+By contrast, examples/properties for developers are
+those which are not useful as user documentation.
+A good example of examples/properties for developers
+is a corner example registered in a bug track system.
 
 BDD tools should be treat both of them and provide functionality
-to disclose visible behaviors as documentation for users.
+to disclose examples/properties for users as user documentation.
 If we can archive this, BDD tools can be used for
@@
 Haddock defines markup syntax in Haskell's comment.
-Haddock provides code block markup indicated with ({\tt >}).
+Haddock provides the code block markup indicated with ({\tt >}).
 For example, the {\tt container} packages has the following code block markup:
@@
 behaviors in Haddock.
-\item Haskell community does not have a BBD tool for visible behaviors.
-\item Haskell community does not have a BBD tool for internal behaviors.
+\item Haskell community does not have a BBD tool for 
+examples/properties for users.
 \end{itemize}
@@
 
-To specify visible behavior, we propose to add two new markups to Haddock and to use {\tt doctest} integrated with QuickCheck.
-To describe internal behavior, we propose to use \hspec{} aslo integrated with QuickCheck.
+To specify examples/properties for users, we propose to add two new markups to Haddock and to use {\tt doctest} integrated with QuickCheck.
+To describe examples/properties for developers, we propose to use \hspec{} also integrated with QuickCheck.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Visible behavior}
-\label{sec:Visible-behavior}
+\subsection{Examples/properties for users}
+\label{sec:for-users}
 
-\begin{quote}
-    @kazu: I think all behavior is user visible, in the way that the user can
-    observe it when he uses the function.  Otherwise we would not be interested
-    in that behavior at all.  What do you think about using \emph{Behavior as
-    user documentation} and \emph{Behavior as code documentation} instead of
-    visible/internal behavior?  Or can we come up with something else?
-\end{quote}
-
-Useful behavior should be documented in Haddock.
+Examples/properties for users should be documented with Haddock markup syntax.
 For this purpose, we propose the followings:
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Internal behavior}
-\label{sec:Internal-behavior}
+\subsection{Examples/properties for developers}
+\label{sec:for-developers}
 
@@
 \noindent Note that the placeholders library provide better keyword such as {\tt notImplemented} and {\tt todo}.
-Next we write visible behaviors in Haddock markup:
+Next we write examples/properties for users in Haddock markup:
 
@@
 \noindent We should execute {\tt doctest} to ensure no syntax errors.
-Also, we specifies internal behaviors with \hspec{}:
+Also, we specifies examples/properties for developers with \hspec{}:
 

### 2012-05-25 paper.tex
adding note.
rephrasing parameter.

@@
 %%\item[unit test] tests for individual programs or libraries.
-\item[example] Typical usage of functions.
+\item[example] Typical usage of functions. (TBD(@sol): more vocal)
 \item[property] Invariants which functions have.
@@
 
-To specify properties, we also propose a new markup. In this paper, we use the {\tt prop>} prefix. We can write the no-parameter equation as follows:
+To specify properties, we also propose a new markup. In this paper, we use the {\tt prop>} prefix. We can write properties which do not take parameters as follows:
 
@@
 
-\noindent Of course, we allowed parameterized properties by
+\noindent Of course, we allowed properties which take parameters by
 using closed lambda terms.
@@
 
-Since no-parameter properties have no parameters,
-it is enough to test them just one.
-However, QuickCheck 2.3 or earlier tests no-parameter properties 100 times.
+It is enough to test properties which do not have parameters just one.
+However, QuickCheck 2.3 or earlier tests 
+properties which do not have parameters 100 times.
 More precisely, the evaluation results are memorized and they are
@@
 The current maintainer of QuickCheck modified so that
-no-parameter properties are tested just once in its source repository.
+properties which do not have parameters
+are tested just once in its source repository.
 This feature will be available in the future release.
@@
 \subsection{Combining multiple specs}
+
+\begin{quote}
+    \emph{should emphasize ``automatically discover tests''}
+\end{quote}
+
+\begin{quote}
+    \emph{talk about CI}
+\end{quote}
 
@@
 We would like to thank
-Nick Smallbone for having improved QuickCheck so that no-parameter properties are tested just once.
+Nick Smallbone for having improved QuickCheck so that 
+properties which do not have parameters are tested just once.
 

### 2012-05-25 paper.tex
brush up.

@@
 integrated with QuickCheck~\cite{quickcheck} as a BDD tool.
-We also make use of \hspec{} inspired by rspec~\cite{rspec}.
+We also make use of \hspec{} inspired by rspec~\cite{rspec} for
+BDD in Haskell.
 
@@
 %% (Section \ref{sec:for-developers} and Section \ref{sec:hspec})
-\item Show best current practice of behavior driven development in Haskell using {\tt doctest} and \hspec{} (Section \ref{ref:Best-current-practice})
+\item Show best current practice of BDD in Haskell using {\tt doctest} and \hspec{} (Section \ref{ref:Best-current-practice})
 \end{itemize}
@@
 BDD tools should be treat both of them and provide functionality
-to disclose examples/properties for users as user documentation.
+to extract examples/properties for users as user documentation.
 If we can archive this, BDD tools can be used for
@@
 
+To describe examples/properties for developers, we can use \hspec{}.
+
 Here are summary of the problems:
@@
 \item Haskell community does not have common syntax to describe
-behaviors in Haddock.
-\item Haskell community does not have a BBD tool for 
+examples/properties with Haddock markup syntax.
+\item Haskell community does not have a BDD tool for 
 examples/properties for users.
@@
 
-Examples and properties in Haddock markup are automatically
+Examples and properties in comments are automatically
 tested by the {\tt doctest} command. We also provided a library
@@
 
-If we have some behavior that does not fit into the user documentation, but
+\begin{quote}
+    \emph{TBD: Behavior should be replaced with examples/properties in this subsection?}
+\end{quote}
+
+If we have examples/properties that does not fit into the user documentation, but
 would rather be part of a design document, we use \hspec{}.
@@
 
+\begin{quote}
+    \emph{TBD: describe the algorithm}
+\end{quote}
+
 \noindent In this case, we should prepend the ``\verb|\xs -> |" string
@@
 
-\subsection{Procedure of BBD}
+\subsection{Procedure of BDD}
 
@@
 
-\noindent Note that the placeholders library provide better keyword such as {\tt notImplemented} and {\tt todo}.
+\noindent Note that the {\tt placeholders} library provide better keyword such as {\tt notImplemented} and {\tt todo}.
 Next we write examples/properties for users in Haddock markup:
@@
 if Haskell compilers can infer correctly.
-Also, we can cut and paste examples and their results
+Also, we can copy-and-paste examples and their results
 from Haskell interpreters to their Haddock markup.

### 2012-05-25 paper.tex
Add affiliation
Minor cleanup of "terminology" section.

@@
 \authorinfo{Simon HENGEL}
-           {TBD: Affiliation}
-           {TBD: Email}
+           {Freelance Software Developer}
+           {sol@typeful.net}
 %% \authorinfo{Trystan SPANGLER}
@@
 \begin{description}
-\item[developer] programmers who develop libraries
-\item[users] programmers who use libraries
+\item[developer]
+    A programmer who develops a library.
+\item[user]
+    A programmer who uses a library.
 %%\item[unit test] tests for individual programs or libraries.
-\item[example] Typical usage of functions. (TBD(@sol): more vocal)
-\item[property] Invariants which functions have.
-\item[behavior] How do functions behave for users. Behavior can be shown by using both examples and properties.
+\item[example]
+    A typical usage example of a function.
+\item[property]
+    An invariant of a function.
+\item[behavior]
+    How does a function behave from a users point of view. Behavior can be
+    shown by using both examples and properties.
 \item[documentation] Explanation of functions.

### 2012-05-25 paper.tex
behaviors -> behavior

@@
 but there are many
-functions whose behaviors cannot be guessed from their names and type
+functions whose behavior cannot be guessed from their names and type
 signatures. A good example is \verb|nub :: [a] -> [a]| in 

### 2012-05-25 paper.tex
Wording of "acknowledgements"
markups -> markup

@@
 
-To specify examples/properties for users, we propose to add two new markups to Haddock and to use {\tt doctest} integrated with QuickCheck.
+To specify examples/properties for users, we propose to add two new markup to Haddock and to use {\tt doctest} integrated with QuickCheck.
 To describe examples/properties for developers, we propose to use \hspec{} also integrated with QuickCheck.
@@
 
-We are grateful to David Waern, 
-one of the Haddock maintainers, 
-for reviewing two new markups to Haddock.
+We are grateful to David Waern, Haddock's maintainer, for reviewing our patches
+and accepting our patches to Haddock.
 We would like to thank

### 2012-05-25 paper.tex
Wording of section on "Haddock"

@@
 
-As we described, Haddock version 2.7.2 or earlier has
-the markup for code block but does not provide
-markups for examples and properties.
-We implemented a markup for examples ({\tt >>>}) and 
-it is available in Haddock version 2.8 or later.
-We also extended Haddock to support a markup for
-properties ({\tt prop>})
-which will be available in the future release.
-Both examples and properties in Haddock markup are pretty printed when producing document
-such as in HTML.
+As we described, Haddock version 2.7.2 or earlier provides markup syntax for
+code blocks, but does not provide any markup syntax for examples or properties.
+We implemented markup support for examples ({\tt >>>}) and it is available in
+Haddock version 2.8 or later.
+We also extended Haddock to support markup for properties ({\tt prop>}) which
+will be available with a future release of Haddock.
+Both examples and properties in Haddock comments are pretty-printed when
+generating documentation, (e.g. as HTML).
 

### 2012-05-25 paper.tex
Wording of "Introduction"

@@
 
-This also means that typical usage examples cannot be found from test code.
+This also means that typical usage examples is not provided by test code.
 
-We agree that type signatures are good abstract document~\cite{free}
-but there are many
-functions whose behavior cannot be guessed from their names and type
-signatures. A good example is \verb|nub :: [a] -> [a]| in 
-the Data.List module.
+We agree that type signatures are good abstract documentation~\cite{free}, but
+there are many functions whose behavior cannot be guessed from their names and
+type signatures. A good example is \verb|nub :: [a] -> [a]| from {\tt Data.List}.
 
@@
 In BDD, test code is written with specification terminology.
-In other words, specification can be used for automatic testing.
+In other words, specifications can be used for automatic testing.
 

### 2012-05-25 paper.tex
Add subsubsections to Implementation -> doctest
More wording changes

@@
 
-This also means that typical usage examples is not provided by test code.
+This also means that typical usage examples are not provided by test code.
 
@@
 there are many functions whose behavior cannot be guessed from their names and
-type signatures. A good example is \verb|nub :: [a] -> [a]| from {\tt Data.List}.
+type signatures.  One example is \verb|nub :: [a] -> [a]| from {\tt Data.List}.
 
@@
 In particular, we introduced {\tt doctest}~\cite{doctest} and
-integrated with QuickCheck~\cite{quickcheck} as a BDD tool.
-We also make use of \hspec{} inspired by rspec~\cite{rspec} for
+integrated it with QuickCheck~\cite{quickcheck} as a BDD tool.
+We also make use of \hspec{} inspired by RSpec~\cite{rspec} for
 BDD in Haskell.
 
-The contribution of this paper is as follows:
+The contributions of this paper are:
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Problems}
+\section{Problem}
 
-In Haskell community, HUnit and QuickCheck are commonly used %% what about SmallCheck?
+In the Haskell community, HUnit and QuickCheck are commonly used %% what about SmallCheck?
 and the {\tt test-framework} package integrates them.
-We can categorize the {\tt test-framework} package into
-a tool of TDD.
-Before our work and \hspec{}, there is no BDD tool commonly used in Haskell community.
+We can categorize {\tt test-framework} as
+a tool for TDD.
+Before our work and \hspec{}, there was no widely used BDD tool for Haskell.
 
 BDD tools used in other programming language communities are
-mainly based on example test
-and do not support property test.
-To our experience, it is not difficult to find properties for
+mainly based on example tests,
+and do not support property tests.
+To our experience is, that it is not difficult to find properties for
 pure Haskell functions.
@@
 A good example of examples/properties for developers
-is a corner example registered in a bug track system.
+is a corner example registered in a bug tracking system.
 
@@
 Haddock defines markup syntax in Haskell's comment.
-Haddock provides the code block markup indicated with ({\tt >}).
-For example, the {\tt container} packages has the following code block markup:
+Haddock provides markup for code blocks, indicated with ({\tt >}).
+For example, the {\tt containers} packages has the following code block markup:
 
@@
 
-\noindent Though the code block itslef is a necessary feature,
-we think this kind usage has problems.
+\noindent Though the code block itself is a useful feature,
+we think this kind of usage has problems.
 If the equations above are examples, we cannot distinguish
@@
 
-Here are summary of the problems:
+Here is a summary of the problem:
 
 \begin{itemize}
-\item Haskell community does not have common syntax to describe
+\item The Haskell community does not have common syntax to describe
 examples/properties with Haddock markup syntax.
-\item Haskell community does not have a BDD tool for 
-examples/properties for users.
+\item The Haskell community does not have a BDD tool for examples/properties
+    for users.
 \end{itemize}
@@
 
-To specify examples/properties for users, we propose to add two new markup to Haddock and to use {\tt doctest} integrated with QuickCheck.
-To describe examples/properties for developers, we propose to use \hspec{} also integrated with QuickCheck.
+To be able to specify examples and properties for users, we propose to add new
+markup to Haddock and to use {\tt doctest} integrated with QuickCheck. To
+describe examples/properties for developers, we propose to use \hspec{} also
+integrated with QuickCheck.
 
@@
 
+\subsubsection{Testing examples}
+
 TBD(@sol): how to obtain examples and properties from Haddock markup.
@@
 and result string from GHCi, exceptions can be handled.
+
+\subsubsection{Testing properties}
 

### 2012-05-26 paper.tex
Remove affiliation
Add \doctest command
This ensures that we always typeset in in the same way.

@@
 \newcommand{\hspec}{Hspec}
+\newcommand{\doctest}{Doctest}
 
@@
 \authorinfo{Simon HENGEL}
-           {Freelance Software Developer}
+           {}
            {sol@typeful.net}
@@
 
-\keywords Haddock, QuickCheck, doctest, Hspec
+\keywords \doctest{}, Haddock, \hspec{}, QuickCheck
 
@@
 integrated it with a Haskell-original test method.
-In particular, we introduced {\tt doctest}~\cite{doctest} and
+In particular, we introduced \doctest{}~\cite{doctest} and
 integrated it with QuickCheck~\cite{quickcheck} as a BDD tool.
@@
 \begin{itemize}
-\item Introduced {\tt doctest} from Python community to Haskell community
+\item Introduced \doctest{} from Python community to Haskell community
 and integrated it with QuickCheck
@@
 %% (Section \ref{sec:for-developers} and Section \ref{sec:hspec})
-\item Show best current practice of BDD in Haskell using {\tt doctest} and \hspec{} (Section \ref{ref:Best-current-practice})
+\item Show best current practice of BDD in Haskell using \doctest{} and \hspec{} (Section \ref{ref:Best-current-practice})
 \end{itemize}
 
-\noindent Our {\tt doctest} implementation
+\noindent Our \doctest{} implementation
 can be used for design, documentation and automatic testing.
@@
 
-A good candidate for this is {\tt doctest} in Python community.
+A good candidate for this is \doctest{} in Python community.
 It extends Python's documentation strings so that they
@@
 To be able to specify examples and properties for users, we propose to add new
-markup to Haddock and to use {\tt doctest} integrated with QuickCheck. To
+markup to Haddock and to use \doctest{} integrated with QuickCheck. To
 describe examples/properties for developers, we propose to use \hspec{} also
@@
 \item Define a new Haddock markup for properties.
-\item Bring {\tt doctest} from the Python community to the Haskell
+\item Bring \doctest{} from the Python community to the Haskell
 community and integrate it with QuickCheck.
@@
 Examples and properties in comments are automatically
-tested by the {\tt doctest} command. We also provided a library
+tested by the \doctest{} command. We also provided a library
 to do this job. Its implementation will discussed in Section~\ref{sec:doctest}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{doctest}
+\subsection{\doctest}
 \label{sec:doctest}
@@
 
-\noindent We should execute {\tt doctest} to ensure no syntax errors.
+\noindent We should execute \doctest{} to ensure no syntax errors.
 Also, we specifies examples/properties for developers with \hspec{}:
@@
 file can contain multiple test suites, so it is possible to define test suites
-for Doctest and \hspec{} in a single cabal file.
+for \doctest{} and \hspec{} in a single cabal file.
 
@@
 
-\noindent A driver for Doctest looks like so:
+\noindent A driver for \doctest{} looks like so:
 

### 2012-05-26 paper.tex
Add test-shouldbe to contributions

@@
 
-The contributions of this paper are:
+Our contributions are:
 
 \begin{itemize}
-\item Introduced \doctest{} from Python community to Haskell community
-and integrated it with QuickCheck
+\item
+    Add support for QuickCheck properties to \doctest{}
     (Section \ref{sec:for-users} and Section \ref{sec:doctest})
+
+\item
+    Introduce a set of combinators to set of expectations about \emph{code
+    examples} for use with \hspec{}.
 %% \item Introduced {\tt Hspec} from Ruby community to Haskell community and
@@
 %% (Section \ref{sec:for-developers} and Section \ref{sec:hspec})
-\item Show best current practice of BDD in Haskell using \doctest{} and \hspec{} (Section \ref{ref:Best-current-practice})
+\item Show best current practice of BDD in Haskell using \doctest{} and
+    \hspec{} (Section \ref{ref:Best-current-practice})
 \end{itemize}

### 2012-05-26 paper.tex
Extend terminology

@@
 \begin{description}
+
 \item[developer]
     A programmer who develops a library.
+
 \item[user]
     A programmer who uses a library.
-%%\item[unit test] tests for individual programs or libraries.
+
 \item[example]
-    A typical usage example of a function.
+    An example shows how a function is used.  It is either a \emph{code
+    example} in the form of an HUnit test case; or an \emph{interactive
+    example}, given as a short GHCi session.
+
 \item[property]
-    An invariant of a function.
+    Property means a QuickCheck property as defined in~\cite{quickcheck}.
+
 \item[behavior]
-    How does a function behave from a users point of view. Behavior can be
-    shown by using both examples and properties.
-\item[documentation] Explanation of functions.
+    Behavior means an invariant of a function.  We use an informal description
+    to refer to a particular behavior, e.g.  "sorts a list" or "launches
+    missiles when run".
+
+\item[evidence]
+    Evidence is either an example or a property.  It provides some confidence,
+    that a function exhibits a certain behavior.
+
+\item[documentation]
+    Explanation of functions.
 \end{description}

### 2012-05-26 paper.tex
Use evidence instead of example/property
Fix typo
Add important points that should go into the "Introduction"

@@
 %%%% Describe the problem.
+\begin{quote}
+    Important points we want to say:
+    \begin{itemize}
+        \item
+            All behavior should be documented, together with some
+            \emph{evidence} for that behavior.
+        \item
+            BDD approaches \emph{Test-driven development (TDD)} and
+            \emph{Acceptance test-driven development (ATDD)} with a new
+            terminology and mindset .  For the purpose of this paper we focus
+            on the TDD aspect of BDD.
+        \item
+            Automating things, and continuous integration is an important
+            aspect of BDD.
+    \end{itemize}
+\end{quote}
 
@@
 \item
-    Introduce a set of combinators to set of expectations about \emph{code
+    Introduce a set of combinators to set expectations about \emph{code
     examples} for use with \hspec{}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Examples/properties for users}
+\subsection{Evidence as user documentation}
 \label{sec:for-users}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Examples/properties for developers}
+\subsection{Evidence as code documentation}
 \label{sec:for-developers}

### 2012-05-26 paper.tex
Apply terminology

@@
 
-If we have examples/properties that does not fit into the user documentation, but
-would rather be part of a design document, we use \hspec{}.
+If we have evidence that does not fit into the user documentation, but would
+rather be part of a design document, we use \hspec{}.
 
-\hspec{} provides an EDSL for defining specs.  A spec organizes behavior in a
+\hspec{} provides an EDSL for defining specs.  A spec organizes evidence in a
 tree structure, and it is defined with \texttt{it} and \texttt{describe}.

### 2012-05-26 paper.tex
Add missing TBD
Minor clarification
Add a hint for vim to set textwidth to 70
Adapt Hspec examples
Use Specs instead of Spec, and hspec instead of hspecX.

@@
+% vim: tw=70
 %-----------------------------------------------------------------------------
@@
 
+\begin{quote}
+\emph{NOTE: @sol has to release a new version of hspec, that includes
+the Spec-to-Specs and hspecX-to-hspec changes.  Otherwise the code in
+this paper will not work!}
+\end{quote}
+
 %%%% Describe the problem.
 \begin{quote}
-    Important points we want to say:
+    TBD: Important points we want to say about BDD:
     \begin{itemize}
@@
 import Test.Hspec.Monadic
-spec :: Specs
+spec :: Spec
 spec = do
@@
 
-A spec can be run with the \texttt{hspecX} function.  When a spec is run, all
+A spec can be run with the \texttt{hspec} function.  When a spec is run, all
 properties and examples are evaluated, and failing properties and examples are
@@
 main :: IO ()
-main = hspecX $ do
+main = hspec $ do
   describe "Foo"     FooSpec.spec
@@
 import Test.Hspec.Monadic
-main = hspecX spec
+main = hspec spec
 \end{verbatim}

### 2012-05-26 paper.tex
Minor formatting change
Remove comment stuff related to Hspec
whitespace change

@@
            {sol@typeful.net}
-%% \authorinfo{Trystan SPANGLER}
-%%            {TBD: Affiliation}
-%%            {TBD: Email}
 \authorinfo{Kazuhiko YAMAMOTO}
@@
 
-\begin{quote}
-\emph{NOTE: @sol has to release a new version of hspec, that includes
-the Spec-to-Specs and hspecX-to-hspec changes.  Otherwise the code in
-this paper will not work!}
-\end{quote}
+\emph{IMPORTANT: @sol has to release a new version of hspec, that
+includes the Spec-to-Specs and hspecX-to-hspec changes.  Otherwise the
+code in this paper will not work!}
 
@@
 \item
-    Add support for QuickCheck properties to \doctest{}
-    (Section \ref{sec:for-users} and Section \ref{sec:doctest})
+    Add support for QuickCheck properties to \doctest{} (Section
+    \ref{sec:for-users} and Section \ref{sec:doctest})
 
 \item
-    Introduce a set of combinators to set expectations about \emph{code
-    examples} for use with \hspec{}.
-%% \item Introduced {\tt Hspec} from Ruby community to Haskell community and
-%% integrated it with QuickCheck
-%% (Section \ref{sec:for-developers} and Section \ref{sec:hspec})
-\item Show best current practice of BDD in Haskell using \doctest{} and
+    Introduce a set of combinators to set expectations about
+    \emph{code examples} for use with \hspec{}.
+
+\item
+    Show best current practice of BDD in Haskell using \doctest{} and
     \hspec{} (Section \ref{ref:Best-current-practice})
+
 \end{itemize}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \subsection{Hspec}
-%% \label{sec:hspec}
-
-%% TBD.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-05-26 paper.tex
Add CI support to contributions
Add TBDs to section "Problem"
Add a TBD note

@@
 \item
+    Improve CI support for \doctest{} and \hspec{}
+
+    TBD:
+    \emph{This relates to hspec-discover and Doctest's Cabal support.}
+
+\item
     Show best current practice of BDD in Haskell using \doctest{} and
@@
 
+\begin{quote}
+    TBD:
+\begin{itemize}
+    \item Motivate \verb|test-shouldbe|
+    \item Motivate \verb|shouldThrow|
+    \item \emph{optional:} Motivate why it's important to integrate \doctest{} with
+        Cabal.
+    \item \emph{optional:} Motivate \verb|hspec-discover|
+\end{itemize}
+\end{quote}
+
 Here is a summary of the problem:
@@
 \begin{itemize}
-\item The Haskell community does not have common syntax to describe
+\item
+    The Haskell community does not have common syntax to describe
     examples/properties with Haddock markup syntax.
-\item The Haskell community does not have a BDD tool for examples/properties
-    for users.
+\item
+    The Haskell community does not have a BDD tool for
+    examples/properties for users.
+\item
+    Hspec does not provide combinators to express expectations with
+    BDD terminology
 \end{itemize}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evidence as code documentation}
+\subsection{Evidence as code documentation (TBD: Is this still
+appropriate here?}
 \label{sec:for-developers}

### 2012-05-26 paper.tex
Turn section on fundamentalism into a quote (for now)
Add new subsections to "Solution"
Extend "Problem"

@@
 
+\item[spec]
+    TBD: BDD terminology for test suite.
+
 \item[documentation]
@@
     \item \emph{optional:} Motivate \verb|hspec-discover|
+        It is common to have several files that contain \emph{specs}.
+        And we want to be able to run them all together.  Hspec allows
+        us combine specs together.  But this has to be done manually.
+        Hspec does not provide a way to automatically discover and run
+        all specs of a project.
 \end{itemize}
@@
 
+\subsection{Combinators for expectations}
+TBD
+\subsection{Automatic discovery of specs}
+TBD
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Note that the authors are not fundamentalist.
-If functions to be defined are simple enough,
-we might implement them first withtout signatures.
-And their signatures can be automatically
-inserted by IDE (such as Emacs/vim with {\tt ghc-mod}) later
-if Haskell compilers can infer correctly.
-Also, we can copy-and-paste examples and their results
-from Haskell interpreters to their Haddock markup.
+\begin{quote}
+    TBD: @kazu: Do we need this to make our case?  I'm asking, because
+    it will require some work to get the language right.
+
+    Note that the authors are not fundamentalist.  If functions to be
+    defined are simple enough, we might implement them first withtout
+    signatures.  And their signatures can be automatically inserted by
+    IDE (such as Emacs/vim with {\tt ghc-mod}) later if Haskell
+    compilers can infer correctly.  Also, we can copy-and-paste
+    examples and their results from Haskell interpreters to their
+    Haddock markup.
+\end{quote}
 

### 2012-05-26 paper.tex
Add TBD to "Introduction"
Fix typo

@@
 \emph{IMPORTANT: @sol has to release a new version of hspec, that
-includes the Spec-to-Specs and hspecX-to-hspec changes.  Otherwise the
+includes the Specs-to-Spec and hspecX-to-hspec changes.  Otherwise the
 code in this paper will not work!}
@@
 In other words, specifications can be used for automatic testing.
+
+\begin{quote}
+    TBD: Important points we want to say here:
+    \begin{itemize}
+        \item
+            We use \doctest{} and \hspec{} to document behavior.
+            (in contrast to other communities)
+        \item
+            Some features are missing from both, \doctest{} and
+            \hspec{}.
+        \item
+            We extend them accordingly, and show how they can be used
+            in combination to design, document and test Haskell code.
+    \end{itemize}
+\end{quote}
 

### 2012-05-26 paper.tex
Formatting
Add rough sketch for expectations

@@
             aspect of BDD.
+        \item
+            BDD uses expectations instead of assertions
     \end{itemize}
@@
 
+\item[expectation]
+    Expectations are used to define the expected outcome of a code
+    example.  They serve the same purpose as assertions in xUnit
+    testing frameworks, but use a different terminology.
+
 \item[property]
-    Property means a QuickCheck property as defined in~\cite{quickcheck}.
+    Property means a QuickCheck property as defined in
+    \cite{quickcheck}.
 
@@
     \item Motivate \verb|test-shouldbe|
+
+        BDD frameworks use \emph{should} or \emph{must} instead of
+        \emph{assert}.
+
+        Here is an example that uses RSpec.
+
+        \begin{verbatim}
+23.should eq(23)
+        \end{verbatim}
+
+        RSpec provides a mechanism to expect exceptions.
+
+        \begin{verbatim}
+expect {launche_missiles}.to raise_error
+        \end{verbatim}
+
+        \hspec does not yet provide a mechanism to set expectations
+        about the outcome of code examples.  It uses HUnit assertions
+        instead.
+
+        HUnit does not provide a mechanism to expect exceptions.
+
     \item Motivate \verb|shouldThrow|
@@
     \item \emph{optional:} Motivate \verb|hspec-discover|
+
         It is common to have several files that contain \emph{specs}.
@@
 
-    it "gives the original list, if applied twice" $ property $
-      \xs -> reverse (reverse xs) == (xs :: [Int])
+    it "gives the original list, if applied twice" $
+      property $ \xs ->
+        (reverse . reverse) xs == (xs :: [Int])
 \end{verbatim}
@@
 \subsection{Combinators for expectations}
-TBD
-\subsection{Automatic discovery of specs}
+We propose to use three combinators to set expectations about the
+outcome of code examples.
+
+\begin{quote}\small\begin{verbatim}
+shouldBe :: (Show a, Eq a) => a -> a -> Assertion
+shouldSatisfy :: Show a => a -> (a -> Bool) -> Assertion
+shouldReturn :: (Show a, Eq a) => IO a -> a -> Assertion
+\end{verbatim}\end{quote}
+
+\subsection{\emph{optional} Automatic discovery of specs}
 TBD

### 2012-05-27 paper.tex
Add reasoning why type system does not always ensure correct semantics

@@
 
-Unfortunately, the strong type system of Haskell does not guarantee correct
-semantic.
+Unfortunately, the type system of Haskell does not always guarantee
+correct semantics.
 
 \begin{quote}
-    \emph{TBD: What about dependent types and GHC's new features?}
+    TBD: Make it plausible, that the type system does not always
+    guarantee correctly semantics.
+    \begin{itemize}
+        \item Many of the standard type classes \emph{should} adhere
+            to certain laws.
+        \item The type system does not enforce those laws.
+        \item It is the programmers responsibility to ensure that
+            instance definitions \emph{behave} as \emph{expected}.
+    \end{itemize}
 \end{quote}
@@
 
-    \item Motivate \verb|shouldThrow|
     \item \emph{optional:} Motivate why it's important to integrate \doctest{} with
         Cabal.
+
     \item \emph{optional:} Motivate \verb|hspec-discover|

### 2012-05-27 paper.tex
Add a comment
Fix typo
Add some text to "Combinators for expectations"

@@
         \begin{verbatim}
-expect {launche_missiles}.to raise_error
+expect {launch_missiles}.to raise_error
         \end{verbatim}
@@
 \subsection{Combinators for expectations}
-We propose to use three combinators to set expectations about the
-outcome of code examples.
+
+We propose four combinators, that can be used to set expectations
+about the outcome of code examples.
+
+A common expectation is, that a value should be equal to an other
+value.
+
+\begin{quote}\small\begin{verbatim}
+x `shouldBe` 23
+\end{verbatim}\end{quote}
+
+\noindent This is just an other name for HUnit's \texttt{@?=}
+operator.  The type is:
 
@@
 shouldBe :: (Show a, Eq a) => a -> a -> Assertion
-shouldSatisfy :: Show a => a -> (a -> Bool) -> Assertion
-shouldReturn :: (Show a, Eq a) => IO a -> a -> Assertion
+\end{verbatim}\end{quote}
+
+\noindent Or we may expect that a value satisfies some predicate.
+
+\begin{quote}\small\begin{verbatim}
+x `shouldSatisfy` (< 23)
+\end{verbatim}\end{quote}
+
+\noindent The value is required to be in the \texttt{Show} class, so that a
+useful error message can be given when the predicate is not satisfied.
+
+\begin{quote}\small\begin{verbatim}
+shouldSatisfy :: Show a => a -> (a -> Bool)
+  -> Assertion
+\end{verbatim}\end{quote}
+
+\noindent Examples are often used to provide evidence for {\tt IO}
+actions.
+
+\begin{quote}\small\begin{verbatim}
+launchMissiles >>=
+  (`shouldBe` (Left "not implemented"))
+\end{verbatim}\end{quote}
+
+\noindent So we propose a shortcut for this common use case.
+
+\begin{quote}\small\begin{verbatim}
+launchMissiles `shouldReturn` (Left "not implemented")
+\end{verbatim}\end{quote}
+
+\begin{quote}\small\begin{verbatim}
+shouldReturn :: (Show a, Eq a) => IO a -> a
+  -> Assertion
+\end{verbatim}\end{quote}
+
+\subsubsection{Expecting exceptions}
+
+HUnit does not provide any versatile mechanism to expect that an
+exception is thrown.  So we propose a combinator for that.
+
+\begin{quote}\small\begin{verbatim}
+shouldThrow :: Exception e => IO a -> (e -> Bool)
+  -> Assertion
+\end{verbatim}\end{quote}
+
+\noindent It takes an {\tt IO} action, and a predicate.  The predicate
+servers two purposes.  It selects the type of the expected exception,
+and constraints the value.
+
+\begin{quote}\small\begin{verbatim}
+launchMissiles `shouldThrow` (== ExitFailure 1)
+\end{verbatim}\end{quote}
+
+\noindent To make it more convenient, to expect any exception of a
+certain type, we provide type-restricted versions of {\tt (const
+True)} for all standard exceptions.
+
+\begin{quote}\small\begin{verbatim}
+anyArithException :: ArithException -> Bool
+anyArithException = const True
+\end{verbatim}\end{quote}
+
+\noindent Which can be used like so:
+
+\begin{quote}\small\begin{verbatim}
+let x = 1 `div` 0 in (x `seq` return x)
+      `shouldThrow` anyArithException
 \end{verbatim}\end{quote}
@@
 \subsection{Procedure of BDD}
+
+\begin{quote}
+@kazu: Can we use \verb|reverse| instead of \verb|rev| here?
+\end{quote}
 

### 2012-05-27 paper.tex
Remove \hspec and \doctest macros

@@
 \usepackage{url}
-
-\newcommand{\hspec}{Hspec}
-\newcommand{\doctest}{Doctest}
 
@@
 
-\keywords \doctest{}, Haddock, \hspec{}, QuickCheck
+\keywords Doctest, Haddock, Hspec, QuickCheck
 
@@
         \item
-            We use \doctest{} and \hspec{} to document behavior.
+            We use Doctest and Hspec to document behavior.
             (in contrast to other communities)
         \item
-            Some features are missing from both, \doctest{} and
-            \hspec{}.
+            Some features are missing from both, Doctest and
+            Hspec.
         \item
@@
 integrated it with a Haskell-original test method.
-In particular, we introduced \doctest{}~\cite{doctest} and
+In particular, we introduced Doctest~\cite{doctest} and
 integrated it with QuickCheck~\cite{quickcheck} as a BDD tool.
-We also make use of \hspec{} inspired by RSpec~\cite{rspec} for
+We also make use of Hspec inspired by RSpec~\cite{rspec} for
 BDD in Haskell.
@@
 \item
-    Add support for QuickCheck properties to \doctest{} (Section
+    Add support for QuickCheck properties to Doctest (Section
     \ref{sec:for-users} and Section \ref{sec:doctest})
@@
     Introduce a set of combinators to set expectations about
-    \emph{code examples} for use with \hspec{}.
+    \emph{code examples} for use with Hspec.
 
 \item
-    Improve CI support for \doctest{} and \hspec{}
+    Improve CI support for Doctest and Hspec
 
@@
 \item
-    Show best current practice of BDD in Haskell using \doctest{} and
-    \hspec{} (Section \ref{ref:Best-current-practice})
+    Show best current practice of BDD in Haskell using Doctest and
+    Hspec (Section \ref{ref:Best-current-practice})
 
@@
 
-\noindent Our \doctest{} implementation
+\noindent Our Doctest implementation
 can be used for design, documentation and automatic testing.
@@
 a tool for TDD.
-Before our work and \hspec{}, there was no widely used BDD tool for Haskell.
+Before our work and Hspec, there was no widely used BDD tool for Haskell.
 
@@
 
-A good candidate for this is \doctest{} in Python community.
+A good candidate for this is Doctest in Python community.
 It extends Python's documentation strings so that they
@@
 
-To describe examples/properties for developers, we can use \hspec{}.
+To describe examples/properties for developers, we can use Hspec.
 
@@
 
-        \hspec does not yet provide a mechanism to set expectations
+        Hspec does not yet provide a mechanism to set expectations
         about the outcome of code examples.  It uses HUnit assertions
@@
 
-    \item \emph{optional:} Motivate why it's important to integrate \doctest{} with
+    \item \emph{optional:} Motivate why it's important to integrate
+        Doctest with
         Cabal.
@@
 To be able to specify examples and properties for users, we propose to add new
-markup to Haddock and to use \doctest{} integrated with QuickCheck. To
-describe examples/properties for developers, we propose to use \hspec{} also
+markup to Haddock and to use Doctest integrated with QuickCheck. To
+describe examples/properties for developers, we propose to use Hspec also
 integrated with QuickCheck.
@@
 \item Define a new Haddock markup for properties.
-\item Bring \doctest{} from the Python community to the Haskell
+\item Bring Doctest from the Python community to the Haskell
 community and integrate it with QuickCheck.
@@
 Examples and properties in comments are automatically
-tested by the \doctest{} command. We also provided a library
+tested by the Doctest command. We also provided a library
 to do this job. Its implementation will discussed in Section~\ref{sec:doctest}.
@@
 If we have evidence that does not fit into the user documentation, but would
-rather be part of a design document, we use \hspec{}.
+rather be part of a design document, we use Hspec.
 
-\hspec{} provides an EDSL for defining specs.  A spec organizes evidence in a
+Hspec provides an EDSL for defining specs.  A spec organizes evidence in a
 tree structure, and it is defined with \texttt{it} and \texttt{describe}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{\doctest}
+\subsection{Doctest}
 \label{sec:doctest}
@@
 
-\noindent We should execute \doctest{} to ensure no syntax errors.
-Also, we specifies examples/properties for developers with \hspec{}:
+\noindent We should execute Doctest to ensure no syntax errors.
+Also, we specifies examples/properties for developers with Hspec:
 
@@
 
-\noindent Also, we should execute this \hspec{} module
+\noindent Also, we should execute this Hspec module
 to ensure no syntax errors.
@@
 file can contain multiple test suites, so it is possible to define test suites
-for \doctest{} and \hspec{} in a single cabal file.
+for Doctest and Hspec in a single cabal file.
 
-A driver for \hspec{} looks like so:
+A driver for Hspec looks like so:
 
@@
 
-\noindent A driver for \doctest{} looks like so:
+\noindent A driver for Doctest looks like so:
 

### 2012-05-27 paper.tex
Add a comment

@@
 we use ``behavior'' in this paper.
+
+\begin{quote}
+    @kazu: We e.g. say ill-behaved Monoid instance.  So is
+    ``behavior'' really a bad word?  But I agree that some people in
+    the community might associated it with something imperative.
+
+    On might also argue that ``should'' is not very mathematical, but
+    it is common language when we talk about mathematics.  E.g. Brent
+    Yorgey uses ``should'' a lot in the Typeclassopedia (e.g. ``\ldots
+    {\tt Alternative} should satisfy the monoid laws\ldots'')
+\end{quote}
 

### 2012-05-27 paper.tex
fix typo
Assertion -> Expectation

@@
 
-    On might also argue that ``should'' is not very mathematical, but
+    One might also argue that ``should'' is not very mathematical, but
     it is common language when we talk about mathematics.  E.g. Brent
@@
 
-We propose four combinators, that can be used to set expectations
-about the outcome of code examples.
+We propose several combinators, that can be used to set expectations
+about the outcome of code examples.  And we build them on top of
+HUnit.
 
-A common expectation is, that a value should be equal to an other
+\begin{quote}\small\begin{verbatim}
+type Expectation = Assertion
+\end{verbatim}\end{quote}
+
+\noindent A common expectation is, that a value should be equal to an other
 value.
@@
 \begin{quote}\small\begin{verbatim}
-shouldBe :: (Show a, Eq a) => a -> a -> Assertion
+shouldBe :: (Show a, Eq a) => a -> a -> Expectation
 \end{verbatim}\end{quote}
@@
 shouldSatisfy :: Show a => a -> (a -> Bool)
-  -> Assertion
+  -> Expectation
 \end{verbatim}\end{quote}
@@
 shouldReturn :: (Show a, Eq a) => IO a -> a
-  -> Assertion
+  -> Expectation
 \end{verbatim}\end{quote}
@@
 shouldThrow :: Exception e => IO a -> (e -> Bool)
-  -> Assertion
+  -> Expectation
 \end{verbatim}\end{quote}

### 2012-05-27 paper.tex
Wording change
Add TBDs

@@
 
-HUnit does not provide any versatile mechanism to expect that an
-exception is thrown.  So we propose a combinator for that.
+HUnit does not provide any versatile mechanism to state that an action
+should throw an exception.  So we propose a combinator for that.
 
@@
       `shouldThrow` anyArithException
+\end{verbatim}\end{quote}
+
+\begin{quote}
+    TBD:
+    \begin{itemize}
+        \item
+            This code snippet also illustrates how to deal with pure
+            functions.
+        \item
+            {\tt seq} forces the exception.
+        \item
+            If {\tt x} is in {\tt NFData}, you may want to use {\tt
+            deepseq} instead.
+        \item
+            The following would not work:
+\begin{quote}\small\begin{verbatim}
+(1 `div` 0) `shouldThrow` anyArithException
+\end{verbatim}\end{quote}
+        \item
+            We think that pure functions that throw exceptions is bad
+            API design.
+    \end{itemize}
+\end{quote}
+
+TBD: Should we propose:
+\begin{quote}\small\begin{verbatim}
+shouldThrow_ :: (NFData a, Exception e) =>
+  a -> (e -> Bool) -> Expectation
 \end{verbatim}\end{quote}

### 2012-05-27 paper.tex
Wording
Combining multiple specs -> Automatic discover of specs

@@
 
-\subsection{Combinators for expectations}
+\subsection{Setting expectations}
 
@@
 
-\subsection{Combining multiple specs}
+\subsection{Automatic discover of specs}
 
@@
 
-\begin{quote}
-\emph{How about moving this to appendix?}
-\end{quote}
-
-\begin{itemize}
-\item We create a spec file for each source file
-\item By convention, spec files end in Spec.hs
+\noindent We think the following are good conventions for organizing specs:
+\begin{itemize}
+    \item create a separate spec file for each source file
+    \item append {\tt Spec.hs} to the name of spec files
 \end{itemize}

### 2012-05-27 paper.tex
Add TBD for implementation of expectation combinators
Wording

@@
         \item
-            BDD approaches \emph{Test-driven development (TDD)} and
-            \emph{Acceptance test-driven development (ATDD)} with a new
-            terminology and mindset .  For the purpose of this paper we focus
-            on the TDD aspect of BDD.
+            \emph{Behavior driven development (BDD)} approaches
+            \emph{test-driven development (TDD)} and \emph{acceptance
+            test-driven development (ATDD)} with a new terminology and
+            mindset.  For the purpose of this paper we only focus on
+            the TDD aspect of BDD.
         \item
@@
 to the property.
+
+\subsection{Combinators for expectations}
+
+TBD: describe the implementation of {\tt shouldBe}, etc.
 

### 2012-05-27 paper.tex
Add missing \noindent statements

@@
 
-First, we introduce a new Haddock markup for the example.
+\noindent First, we introduce a new Haddock markup for the example.
 In this paper, we use {\tt >>>} which is the same as Python's one.
@@
 
-To specify properties, we also propose a new markup. In this paper, we use the {\tt prop>} prefix. We can write properties which do not take parameters as follows:
+\noindent To specify properties, we also propose a new markup. In this paper, we
+use the {\tt prop>} prefix. We can write properties which do not take
+parameters as follows:
 
@@
 
-Examples and properties in comments are automatically
+\noindent Examples and properties in comments are automatically
 tested by the Doctest command. We also provided a library

### 2012-05-28 paper.tex
Add testpack to "related work"

@@
 \section{Open problems and related work}
-
-literate programming.
-
-hunit.
-
-test-framework.
-
-smallCheck.
-
-And others?
+\begin{itemize}
+    \item literate programming.
+    \item hunit.
+    \item test-framework.
+    \item smallCheck.
+    \item And others?
+    \item The {\tt testpack} package provides a function that is
+        similar in purpose to {\tt shouldThrow}.
+
+        @kazu: Do you think we should discuss this; or is it safe to
+        ignore.
+
+\begin{quote}\small\begin{verbatim}
+assertRaises :: (Show a, Exception e, Show e, Eq e) =>
+  String -> e -> IO a -> IO ()
+\end{verbatim}\end{quote}
+        But it has several shortcomings:
+        \begin{itemize}
+            \item
+                It only works for exceptions with an {\tt Eq}
+                instance, but several standard exceptions (most
+                prominently {\tt ErrorCall}) have no {\tt Eq}
+                instance.
+            \item
+                It places an unnecessary {\tt Show} constraint on the
+                result of the {\tt IO} action, further limiting it's
+                applicability.
+            \item
+                It only allows to expect specific values; expecting
+                any exception of a certain type is not possible.
+        \end{itemize}
+\end{itemize}
 

### 2012-05-28 paper.tex
using bibtex.

@@
 
-\item @kazu: Can you add a reference to:
-\url{http://www.haskell.org/cabal/users-guide/developing-packages.html#test-suites}
-
 \item It is important that testing is automated, including running of test
@@
 
-\noindent Cabals test feature can be used to automatically run tests.  A cabal
+\noindent Cabals test feature can be used to automatically run tests~\cite{cabal}.  A cabal
 file can contain multiple test suites, so it is possible to define test suites
@@
 
-% We recommend abbrvnat bibliography style.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \bibliographystyle{abbrvnat}
-
-% The bibliography should be embedded for final submission.
-
-\begin{thebibliography}{}
-\softraggedright
-
-\bibitem[Wadler 1989]{free}
-Wadler P.,
-``Theorems for free!'',
-In proceeding of 4th International Conference on Functional Programming and Computer Architecture, 1989.
-
-\bibitem[Claessen et al. 2000]{quickcheck}
-Claessen K. and Hughes J.,
-``QuickCheck: a lightweight tool for random testing of Haskell programs'',
-ICFP '00 Proceedings of the fifth ACM SIGPLAN international conference on Functional programming,
-pp 268 -- 279,
-2000
-
-\bibitem[Marlow et al. 2010]{haddock}
-Marlow S. and Waern D.,
-``Haddock User Guide'',
-URL http://www.haskell.org/haddock/doc/html/,
-2010
-
-\bibitem[Python 2012]{doctest}
-Python Software Foundation,
-``Python documentation'',
-URL http://docs.python.org/library/doctest.html,
-2012
-
-\bibitem[Chelimsky et al. 2010]{rspec}
-Chelimsky D.,
-Astels D., Dennis Z., Hellesoy A., Helmkamp B., and North D.,
-``The RSpec Book: Behaviour-Driven
-Development with RSpec, Cucumber, and Friends'',
-2010
-%% Hellesoy -- o/
-
-\end{thebibliography}
+\bibliography{refs}
 

### 2012-05-28 paper.tex
ref to Haskell Report.
reference fix.

@@
 
-Haskell has a very strong type system, which ensures that most type errors are
+Haskell~\cite{haskell} has a very strong type system, which ensures that most type errors are
 detected at compile time rather than runtime.  Thanks to this, we assume that
@@
 
-The word ``behavior'' is associated with actions.
+The word ``behavior'' is associated with imperative actions.
 Since pure functions of Haskell are static,
 this word might be improper for Haskell.
-However, to avoid to define a new terminology,
-we use ``behavior'' in this paper.
-
-\begin{quote}
-    @kazu: We e.g. say ill-behaved Monoid instance.  So is
-    ``behavior'' really a bad word?  But I agree that some people in
-    the community might associated it with something imperative.
-
-    One might also argue that ``should'' is not very mathematical, but
-    it is common language when we talk about mathematics.  E.g. Brent
-    Yorgey uses ``should'' a lot in the Typeclassopedia (e.g. ``\ldots
-    {\tt Alternative} should satisfy the monoid laws\ldots'')
-\end{quote}
+However, this word is already used in Haskell community~\cite{typeclassopedia}.
+So, we use ``behavior'' instead of defining a new terminology in this paper.
+
+%% \begin{quote}
+%%     @kazu: We e.g. say ill-behaved Monoid instance.  So is
+%%     ``behavior'' really a bad word?  But I agree that some people in
+%%     the community might associated it with something imperative.
+
+%%     One might also argue that ``should'' is not very mathematical, but
+%%     it is common language when we talk about mathematics.  E.g. Brent
+%%     Yorgey uses ``should'' a lot in the Typeclassopedia (e.g. ``\ldots
+%%     {\tt Alternative} should satisfy the monoid laws\ldots'')
+%% \end{quote}
 

### 2012-05-28 paper.tex
using "evidence".

@@
 
-We can categorize examples/properties into two types: for users and
+We can categorize evidences into two types: for users and
 and for developers.
-Examples/properties for users are those which are useful as user documentation.
-By contrast, examples/properties for developers are
+evidences for users are those which are useful as user documentation.
+By contrast, evidences for developers are
 those which are not useful as user documentation.
-A good example of examples/properties for developers
+A good example of evidences for developers
 is a corner example registered in a bug tracking system.
@@
 BDD tools should be treat both of them and provide functionality
-to extract examples/properties for users as user documentation.
+to extract evidences for users as user documentation.
 If we can archive this, BDD tools can be used for
@@
 
-To describe examples/properties for developers, we can use Hspec.
+To describe evidences for developers, we can use Hspec.
 
@@
     The Haskell community does not have common syntax to describe
-    examples/properties with Haddock markup syntax.
+    evidences with Haddock markup syntax.
 \item
     The Haskell community does not have a BDD tool for
-    examples/properties for users.
+    evidences for users.
 \item
@@
 
-To be able to specify examples and properties for users, we propose to add new
+To be able to specify evidences for users, we propose to add new
 markup to Haddock and to use Doctest integrated with QuickCheck. To
-describe examples/properties for developers, we propose to use Hspec also
+describe evidences for developers, we propose to use Hspec also
 integrated with QuickCheck.
@@
 
-Examples/properties for users should be documented with Haddock markup syntax.
+evidences for users should be documented with Haddock markup syntax.
 For this purpose, we propose the followings:
@@
 
-\noindent Examples and properties in comments are automatically
+\noindent Evidences in comments are automatically
 tested by the Doctest command. We also provided a library
@@
 \begin{quote}
-    \emph{TBD: Behavior should be replaced with examples/properties in this subsection?}
+    \emph{TBD: Behavior should be replaced with evidences in this subsection?}
 \end{quote}
@@
 \noindent The {\tt it} function combines a text description for a behavior and
-a property or an example that gives some confidence in that behavior into a
+an evidence that gives some confidence in that behavior into a
 spec.
@@
 A spec can be run with the \texttt{hspec} function.  When a spec is run, all
-properties and examples are evaluated, and failing properties and examples are
+evidences are evaluated, and failing evidences are
 reported to the user (FIXME: user does not fit our definition of \emph{user}
@@
 As we described, Haddock version 2.7.2 or earlier provides markup syntax for
-code blocks, but does not provide any markup syntax for examples or properties.
+code blocks, but does not provide any markup syntax for evidences.
 We implemented markup support for examples ({\tt >>>}) and it is available in
@@
 
-TBD(@sol): how to obtain examples and properties from Haddock markup.
+TBD(@sol): how to obtain evidences from Haddock markup.
 Old version uses Haddock API but now uses GHC API directly. Why?
@@
 \noindent Note that the {\tt placeholders} library provide better keyword such as {\tt notImplemented} and {\tt todo}.
-Next we write examples/properties for users in Haddock markup:
+Next we write evidences for users in Haddock markup:
 
@@
 \noindent We should execute Doctest to ensure no syntax errors.
-Also, we specifies examples/properties for developers with Hspec:
+Also, we specifies evidences for developers with Hspec:
 

### 2012-05-28 paper.tex
adding terminology.
add comment.
Talking HP as evidence.

@@
 
-Haskell programmers, however, tend to rely on the type system too much.  They
-tend to write less test code than programmers of other programming languages.
-
-\begin{quote}
-    \emph{we need evidence to support this.}
-\end{quote}
-
-This also means that typical usage examples are not provided by test code.
+Haskell programmers, however, is apt to rely on the type system too much
+and tend not to write test code.
+For example, Haskell Platform 2011.4.0.0 contains
+TBD packages but TBD of them have no test code.
+Lacking test code also means that typical usage examples are not provided.
 
@@
 
-\item[documentation]
-    Explanation of functions.
+\item[user documentation]
+    TBD: Explanation of functions.
+\item[code documentation]
+    TBD: Explanation of functions.
+
 \end{description}
@@
 @kazu: Can we use \verb|reverse| instead of \verb|rev| here?
+\end{quote}
+
+\begin{quote}
+@sol: I used \verb|rev| just because it is short.
+Actually I used \verb|reverse| once but it is overflowed
+in the right hand side.
 \end{quote}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusion}
-
-\appendix
-
-\section{Package Survey}
-
-TBD
+%% \section{Conclusion}
+
+%% \appendix
+
+%% \section{Package Survey}
+
+%% TBD
 

### 2012-05-28 paper.tex
clearly say that we implemented Doctest in Haskell.
adding CI ref and describe monad low.

@@
         \item
-            Automating things, and continuous integration is an important
+            Automating things, and continuous integration~\cite{ci} is an important
             aspect of BDD.
@@
 correct semantics.
-
-\begin{quote}
-    TBD: Make it plausible, that the type system does not always
-    guarantee correctly semantics.
-    \begin{itemize}
-        \item Many of the standard type classes \emph{should} adhere
-            to certain laws.
-        \item The type system does not enforce those laws.
-        \item It is the programmers responsibility to ensure that
+For example, it does not ensure that division by zero does not occur.
+Another example is lows of type classes. 
+For instance, the {\tt Monad} class has three Monad lows 
+that its instance \emph{should} satisfy
+but
+it is the programmers responsibility to ensure that
 instance definitions \emph{behave} as \emph{expected}.
-    \end{itemize}
-\end{quote}
 
@@
 \item
-    Add support for QuickCheck properties to Doctest (Section
+    Implement Doctest in Haskell and add support for QuickCheck properties to Doctest (Section
     \ref{sec:for-users} and Section \ref{sec:doctest})

### 2012-05-28 paper.tex
brushing up intro.
improving intro.

@@
 %%%% Describe the problem.
-\begin{quote}
-    TBD: Important points we want to say about BDD:
-    \begin{itemize}
-        \item
-            All behavior should be documented, together with some
-            \emph{evidence} for that behavior.
-        \item
-            \emph{Behavior driven development (BDD)} approaches
-            \emph{test-driven development (TDD)} and \emph{acceptance
-            test-driven development (ATDD)} with a new terminology and
-            mindset.  For the purpose of this paper we only focus on
-            the TDD aspect of BDD.
-        \item
-            Automating things, and continuous integration~\cite{ci} is an important
-            aspect of BDD.
-        \item
-            BDD uses expectations instead of assertions
-    \end{itemize}
-\end{quote}
 
@@
 
-\begin{quote}
-    TBD: Important points we want to say here:
-    \begin{itemize}
-        \item
-            We use Doctest and Hspec to document behavior.
-            (in contrast to other communities)
-        \item
-            Some features are missing from both, Doctest and
-            Hspec.
-        \item
-            We extend them accordingly, and show how they can be used
-            in combination to design, document and test Haskell code.
-    \end{itemize}
-\end{quote}
+In this paper, we explore how to apply BDD to Haskell.
+BDD approaches
+TDD and acceptance test-driven development (ATDD)
+with a new terminology and mindset. 
+For the purpose of this paper we only focus on the TDD aspect of BDD.
 
-In this paper, we explore how to apply BDD to Haskell.
-We brought a test method from other programming language community and
-integrated it with a Haskell-original test method.
+In BDD, all behavior should be documented, together with some
+\emph{evidence} for that behavior.
+for this purpose, 
+we brought a test method from other programming language community and
+extended missing features.
 In particular, we introduced Doctest~\cite{doctest} and
-integrated it with QuickCheck~\cite{quickcheck} as a BDD tool.
-We also make use of Hspec inspired by RSpec~\cite{rspec} for
-BDD in Haskell.
+integrated it with QuickCheck~\cite{quickcheck}
+for user documentation.
+For code documentation, we also make use of Hspec inspired by RSpec~\cite{rspec}
+and we proposed some combinators to it.
+
+Automating such as test discovery and continuous integration (CI) ~\cite{ci} 
+is an important aspect of BDD. TBD: need to say somting.
 
@@
     Introduce a set of combinators to set expectations about
-    \emph{code examples} for use with Hspec.
+    \emph{code examples} for use with Hspec. (Section \ref{sec:for-developers})
 
 \item
-    Improve CI support for Doctest and Hspec
+    Improve CI support for Doctest and Hspec (Section \ref{sec:TBD})
 
@@
 
-\noindent Our Doctest implementation
+\noindent Our BDD procedure
 can be used for design, documentation and automatic testing.
@@
 motivates Haskellers to write test code.
+
+\begin{quote}
+    TBD: Important points we want to say about BDD:
+    \begin{itemize}
+        \item
+            BDD uses expectations instead of assertions
+    \end{itemize}
+\end{quote}
 
@@
 \begin{quote}
-    \emph{talk about CI}
-\end{quote}
-
-\begin{quote}
 \emph{This is optional, so we can skip it, if necessary.}

### 2012-05-28 paper.tex
typo fix.

@@
 For example, it does not ensure that division by zero does not occur.
-Another example is lows of type classes. 
-For instance, the {\tt Monad} class has three Monad lows 
+Another example is laws of type classes. 
+For instance, the {\tt Monad} class has three Monad laws 
 that its instance \emph{should} satisfy
@@
 In BDD, all behavior should be documented, together with some
-\emph{evidence} for that behavior.
+\emph{evidence}\/s for that behavior.
 for this purpose, 

### 2012-05-28 paper.tex
wordings.

@@
 \emph{evidence}\/s for that behavior.
-for this purpose, 
+For this purpose, 
 we brought a test method from other programming language community and
@@
 \item[user documentation]
-    TBD: Explanation of functions.
+    Explanation of functions for users.
 \item[code documentation]
-    TBD: Explanation of functions.
+    Explanation of functions for developers.
 
@@
 and for developers.
-evidences for users are those which are useful as user documentation.
+Evidences for users are those which are useful as user documentation.
 By contrast, evidences for developers are
-those which are not useful as user documentation.
+those which are not useful as user documentation
+but should be written as code documentation.
 A good example of evidences for developers
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evidence as user documentation}
+\subsection{Evidences as user documentation}
 \label{sec:for-users}
 
-evidences for users should be documented with Haddock markup syntax.
+Evidences for users should be documented with Haddock markup syntax.
 For this purpose, we propose the followings:
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evidence as code documentation (TBD: Is this still
+\subsection{Evidences as code documentation (TBD: Is this still
 appropriate here?}

### 2012-05-28 paper.tex
restructuring sections.

@@
 
+\subsection{Doctest and Haddock}
+
 A good candidate for this is Doctest in Python community.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Hspec}
+
 To describe evidences for developers, we can use Hspec.
+Hspec provides an EDSL for defining specs.  A spec organizes evidence in a
+tree structure, and it is defined with \texttt{it} and \texttt{describe}.
+
+\begin{quote}
+\small
+\begin{verbatim}
+import Test.Hspec.Monadic
+spec :: Spec
+spec = do
+  describe "reverse" $ do
+    it "reverses a list" $
+      reverse [1 :: Int, 2, 3] @?= [3, 2, 1]
+
+    it "gives the original list, if applied twice" $
+      property $ \xs ->
+        (reverse . reverse) xs == (xs :: [Int])
+\end{verbatim}
+\end{quote}
+
+\noindent The {\tt it} function combines a text description and
+an evidence and makes a spec.
+Examples can be arbitrary HUnit assertions.  Properties must be of type {\tt
+Property}; QuickCheck's \texttt{property} function can be used to turn
+arbitrary properties into a \texttt{Property}.
+
+The \texttt{describe} function combines a list of specs into a larger spec.  It
+can be used to organize a spec as a tree.
+Internally, a \texttt{Writer} is used to build up specs in a monadic way.
+
+A spec can be run with the \texttt{hspec} function.  When a spec is run, all
+evidences are evaluated, and failing evidences are reported to developers.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Problem Summary}
+
 Here is a summary of the problem:
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evidences as user documentation}
+\subsection{Evidences for users}
 \label{sec:for-users}
@@
 \end{itemize}
+
+%%%%%%%%%%%%%%%%
+\subsubsection{Markup for Examples}
 
@@
 \end{quote}
+
+%%%%%%%%%%%%%%%%
+\subsubsection{Markup for Properties}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evidences as code documentation (TBD: Is this still
-appropriate here?}
-\label{sec:for-developers}
+\subsection{Evidences for developers}
 
-\begin{quote}
-    \emph{TBD: Behavior should be replaced with evidences in this subsection?}
-\end{quote}
+We propose following extensions for Hspec to 
+describe evidences for developers.
 
-If we have evidence that does not fit into the user documentation, but would
-rather be part of a design document, we use Hspec.
-
-Hspec provides an EDSL for defining specs.  A spec organizes evidence in a
-tree structure, and it is defined with \texttt{it} and \texttt{describe}.
-
-\begin{quote}
-\small
-\begin{verbatim}
-import Test.Hspec.Monadic
-spec :: Spec
-spec = do
-  describe "reverse" $ do
-    it "reverses a list" $
-      reverse [1 :: Int, 2, 3] @?= [3, 2, 1]
-
-    it "gives the original list, if applied twice" $
-      property $ \xs ->
-        (reverse . reverse) xs == (xs :: [Int])
-\end{verbatim}
-\end{quote}
-
-\noindent The {\tt it} function combines a text description for a behavior and
-an evidence that gives some confidence in that behavior into a
-spec.
-Examples can be arbitrary HUnit assertions.  Properties must be of type {\tt
-Property}; QuickCheck's \texttt{property} function can be used to turn
-arbitrary properties inot a \texttt{Property}.
-
-The \texttt{describe} function combines a list of specs into a larger spec.  It
-can be used to organize a spec as a tree.
-Internally, a \texttt{Writer} is used to build up specs in a monadic way.
-
-\begin{quote}
-    \emph{TBD: provide an example?}
-\end{quote}
-
-A spec can be run with the \texttt{hspec} function.  When a spec is run, all
-evidences are evaluated, and failing evidences are
-reported to the user (FIXME: user does not fit our definition of \emph{user}
-here!).
-
-\begin{quote}
-    \emph{TBD: should user be replaced with developer?}
-\end{quote}
-
-\subsection{Setting expectations}
+%%%%%%%%%%%%%%%%
+\subsubsection{Setting expectations}
 
@@
 
+%%%%%%%%%%%%%%%%
 \subsubsection{Expecting exceptions}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{\emph{optional} Automatic discovery of specs}

### 2012-05-28 paper.tex
brushing up.

@@
 integrated it with QuickCheck~\cite{quickcheck}
-for user documentation.
-For code documentation, we also make use of Hspec inspired by RSpec~\cite{rspec}
+for evidences for users.
+To specify evidences for developers,
+we make use of Hspec inspired by RSpec~\cite{rspec}
 and we proposed some combinators to it.
@@
 Automating such as test discovery and continuous integration (CI) ~\cite{ci} 
-is an important aspect of BDD. TBD: need to say somting.
+is an important aspect of BDD.
+We improve CI suport for Doctest and Hspec.
 
@@
     Introduce a set of combinators to set expectations about
-    \emph{code examples} for use with Hspec. (Section \ref{sec:for-developers})
+    code examples (simply ``code'' hereafter) for use with Hspec. (Section \ref{sec:for-developers})
 
@@
 motivates Haskellers to write test code.
-
-\begin{quote}
-    TBD: Important points we want to say about BDD:
-    \begin{itemize}
-        \item
-            BDD uses expectations instead of assertions
-    \end{itemize}
-\end{quote}
 
@@
     Expectations are used to define the expected outcome of a code
-    example.  They serve the same purpose as assertions in xUnit
+    example.  They serve the same purpose as \emph{assertion}\/s in xUnit
     testing frameworks, but use a different terminology.
@@
 
-A good candidate for this is Doctest in Python community.
+A good candidate to describe evidences for users is Doctest in Python community.
 It extends Python's documentation strings so that they
@@
         Hspec does not yet provide a mechanism to set expectations
-        about the outcome of code examples.  It uses HUnit assertions
+        about the outcome of examples.  It uses HUnit assertions
         instead.
@@
 We propose several combinators, that can be used to set expectations
-about the outcome of code examples.  And we build them on top of
+about the outcome of examples.  And we build them on top of
 HUnit.
@@
 
-\subsection{Combinators for expectations}
+\subsection{Hspec combinators for expectations}
 

### 2012-05-28 paper.tex
evidences -> evidence.

@@
 integrated it with QuickCheck~\cite{quickcheck}
-for evidences for users.
-To specify evidences for developers,
+for evidence for users.
+To specify evidence for developers,
 we make use of Hspec inspired by RSpec~\cite{rspec}
@@
 
-We can categorize evidences into two types: for users and
+We can categorize evidence into two types: for users and
 and for developers.
-Evidences for users are those which are useful as user documentation.
-By contrast, evidences for developers are
+Evidence for users are those which are useful as user documentation.
+By contrast, evidence for developers are
 those which are not useful as user documentation
 but should be written as code documentation.
-A good example of evidences for developers
+A good example of evidence for developers
 is a corner example registered in a bug tracking system.
@@
 BDD tools should be treat both of them and provide functionality
-to extract evidences for users as user documentation.
+to extract evidence for users as user documentation.
 If we can archive this, BDD tools can be used for
@@
 
-A good candidate to describe evidences for users is Doctest in Python community.
+A good candidate to describe evidence for users is Doctest in Python community.
 It extends Python's documentation strings so that they
@@
 
-To describe evidences for developers, we can use Hspec.
+To describe evidence for developers, we can use Hspec.
 Hspec provides an EDSL for defining specs.  A spec organizes evidence in a
@@
 A spec can be run with the \texttt{hspec} function.  When a spec is run, all
-evidences are evaluated, and failing evidences are reported to developers.
+evidence are evaluated, and failing evidence are reported to developers.
 
@@
     The Haskell community does not have common syntax to describe
-    evidences with Haddock markup syntax.
+    evidence with Haddock markup syntax.
 \item
     The Haskell community does not have a BDD tool for
-    evidences for users.
+    evidence for users.
 \item
@@
 
-To be able to specify evidences for users, we propose to add new
+To be able to specify evidence for users, we propose to add new
 markup to Haddock and to use Doctest integrated with QuickCheck. To
-describe evidences for developers, we propose to use Hspec also
+describe evidence for developers, we propose to use Hspec also
 integrated with QuickCheck.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evidences for users}
+\subsection{Evidence for users}
 \label{sec:for-users}
 
-Evidences for users should be documented with Haddock markup syntax.
+Evidence for users should be documented with Haddock markup syntax.
 For this purpose, we propose the followings:
@@
 
-\noindent Evidences in comments are automatically
+\noindent Evidence in comments are automatically
 tested by the Doctest command. We also provided a library
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evidences for developers}
+\subsection{Evidence for developers}
 
 We propose following extensions for Hspec to 
-describe evidences for developers.
+describe evidence for developers.
 
@@
 As we described, Haddock version 2.7.2 or earlier provides markup syntax for
-code blocks, but does not provide any markup syntax for evidences.
+code blocks, but does not provide any markup syntax for evidence.
 We implemented markup support for examples ({\tt >>>}) and it is available in
@@
 
-TBD(@sol): how to obtain evidences from Haddock markup.
+TBD(@sol): how to obtain evidence from Haddock markup.
 Old version uses Haddock API but now uses GHC API directly. Why?
@@
 \noindent Note that the {\tt placeholders} library provide better keyword such as {\tt notImplemented} and {\tt todo}.
-Next we write evidences for users in Haddock markup:
+Next we write evidence for users in Haddock markup:
 
@@
 \noindent We should execute Doctest to ensure no syntax errors.
-Also, we specifies evidences for developers with Hspec:
+Also, we specifies evidence for developers with Hspec:
 

### 2012-05-28 paper.tex
prefer "user documentation" and "code documentation" over "evidence for user" and "evidence for developer"
Need to check this again.

@@
 integrated it with QuickCheck~\cite{quickcheck}
-for evidence for users.
-To specify evidence for developers,
+for user documentation.
+For code documentation,
 we make use of Hspec inspired by RSpec~\cite{rspec}
@@
 \noindent Note that the {\tt placeholders} library provide better keyword such as {\tt notImplemented} and {\tt todo}.
-Next we write evidence for users in Haddock markup:
+Next we write user documentation in Haddock markup:
 
@@
 \noindent We should execute Doctest to ensure no syntax errors.
-Also, we specifies evidence for developers with Hspec:
+Also, we write code documentation with Hspec:
 

### 2012-05-28 paper.tex
evidences -> evidence.
label and ref.

@@
 In BDD, all behavior should be documented, together with some
-\emph{evidence}\/s for that behavior.
+\emph{evidence} for that behavior.
 For this purpose, 
@@
     Implement Doctest in Haskell and add support for QuickCheck properties to Doctest (Section
-    \ref{sec:for-users} and Section \ref{sec:doctest})
+    \ref{sec:user-doc} and Section \ref{sec:doctest})
 
@@
     Introduce a set of combinators to set expectations about
-    code examples (simply ``code'' hereafter) for use with Hspec. (Section \ref{sec:for-developers})
+    code examples for use with Hspec. (Section \ref{sec:code-doc} and Section \ref{sec:hspec-comb})
 
@@
 \item
-    The Haskell community does not have a BDD tool for
-    evidence for users.
+    The Haskell community does not have a BDD tool for user documentation.
 \item
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evidence for users}
-\label{sec:for-users}
+\subsection{User documentation}
+\label{sec:user-doc}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evidence for developers}
+\subsection{Code documentation}
+\label{sec:code-doc}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Hspec combinators for expectations}
+\label{sec:hspec-comb}
 

### 2012-05-28 paper.tex
TDD: what's it?

@@
 
-In other programming language communities,
-behavior driven development (BDD) is getting much attention, and is
-slowly replacing test driven development (TDD).
-In BDD, test code is written with specification terminology.
-In other words, specifications can be used for automatic testing.
+In other programming language communities, behavior driven development
+(BDD) is getting much attention, and is slowly replacing test driven
+development (TDD).  TDD is an iterative style of software development
+where the developer first writes a few test cases, then implements the
+software in the simplest possible way which makes the test cases pass.
+That could mean yielding a constant value at first.  Subsequently,
+more test cases are written to reqire more details and drive the
+implementation closer to the desired goal.  Rinse and repeat.  BDD
+includes TDD, amongst other software development methodologies, but
+with a terminology and mindset  shifted towards specification instead
+of testing.  In BDD, specifications can be used for automatic testing.
 
 In this paper, we explore how to apply BDD to Haskell.
-BDD approaches
-TDD and acceptance test-driven development (ATDD)
-with a new terminology and mindset. 
 For the purpose of this paper we only focus on the TDD aspect of BDD.

### 2012-05-28 paper.tex
writing related work.

@@
 \section{Open problems and related work}
+
+SmallCheck~\cite{smallcheck} is a tool for
+automatic exhaustive testing at a certain level.
+Both Doctest and Hspec do not have support for SmallCheck
+at this moment.
+
+The {\tt test-framework} family is a TDD tool, 
+which currently integrates HUnit, QuickCheck, SmallCheck, and Doctest.
+Automatic test discovery based on prefixes of function name
+and is implemented using Template Haskell.
+
 \begin{itemize}
-    \item literate programming.
-    \item hunit.
-    \item test-framework.
-    \item smallCheck.
-    \item And others?
     \item The {\tt testpack} package provides a function that is
@@
         ignore.
+
+        @sol: Let's add this.
 

### 2012-05-28 paper.tex
small change for intro.
Do not use "evidence for developers"

@@
 we make use of Hspec inspired by RSpec~\cite{rspec}
-and we proposed some combinators to it.
+and we proposed some combinators to set expectations
+about \emph{code example}\/s.
 
@@
 
-To describe evidence for developers, we can use Hspec.
-Hspec provides an EDSL for defining specs.  A spec organizes evidence in a
-tree structure, and it is defined with \texttt{it} and \texttt{describe}.
+If we have evidence that does not fit into the user documentation, but
+would rather be part of a design document, we use Hspec.  Hspec
+provides an EDSL for defining specs.  A spec organizes evidence in a
+tree structure, and it is defined with \texttt{it} and
+\texttt{describe}.
 
@@
 \subsection{QuickCheck}
+
+\emph{TBD What does this have to do with the topic of the paper?}
 

### 2012-05-28 paper.tex
The section about QuickCheck is not related to the paper's topic

--- 2012-05-28 paper.tex
@@
 
+\emph{TBD What does this have to do with the topic of the paper?}
+
 It is enough to test properties which do not have parameters just one.

### 2012-05-28 paper.tex
Use property or example instead of evidence, it fit's better here

--- 2012-05-28 paper.tex
@@
 
-\noindent The {\tt it} function combines a text description and
-an evidence and makes a spec.
+\noindent The {\tt it} function combines a text description for a
+behavior and a property or an example that gives some confidence in
+that behavior into a spec.
 Examples can be arbitrary HUnit assertions.  Properties must be of type {\tt
@@
 
-A spec can be run with the \texttt{hspec} function.  When a spec is run, all
-evidence are evaluated, and failing evidence are reported to developers.
+A spec can be run with the \texttt{hspec} function.  When a spec is
+run, all properties and examples are evaluated, and failing properties
+and examples are reported to the developer.
 

### 2012-05-28 paper.tex
adding ref for TH.
removing the section about quickcheck.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{QuickCheck}
-
-\emph{TBD What does this have to do with the topic of the paper?}
-
-It is enough to test properties which do not have parameters just one.
-However, QuickCheck 2.3 or earlier tests
-properties which do not have parameters 100 times.
-More precisely, the evaluation results are memorized and they are
-compared with {\tt True} 100 times.
-The current maintainer of QuickCheck modified so that
-properties which do not have parameters
-are tested just once in its source repository.
-This feature will be available in the future release.
+%% \subsection{QuickCheck}
+
+%% \emph{TBD What does this have to do with the topic of the paper?}
+%% \emph{Acknowledgment is enough.}
+
+%% It is enough to test properties which do not have parameters just one.
+%% However, QuickCheck 2.3 or earlier tests
+%% properties which do not have parameters 100 times.
+%% More precisely, the evaluation results are memorized and they are
+%% compared with {\tt True} 100 times.
+%% The current maintainer of QuickCheck modified so that
+%% properties which do not have parameters
+%% are tested just once in its source repository.
+%% This feature will be available in the future release.
 
@@
 Automatic test discovery based on prefixes of function name
-and is implemented using Template Haskell.
+and is implemented using Template Haskell~\cite{template-haskell}.
+All examples and properties must be written in a single
+source file.
 

### 2012-05-28 paper.tex
Add short sketch for algorithm to close properties

@@
 
-\begin{quote}
-    \emph{TBD: describe the algorithm}
-\end{quote}
-
 \noindent In this case, we should prepend the ``\verb|\xs -> |" string
 to the property.
+
+\begin{quote}
+    TBD: describe the algorithm
+
+    Below is the schematic algorithm, from my original mail to kazu:
+\end{quote}
+
+\noindent \verb|xs| is a free variable here.  So we would need to
+close that term by transforming it to:
+
+\begin{quote}\small\begin{verbatim}
+\xs -> sort xs == sort (sort (xs::[Int]))
+\end{verbatim}\end{quote}
+
+\noindent I imagine this could work by performing those steps:
+
+\begin{enumerate}
+
+    \item
+        parse the property with e.g. haskell-src (never used it, so
+        I'm not sure if it allows you to parse code snippets)
+
+    \item
+        Load the module containing the DocProperty with hint, and try
+        to look up all variables used by the property (\verb|sort| and
+        \verb|xs| for the example above).
+
+    \item
+        Close the term over all variables, for which lookup failed,
+        load it with hint and test it with quickCheck.
+\end{enumerate}
 

### 2012-05-28 paper.tex
Extend TBD note
Whitespace changes

@@
 
-To test properties, we also uses GHCi.
-Each property is enclosed with
-the ``\verb|quickCheck (|'' and ``\verb|)|'' string
-and GHCi evaluates it.
-Success or failures is decided according to result string from GHCi.
-
-If properties contain free variables,
-we need to add lambda abstraction prefix before prepending
-the `quickCheck'' string.
-To check if properties contain free variables,
-we also use GHCi for easy implementation at this moment.
-GHCi reports free variables if found:
+To test properties, we also use GHCi.  Each property is enclosed
+within ``\verb|quickCheck (|'' and ``\verb|)|'' string and GHCi
+evaluates it.  Success or failures is decided according to result
+string from GHCi.
+
+If properties contain free variables, we need to add lambda
+abstraction prefix before prepending the `quickCheck'' string.  To
+check if properties contain free variables, we also use GHCi for easy
+implementation at this moment.  GHCi reports free variables if found:
 
@@
 
-    Below is the schematic algorithm, from my original mail to kazu:
+    Below is the schematic algorithm, from my original mail to kazu.
+    Can we somehow get rid of haskell-src, and hint, and only describe
+    the algorithm?
 \end{quote}

### 2012-05-28 paper.tex
Add result of package survey to paper (8 out of 23 have tests)
More survey results included for now (we can always strip this)

@@
 
-Haskell programmers, however, is apt to rely on the type system too much
-and tend not to write test code.
-For example, Haskell Platform 2011.4.0.0 contains
-TBD packages but TBD of them have no test code.
-Lacking test code also means that typical usage examples are not provided.
+Haskell programmers, however, is apt to rely on the type system too
+much and tend not to write test code.  For example, the Haskell
+Platform 2011.4.0.0 contains 23 packages (not counting the 19 core
+libraries that come with GHC), but only 8 of them a test suite (5
+provide example code, 10 have no tests nor examples).  Lacking test
+code also means that typical usage examples are not provided.
 

### 2012-05-28 paper.tex
Change comments, so that it is clear who said what
This works better, if three people work on the paper.

@@
 
-\emph{IMPORTANT: @sol has to release a new version of hspec, that
+\emph{IMPORTANT: sol: I have to release a new version of hspec, that
 includes the Specs-to-Spec and hspecX-to-hspec changes.  Otherwise the
@@
 %% \begin{quote}
-%%     @kazu: We e.g. say ill-behaved Monoid instance.  So is
+%%     sol: We e.g. say ill-behaved Monoid instance.  So is
 %%     ``behavior'' really a bad word?  But I agree that some people in
@@
 
-TBD(@sol): how to obtain evidence from Haddock markup.
+TBD: kazu: how to obtain evidence from Haddock markup.
 Old version uses Haddock API but now uses GHC API directly. Why?
 
-TBD(@sol): escape from HUnit. Why?
-
-TBD(@sol): Old version executes one GHCi for each function examples.
+TBD: kazu: escape from HUnit. Why?
+
+TBD: kazu: Old version executes one GHCi for each function examples.
 But new version executes one GHCI per module.
@@
 \begin{quote}
-@kazu: Can we use \verb|reverse| instead of \verb|rev| here?
-\end{quote}
-
-\begin{quote}
-@sol: I used \verb|rev| just because it is short.
+sol: Can we use \verb|reverse| instead of \verb|rev| here?
+\end{quote}
+
+\begin{quote}
+kazu: I used \verb|rev| just because it is short.
 Actually I used \verb|reverse| once but it is overflowed
@@
 \begin{quote}
-    TBD: @kazu: Do we need this to make our case?  I'm asking, because
+    TBD: sol: Do we need this to make our case?  I'm asking, because
     it will require some work to get the language right.
@@
 
+\begin{quote}
+    sol: I think you can have several files with tests, but you have
+    to combine the test suites manually.
+\end{quote}
+
 \begin{itemize}
@@
 
-        @kazu: Do you think we should discuss this; or is it safe to
+        sol: Do you think we should discuss this; or is it safe to
         ignore.
 
-        @sol: Let's add this.
+        kazu: Let's add this.
 

### 2012-05-28 paper.tex
Wording

@@
 
-To test properties, we also use GHCi.  Each property is enclosed
-within ``\verb|quickCheck (|'' and ``\verb|)|'' string and GHCi
-evaluates it.  Success or failures is decided according to result
-string from GHCi.
-
-If properties contain free variables, we need to add lambda
+To test properties, we use GHCi.  Each property is enclosed within
+``\verb|quickCheck (|'' and ``\verb|)|'', and GHCi is used to evaluate
+that.  GHCi output is used to decide, whether the property succeeded,
+or failed.
+
+If a property contains free variables, we need to add lambda
 abstraction prefix before prepending the `quickCheck'' string.  To

### 2012-05-28 paper.tex
Use (sort . sort) instead of (sort (sort ..))

@@
 \begin{verbatim}
-prop> \xs -> sort xs == sort (sort (xs::[Int]))
+prop> \xs -> sort xs == (sort . sort) (xs :: [Int])
 \end{verbatim}
@@
 \begin{verbatim}
-prop> sort xs == sort (sort (xs::[Int]))
+prop> sort xs == (sort . sort) (xs :: [Int])
 \end{verbatim}
@@
 \begin{verbatim}
-Prelude Data.List> sort xs == sort (sort (xs::[Int]))
+Prelude Data.List> sort xs == (sort . sort) (xs :: [Int]))
 <interactive>:1:6: Not in scope: `xs'
@@
 \begin{quote}\small\begin{verbatim}
-\xs -> sort xs == sort (sort (xs::[Int]))
+\xs -> sort xs == (sort . sort) (xs :: [Int])
 \end{verbatim}\end{quote}

### 2012-05-28 paper.tex
Remove trailing whitespace
Introduce implicitly quantified properties
Describe how properties are tested

@@
 
-\noindent Evidence in comments are automatically
+\noindent We call such properties \emph{implicitly quantified}.  The
+given property is implicitly quantified over \verb|xs|.
+
+Evidence in comments are automatically
 tested by the Doctest command. We also provided a library
@@
 
-To test properties, we use GHCi.  Each property is enclosed within
-``\verb|quickCheck (|'' and ``\verb|)|'', and GHCi is used to evaluate
-that.  GHCi output is used to decide, whether the property succeeded,
-or failed.
-
-If a property contains free variables, we need to add lambda
-abstraction prefix before prepending the `quickCheck'' string.  To
-check if properties contain free variables, we also use GHCi for easy
-implementation at this moment.  GHCi reports free variables if found:
-
-\begin{quote}
-\small
-\begin{verbatim}
-Prelude Data.List> sort xs == (sort . sort) (xs :: [Int]))
-<interactive>:1:6: Not in scope: `xs'
-<interactive>:1:24: Not in scope: `xs'
-\end{verbatim}
-\end{quote}
-
-\noindent In this case, we should prepend the ``\verb|\xs -> |" string
-to the property.
-
-\begin{quote}
-    TBD: describe the algorithm
-
-    Below is the schematic algorithm, from my original mail to kazu.
-    Can we somehow get rid of haskell-src, and hint, and only describe
-    the algorithm?
-\end{quote}
-
-\noindent \verb|xs| is a free variable here.  So we would need to
-close that term by transforming it to:
+\begin{quote}
+    sol: Should we use verify instead of test here?
+\end{quote}
+
+\noindent QuickCheck and GHCi are used to test properties.  Before we
+test a property, we bring the module that contains the Haddock comment
+from which we extracted the property into scope.
+
+Implicitly quantified properties need to be closed before they can be
+tested.  This requires the following steps.
+\begin{enumerate}
+    \item
+        Find all free variables of the property
+    \item
+        Try to lookup each free variable in the module context
+    \item
+        Close the property over all free variables, for which lookup
+        failed
+\end{enumerate}
+
+\noindent Here is an example:
+
+\begin{quote}\small\begin{verbatim}
+sort xs == (sort . sort) (xs :: [Int]))
+\end{verbatim}\end{quote}
+
+\noindent Both \verb|xs| and \verb|sort| are free variables, but
+\verb|sort| is bound by the module context.  So we need to close the
+term over \verb|xs|.
 
@@
 \end{verbatim}\end{quote}
-
-\noindent I imagine this could work by performing those steps:
-
-\begin{enumerate}
-
-    \item
-        parse the property with e.g. haskell-src (never used it, so
-        I'm not sure if it allows you to parse code snippets)
-
-    \item
-        Load the module containing the DocProperty with hint, and try
-        to look up all variables used by the property (\verb|sort| and
-        \verb|xs| for the example above).
-
-    \item
-        Close the term over all variables, for which lookup failed,
-        load it with hint and test it with quickCheck.
-\end{enumerate}
 

### 2012-05-29 paper.tex
adding comments.
revert mis operation: Conclusion is back.
using quantified.

@@
 
+\item[non-quantified property]
+    Property which does not take parameters.
+
+\item[quantified property]
+    Property which takes parameters.
+
 \item[behavior]
@@
 
+\begin{quote}
+\emph{kazu: if this subsection is long, we can remove the following paragraph.}
+\end{quote}
+
 The \texttt{describe} function combines a list of specs into a larger spec.  It
@@
 \noindent To specify properties, we also propose a new markup. In this paper, we
-use the {\tt prop>} prefix. We can write properties which do not take
-parameters as follows:
+use the {\tt prop>} prefix. We can write non-quantified properties as follows:
 
@@
 
-\noindent Of course, we allowed properties which take parameters by
-using closed lambda terms.
+\noindent Of course, we allowed quantified properties by using closed
+lambda terms.
 Type signatures can be specified for each parameter:
@@
 \end{verbatim}\end{quote}
+
+\begin{quote}
+\emph{kazu: are parentheses necessary? how about following?}
+\begin{verbatim}
+  launchMissiles >>=
+  (`shouldBe` Left "not implemented")
+  launchMissiles `shouldReturn` Left "not implemented"
+\end{verbatim}
+\end{quote}
 
@@
 
-%% It is enough to test properties which do not have parameters just one.
+%% It is enough to test non-quantified properties just one.
 %% However, QuickCheck 2.3 or earlier tests
-%% properties which do not have parameters 100 times.
+%% non-quantified properties 100 times.
 %% More precisely, the evaluation results are memorized and they are
@@
 %% The current maintainer of QuickCheck modified so that
-%% properties which do not have parameters
+%% non-quantified properties
 %% are tested just once in its source repository.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \section{Conclusion}
-
-%% \appendix
-
-%% \section{Package Survey}
-
-%% TBD
+\section{Conclusion}
+
+TBD
 
@@
 Nick Smallbone for having improved QuickCheck so that
-properties which do not have parameters are tested just once.
+non-quantified properties are tested just once.
 

### 2012-05-29 paper.tex
talk about Hspec.

@@
 \begin{quote}
-\emph{kazu: if this subsection is long, we can remove the following paragraph.}
-\end{quote}
-
-The \texttt{describe} function combines a list of specs into a larger spec.  It
-can be used to organize a spec as a tree.
-Internally, a \texttt{Writer} is used to build up specs in a monadic way.
-
+\emph{kazu: How implemneted the ``property'' function? Trystan or Simon?}
+\end{quote}
+
+The \texttt{describe} function combines a list of specs into a larger spec.
+%% It can be used to organize a spec as a tree.
+%% Internally, a \texttt{Writer} is used to build up specs in a monadic way.
 A spec can be run with the \texttt{hspec} function.  When a spec is
@@
 
+One problem of the original Hspec is that it still uses HUnit
+assertions (e.g. verb|@?=| in the example above) , which are TDD
+terminologies. In BDD, they are replaced with design terminologies,
+{\it expectation}.
+The following Rspec example shows the ``should'' expectation keyword:
+
+\begin{quote}
+\small
+\begin{verbatim}
+23.should eq(23)
+\end{verbatim}
+\end{quote}
+
 \begin{quote}
@@
 \begin{itemize}
-    \item Motivate \verb|test-shouldbe|
-
-        BDD frameworks use \emph{should} or \emph{must} instead of
-        \emph{assert}.
-
-        Here is an example that uses RSpec.
-
-        \begin{verbatim}
-23.should eq(23)
-        \end{verbatim}
-
-        RSpec provides a mechanism to expect exceptions.
+    \item RSpec provides a mechanism to expect exceptions.
 

### 2012-05-29 paper.tex
Remove redundant parentheses

@@
 launchMissiles >>=
-  (`shouldBe` (Left "not implemented"))
+  (`shouldBe` Left "not implemented")
 \end{verbatim}\end{quote}
@@
 \begin{quote}\small\begin{verbatim}
-launchMissiles `shouldReturn` (Left "not implemented")
+launchMissiles `shouldReturn` Left "not implemented"
 \end{verbatim}\end{quote}
@@
 \end{verbatim}\end{quote}
-
-\begin{quote}
-\emph{kazu: are parentheses necessary? how about following?}
-\begin{verbatim}
-  launchMissiles >>=
-  (`shouldBe` Left "not implemented")
-  launchMissiles `shouldReturn` Left "not implemented"
-\end{verbatim}
-\end{quote}
 

### 2012-05-29 paper.tex
adding \.
Remove comment about verify
QuickCheck can not verify anything, it can only try to find a counter
example.
brushing up Hspec section.

@@
 
-One problem of the original Hspec is that it still uses HUnit
-assertions (e.g. verb|@?=| in the example above) , which are TDD
-terminologies. In BDD, they are replaced with design terminologies,
-{\it expectation}.
+The problem of the original Hspec is that it does not provide
+a mechanism to express \emph{expectation}\/s with
+BDD terminology.
+For instance,
+it still uses HUnit
+assertions (e.g. \verb|@?=| in the example above), which are TDD
+terminologies. In BDD, they are replaced with design terminologies.
 The following Rspec example shows the ``should'' expectation keyword:
@@
 23.should eq(23)
+\end{verbatim}
+\end{quote}
+
+\noindent Another example is exceptions. Hspec does not yet provide a
+mechanism to express expectation on exceptions. The following is
+how to express expectation on an exception in Rspec:
+
+\begin{quote}
+\small
+\begin{verbatim}
+expect {launch_missiles}.to raise_error
 \end{verbatim}
@@
 \begin{itemize}
-    \item RSpec provides a mechanism to expect exceptions.
-
-        \begin{verbatim}
-expect {launch_missiles}.to raise_error
-        \end{verbatim}
-
-        Hspec does not yet provide a mechanism to set expectations
-        about the outcome of examples.  It uses HUnit assertions
-        instead.
-
-        HUnit does not provide a mechanism to expect exceptions.
-
     \item \emph{optional:} Motivate why it's important to integrate
@@
     Hspec does not provide combinators to express expectations with
-    BDD terminology
+    BDD terminology.
 \end{itemize}
@@
 \subsubsection{Testing properties}
-
-\begin{quote}
-    sol: Should we use verify instead of test here?
-\end{quote}
 

### 2012-05-29 paper.tex
Add important points we want to say in intro
Whitespace change

@@
 \emph{evidence} for that behavior.
-For this purpose,
-we brought a test method from other programming language community and
-extended missing features.
-In particular, we introduced Doctest~\cite{doctest} and
-integrated it with QuickCheck~\cite{quickcheck}
-for user documentation.
-For code documentation,
-we make use of Hspec inspired by RSpec~\cite{rspec}
-and we proposed some combinators to set expectations
-about \emph{code example}\/s.
+
+For this purpose, we brought a test method from other programming
+language community and extended missing features.  In particular, we
+introduced Doctest~\cite{doctest} and integrated it with
+QuickCheck~\cite{quickcheck} for user documentation.
+
+For code documentation, we make use of Hspec inspired by
+RSpec~\cite{rspec} and we proposed some combinators to set
+expectations about \emph{code example}\/s.
 
@@
 \section{Problem}
+
+\begin{quote}\begin{itemize}
+    \item
+        We want to use properties and examples to provide evidence for
+        behavior.
+    \item
+        Haddock already has markup support for interactive examples,
+        and Doctest can verify them.
+    \item
+        It is not possible to use properties in Haddock comments and
+        automatically test them.
+\end{itemize}\end{quote}
 
@@
 arbitrary properties into a \texttt{Property}.
-
-\begin{quote}
-\emph{kazu: How implemneted the ``property'' function? Trystan or Simon?}
-\end{quote}
 

### 2012-05-29 paper.tex
Introduction (wip)

@@
 In BDD, all behavior should be documented, together with some
-\emph{evidence} for that behavior.
+\emph{evidence} for that behavior.  Hspec inspired by
+RSpec~\cite{rspec} allows to document code with \emph{properties} and
+\emph{code examples}, and create an informal specification document
+from that.
+
+We made two observations:
+\begin{enumerate}
+    \item
+        Parts of a specification can be useful as user documentation.
+    \item
+        Doctest, a tool inspired by a Python module with the same
+        name~\cite{doctest}, already allows to verify
+        \emph{interactive examples} in Haddock comments.
+\end{enumerate}
+
+\noindent Doctest is not commonly described as a BDD tool, but we
+think it fits the picture.
+We show how to use Doctest and Hspec in combination to design,
+document and test Haskell code.
 
@@
 
-For code documentation, we make use of Hspec inspired by
-RSpec~\cite{rspec} and we proposed some combinators to set
-expectations about \emph{code example}\/s.
+\begin{itemize}
+    \item
+        We proposed some combinators to set expectations about
+        \emph{code examples}.
+\end{itemize}
 

### 2012-05-29 paper.tex
removing comment.
s/rev/reverse/

@@
 \usepackage{url}
+\usepackage{alltt}
 
@@
 
-\begin{quote}
-sol: Can we use \verb|reverse| instead of \verb|rev| here?
-\end{quote}
-
-\begin{quote}
-kazu: I used \verb|rev| just because it is short.
-Actually I used \verb|reverse| once but it is overflowed
-in the right hand side.
-\end{quote}
-
 First, we write a signature and leave the definition of the function undefined.
@@
 \begin{verbatim}
-rev :: [a] -> [a]
-rev = undefined
+reverse :: [a] -> [a]
+reverse = undefined
 \end{verbatim}
@@
 \small
-\begin{verbatim}
+\begin{alltt}
 -- |
--- 'rev' @xs@ returns the elements of @xs@
+-- 'reverse' @xs@ returns the elements of @xs@
 -- in reverse order. @xs@ must be finite.
 --
--- >>> rev [1,2,3]
+-- >>> reverse [1,2,3]
 -- [3,2,1]
 --
--- prop> rev [] == rev []
--- prop> rev (xs ++ ys) == rev ys ++ rev (xs::[Int])
--- prop> rev (rev xs) == (xs::[Int])
-
-rev :: [a] -> [a]
-rev = undefined
-\end{verbatim}
+-- prop> reverse [] == reverse []
+-- prop> reverse (xs ++ ys) == reverse ys ++ \(\hookleftarrow\)
+                               reverse (xs::[Int])
+-- prop> reverse (reverse xs) == (xs::[Int])
+
+reverse :: [a] -> [a]
+reverse = undefined
+\end{alltt}
 \end{quote}
@@
 \begin{verbatim}
-describe "rev" $ do
+describe "reverse" $ do
   it "returns the first element in the last" $
     property $ \xs -> not (null xs) ==>
-      head (rev xs) == last (xs :: [Int])
+      head (reverse xs) == last (xs :: [Int])
 
@@
     property $ \xs -> not (null xs) ==>
-      last (rev xs) == head (xs :: [Int])
+      last (reverse xs) == head (xs :: [Int])
 \end{verbatim}
@@
 \begin{verbatim}
-rev :: [a] -> [a]
-rev = foldl (flip (:)) []
+reverse :: [a] -> [a]
+reverse = foldl (flip (:)) []
 \end{verbatim}

### 2012-05-29 paper.tex
Remove redundant text from "Introduction"

@@
 In BDD, all behavior should be documented, together with some
-\emph{evidence} for that behavior.  Hspec inspired by
-RSpec~\cite{rspec} allows to document code with \emph{properties} and
-\emph{code examples}, and create an informal specification document
-from that.
+\emph{evidence} for that behavior.  Hspec, inspired by
+RSpec~\cite{rspec}, allows to document Haskell code with
+\emph{properties} and \emph{code examples}, verifies examples and
+tests properties, and creates an informal specification document.
 
@@
 
-For this purpose, we brought a test method from other programming
-language community and extended missing features.  In particular, we
-introduced Doctest~\cite{doctest} and integrated it with
-QuickCheck~\cite{quickcheck} for user documentation.
+Automation, such as test discovery and continuous integration (CI)
+~\cite{ci} is an important aspect of BDD.
+
+Our contributions are:
 
@@
 \item
-        We proposed some combinators to set expectations about
-        \emph{code examples}.
-\end{itemize}
-
-Automating such as test discovery and continuous integration (CI) ~\cite{ci}
-is an important aspect of BDD.
-We improve CI suport for Doctest and Hspec.
-
-Our contributions are:
-
-\begin{itemize}
-\item
-    Implement Doctest in Haskell and add support for QuickCheck properties to Doctest (Section
+    Add support for QuickCheck properties to Doctest (Section
     \ref{sec:user-doc} and Section \ref{sec:doctest})

### 2012-05-29 paper.tex
Add comment
Fix syntax error
Sate, that Doctest and Hspec miss some features

@@
 
-Automation, such as test discovery and continuous integration (CI)
-~\cite{ci} is an important aspect of BDD.
-
-Our contributions are:
+\begin{quote}
+    \emph{sol: @mkl, any idea where to put this?}
+
+\emph{Automation, such as test discovery and continuous integration (CI)
+~\cite{ci} is an important aspect of BDD.}
+\end{quote}
+
+Both, Doctest and Hspec miss some features.  So we extend them as
+needed.  Our contributions are:
 

### 2012-05-29 paper.tex
Add a comment
initial abstruct.

@@
 
-TBD: Asbtract will be written at the final moment.
-
-State the problem.
-
-Say why it’s an interesting problem.
-
-Say what your solution achieves.
-
-Say what follows from your solution.
+Many Haskell programmers tend not to write test code even though the
+strong type system of Haskell cannot ensure correct semantics.
+Without test code, it is difficult for users to understand usage of
+functions and it is difficult for developers to maintain semantics
+while progressing software. To motivate Haskell programmers to write
+test code, we propose to carry out behavior driven development with
+two tools: doctest integrated with QuickCheck for user documentation
+and Hspec with combinators extended for developer documentation. With
+our method, specification can be used for documentation and automatic
+testing.
+
+%% State the problem.
+%% Say why it's an interesting problem.
+%% Say what your solution achieves.
+%% Say what follows from your solution.
 
@@
 of testing.  In BDD, specifications can be used for automatic testing.
+
+\begin{quote}\begin{itemize}
+    \item sol: Should we state that BDD-style TDD has been pioneered
+        by RSpec?
+\end{itemize}\end{quote}
 

### 2012-05-29 paper.tex
removing a comment according to our discussion.

@@
 
-\begin{quote}
-    TBD: sol: Do we need this to make our case?  I'm asking, because
-    it will require some work to get the language right.
-
-    Note that the authors are not fundamentalist.  If functions to be
-    defined are simple enough, we might implement them first withtout
-    signatures.  And their signatures can be automatically inserted by
-    IDE (such as Emacs/vim with {\tt ghc-mod}) later if Haskell
-    compilers can infer correctly.  Also, we can copy-and-paste
-    examples and their results from Haskell interpreters to their
-    Haddock markup.
-\end{quote}
-
 \subsection{Automatic discover of specs}

### 2012-05-29 paper.tex
Problem wip

@@
 functions and it is difficult for developers to maintain semantics
-while progressing software. To motivate Haskell programmers to write
-test code, we propose to carry out behavior driven development with
-two tools: doctest integrated with QuickCheck for user documentation
-and Hspec with combinators extended for developer documentation. With
-our method, specification can be used for documentation and automatic
-testing.
+while extending software. To enhance the value of test code, we
+propose to carry out behavior driven development with two tools:
+doctest integrated with QuickCheck for user documentation and Hspec
+with combinators extended for developer documentation. With our
+method, specification can be used for documentation and automatic
+testing. This kill-three-birds-with-one-stone aspect would motivate
+Haskellers to write test code.
 
@@
 Another example is laws of type classes.
-For instance, the {\tt Monad} class has three Monad laws
+For instance, the {\tt Monoid} class has three Monoid laws
 that its instance \emph{should} satisfy
@@
 \item[evidence]
-    Evidence is either an example or a property.  It provides some confidence,
-    that a function exhibits a certain behavior.
+    Evidence provides some confidence, that a function exhibits a
+    certain behavior.  In Haskell we use examples and properties as
+    evidence.
 
@@
     \item
-        We want to use properties and examples to provide evidence for
-        behavior.
-    \item
-        Haddock already has markup support for interactive examples,
-        and Doctest can verify them.
-    \item
-        It is not possible to use properties in Haddock comments and
-        automatically test them.
-\end{itemize}\end{quote}
-
-In the Haskell community, HUnit and QuickCheck are commonly used %% what about SmallCheck?
-and the {\tt test-framework} package integrates them.
-We can categorize {\tt test-framework} as
-a tool for TDD.
-Before our work and Hspec, there was no widely used BDD tool for Haskell.
-
-BDD tools used in other programming language communities are
-mainly based on example tests,
-and do not support property tests.
-To our experience is, that it is not difficult to find properties for
-pure Haskell functions.
-So, when we bring BDD tools to
-Haskell community, we should integrate them with properties test.
-
-We can categorize evidence into two types: for users and
-and for developers.
-Evidence for users are those which are useful as user documentation.
-By contrast, evidence for developers are
-those which are not useful as user documentation
-but should be written as code documentation.
-A good example of evidence for developers
-is a corner example registered in a bug tracking system.
+        BDD tools for other programming languages use code examples as
+        evidence for behavior.
+    \item
+        When we apply BDD to Haskell we want to use properties and
+        examples to provide evidence for behavior.
+    \item
+        Hspec already has support for both.
+    \item
+        Haddock has markup support for interactive examples, and
+        Doctest can verify them.
+    \item
+        But it is not possible to use properties in Haddock comments
+        and automatically test them.
+\end{itemize} \end{quote}
+
+\begin{quote}
+Hspec can be used to document the behavior of a system in a way that
+is similar to a specification.  All facets of the system are described
+with examples and properties, including every possible corner case.
+
+We think that some evidence is useful as user documentation, hence we
+want to be able to include it in Haddock comments.  But there is also
+evidence that is not useful as user documentation.  E.g. that a
+particular \texttt{Monoid} instance is a monoid in the mathematical
+sense is self-evident.  So we do not want to include any examples or
+behaviors for that into user documentation, but we still want to be
+able to automatically test it.
+
+We propose to use Doctest to test evidence that is useful as user
+documentation, and Hspec for evidence that is not useful as user
+documentation.
+\end{quote}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Problem Summary}
+\subsection{Problem summary}
 
@@
 %%%%%%%%%%%%%%%%
-\subsubsection{Markup for Examples}
+\subsubsection{Markup for examples}
 
@@
 %%%%%%%%%%%%%%%%
-\subsubsection{Markup for Properties}
+\subsubsection{Markup for properties}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Open problems and related work}
+\section{Open problems} %% Future work?
+
+\subsection{Reusable spec}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Related work}
 

### 2012-05-29 paper.tex
Use assertEqual instead of @=?

--- 2012-05-29 paper.tex
@@
     it "reverses a list" $
-      reverse [1 :: Int, 2, 3] @?= [3, 2, 1]
+      assertEqual "" [3, 2, 1]
+                     (reverse [1 :: Int, 2, 3])
 
@@
 it still uses HUnit
-assertions (e.g. \verb|@?=| in the example above), which are TDD
+assertions, which are TDD
 terminologies. In BDD, they are replaced with design terminologies.

### 2012-05-29 paper.tex
Wording
using small letters in section titles.

@@
 In BDD, all behavior should be documented, together with some
-\emph{evidence} for that behavior.  Hspec, inspired by
-RSpec~\cite{rspec}, allows to document Haskell code with
-\emph{properties} and \emph{code examples}, verifies examples and
-tests properties, and creates an informal specification document.
+\emph{evidence} for that behavior.  Most existing BDD tools use
+\emph{code examples} as evidence for behavior.
+Hspec, a BDD tool for Haskell inspired by RSpec~\cite{rspec}, supports
+both code examples and \emph{properties} as evidence.
+It can verify examples and tests properties, and it can create an
+informal specification document from that.
 
@@
 Hspec can be used to document the behavior of a system in a way that
-is similar to a specification.  All facets of the system are described
-with examples and properties, including every possible corner case.
+is similar to a specification.  All facets of the system should be
+described with examples and properties, including every possible
+corner case.
 

### 2012-05-29 paper.tex
describing reusable spec.

--- 2012-05-29 paper.tex
@@
 
+\begin{itemize}
+\item Some type classes have their laws.
+\item We can use Hspec to enforce such properties.
+\item If such properties are reusable among instances,
+we can avoid boilerplates.
+\item reusable spec can do this.
+\end{itemize}
+
+\begin{quote}
+\small
+\begin{verbatim}
+shouldSatisfyMonoidLaws ::
+  (Eq a, Show a, Monoid a, Arbitrary a) => a -> Specs
+shouldSatisfyMonoidLaws t = do
+  describe "mempty" $ do
+    it "is a left identity" $ property $ \x ->
+      mempty <> x == x `asTypeOf` t
+
+    it "is a right identity" $ property $ \x ->
+      x <> mempty == x `asTypeOf` t
+
+  describe "<>" $ do
+    it "is associative" $ property $ \x y z ->
+      (x <> y) <> z == x <> (y <> z) `asTypeOf` t
+\end{verbatim}
+\end{quote}
+
+\noindent The following is an example to apply this laws
+to lists.
+
+\begin{quote}
+\small
+\begin{verbatim}
+describe "List as a Monoid" $
+  shouldSatisfyMonoidLaws (undefined :: [Int])
+\end{verbatim}
+\end{quote}
+
+\begin{itemize}
+\item we should prepare this kind of laws.
+\end{itemize}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-05-29 paper.tex
More work on "Problem"

@@
 Many Haskell programmers tend not to write test code even though the
-strong type system of Haskell cannot ensure correct semantics.
+strong type system of Haskell cannot always ensure correct semantics.
 Without test code, it is difficult for users to understand usage of
@@
 
-\begin{quote}\begin{itemize}
-    \item
-        BDD tools for other programming languages use code examples as
-        evidence for behavior.
-    \item
-        When we apply BDD to Haskell we want to use properties and
-        examples to provide evidence for behavior.
-    \item
-        Hspec already has support for both.
-    \item
-        Haddock has markup support for interactive examples, and
-        Doctest can verify them.
-    \item
-        But it is not possible to use properties in Haddock comments
-        and automatically test them.
-\end{itemize} \end{quote}
-
-\begin{quote}
 Hspec can be used to document the behavior of a system in a way that
@@
 particular \texttt{Monoid} instance is a monoid in the mathematical
-sense is self-evident.  So we do not want to include any examples or
-behaviors for that into user documentation, but we still want to be
-able to automatically test it.
-
-We propose to use Doctest to test evidence that is useful as user
-documentation, and Hspec for evidence that is not useful as user
-documentation.
-\end{quote}
-
+sense is self-evident.  So we do not want to include any evidence for
+that into user documentation, but we still want to automatically test
+it.
+
+We propose to use Doctest and Haddock to test evidence that is useful
+as user documentation, and Hspec for evidence that is not useful as
+user documentation.
+
+\begin{quote}
 BDD tools should be treat both of them and provide functionality
@@
 triple purpose: design, documentation and automatic testing.
+\end{quote}
 
 \subsection{Doctest and Haddock}
+
+\begin{quote}\begin{itemize}
+    \item
+        Haddock has markup support for interactive examples, and
+        Doctest can verify them.
+    \item
+        But it is not possible to use properties in Haddock comments
+        and automatically test them.
+\end{itemize} \end{quote}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Open problems} %% Future work?
+\section{Open problems and future work}
 

### 2012-05-29 paper.tex
section title change.
partially denied.
Wording

@@
 is similar to a specification.  All facets of the system should be
-described with examples and properties, including every possible
-corner case.
+described with evidence, including every possible corner case.
 
@@
 
-We propose to use Doctest and Haddock to test evidence that is useful
-as user documentation, and Hspec for evidence that is not useful as
-user documentation.
+We propose to use Doctest and Haddock for evidence that is useful as
+user documentation, and Hspec for evidence that is not useful as user
+documentation.
 
@@
 \begin{itemize}
-\item We use a preprocessor (http://hackage.haskell.org/package/hspec-discover) to automatically generate this boilerplate.
+\item We use a preprocessor (the {\tt hspec-discover} package) to automatically generate this boilerplate.
 \end{itemize}

### 2012-05-29 paper.tex
using the same style.
Wording of Problem -> Doctest and Haddock

@@
 
-\begin{quote}\begin{itemize}
-    \item
-        Haddock has markup support for interactive examples, and
-        Doctest can verify them.
-    \item
-        But it is not possible to use properties in Haddock comments
-        and automatically test them.
-\end{itemize} \end{quote}
-
-A good candidate to describe evidence for users is Doctest in Python community.
-It extends Python's documentation strings so that they
-can contain examples.
-Examples can be tested by the Python interpreter.
-
 The standard documentation mechanism for Haskell is
 Haddock~\cite{haddock}.
-Haddock defines markup syntax in Haskell's comment.
-Haddock provides markup for code blocks, indicated with ({\tt >}).
-For example, the {\tt containers} packages has the following code block markup:
+Haddock provides markup syntax for code blocks, indicated with ({\tt
+>}).
+For example, the {\tt containers} packages contains a Haddock comment
+with the following code block.
 
@@
 from example code.
+
+\begin{quote}\begin{itemize}
+    \item
+        Haddock has markup support for interactive examples, and
+        Doctest can verify them.
+    \item
+        But it is not possible to use properties in Haddock comments
+        and automatically test them.
+\end{itemize} \end{quote}
 

### 2012-05-29 paper.tex
More wording
Improve Problem -> Hspec

@@
 
-\noindent Though the code block itself is a useful feature,
-we think this kind of usage has problems.
+\noindent Though code blocks in Haddock comments are a useful feature,
+we still think that this particular usage has problems.
 If the equations above are examples, we cannot distinguish
@@
 import Test.Hspec.Monadic
-spec :: Spec
-spec = do
+main :: IO ()
+main = spec $ do
   describe "reverse" $ do
     it "reverses a list" $
-      assertEqual "" [3, 2, 1]
-                     (reverse [1 :: Int, 2, 3])
+      assertEqual "" [3, 2, 1] (reverse [1, 2, 3])
 
@@
 \noindent The {\tt it} function combines a text description for a
-behavior and a property or an example that gives some confidence in
-that behavior into a spec.
-Examples can be arbitrary HUnit assertions.  Properties must be of type {\tt
+behavior and some evidence into a spec.
+Code examples can be arbitrary HUnit assertions.  Properties must be of type {\tt
 Property}; QuickCheck's \texttt{property} function can be used to turn
@@
 A spec can be run with the \texttt{hspec} function.  When a spec is
-run, all properties and examples are evaluated, and failing properties
-and examples are reported to the developer.
+run, all evidence is tested, and behavior with evidence that does not
+hold is reported to the developer.
 

### 2012-05-29 paper.tex
Remove a note
"Introduction"

@@
 
-\emph{IMPORTANT: sol: I have to release a new version of hspec, that
-includes the Specs-to-Spec and hspecX-to-hspec changes.  Otherwise the
-code in this paper will not work!}
-
 %%%% Describe the problem.
@@
 Platform 2011.4.0.0 contains 23 packages (not counting the 19 core
-libraries that come with GHC), but only 8 of them a test suite (5
+libraries that come with GHC), but only 8 of them have a test suite (5
 provide example code, 10 have no tests nor examples).  Lacking test
@@
 both code examples and \emph{properties} as evidence.
-It can verify examples and tests properties, and it can create an
+It can verify examples and test properties, and it can create an
 informal specification document from that.
@@
 
-Both, Doctest and Hspec miss some features.  So we extend them as
-needed.  Our contributions are:
+Both, Doctest and Hspec are missing some features.  So we extend them
+as needed.  In particular, our contributions are:
 
@@
     Introduce a set of combinators to set expectations about
-    code examples for use with Hspec. (Section \ref{sec:code-doc} and Section \ref{sec:hspec-comb})
+    code examples for use with Hspec (Section \ref{sec:code-doc} and Section \ref{sec:hspec-comb})
 
@@
 \item
-    Show best current practice of BDD in Haskell using Doctest and
+    Show best current practice of BDD in Haskell, using Doctest and
     Hspec (Section \ref{ref:Best-current-practice})
@@
 
-\item[developer]
-    A programmer who develops a library.
-
-\item[user]
-    A programmer who uses a library.
-
 \item[example]
@@
 
-\item[non-quantified property]
-    Property which does not take parameters.
-
-\item[quantified property]
-    Property which takes parameters.
-
 \item[behavior]
@@
     TBD: BDD terminology for test suite.
+\end{description}
+
+\begin{quote}
+    \emph{sol: Can we remove what follows??}
+\end{quote}
+
+\begin{description}
+\item[developer]
+    A programmer who develops a library.
+
+\item[user]
+    A programmer who uses a library.
 
@@
     Explanation of functions for developers.
+
+\item[non-quantified property]
+    Property which does not take parameters.
+
+\item[quantified property]
+    Property which takes parameters.
 

### 2012-05-29 paper.tex
Add a comment

@@
 \begin{quote}
+    sol: remove?
+
 BDD tools should be treat both of them and provide functionality
@@
 we still think that this particular usage has problems.
+
+\begin{quote}
+    \emph{sol: I think the comment contains properties.  So the
+    argumentation makes not really sense?}
+\end{quote}
+
 If the equations above are examples, we cannot distinguish

### 2012-05-29 paper.tex
Uncomment comment, so that Markus can read it

@@
 
-%% \begin{quote}
-%%     sol: We e.g. say ill-behaved Monoid instance.  So is
-%%     ``behavior'' really a bad word?  But I agree that some people in
-%%     the community might associated it with something imperative.
-
-%%     One might also argue that ``should'' is not very mathematical, but
-%%     it is common language when we talk about mathematics.  E.g. Brent
-%%     Yorgey uses ``should'' a lot in the Typeclassopedia (e.g. ``\ldots
-%%     {\tt Alternative} should satisfy the monoid laws\ldots'')
-%% \end{quote}
+\begin{quote}
+    sol: We e.g. say ill-behaved Monoid instance.  So is
+    ``behavior'' really a bad word?  But I agree that some people in
+    the community might associated it with something imperative.
+
+    One might also argue that ``should'' is not very mathematical, but
+    it is common language when we talk about mathematics.  E.g. Brent
+    Yorgey uses ``should'' a lot in the Typeclassopedia (e.g. ``\ldots
+    {\tt Alternative} should satisfy the monoid laws\ldots'')
+\end{quote}
 

### 2012-05-29 paper.tex
Comment change
Change a comment

@@
     sol: We e.g. say ill-behaved Monoid instance.  So is
-    ``behavior'' really a bad word?  But I agree that some people in
+    ``behavior'' really a ``bad'' word?  But I agree that some people in
     the community might associated it with something imperative.
@@
     it is common language when we talk about mathematics.  E.g. Brent
-    Yorgey uses ``should'' a lot in the Typeclassopedia (e.g. ``\ldots
-    {\tt Alternative} should satisfy the monoid laws\ldots'')
+    Yorgey uses both ``should'' and ``behave'' a lot in the
+    Typeclassopedia (e.g. ``\ldots {\tt Alternative} should satisfy
+    the monoid laws\ldots'')
 \end{quote}
@@
 \begin{quote}
-    \emph{sol: I think the comment contains properties.  So the
-    argumentation makes not really sense?}
+    \emph{sol: I think the Haddock comment only contains properties.
+    So the argumentation is shaky?}
 \end{quote}

### 2012-05-29 paper.tex
Specs are forests!
Rspec -> RSpec + better example

@@
 provides an EDSL for defining specs.  A spec organizes evidence in a
-tree structure, and it is defined with \texttt{it} and
+tree structure (or a forest, to be accurate), and it is defined with \texttt{it} and
 \texttt{describe}.
@@
 terminologies. In BDD, they are replaced with design terminologies.
-The following Rspec example shows the ``should'' expectation keyword:
-
-\begin{quote}
-\small
-\begin{verbatim}
-23.should eq(23)
-\end{verbatim}
-\end{quote}
+The following RSpec example shows the ``should'' expectation keyword:
+
+\begin{quote}\small\begin{verbatim}
+[1, 2, 3].reverse.should eq([3, 2, 1])
+\end{verbatim}\end{quote}
 
@@
 mechanism to express expectation on exceptions. The following is
-how to express expectation on an exception in Rspec:
+how to express expectation on an exception in RSpec:
 

### 2012-05-29 paper.tex
Work on Problem -> Hspec

@@
     Expectations are used to define the expected outcome of a code
-    example.  They serve the same purpose as \emph{assertion}\/s in xUnit
+    example.  They serve the same purpose as \emph{assertions} in xUnit
     testing frameworks, but use a different terminology.
@@
 The \texttt{describe} function combines a list of specs into a larger spec.
-%% It can be used to organize a spec as a tree.
-%% Internally, a \texttt{Writer} is used to build up specs in a monadic way.
+It can be used to organize a spec as a tree.
+Internally, a \texttt{Writer} is used to build up specs in a monadic way.
+
 A spec can be run with the \texttt{hspec} function.  When a spec is
-run, all evidence is tested, and behavior with evidence that does not
+run, all evidence is tested.  Behavior with evidence that does not
 hold is reported to the developer.
 
-The problem of the original Hspec is that it does not provide
-a mechanism to express \emph{expectation}\/s with
-BDD terminology.
-For instance,
-it still uses HUnit
-assertions, which are TDD
-terminologies. In BDD, they are replaced with design terminologies.
-The following RSpec example shows the ``should'' expectation keyword:
+\subsubsection{Hspec not provide a mechanism for expectations}
+BDD frameworks use \emph{should} or \emph{must} instead of
+\emph{assert} to express expectations.  Here is an example that uses
+RSpec.
 
@@
 
-\noindent Another example is exceptions. Hspec does not yet provide a
-mechanism to express expectation on exceptions. The following is
-how to express expectation on an exception in RSpec:
-
-\begin{quote}
-\small
-\begin{verbatim}
+\noindent Hspec does not yet provide a mechanism to express
+\emph{expectations} with that terminology.  The developer has to use
+HUnit assertions instead.
+
+\subsubsection{Hspec provides no mechanism to expect exceptions}
+RSpec provides a mechanism to expect exceptions.
+
+\begin{quote}\small\begin{verbatim}
 expect {launch_missiles}.to raise_error
-\end{verbatim}
-\end{quote}
+\end{verbatim}\end{quote}
+
+\noindent Neither Hspec nor HUnit provide a convenient equivalent.
 

### 2012-05-29 paper.tex
Add TBD
Add a comment

@@
 If we have evidence that does not fit into the user documentation, but
-would rather be part of a design document, we use Hspec.  Hspec
-provides an EDSL for defining specs.  A spec organizes evidence in a
-tree structure (or a forest, to be accurate), and it is defined with \texttt{it} and
-\texttt{describe}.
+would rather be part of a design document, we use Hspec.
+
+\begin{quote}
+    Hspec was originally implemented by Trystan Spangler, and is now
+    maintained and actively/further developed by one of the authors.
+
+    \emph{sol: Do you think we should include this here? Is it
+    relevant? If it is relevant, it would probably better fit in to
+    the ``Introduction'', right?}
+\end{quote}
+
+Hspec provides an EDSL for defining specs.  A spec organizes evidence
+in a tree structure (or a forest, to be accurate), and it is defined
+with \texttt{it} and \texttt{describe}.
 
@@
 
-\subsubsection{Hspec not provide a mechanism for expectations}
+%%%%%%%%%%%%%%%%
+\subsubsection{Hspec does not provide a mechanism for expectations}
 BDD frameworks use \emph{should} or \emph{must} instead of
@@
 
+%%%%%%%%%%%%%%%%
 \subsubsection{Hspec provides no mechanism to expect exceptions}
@@
 \subsection{Problem summary}
+
+\begin{quote}
+    \emph{TBD}
+\end{quote}
 

### 2012-05-29 paper.tex
Sum up Doctests implementation

@@
 
-TBD: kazu: how to obtain evidence from Haddock markup.
-Old version uses Haddock API but now uses GHC API directly. Why?
-
-TBD: kazu: escape from HUnit. Why?
-
-TBD: kazu: Old version executes one GHCi for each function examples.
-But new version executes one GHCI per module.
-Drastic speed up. Do We need a graph for execution time?
-
-Since test is done by comparing result string in Haddock notation
-and result string from GHCi, exceptions can be handled.
+\begin{itemize}
+    \item
+        Doctest uses the GHC API to extract Haddock comments.
+    \item
+        It uses the GHC API directly, so that it can more easily
+        support old GHC versions.  Each release of Haddock only
+        supports the latest GHC version.
+    \item
+        Doctest use GHCi to verify examples.
+    \item
+        It reuses the same GHCi session for all examples.
+    \item
+        Examples from the same Haddock comment are grouped together,
+        and share the same scope.
+    \item
+        The scope is cleared with \verb|:reload| before a new
+        Example group is started.
+    \item
+\end{itemize}
+
+Since test is done by comparing result string in Haddock notation and
+result string from GHCi, exceptions can be handled.
 

### 2012-05-29 paper.tex
Add Doctest as a contribution
Add doctust as a contribution (again!)
Solution wip
Comment for @kazu

@@
 \item
+    We originally implemented Doctest, back in 2010.
+\item
     Add support for QuickCheck properties to Doctest (Section
@@
 
-\begin{quote}\begin{itemize}
-    \item
-        Haddock has markup support for interactive examples, and
-        Doctest can verify them.
-    \item
-        But it is not possible to use properties in Haddock comments
-        and automatically test them.
-\end{itemize} \end{quote}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{quote}
-    \emph{TBD}
+    TBD: sol: @kazu, can you update this?
 \end{quote}
@@
 \label{sec:user-doc}
+
+\begin{quote}\begin{itemize}
+    \item
+        We already added markup support for interactive examples to
+        Haddock, back in 2010; and Doctest can verify them.
+    \item
+        But it is not possible to use properties in Haddock comments
+        and automatically test them.
+\end{itemize} \end{quote}
 

### 2012-05-29 paper.tex
Wording, and a comment for @kazu

@@
 Haddock~\cite{haddock}.
-Haddock provides markup syntax for code blocks, indicated with ({\tt
->}).
-For example, the {\tt containers} packages contains a Haddock comment
-with the following code block.
+Haddock has markup support for code blocks.  And it is widely used to
+include properties and examples in documentation.
+
+\begin{quote}
+    sol: @kazu should we cite package documentation, to support this?
+    I'm aware of:
+    \begin{itemize}
+        \item filepath
+        \item containers
+    \end{itemize}
+\end{quote}
+
+Here is an example from the {\tt containers} packages.
 

### 2012-05-29 paper.tex
Move comment around
Add more comments for @kazu

@@
 
-\begin{quote}\begin{itemize}
-    \item
-        We already added markup support for interactive examples to
-        Haddock, back in 2010; and Doctest can verify them.
-    \item
-        But it is not possible to use properties in Haddock comments
-        and automatically test them.
-\end{itemize} \end{quote}
-
 Evidence for users should be documented with Haddock markup syntax.
@@
 \subsubsection{Markup for examples}
+
+\begin{quote}\begin{itemize}
+    \item
+        We already added markup support for interactive examples to
+        Haddock, back in 2010; and Doctest can verify them.
+    \item
+        But it is not possible to use properties in Haddock comments
+        and automatically test them.
+\end{itemize} \end{quote}
 
@@
 
-TBD: describe the implementation of {\tt shouldBe}, etc.
+\begin{quote}
+    sol: @kazu: Can you add the documentation of \verb|test-shouldbe|
+    to the bib, and cite it here?
+\end{quote}
+
+\noindent All combinators introduced in \ref{} have been
+implemented~\cite{TODO}; and they will be incorporated into a future
+release of Hspec.
+The implementation is trivial, so we will not further discuss it.
+
+\begin{quote}
+    TBD: The implementation of \verb|shouldThrow| may still be
+    interesting.  We can describe it, if we have time left.
+\end{quote}
 
@@
 
+\begin{quote}
     TBD
+
+    sol: @kazu, can you work on this, if you have time.  I think an
+    ``itemized'' list is better than text.  It more flexible for now,
+    and @mkl or @sol can do the final wording)
+\end{quote}
 

### 2012-05-29 paper.tex
Fix typo
Add comment for @mkl
Add more on tests as documentation

@@
 strong type system of Haskell cannot always ensure correct semantics.
-Without test code, it is difficult for users to understand usage of
-functions and it is difficult for developers to maintain semantics
-while extending software. To enhance the value of test code, we
+Test code helps users to understand the usage of
+functions and developers to maintain semantics
+while extending software. To ease the production and enhance the value of test code, we
 propose to carry out behavior driven development with two tools:
@@
 libraries that come with GHC), but only 8 of them have a test suite (5
-provide example code, 10 have no tests nor examples).  Lacking test
-code also means that typical usage examples are not provided.
+provide example code, 10 have no tests nor examples).
+A lack of tests is not only a quality and maintainability issue.  It
+is also a lack of documentation.  Tests document how code is supposed
+to be used; and tests documents the behavior of code.
 
@@
 \begin{quote}\begin{itemize}
-    \item sol: Should we state that BDD-style TDD has been pioneered
-        by RSpec?
+    \item
+        sol: Should we state that BDD-style TDD has been pioneered by
+        RSpec?  By now I think it would be a good idea, that way it
+        feels more natural to the reader that we always look at RSpec.
+
+        @mkl: Where would that fit in?
 \end{itemize}\end{quote}
@@
     sol: @kazu, can you work on this, if you have time.  I think an
-    ``itemized'' list is better than text.  It more flexible for now,
-    and @mkl or @sol can do the final wording)
+    ``itemized'' list is better than text.  It's more flexible for
+    now, and @mkl or @sol can do the final wording)
 \end{quote}

### 2012-05-29 paper.tex
abstract: turn negative sentences into positive ones
introduction: BDD-style TDD has been pioneerd by RSpec

@@
 
-\begin{quote}\begin{itemize}
-    \item
-        sol: Should we state that BDD-style TDD has been pioneered by
-        RSpec?  By now I think it would be a good idea, that way it
-        feels more natural to the reader that we always look at RSpec.
-
-        @mkl: Where would that fit in?
-\end{itemize}\end{quote}
-
-In this paper, we explore how to apply BDD to Haskell.
-For the purpose of this paper we only focus on the TDD aspect of BDD.
+BDD-style TDD has been pioneered by RSpec~\cite{rspec}.  In this
+paper, we explore how to apply the TDD aspect of BDD to Haskell.
 
@@
 \emph{code examples} as evidence for behavior.
-Hspec, a BDD tool for Haskell inspired by RSpec~\cite{rspec}, supports
+Hspec, a BDD tool for Haskell inspired by RSpec, supports
 both code examples and \emph{properties} as evidence.

### 2012-05-29 paper.tex
a markup -> markup

@@
 implementation closer to the desired goal.  Rinse and repeat.  BDD
-includes TDD, amongst other software development methodologies, but
+includes TDD, amongst other software development methods, but
 with a terminology and mindset  shifted towards specification instead
@@
 \begin{itemize}
-\item Define a new Haddock markup for examples.
-\item Define a new Haddock markup for properties.
+\item Define new Haddock markup for examples.
+\item Define new Haddock markup for properties.
 \item Bring Doctest from the Python community to the Haskell
@@
 
-\noindent First, we introduce a new Haddock markup for the example.
+\noindent First, we introduce new Haddock markup for the example.
 In this paper, we use {\tt >>>} which is the same as Python's one.
@@
 
-\noindent To specify properties, we also propose a new markup. In this paper, we
+\noindent To specify properties, we also propose new markup. In this paper, we
 use the {\tt prop>} prefix. We can write non-quantified properties as follows:

### 2012-05-29 paper.tex
introduction: s/methodologies/methods/
introduction: language review: Haskell programmers, however ...

@@
 
-Haskell programmers, however, is apt to rely on the type system too
-much and tend not to write test code.  For example, the Haskell
-Platform 2011.4.0.0 contains 23 packages (not counting the 19 core
-libraries that come with GHC), but only 8 of them have a test suite (5
-provide example code, 10 have no tests nor examples).
+Haskell programmers, however, are apt to rely on the type system too
+much and tend not to write test code.  For example, Haskell
+Platform 2011.4.0.0 contains 23 packages, not counting the 19 core
+libraries that come with GHC.  Only 8 of them have a test suite, 5
+provide example code, and 10 have neither tests nor examples.
 A lack of tests is not only a quality and maintainability issue.  It
-is also a lack of documentation.  Tests document how code is supposed
-to be used; and tests documents the behavior of code.
+also is a lack of documentation.  Tests document how code is supposed
+to be used, and tests document the behavior of code.
 

### 2012-05-29 paper.tex
intro: BDD also includes test discovery and continuous integration

@@
 implementation closer to the desired goal.  Rinse and repeat.  BDD
-includes TDD, amongst other software development methods, but
-with a terminology and mindset  shifted towards specification instead
-of testing.  In BDD, specifications can be used for automatic testing.
+includes TDD, but with a terminology and mindset  shifted towards
+specification instead of testing. BDD also includes other software
+development methods, such as automatic test discovery and continuous
+integration~\cite{ci}.  In BDD, specifications can be used for
+automatic testing.
 
@@
 document and test Haskell code.
-
-\begin{quote}
-    \emph{sol: @mkl, any idea where to put this?}
-
-\emph{Automation, such as test discovery and continuous integration (CI)
-~\cite{ci} is an important aspect of BDD.}
-\end{quote}
 

### 2012-05-29 paper.tex
Add a note
Extend "Open problems and future work"

@@
 
+\subsection{Source locations for failing examples}
+
+TBD:
+
+We do not get source locations for failing tests. Doctest solves this.
+For Hspec/QuickCheck/HUnit something like JHC's
+\verb|SRCLOC_ANNOTATE|
+pragma~\cite{todo}
+could help.
+
+\begin{quote}
+    sol: @kazu should we cite
+    \url{http://repetae.net/computer/jhc/jhc.shtml}?
+\end{quote}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 at this moment.
+\begin{itemize}
+    \item
+        \emph{sol: Hspec's mechanism for evidence is extensible.  It
+        would be easy to add support for SmallCheck.}
+\end{itemize}
 

### 2012-05-29 paper.tex
wording: set expectations -> express expectations

@@
 \item
-    Introduce a set of combinators to set expectations about
+    Introduce a set of combinators to express expectations about
     code examples for use with Hspec (Section \ref{sec:code-doc} and Section \ref{sec:hspec-comb})
@@
 
-We propose several combinators, that can be used to set expectations
-about the outcome of examples.  And we build them on top of
-HUnit.
+We propose several combinators which can be used to express expectations
+about the outcome of examples.  These are built on top of HUnit.
 
@@
 
+\subsection{TBD}
+
+\begin{itemize}
+    \item \emph{optional}: Bridging the gap between HUnit and QuickCheck
+    \item
+        Feedback for failing QuickCheck properties is not always as
+        useful as it could be -> combinators that use QuickCheck's
+        Result type could help
+    \item
+        Are more high-lever approaches to BDD, akin to Ruby's
+        Cucumber, applicable to Haskell? (Observation: Cucumber's
+        approach is tightly tied to OO)
+\end{itemize}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-05-29 paper.tex
terminology: example, expectation, property

--- 2012-05-29 paper.tex
@@
 \item[example]
-    An example shows how a function is used.  It is either a \emph{code
-    example} in the form of an HUnit test case; or an \emph{interactive
-    example}, given as a short GHCi session.
+    An example is a piece of source code that shows how a function is
+    used.  It consists of the function call with concrete values as
+    parameters, together with the expected outcome.  The outcome can
+    be given as a pure value, or in the case of impure code, as a
+    short GHCi session.
 
 \item[expectation]
-    Expectations are used to define the expected outcome of a code
-    example.  They serve the same purpose as \emph{assertions} in xUnit
-    testing frameworks, but use a different terminology.
+    An expectation is a proposition that defines the anticipated
+    outcome of a function call.  Expectations serve the same purpose
+    as \emph{assertions} in classical unit testing frameworks, but use
+    a different terminology. For instance, instead of
+    \begin{quote}
+    \small
+    \begin{verbatim}assertEqual (gcd 13 7) 1\end{verbatim}
+    \end{quote}
+    we write
+    \begin{quote}
+    \small
+    \begin{verbatim}gcd 13 7 `shouldBe` 1\end{verbatim}
+    \end{quote}
 
 \item[property]
-    Property means a QuickCheck property as defined in
-    \cite{quickcheck}.
+    A property is a piece of code similar to an example, except that
+    some parameters of the function call are abstracted from.  The
+    test framework runs a property several times with different
+    parameters, resulting in several outcomes. The expectation is
+    therefore not given as a concrete value, but as a predicate which
+    all outcomes should satisfy.
 
@@
 \item[evidence]
-    Evidence provides some confidence, that a function exhibits a
+    Evidence provides some confidence that a function exhibits a
     certain behavior.  In Haskell we use examples and properties as

### 2012-05-30 paper.tex
let "make" work.
supply "in Haskell".

@@
 \item
-    We originally implemented Doctest, back in 2010.
+    We originally implemented Doctest in Haskell, back in 2010.
 \item
@@
 \noindent All combinators introduced in \ref{} have been
-implemented~\cite{TODO}; and they will be incorporated into a future
+implemented~\cite{todo}; and they will be incorporated into a future
 release of Hspec.

### 2012-05-30 paper.tex
move Hspec explanation to footnote.

@@
 If we have evidence that does not fit into the user documentation, but
-would rather be part of a design document, we use Hspec.
-
-\begin{quote}
+would rather be part of a design document, we use Hspec\footnote{%
 Hspec was originally implemented by Trystan Spangler, and is now
-    maintained and actively/further developed by one of the authors.
-
-    \emph{sol: Do you think we should include this here? Is it
-    relevant? If it is relevant, it would probably better fit in to
-    the ``Introduction'', right?}
-\end{quote}
+maintained and actively/further developed by one of the authors.}.
 

### 2012-05-30 paper.tex
adding my comment.
add comment.
making paragraphs be the almost same size.
moving citation for Haddock to the first appearance.

@@
 more test cases are written to reqire more details and drive the
-implementation closer to the desired goal.  Rinse and repeat.  BDD
-includes TDD, but with a terminology and mindset  shifted towards
+implementation closer to the desired goal.  Rinse and repeat.
+
+BDD includes TDD, but with a terminology and mindset shifted towards
 specification instead of testing. BDD also includes other software
@@
 integration~\cite{ci}.  In BDD, specifications can be used for
-automatic testing.
-
-BDD-style TDD has been pioneered by RSpec~\cite{rspec}.  In this
-paper, we explore how to apply the TDD aspect of BDD to Haskell.
+automatic testing.  BDD-style TDD has been pioneered by
+RSpec~\cite{rspec}.  In this paper, we explore how to apply the TDD
+aspect of BDD to Haskell.
 
@@
         name~\cite{doctest}, already allows to verify
-        \emph{interactive examples} in Haddock comments.
+        \emph{interactive examples} in Haddock~\cite{haddock} comments.
 \end{enumerate}
@@
     \emph{sol: Can we remove what follows??}
+    \emph{kazu: I want to keep them for readability??}
 \end{quote}
@@
     the community might associated it with something imperative.
+
+    kazu: ``behavior'' sounded strange at least to me at the first time.
+    We alreay supported this word with typeclassopedia. So, I think
+    this paragraph is completed.
 
@@
 The standard documentation mechanism for Haskell is
-Haddock~\cite{haddock}.
+Haddock.
 Haddock has markup support for code blocks.  And it is widely used to

### 2012-05-30 paper.tex
adding my comment.
supply markup credits to our contributions.

@@
 \item
-    We originally implemented Doctest in Haskell, back in 2010.
-\item
-    Add support for QuickCheck properties to Doctest (Section
+    We originally implemented Doctest in Haskell and
+    extended Haddock markup to support examples, back in 2010
+\item
+    Add support for QuickCheck properties to Doctest and
+    extend Haddock markup to support properties
+    (Section
     \ref{sec:user-doc} and Section \ref{sec:doctest})
@@
 and accepting our patches to Haddock.
+
+\begin{quote}
+  kazu: this seems redundant: how about "\emph{for reviewing and accepting our patches}"?
+\end{quote}
+
 We would like to thank

### 2012-05-30 paper.tex
brush up a comment.
adding a comment.

@@
 \begin{quote}
-    \emph{sol: Can we remove what follows??}
-    \emph{kazu: I want to keep them for readability??}
+    \emph{sol: Can we remove what follows??}\\
+    \emph{kazu: I want to keep them for readability.}
 \end{quote}
@@
     We alreay supported this word with typeclassopedia. So, I think
-    this paragraph is completed.
+    this paragraph is completed. Should we replace ``is associated'' with
+    ``might be associated''?
 
@@
     sol: remove?
+    kazu: yes. But I would add a sentence such as ``This section
+    will discuss problems when we try to use these tools for BDD''.
 

### 2012-05-30 paper.tex
adding filepath. adding a comment.

@@
 include properties and examples in documentation.
-
-\begin{quote}
-    sol: @kazu should we cite package documentation, to support this?
-    I'm aware of:
-    \begin{itemize}
-        \item filepath
-        \item containers
-    \end{itemize}
-\end{quote}
-
-Here is an example from the {\tt containers} packages.
+For instance, the {\tt filepath} package and
+the {\tt containers} package adopt this style.
+Here is an example from the latter.
 
@@
     \emph{sol: I think the Haddock comment only contains properties.
-    So the argumentation is shaky?}
+    So the argumentation is shaky?} \\
+    \emph{kazu: My point is that we cannot tell which the author of packages
+   want to express, examples or properties.}
 \end{quote}

### 2012-05-30 paper.tex
adding comments.
brush up comment.
adding a comment.

@@
 
+\begin{quote}
+kazu: the two paragraphs above doe not seem logical to me. I would change these
+as follows:
+
+BDD tools can be used to document the behavior of a system in a way that
+is similar to a specification.  All facets of the system should be
+described with evidence, including every possible corner case.
+
+We think that some evidence is useful as user documentation, hence we
+want to be able to include it in Haddock comments.  But others are not
+useful as user documentation.  E.g. that a particular \texttt{Monoid}
+instance is a monoid in the mathematical sense is self-evident.
+They should be included in code document. Both kinds of evidence
+should be automatically tested.
+\end{quote}
+
 We propose to use Doctest and Haddock for evidence that is useful as
@@
 
+\begin{quote}
+    \emph{kazu: We should explicitly repeat that we propose to Doctest as a BDD tool here..}
+\end{quote}
+
 To be able to specify evidence for users, we propose to add new
@@
 
-\begin{quote}\begin{itemize}
+\begin{quote}
+\begin{itemize}
     \item
@@
         and automatically test them.
-\end{itemize} \end{quote}
+
+    \emph{kazu: I guess that it's enough to describe this in the Implementation Section.}
+\end{itemize}
+\end{quote}
 

### 2012-05-30 paper.tex
brush up my comment.

@@
 
-\begin{quote}
-kazu: the two paragraphs above doe not seem logical to me. I would change these
+We propose to use Doctest and Haddock for evidence that is useful as
+user documentation, and Hspec for evidence that is not useful as user
+documentation.
+
+\begin{quote}
+kazu: I'm uncomfortable with the three paragraphs above. I would change these
 as follows:
@@
 should be automatically tested.
-\end{quote}
 
@@
 user documentation, and Hspec for evidence that is not useful as user
-documentation.
-
-\begin{quote}
-    sol: remove?
-    kazu: yes. But I would add a sentence such as ``This section
-    will discuss problems when we try to use these tools for BDD''.
-
-BDD tools should be treat both of them and provide functionality
-to extract evidence for users as user documentation.
-If we can archive this, BDD tools can be used for
-triple purpose: design, documentation and automatic testing.
+documentation. The following subsections disccuss problems when
+we try to carry out our proposal of BDD.
 \end{quote}

### 2012-05-30 paper.tex
better explanation.
Fix typo

@@
     We originally implemented Doctest in Haskell and
-    extended Haddock markup to support examples, back in 2010
+    extended Haddock's markup to support examples, back in 2010
 \item
@@
 
-The standard documentation mechanism for Haskell is
-Haddock.
+For Doctest, we should use Haddock markup to express evidence.
 Haddock has markup support for code blocks.  And it is widely used to

### 2012-05-30 paper.tex
Remove "in Haskell"
That would need to be "for Haskell", but "for Haskell" is ambiguous.  An
option that would work is "Haskell's version of Doctest", so we would
need to use that, if we really need this.
Add comment

@@
 \item
-    We originally implemented Doctest in Haskell and
+    We originally implemented Doctest and
     extended Haddock's markup to support examples, back in 2010
@@
 \begin{quote}
-    \emph{sol: Can we remove what follows??}\\
+    \emph{sol: Can we remove what follows?}
+
     \emph{kazu: I want to keep them for readability.}
+
+    \emph{sol: For me it really depends on how often we use something.
+    If we only use it once or twice, I'd try make it clear from the
+    context what is meant.}
 \end{quote}
@@
     ``might be associated''?
+
+    sol: @kazu, this was primarily meant for @mkl as reference.
 

### 2012-05-30 paper.tex
Make acks more concise (@kazu, yes, I agree)

@@
 
-We are grateful to David Waern, Haddock's maintainer, for reviewing our patches
-and accepting our patches to Haddock.
-
-\begin{quote}
-  kazu: this seems redundant: how about "\emph{for reviewing and accepting our patches}"?
-\end{quote}
+We are grateful to David Waern, Haddock's maintainer for reviewing and
+accepting our patches to Haddock.
 

### 2012-05-30 paper.tex
better introduction of Solution.

@@
 \section{Problem}
+\label{sec:problem}
 
@@
 
-\begin{quote}
-    \emph{kazu: We should explicitly repeat that we propose to Doctest as a BDD tool here..}
-\end{quote}
-
-To be able to specify evidence for users, we propose to add new
-markup to Haddock and to use Doctest integrated with QuickCheck. To
-describe evidence for developers, we propose to use Hspec also
-integrated with QuickCheck.
+As we explained in Section \ref{sec:problem},
+we use Doctest for user documentation and Hspec for code documentation.
+The following subsections explain two new markups of Haddock
+and evidence combinators for Hspec.
 

### 2012-05-30 paper.tex
A suggestion
Fix typo
Remove redundant sentence.
adding text and bytestring to support observation.

@@
 \item
-    We originally implemented Doctest and
-    extended Haddock's markup to support examples, back in 2010
+    We originally implemented Haskell's version of Doctest, and extended
+    Haddock's markup to support examples, back in 2010
 \item
@@
 
-For Doctest, we should use Haddock markup to express evidence.
 Haddock has markup support for code blocks.  And it is widely used to
 include properties and examples in documentation.
-For instance, the {\tt filepath} package and
-the {\tt containers} package adopt this style.
+For instance, some packages, including {\tt filepath}, 
+{\tt containers}, {\tt bytestring} and {\tt text}, adopt this style.
 Here is an example from the latter.

### 2012-05-30 paper.tex
fixing mistake.

@@
 include properties and examples in documentation.
-For instance, some packages, including {\tt filepath}, 
-{\tt containers}, {\tt bytestring} and {\tt text}, adopt this style.
-Here is an example from the latter.
+For instance, some packages, including {\tt containers}, {\tt filepath}, 
+{\tt bytestring} and {\tt text}, adopt this style.
+Here is an example from the first one.
 

### 2012-05-30 paper.tex
brush up the Haddock explanation.

@@
 
-As we described, Haddock version 2.7.2 or earlier provides markup syntax for
-code blocks, but does not provide any markup syntax for evidence.
-We implemented markup support for examples ({\tt >>>}) and it is available in
-Haddock version 2.8 or later.
+Haddock version 2.7.2 or earlier provided markup syntax for
+code blocks, but did not provide any markup syntax for evidence.
+As we described, we implemented markup support for examples ({\tt >>>}) in 2010.
+It is available in Haddock version 2.8 or later.
 We also extended Haddock to support markup for properties ({\tt prop>}) which

### 2012-05-30 paper.tex
typo fix.

@@
 propose to carry out behavior driven development with two tools:
-doctest integrated with QuickCheck for user documentation and Hspec
+Doctest integrated with QuickCheck for user documentation and Hspec
 with combinators extended for developer documentation. With our

### 2012-05-30 paper.tex
Fix typo

@@
 That could mean yielding a constant value at first.  Subsequently,
-more test cases are written to reqire more details and drive the
+more test cases are written to require more details and drive the
 implementation closer to the desired goal.  Rinse and repeat.
@@
 \item
-    We originally implemented Haskell's version of Doctest, and extended
+    We originally implemented Haskell's version of Doctest\footnote{%
+    http://hackage.haskell.org/package/doctest}, and extended
     Haddock's markup to support examples, back in 2010
@@
 
-\begin{quote}
-    sol: @kazu: Can you add the documentation of \verb|test-shouldbe|
-    to the bib, and cite it here?
-\end{quote}
-
-\noindent All combinators introduced in \ref{} have been
-implemented~\cite{todo}; and they will be incorporated into a future
+\noindent All combinators introduced in \ref{sec:code-doc} have been
+implemented in the {\tt hspec-shouldbe} package\footnote{%
+http://hackage.haskell.org/package/hspec-shouldbe};
+and they will be incorporated into a future
 release of Hspec.
@@
 
-\noindent Note that the {\tt placeholders} library provide better keyword such as {\tt notImplemented} and {\tt todo}.
+\noindent Note that the {\tt placeholders} library provides better keywords such as {\tt notImplemented} and {\tt todo} and they tells source locations when executed.
 Next we write user documentation in Haddock markup:
@@
 \begin{itemize}
-\item We use a preprocessor (the {\tt hspec-discover} package) to automatically generate this boilerplate.
+\item We use a preprocessor (the {\tt hspec-discover} package\footnote{%
+http://hackage.haskell.org/package/hspec-discover}) to automatically generate this boilerplate.
 \end{itemize}
@@
 \verb|SRCLOC_ANNOTATE|
-pragma~\cite{todo}
-could help.
+pragma\footnote{%
+http://repetae.net/computer/jhc/jhc.shtml} could help.
 

### 2012-05-30 paper.tex
resolving citation.

@@
 That could mean yielding a constant value at first.  Subsequently,
-more test cases are written to require more details and drive the
+more test cases are written to reqire more details and drive the
 implementation closer to the desired goal.  Rinse and repeat.
@@
 
-\noindent Note that the {\tt placeholders} library provides better keywords such as {\tt notImplemented} and {\tt todo} and they tells source locations when executed.
+\noindent Note that the {\tt placeholders} library provide better keyword such as {\tt notImplemented} and {\tt todo}.
 Next we write user documentation in Haddock markup:
@@
 \begin{itemize}
-\item We use a preprocessor (the {\tt hspec-discover} package\footnote{%
-http://hackage.haskell.org/package/hspec-discover}) to automatically generate this boilerplate.
+\item We use a preprocessor (the {\tt hspec-discover} package) to automatically generate this boilerplate.
 \end{itemize}

### 2012-05-30 paper.tex
Put \cite for Haddock back to where it was before
This is not ideal, but it as least works.
adding explanation for placeholders.
footnote for hspec-discover.
adding a comment on citation.
Incorporated suggestions made by @kazu into "Problem" section

--- 2012-05-30 paper.tex
@@
 
+\begin{quote}
+    \emph{kazu: we should talk about Python's doctest,
+    Haskell's doctest and Haddock here and put
+    citation on them.}
+\end{quote}
+
 We made two observations:
@@
         name~\cite{doctest}, already allows to verify
-        \emph{interactive examples} in Haddock~\cite{haddock} comments.
+        \emph{interactive examples} in Haddock comments.
 \end{enumerate}
@@
 
-Hspec can be used to document the behavior of a system in a way that
+BDD tools can be used to document the behavior of a system in a way that
 is similar to a specification.  All facets of the system should be
@@
 
-We think that some evidence is useful as user documentation, hence we
-want to be able to include it in Haddock comments.  But there is also
-evidence that is not useful as user documentation.  E.g. that a
-particular \texttt{Monoid} instance is a monoid in the mathematical
-sense is self-evident.  So we do not want to include any evidence for
-that into user documentation, but we still want to automatically test
-it.
+We think that some evidence is useful as user documentation.
+But there is also evidence, that is not
+useful as user documentation.  E.g. that a particular \texttt{Monoid}
+instance is a monoid in the mathematical sense is self-evident.
 
@@
 \begin{quote}
-kazu: I'm uncomfortable with the three paragraphs above. I would change these
-as follows:
-
-BDD tools can be used to document the behavior of a system in a way that
-is similar to a specification.  All facets of the system should be
-described with evidence, including every possible corner case.
-
-We think that some evidence is useful as user documentation, hence we
-want to be able to include it in Haddock comments.  But others are not
-useful as user documentation.  E.g. that a particular \texttt{Monoid}
-instance is a monoid in the mathematical sense is self-evident.
-They should be included in code document. Both kinds of evidence
-should be automatically tested.
-
-We propose to use Doctest and Haddock for evidence that is useful as
-user documentation, and Hspec for evidence that is not useful as user
-documentation. The following subsections disccuss problems when
-we try to carry out our proposal of BDD.
+    TBD: Not correct English, so we have to find an other way to say
+    this.
+
+    The following subsections disccuss problems when we try to carry
+    out our proposal of BDD.
 \end{quote}
@@
 
-Haddock has markup support for code blocks.  And it is widely used to
+Haddock has markup support for code blocks~\cite{haddock}.  And it is widely used to
 include properties and examples in documentation.

### 2012-05-30 paper.tex
Fix grammar issue
Put evidence into footnote

@@
 
-Haddock has markup support for code blocks~\cite{haddock}.  And it is widely used to
-include properties and examples in documentation.
-For instance, some packages, including {\tt containers}, {\tt filepath}, 
-{\tt bytestring} and {\tt text}, adopt this style.
-Here is an example from the first one.
+Haddock has markup support for code blocks, and they are widely used
+to include properties and examples into documentation\footnote{
+See e.g. the documentation of
+{\tt containers},
+{\tt filepath},
+{\tt bytestring} or
+{\tt text}}.
+Here is an example from the {\tt containers} package.
 

### 2012-05-30 paper.tex
State again, that Haddock is the standard way to document Haskell code
describing reusable spec.

@@
 
+The standard documentation mechanism for Haskell is
+Haddock~\cite{haddock}.
 Haddock has markup support for code blocks, and they are widely used
@@
 
-\begin{itemize}
-\item Some type classes have their laws.
-\item We can use Hspec to enforce such properties.
-\item If such properties are reusable among instances,
-we can avoid boilerplates.
-\item reusable spec can do this.
-\end{itemize}
+Reusable spec is an idea to scrap boilerplate of evidence.
+For instance, without reusable spec,
+each developer of an instance of a type class should
+express evidence for its laws.
+Since instances of a type class share API,
+we can avoid such boilerplate.
+The following is an example to implement
+evidence for Monoid laws:
 
@@
 
-\noindent The following is an example to apply this laws
-to lists.
+\noindent This can be used for any instances of Monoid.
+Here is an example to apply it to a list:
 
@@
 
-\begin{itemize}
-\item we should prepare this kind of laws.
-\end{itemize}
+\noindent We have already implemented this experimentally.
+We should ensure the practicality of reusable spec and
+prepare reusable spec to cover variety of type classes.
 

### 2012-05-30 paper.tex
adding subsection separators.
Try to make a better case against code blocks

@@
 {\tt text}}.
-Here is an example from the {\tt containers} package.
-
-\begin{quote}
-\small
-\begin{verbatim}
--- | /O(1)/. Is the map empty?
+Here is an example from the {\tt filepath} package.
+
+\begin{quote}
+\small
+\begin{verbatim}
+-- | Does the given filename have an extension?
 --
--- > Data.Map.null (empty)           == True
--- > Data.Map.null (singleton 1 'a') == False
-
-null :: Map k a -> Bool
-null Tip      = True
-null (Bin {}) = False
+-- > null (takeExtension x) == not (hasExtension x)
+hasExtension :: FilePath -> Bool
+hasExtension = any isExtSeparator . takeFileName
 \end{verbatim}
@@
 \noindent Though code blocks in Haddock comments are a useful feature,
-we still think that this particular usage has problems.
-
-\begin{quote}
-    \emph{sol: I think the Haddock comment only contains properties.
-    So the argumentation is shaky?} \\
-    \emph{kazu: My point is that we cannot tell which the author of packages
-   want to express, examples or properties.}
-\end{quote}
-
-If the equations above are examples, we cannot distinguish
-their results from their expression to be evaluated.
-If they are properties, they should be distinguished
-from example code.
+we still think that this particular usage has problems.  The code
+block contains a property.  And it would be useful to have a tool that
+automatically extracts and tests the property.
+But writing such a tool is hard, because it is not easy to decide if
+the code block is a property, an example, or some random code snippet.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Reusable spec}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Source locations for failing examples}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{TBD}

### 2012-05-30 paper.tex
Wording
Minor wording (if -> whether)
Resolve a TBD

@@
 user documentation, and Hspec for evidence that is not useful as user
-documentation.
-
-\begin{quote}
-    TBD: Not correct English, so we have to find an other way to say
-    this.
-
-    The following subsections disccuss problems when we try to carry
-    out our proposal of BDD.
-\end{quote}
+documentation.  The following subsections discuss problems that we
+have to solve before we can do this.
 
@@
 automatically extracts and tests the property.
-But writing such a tool is hard, because it is not easy to decide if
-the code block is a property, an example, or some random code snippet.
+But writing such a tool is hard, because it is not easy to decide
+whether the code block is a property, an example, or something else.
 

### 2012-05-30 paper.tex
State that evidence should be testable
Revert "terminology: example, expectation, property"
This reverts commit 6f683578e639582483c96cd5c53471d261983c33.

We use /code example/ and /interactive example in the text.  So I think
it is a good idea to keep them in the terminology.

@@
 
-In BDD, all behavior should be documented, together with some
+In BDD, all behavior should be documented, together with some testable
 \emph{evidence} for that behavior.  Most existing BDD tools use
@@
 \item[example]
-    An example is a piece of source code that shows how a function is
-    used.  It consists of the function call with concrete values as
-    parameters, together with the expected outcome.  The outcome can
-    be given as a pure value, or in the case of impure code, as a
-    short GHCi session.
+    An example shows how a function is used.  It is either a \emph{code
+    example} in the form of an HUnit test case; or an \emph{interactive
+    example}, given as a short GHCi session.
 
 \item[expectation]
-    An expectation is a proposition that defines the anticipated
-    outcome of a function call.  Expectations serve the same purpose
-    as \emph{assertions} in classical unit testing frameworks, but use
-    a different terminology. For instance, instead of
-    \begin{quote}
-    \small
-    \begin{verbatim}assertEqual (gcd 13 7) 1\end{verbatim}
-    \end{quote}
-    we write
-    \begin{quote}
-    \small
-    \begin{verbatim}gcd 13 7 `shouldBe` 1\end{verbatim}
-    \end{quote}
+    Expectations are used to define the expected outcome of a code
+    example.  They serve the same purpose as \emph{assertions} in xUnit
+    testing frameworks, but use a different terminology.
 
 \item[property]
-    A property is a piece of code similar to an example, except that
-    some parameters of the function call are abstracted from.  The
-    test framework runs a property several times with different
-    parameters, resulting in several outcomes. The expectation is
-    therefore not given as a concrete value, but as a predicate which
-    all outcomes should satisfy.
+    Property means a QuickCheck property as defined in
+    \cite{quickcheck}.
 
@@
 \item[evidence]
-    Evidence provides some confidence that a function exhibits a
+    Evidence provides some confidence, that a function exhibits a
     certain behavior.  In Haskell we use examples and properties as

### 2012-05-30 paper.tex
using the example of Sec 3.1 to explain prop> in step by step manner.
adding more explanations to Sec 4.1.2.

@@
 \subsection{Doctest and Haddock}
+\label{sec:doctest-haddock}
 
@@
 \noindent Of course, we allowed quantified properties by using closed
-lambda terms.
-Type signatures can be specified for each parameter:
+lambda terms. The property example in Section~\ref{sec:doctest-haddock}
+can be written as follows:
+
+\begin{quote}
+\small
+\begin{alltt}
+prop> \verb|\|x -> null (takeExtension x) == \(\hookleftarrow\)
+            not (hasExtension x)
+\end{alltt}
+\end{quote}
+
+\noindent If a default type is not suitable, expression type--signatures can be specified for each parameter:
 

### 2012-05-30 paper.tex
add more explanation in Sec 5.2.2.
better wordings.

@@
 \subsubsection{Markup for properties}
+\label{sec:markups}
 
@@
 
-\noindent Here is an example:
+\noindent Here is the example in Section~\ref{sec:markups} again:
 
@@
 
-\noindent Note that the {\tt placeholders} library provides better keywords such as {\tt notImplemented} and {\tt todo} and they tells source locations when executed.
+\noindent Note that the {\tt placeholders} library provides better keywords such as {\tt notImplemented} and {\tt todo}, which tells source locations when executed.
 Next we write user documentation in Haddock markup:

### 2012-05-30 paper.tex
describing the line-continue symbol.

@@
 
-\noindent If a default type is not suitable, expression type--signatures can be specified for each parameter:
+\noindent Note that the $\hookleftarrow$ symbol denotes that two lines are actually written in a single line.
+If a default type is not suitable, expression type--signatures can be specified for each parameter:
 

### 2012-05-30 paper.tex
removing sol's comment and resolves overflow.

@@
 We do not get source locations for failing tests. Doctest solves this.
-For Hspec/QuickCheck/HUnit something like JHC's
+Something like JHC's
 \verb|SRCLOC_ANNOTATE|
 pragma\footnote{%
-http://repetae.net/computer/jhc/jhc.shtml} could help.
-
-\begin{quote}
-    sol: @kazu should we cite
-    \url{http://repetae.net/computer/jhc/jhc.shtml}?
-\end{quote}
+http://repetae.net/computer/jhc/jhc.shtml} could help
+for Hspec/QuickCheck/HUnit.
 

### 2012-05-30 paper.tex
cosmetic change to rescure highlighting of Emacs.
outline of Conclusion.

@@
 
-\begin{quote}\small\begin{verbatim}
+\begin{quote}
+\small
+\begin{verbatim}
 assertRaises :: (Show a, Exception e, Show e, Eq e) =>
@@
 
-\begin{quote}
-    TBD
-
-    sol: @kazu, can you work on this, if you have time.  I think an
-    ``itemized'' list is better than text.  It's more flexible for
-    now, and @mkl or @sol can do the final wording)
-\end{quote}
+TBD:
+
+\begin{itemize}
+\item We discussed how to carry out BDD in Haskell.
+\item We proposed to use Doctest for user documentation
+and Hspec for code documentation.
+\item We enhanced Haddock to support examples and properties, 
+and implemented Doctest integrated with QuickCheck.
+\item We also supplied expectation combinators to Hspec.
+\item With these two tools, specification can also be used for
+documentation and automatic testing.
+\item test discovery
+\item This enhances the value of test code and would motivate
+Haskell programmers to write test code.
+\end{itemize}
 

### 2012-05-30 paper.tex
correcting wrong wording.

@@
 Doctest integrated with QuickCheck for user documentation and Hspec
-with combinators extended for developer documentation. With our
+with combinators extended for code documentation. With our
 method, specification can be used for documentation and automatic

### 2012-05-30 paper.tex
correct hyphenation.
adding comment.
adding "only".

@@
 
+\begin{quote}
+    \emph{kazu: I would like to exchange 2nd sentence and 3rd sentence:}
+
+BDD includes TDD, but with a terminology and mindset shifted towards
+specification instead of testing. In BDD, specifications can be used
+for automatic testing.  BDD also includes other software development
+methods, such as automatic test discovery and continuous
+integration~\cite{ci}.  BDD-style TDD has been pioneered by
+RSpec~\cite{rspec}.  In this paper, we explore how to apply the TDD
+aspect of BDD to Haskell.
+
+\end{quote}
+
 In BDD, all behavior should be documented, together with some testable
-\emph{evidence} for that behavior.  Most existing BDD tools use
+\emph{evidence} for that behavior.  Most existing BDD tools only use
 \emph{code examples} as evidence for behavior.
@@
 
-\begin{quote}
-    \emph{kazu: we should talk about Python's doctest,
-    Haskell's doctest and Haddock here and put
-    citation on them.}
-\end{quote}
-
 We made two observations:
@@
 as needed.  In particular, our contributions are:
+
+\begin{quote}
+    \emph{kazu: I would like to change blocks above (from ``We made two ...'') with:}
+
+Hspec is thus useful for code documentation but we observed that parts
+of a specification can be useful as user documentation. We pay attention
+to Python's Doctest~\cite{doctest} for this purpose.
+Doctest is not commonly described as a BDD tool, but we think
+it can be used for BDD.
+There is Haskell version of Doctest\footnote{%
+http://hackage.haskell.org/package/doctest} coped with Haddock~\cite{haddock},
+standard documentation mechanism for Haskell.
+It extracts \emph{interactive examples} in Haddock comments and
+verify them automatically.
+
+We will show how to use Doctest and Hspec in combination to design,
+document and test Haskell code.
+Both, Doctest and Hspec are missing some features.  So we extend them
+as needed.  In particular, our contributions are:
+\end{quote}
 
@@
 \noindent Note that the $\hookleftarrow$ symbol denotes that two lines are actually written in a single line.
-If a default type is not suitable, expression type--signatures can be specified for each parameter:
+If a default type is not suitable, expression type-signatures can be specified for each parameter:
 

### 2012-05-30 paper.tex
adding footnote for Hspec.
Fix grammar issues

@@
 \emph{code examples} as evidence for behavior.
-Hspec, a BDD tool for Haskell inspired by RSpec, supports
+Hspec\footnote{%
+http://hackage.haskell.org/package/hspec},
+a BDD tool for Haskell inspired by RSpec, supports
 both code examples and \emph{properties} as evidence.
@@
 it can be used for BDD.
-There is Haskell version of Doctest\footnote{%
-http://hackage.haskell.org/package/doctest} coped with Haddock~\cite{haddock},
-standard documentation mechanism for Haskell.
+There is a Haskell version of Doctest\footnote{%
+http://hackage.haskell.org/package/doctest}.  It is integrated with with Haddock~\cite{haddock},
+the standard documentation mechanism for Haskell.
 It extracts \emph{interactive examples} in Haddock comments and

### 2012-05-30 paper.tex
Emphasize /code documentation/ on first usage
Shorten "Problem" -> "Hspec"

@@
 
-Hspec is thus useful for code documentation but we observed that parts
+Hspec is thus useful for \emph{code documentation} but we observed that parts
 of a specification can be useful as user documentation. We pay attention
@@
 \noindent The {\tt it} function combines a text description for a
-behavior and some evidence into a spec.
+behavior and some evidence (a code example or a property) into a spec.
 Code examples can be arbitrary HUnit assertions.  Properties must be of type {\tt
@@
 The \texttt{describe} function combines a list of specs into a larger spec.
-It can be used to organize a spec as a tree.
-Internally, a \texttt{Writer} is used to build up specs in a monadic way.
-
+% It can be used to organize a spec as a tree.
+% Internally, a \texttt{Writer} is used to build up specs in a monadic way.
 A spec can be run with the \texttt{hspec} function.  When a spec is
@@
 %%%%%%%%%%%%%%%%
-\subsubsection{Hspec does not provide a mechanism for expectations}
+% \subsubsection{Hspec does not provide a mechanism for expectations}
 BDD frameworks use \emph{should} or \emph{must} instead of
@@
 %%%%%%%%%%%%%%%%
-\subsubsection{Hspec provides no mechanism to expect exceptions}
-RSpec provides a mechanism to expect exceptions.
+% \subsubsection{Hspec provides no mechanism to expect exceptions}
+Moreover RSpec provides a mechanism to expect exceptions.
 

### 2012-05-30 paper.tex
minor fix in comment of Intro.

@@
 There is a Haskell version of Doctest\footnote{%
-http://hackage.haskell.org/package/doctest}.  It is integrated with with Haddock~\cite{haddock},
+http://hackage.haskell.org/package/doctest}, which is integrated with Haddock~\cite{haddock},
 the standard documentation mechanism for Haskell.

### 2012-05-30 paper.tex
Add a comment
adopt comment's one.

@@
 
+    sol: this paragraph still has issues
+    \begin{itemize}
+        \item test discovery is not an aspect of BDD, but an emphasis
+            on automating things, so that they e.g. work with CI is.
+        \item @mkl your previous version stated, that BDD includes
+            other things (ATDD!), but we only focus on TDD.  I think
+            this got lost.
+    \end{itemize}
+
 BDD includes TDD, but with a terminology and mindset shifted towards
@@
 informal specification document from that.
-
-We made two observations:
-\begin{enumerate}
-    \item
-        Parts of a specification can be useful as user documentation.
-    \item
-        Doctest, a tool inspired by a Python module with the same
-        name~\cite{doctest}, already allows to verify
-        \emph{interactive examples} in Haddock comments.
-\end{enumerate}
-
-\noindent Doctest is not commonly described as a BDD tool, but we
-think it fits the picture.
-We show how to use Doctest and Hspec in combination to design,
-document and test Haskell code.
-
-Both, Doctest and Hspec are missing some features.  So we extend them
-as needed.  In particular, our contributions are:
-
-\begin{quote}
-    \emph{kazu: I would like to change blocks above (from ``We made two ...'') with:}
 
@@
 as needed.  In particular, our contributions are:
-\end{quote}
-
-\begin{itemize}
-\item
-    We originally implemented Haskell's version of Doctest\footnote{%
-    http://hackage.haskell.org/package/doctest}, and extended
+
+\begin{itemize}
+\item
+    We originally implemented Haskell's version of Doctest, and extended
     Haddock's markup to support examples, back in 2010

### 2012-05-30 paper.tex
removing unnecessary sentence anymore.
Wording

@@
 Hspec is thus useful for \emph{code documentation} but we observed that parts
-of a specification can be useful as user documentation. We pay attention
-to Python's Doctest~\cite{doctest} for this purpose.
-Doctest is not commonly described as a BDD tool, but we think
-it can be used for BDD.
-There is a Haskell version of Doctest\footnote{%
+of a specification can be useful as user documentation. To that end, we think
+Python's {\tt doctest} module~\cite{doctest} is interesting.
+It is not commonly described as a BDD tool, but we think it can be
+used for BDD.
+There is a Haskell version called Doctest\footnote{%
 http://hackage.haskell.org/package/doctest}, which is integrated with Haddock~\cite{haddock},
@@
 
-The standard documentation mechanism for Haskell is
-Haddock~\cite{haddock}.
 Haddock has markup support for code blocks, and they are widely used

### 2012-05-30 paper.tex
adding comment in Sec 3.1.

@@
     \begin{itemize}
-        \item test discovery is not an aspect of BDD, but an emphasis
-            on automating things, so that they e.g. work with CI is.
+        \item test discovery is not an aspect of BDD, only an emphasis
+            on automating things is an aspect of BDD, so that thing
+            e.g. work with CI.
         \item @mkl your previous version stated, that BDD includes
             other things (ATDD!), but we only focus on TDD.  I think
-            this got lost.
+            meaning got lost now.
     \end{itemize}
@@
 the standard documentation mechanism for Haskell.
-It extracts \emph{interactive examples} in Haddock comments and
-verify them automatically.
+It extracts \emph{interactive examples} from Haddock comments and
+verifies them automatically.
 
@@
 whether the code block is a property, an example, or something else.
+
+\begin{quote}
+    \emph{kazu: I would like to add the following:}
+
+So, we need common syntax to describe examples and properties
+independently in Haddock. Doctest should automatically verify
+examples and test properties.
+\end{quote}
 

### 2012-05-30 paper.tex
Grammar
Update comment
"BDD includes TDD, amongst other software development metho..", again

@@
 
-BDD includes TDD, but with a terminology and mindset shifted towards
-specification instead of testing. BDD also includes other software
-development methods, such as automatic test discovery and continuous
-integration~\cite{ci}.  In BDD, specifications can be used for
+BDD includes TDD, amongst other software development methodologies,
+but with a terminology and mindset shifted towards
+specification instead of testing.
+In BDD, specifications can be used for
 automatic testing.  BDD-style TDD has been pioneered by
@@
 \begin{quote}
-    \emph{kazu: I would like to exchange 2nd sentence and 3rd sentence:}
-
-    sol: this paragraph still has issues
+    BDD also includes other software development methods, such as
+    automatic test discovery and continuous integration~\cite{ci}.
+
+    sol: I removed this, and reverted the paragraph to it's previous
+    version, because I think there were several issues.
+
     \begin{itemize}
         \item test discovery is not an aspect of BDD, only an emphasis
-            on automating things is an aspect of BDD, so that thing
-            e.g. work with CI.
+            on automating things is an aspect of BDD, so that
+            \emph{testing is more fun} and \emph{things e.g. work with
+            CI}.  This are two things, but both justify automation.
+            We only need this to justify our discussion of Cabal
+            integration and automatic test discovery.  If we remove
+            those, we do not state that ``automation is an important
+            aspect of BDD''.  This may be a solution.
         \item @mkl your previous version stated, that BDD includes
-            other things (ATDD!), but we only focus on TDD.  I think
-            meaning got lost now.
+            other things (ATDD!), but that we do not cover them.  This
+            is very important.  And I think meaning got lost with the
+            change, hence the revert.
     \end{itemize}

### 2012-05-30 paper.tex
interesting -> of interest.
making ack one paragraph.
Remove leftover

@@
     \end{itemize}
-
-BDD includes TDD, but with a terminology and mindset shifted towards
-specification instead of testing. In BDD, specifications can be used
-for automatic testing.  BDD also includes other software development
-methods, such as automatic test discovery and continuous
-integration~\cite{ci}.  BDD-style TDD has been pioneered by
-RSpec~\cite{rspec}.  In this paper, we explore how to apply the TDD
-aspect of BDD to Haskell.
-
 \end{quote}
@@
 Hspec is thus useful for \emph{code documentation} but we observed that parts
-of a specification can be useful as user documentation. To that end, we think
-Python's {\tt doctest} module~\cite{doctest} is interesting.
+of a specification can be useful as user documentation. To that end, 
+Python's {\tt doctest} module~\cite{doctest} is of interest for us.
 It is not commonly described as a BDD tool, but we think it can be
@@
 accepting our patches to Haddock.
-
 We would like to thank

### 2012-05-30 paper.tex
Make list of contributions more consistent

@@
 \item
-    We originally implemented Haskell's version of Doctest, and extended
-    Haddock's markup to support examples, back in 2010
-\item
-    Add support for QuickCheck properties to Doctest and
-    extend Haddock markup to support properties
+    We originally implemented Haskell's version of Doctest, and
+    extended Haddock's markup to support examples, back in 2010
+\item
+    We added support for QuickCheck properties to Doctest and extended
+    Haddock's markup to support properties
     (Section
@@
 \item
-    Introduce a set of combinators to express expectations about
-    code examples for use with Hspec (Section \ref{sec:code-doc} and Section \ref{sec:hspec-comb})
-
-\item
-    Improve CI support for Doctest and Hspec (Section \ref{sec:TBD})
+    We introduce a set of combinators to express expectations about
+    code examples for use with Hspec (Section \ref{sec:code-doc} and
+    Section \ref{sec:hspec-comb})
+
+\item
+    We improve CI support for Doctest and Hspec (Section
+    \ref{sec:TBD})
 
@@
 \item
-    Show best current practice of BDD in Haskell, using Doctest and
+    We show best current practice of BDD in Haskell, using Doctest and
     Hspec (Section \ref{ref:Best-current-practice})

### 2012-05-30 paper.tex
Grammar

@@
 
-The word ``behavior'' is associated with imperative actions.
-Since pure functions of Haskell are static,
-this word might be improper for Haskell.
-However, this word is already used in Haskell community~\cite{typeclassopedia}.
-So, we use ``behavior'' instead of defining a new terminology in this paper.
+The word ``behavior'' is associated with imperative actions.  Since
+pure functions of Haskell are static, this word might be improper for
+Haskell.  However, this word is already used in the Haskell
+community~\cite{typeclassopedia}.  So, we use ``behavior'' instead of
+defining terminology in this paper.
 

### 2012-05-30 paper.tex
Grammar and wording

@@
 
-\noindent First, we introduce new Haddock markup for the example.
+\noindent First, we introduce new Haddock markup for examples.
 In this paper, we use {\tt >>>} which is the same as Python's one.
@@
 
-\noindent Of course, we allowed quantified properties by using closed
-lambda terms. The property example in Section~\ref{sec:doctest-haddock}
+\noindent Of course, we allow quantified properties by using a
+lambda expression. The property example in Section~\ref{sec:doctest-haddock}
 can be written as follows:
@@
 
-Evidence in comments are automatically
+Evidence in comments is automatically
 tested by the Doctest command. We also provided a library
@@
 
-We propose following extensions for Hspec to
+We propose the following extensions for Hspec to
 describe evidence for developers.

### 2012-05-30 paper.tex
Extend "setting expectations"
State, that support for properties is fully functional
Wording

@@
 
-\noindent Or we may expect that a value satisfies some predicate.
+\noindent Or we may expect that a value should satisfies some
+predicate.
 
@@
 \end{verbatim}\end{quote}
+
+\noindent Where \verb|shouldReturn| has the following type:
 
@@
 \end{verbatim}\end{quote}
+
+The implementation is fully functional, and it will be included in the
+next release of Doctest.
 
@@
 
-\noindent This can be used for any instances of Monoid.
-Here is an example to apply it to a list:
+\noindent This can be used to document and automatically test
+arbitrary Monoid instances.
+Here is an example how to use it for lists:
 

### 2012-05-30 paper.tex
Update comment
Add missing commas

@@
             other things (ATDD!), but that we do not cover them.  This
-            is very important.  And I think meaning got lost with the
-            change, hence the revert.
+            is very important.  And I think that meaning got lost with
+            the change, hence the revert.
     \end{itemize}
@@
 
-Hspec is thus useful for \emph{code documentation} but we observed that parts
+Hspec is thus useful for \emph{code documentation}, but we observed that parts
 of a specification can be useful as user documentation. To that end, 
@@
 
-We will show how to use Doctest and Hspec in combination to design,
+We will show how to use Doctest and Hspec in combination, to design,
 document and test Haskell code.

### 2012-05-30 paper.tex
Update comment
Fix typo
Add an other comment

@@
 In BDD, specifications can be used for
-automatic testing.  BDD-style TDD has been pioneered by
+automatic testing.
+
+\begin{quote}
+    sol: maybe like so?  In addition to \emph{automating things}, this
+    would also motivate \ref{sec:code-doc}, which is IMHO not
+    motivated very well.
+
+    BDD is meant to be an enjoyable experience for the developer.  It
+    provides compared to TDD it provides a much nicer syntax that
+    makes tests very easy to read.  And like TDD it puts emphasis on
+    test automation, to take load from the shoulders of developers.
+\end{quote}
+
+BDD-style TDD has been pioneered by
 RSpec~\cite{rspec}.  In this paper, we explore how to apply the TDD
@@
             \emph{testing is more fun} and \emph{things e.g. work with
-            CI}.  This are two things, but both justify automation.
+            CI}.  This are two different things, but both justify automation.
             We only need this to justify our discussion of Cabal
             integration and automatic test discovery.  If we remove
-            those, we do not state that ``automation is an important
+            those, we do not need to state that ``automation is an important
             aspect of BDD''.  This may be a solution.
@@
 
-\subsection{Automatic discover of specs}
+\subsection{Automatic discovery of specs}
 

### 2012-05-30 paper.tex
Update comment

@@
 
-    BDD is meant to be an enjoyable experience for the developer.  It
-    provides compared to TDD it provides a much nicer syntax that
+    BDD is meant to be an enjoyable experience for the developer.
+    Compared to TDD it provides a much nicer syntax that
     makes tests very easy to read.  And like TDD it puts emphasis on
-    test automation, to take load from the shoulders of developers.
+    test automation, to take work load from the shoulders of
+    developers.
 \end{quote}

### 2012-05-30 paper.tex
Remove trailing whitespace

@@
 
-Many Haskell programmers tend not to write test code even though the
-strong type system of Haskell cannot always ensure correct semantics.
-Test code helps users to understand the usage of
-functions and developers to maintain semantics
-while extending software. To ease the production and enhance the value of test code, we
-propose to carry out behavior driven development with two tools:
-Doctest integrated with QuickCheck for user documentation and Hspec
-with combinators extended for code documentation. With our
-method, specification can be used for documentation and automatic
-testing. This kill-three-birds-with-one-stone aspect would motivate
-Haskellers to write test code.
+When wading through the numerous libraries written by
+the Haskell community, one sometimes gets the impression that the type
+system is being too much relied on, at the expense of testing.
+While the type system cannot always guarantee semantic correctness,
+test code has more applications than getting the code correct.
+Test code assists developers in maintaining correctness when extending
+software, and shows users how an API is supposed to be used.
+To ease the production and enhance the value of test code, we advocate
+behavior driven development with two tools:
+QuickCheck integrated in Doctest for user documentation and Hspec
+with new mnemonic combinators for code documentation.
+Our method utilizes specification that can also be used for
+documentation and automatic testing. We hope that this three-in-one
+quality motivates Haskellers to write more test code.
 
@@
 
+\begin{quote}
+    \emph{kazu: Monoid is already described in Introduction. Why is this repeated? And ``code documentation'' is aslo already dicussed in Introduction. I think we can use ``code documentation'' here.}
+\end{quote}
+
 We propose to use Doctest and Haddock for evidence that is useful as
@@
 
-\begin{quote}
-    \emph{kazu: I would like to add the following:}
-
 So, we need common syntax to describe examples and properties
@@
 examples and test properties.
-\end{quote}
 
@@
 
-\begin{quote}
-    TBD: sol: @kazu, can you update this?
-\end{quote}
-
 Here is a summary of the problem:
@@
 \item
-    The Haskell community does not have common syntax to describe
-    evidence with Haddock markup syntax.
-\item
-    The Haskell community does not have a BDD tool for user documentation.
-\item
+    For user documentation,
+    Haddock does not have markups to describe evidence.
+\item
+    For code documentation,
     Hspec does not provide combinators to express expectations with
@@
 community and integrate it with QuickCheck.
+Examples and properties in Haddock comments is automatically
+tested by the Doctest command.
+Its implementation will be discussed in Section~\ref{sec:doctest}.
+
 \end{itemize}
@@
 given property is implicitly quantified over \verb|xs|.
-
-Evidence in comments is automatically
-tested by the Doctest command. We also provided a library
-to do this job. Its implementation will discussed in Section~\ref{sec:doctest}.
 

### 2012-05-30 paper.tex
abstract: mkl's proposition

--- 2012-05-30 paper.tex
@@
 
+\emph{mkl proposes:}
+When wading through the numerous libraries written by
+the Haskell community, one sometimes gets the impression that the type
+system is being too much relied on, at the expense of testing.
+While the type system cannot always guarantee semantic correctness,
+test code has more applications than getting the code correct.
+Test code assists developers in maintaining correctness when extending
+software, and shows users how an API is supposed to be used.
+To ease the production and enhance the value of test code, we advocate
+behavior driven development with two tools:
+QuickCheck integrated in Doctest for user documentation and Hspec
+with new mnemonic combinators for code documentation.
+Our method utilizes specification that can also be used for
+documentation and automatic testing. We hope that this three-in-one
+quality motivates Haskellers to write more test code.
+
 %% State the problem.

### 2012-05-31 paper.tex
new abstract.

--- 2012-05-30 paper.tex
@@
 
+\begin{quote}
+    \emph{kazu: I would like to add the following:}
+
 So, we need common syntax to describe examples and properties
@@
 examples and test properties.
+\end{quote}
 
@@
 
+\begin{quote}
+    TBD: sol: @kazu, can you update this?
+\end{quote}
+
 Here is a summary of the problem:
@@
 \item
-    For user documentation,
-    Haddock does not have markups to describe evidence.
-\item
-    For code documentation,
+    The Haskell community does not have common syntax to describe
+    evidence with Haddock markup syntax.
+\item
+    The Haskell community does not have a BDD tool for user documentation.
+\item
     Hspec does not provide combinators to express expectations with

### 2012-05-31 paper.tex
Motivate automatic spec discovery

@@
 
-\begin{quote}
-    TBD:
-\begin{itemize}
-    \item \emph{optional:} Motivate why it's important to integrate
-        Doctest with
-        Cabal.
-
-    \item \emph{optional:} Motivate \verb|hspec-discover|
-
-        It is common to have several files that contain \emph{specs}.
-        And we want to be able to run them all together.  Hspec allows
-        us combine specs together.  But this has to be done manually.
-        Hspec does not provide a way to automatically discover and run
-        all specs of a project.
-\end{itemize}
-\end{quote}
+\subsection{Automatic discovery of specs}
+
+It is a common convention to have one spec file for each source file.
+To run all specs for a given project together, it is necessary to
+combine all the specs from all spec files into a single spec.  For
+example, if we have three modules, \verb|Foo|, \verb|Foo.Bar| and
+\verb|Baz|, and corresponding specs in \verb|FooSpec|,
+\verb|Foo.BarSpec| and \verb|BazSpec|, we are required to write the
+following boilerplate:
+
+\begin{quote}
+\small
+\begin{verbatim}
+import Test.Hspec.Monadic
+
+import qualified FooSpec
+import qualified Foo.BarSpec
+import qualified BazSpec
+
+main :: IO ()
+main = hspec $ do
+  describe "Foo"     FooSpec.spec
+  describe "Foo.Bar" Foo.BarSpec.spec
+  describe "Baz"     BazSpec.spec
+\end{verbatim}
+\end{quote}
+
+\noindent This is neither challenging nor interesting, so it should be
+automated.  But Hspec does not provide a solution for that.
 
@@
 
-\begin{quote}
-\small
-\begin{verbatim}
-import Test.Hspec.Monadic
-
-import qualified FooSpec
-import qualified Foo.BarSpec
-import qualified BazSpec
-
-main :: IO ()
-main = hspec $ do
-  describe "Foo"     FooSpec.spec
-  describe "Foo.Bar" Foo.BarSpec.spec
-  describe "Baz"     BazSpec.spec
-\end{verbatim}
-\end{quote}
-
 \begin{itemize}

### 2012-05-31 paper.tex
writing Sec 5.2 (Doctest).

@@
 
+To extract evidence from Haddock markup,
+Doctest used Haddock API in the past.
+Since Each release of Haddock supports only the latest version of GHC,
+it is difficult to support old versions of GHC with Haddock API.
+Doctest is now directly using GHC API for portability.
+
 \subsubsection{Testing examples}
 
-\begin{itemize}
-    \item
-        Doctest uses the GHC API to extract Haddock comments.
-    \item
-        It uses the GHC API directly, so that it can more easily
-        support old GHC versions.  Each release of Haddock only
-        supports the latest GHC version.
-    \item
-        Doctest use GHCi to verify examples.
-    \item
-        It reuses the same GHCi session for all examples.
-    \item
+To verify examples, Doctest uses GHCi.
+Test is done by comparing result string in Haddock notation and
+result string from GHCi.
+Since verification based on string comparison, exceptions can be handled.
+
+Doctest reuses the same GHCi session for all examples in a module.
 Examples from the same Haddock comment are grouped together,
 and share the same scope.
-    \item
 The scope is cleared with \verb|:reload| before a new
 Example group is started.
-    \item
-\end{itemize}
-
-Since test is done by comparing result string in Haddock notation and
-result string from GHCi, exceptions can be handled.
 

### 2012-05-31 paper.tex
Doctest reuses the GHCi session for all examples
(not just all examples in a module)
adopting assertRaises.

@@
 
-Doctest reuses the same GHCi session for all examples in a module.
+Doctest reuses the same GHCi session for all examples.
 Examples from the same Haddock comment are grouped together,
@@
 
-\begin{itemize}
-    \item The {\tt testpack} package provides a function that is
-        similar in purpose to {\tt shouldThrow}.
-
-        sol: Do you think we should discuss this; or is it safe to
-        ignore.
-
-        kazu: Let's add this.
+The {\tt testpack} package provides a function to check
+exception, similar to {\tt shouldThrow}.
 
@@
 \end{verbatim}\end{quote}
-        But it has several shortcomings:
-        \begin{itemize}
-            \item
-                It only works for exceptions with an {\tt Eq}
-                instance, but several standard exceptions (most
-                prominently {\tt ErrorCall}) have no {\tt Eq}
-                instance.
-            \item
+
+\noindent But it has several shortcomings:
+It only works for exceptions with an {\tt Eq} instance,
+but several standard exceptions (most
+prominently {\tt ErrorCall}) have no {\tt Eq} instance.
 It places an unnecessary {\tt Show} constraint on the
-                result of the {\tt IO} action, further limiting it's
-                applicability.
-            \item
+result of the {\tt IO} action, further limiting it's applicability.
 It only allows to expect specific values; expecting
 any exception of a certain type is not possible.
-        \end{itemize}
-\end{itemize}
 

### 2012-05-31 paper.tex
adopting sol's comment for SmallCheck.
skelton for shouldThrow.

@@
 
+\begin{quote}
+\small
+\begin{verbatim}
+shouldThrow ::
+  Exception e => IO a -> Selector e -> Assertion
+\end{verbatim}
+\end{quote}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 at this moment.
-\begin{itemize}
-    \item
-        \emph{sol: Hspec's mechanism for evidence is extensible.  It
-        would be easy to add support for SmallCheck.}
-\end{itemize}
+Hspec's mechanism for evidence is extensible.
+If necessary, it would be easy to support SmallCheck.
 
@@
 The {\tt testpack} package provides a function to check
-exception, similar to {\tt shouldThrow}.
+exception, similar to {\tt shouldThrow} described in Section~\ref{sec:hspec-comb}.
 
@@
   String -> e -> IO a -> IO ()
-\end{verbatim}\end{quote}
+\end{verbatim}
+\end{quote}
 

### 2012-05-31 paper.tex
Describe solution for spec discovery

@@
 \subsection{Automatic discovery of specs}
+\label{sec:problem-discovery}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{\emph{optional} Automatic discovery of specs}
-TBD
+\subsection{Automatic discovery of specs}
+We want to automatically discover all specs of a project, so that we
+can run them all together.  This requires to automatically generate
+boilerplate code, similar to what is shown in
+\ref{sec:problem-discovery}.
+
+We make creative use GHC's mechanism for custom preprocessors for
+that.  With our solution, the developer creates test driver, that only
+contains a single line:
+
+\begin{quote}
+\small
+\begin{verbatim}
+{-# OPTIONS_GHC -F -pgmF hspec-discover #-}
+\end{verbatim}
+\end{quote}
+
+This tells GHC to run a preprocessor, \verb|hspec-discover| on the
+source file.  \verb|hspec-discover| finds all specs for a project, and
+generates a suitable test driver.
+
+This approach relies on the following conventions
+\begin{enumerate}
+    \item
+        Spec files have to be placed into the same directory as the
+        test driver, or into a subdirectory.
+    \item
+        The name of a spec file has to end in \verb|Spec.hs|, and the
+        module name has to match the file name.
+    \item
+        Each spec file has to provide a top-level binding \verb|spec|,
+        of type \verb|Spec|.
+\end{enumerate}
+
+It would be possible to lift some of those requirements.  But we think
+they are reasonable conventions, anyway.
 
@@
 \end{itemize}
-
-\begin{quote}
-\small
-\begin{verbatim}
-{-# OPTIONS_GHC -F -pgmF hspec-discover #-}
-\end{verbatim}
-\end{quote}
 

### 2012-05-31 paper.tex
Rough sketch for implementation of hspec-discover

@@
 
+\newpage % FIXME
+\subsection{hspec-discover}
+
+As discussed in \ref{} we use a preprocessor to automatically generate
+test drivers.  GHC invokes preprocessors with three arguments.
+
+\begin{description}
+    \item[{\tt src}]
+        The original source file.  A preprocessor usually does not
+        look at the content of this file.  It is only provided, so
+        that it can be used in error messages.
+    \item[{\tt cur}]
+        The file that holds the input for the preprocessor.
+    \item[{\tt dst}]
+        A path, where the preprocessor should write it's result to.
+\end{description}
+
+\noindent A typical preprocessor reads the content of {\tt cur},
+processes it, and writes the result to {\tt dst}.  If something goes
+wrong, it uses {\tt src} in error messages.
+
+\verb|hspec-discover| ignores {\tt cur} altogether.  It uses the
+directory component of {\tt src}, to find all files that are in the
+same directory as {\tt src}, or any subdirectories.  It than generates
+a test driver from all files that end in {\tt Spec.hs}.
+
+A fully functional implementations is available from
+Hackage\footnote{http://hackage.haskell.org/package/hspec-discover},
+and it will be incorporate into a future release of Hspec.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-05-31 paper.tex
Add ref

@@
 \subsection{Automatic discovery of specs}
+\label{sec:solution-discovery}
 We want to automatically discover all specs of a project, so that we
@@
 \subsection{hspec-discover}
-
-As discussed in \ref{} we use a preprocessor to automatically generate
+\label{sec:implementation-discovery}
+
+As discussed in \ref{sec:solution-discovery} we use a preprocessor to automatically generate
 test drivers.  GHC invokes preprocessors with three arguments.

### 2012-05-31 paper.tex
Layout experiment, not sure if that is a good idea.

@@
 
-As discussed in \ref{sec:solution-discovery} we use a preprocessor to automatically generate
-test drivers.  GHC invokes preprocessors with three arguments.
+As discussed in \ref{sec:solution-discovery}, we use a preprocessor to
+automatically generate test drivers.  GHC invokes preprocessors with
+three arguments.
 
@@
     \item[{\tt src}]
-        The original source file.  A preprocessor usually does not
+        -- the original source file\\A preprocessor usually does not
         look at the content of this file.  It is only provided, so
@@
     \item[{\tt cur}]
-        The file that holds the input for the preprocessor.
+        -- the file that holds the input for the preprocessor
     \item[{\tt dst}]
-        A path, where the preprocessor should write it's result to.
+        -- a path, where the preprocessor should write it's result to
 \end{description}
@@
 BDD in Haskell.
-
-\begin{quote}
-    \emph{TBD: one more example relating to IO is desired.}
-\end{quote}
 

### 2012-05-31 paper.tex
removing TBD: we have already discussed IO much.
Resolve a TBD

@@
 \begin{quote}\small\begin{verbatim}
-let x = 1 `div` 0 in (x `seq` return x)
-      `shouldThrow` anyArithException
-\end{verbatim}\end{quote}
-
-\begin{quote}
-    TBD:
-    \begin{itemize}
-        \item
-            This code snippet also illustrates how to deal with pure
-            functions.
-        \item
-            {\tt seq} forces the exception.
-        \item
-            If {\tt x} is in {\tt NFData}, you may want to use {\tt
-            deepseq} instead.
-        \item
-            The following would not work:
-\begin{quote}\small\begin{verbatim}
-(1 `div` 0) `shouldThrow` anyArithException
-\end{verbatim}\end{quote}
-        \item
-            We think that pure functions that throw exceptions is bad
-            API design.
-    \end{itemize}
-\end{quote}
-
-TBD: Should we propose:
-\begin{quote}\small\begin{verbatim}
-shouldThrow_ :: (NFData a, Exception e) =>
-  a -> (e -> Bool) -> Expectation
+evaluate (1 `div` 0) `shouldThrow` anyArithException
 \end{verbatim}\end{quote}

### 2012-05-31 paper.tex
More on `shouldThrow`
Use straightforward instead of trivial
Add comment for @kazu

@@
 automatic testing.
-
-\begin{quote}
-    sol: maybe like so?  In addition to \emph{automating things}, this
-    would also motivate \ref{sec:code-doc}, which is IMHO not
-    motivated very well.
-
 BDD is meant to be an enjoyable experience for the developer.
@@
 developers.
-\end{quote}
-
 BDD-style TDD has been pioneered by
@@
 aspect of BDD to Haskell.
-
-\begin{quote}
-    BDD also includes other software development methods, such as
-    automatic test discovery and continuous integration~\cite{ci}.
-
-    sol: I removed this, and reverted the paragraph to it's previous
-    version, because I think there were several issues.
-
-    \begin{itemize}
-        \item test discovery is not an aspect of BDD, only an emphasis
-            on automating things is an aspect of BDD, so that
-            \emph{testing is more fun} and \emph{things e.g. work with
-            CI}.  This are two different things, but both justify automation.
-            We only need this to justify our discussion of Cabal
-            integration and automatic test discovery.  If we remove
-            those, we do not need to state that ``automation is an important
-            aspect of BDD''.  This may be a solution.
-        \item @mkl your previous version stated, that BDD includes
-            other things (ATDD!), but that we do not cover them.  This
-            is very important.  And I think that meaning got lost with
-            the change, hence the revert.
-    \end{itemize}
-\end{quote}
 
@@
 \begin{quote}
-\small
-\begin{verbatim}
->>> length []
-0
+    sol: @kazu should we make the Haddock examples in this section
+    valid Haddock comments?
+\end{quote}
+
+\begin{quote}
+\small
+\begin{verbatim}
+-- |
+-- >>> length []
+-- 0
 \end{verbatim}
@@
 launchMissiles `shouldThrow` (== ExitFailure 1)
+\end{verbatim}\end{quote}
+
+\noindent And it can be used with predicates from {\tt
+System.IO.Error}, to require specific {\tt IOException}, e.g.  with
+{\tt isPermissionError}:
+
+\begin{quote}\small\begin{verbatim}
+launchMissiles `shouldThrow` isPermissionError
 \end{verbatim}\end{quote}
@@
 release of Hspec.
-The implementation is trivial, so we will not further discuss it.
+The implementation is straightforward, so we will not further discuss
+it.
 

### 2012-05-31 paper.tex
Fix semantic
Add TBD
Many fixes and wording improvements

@@
 
+\begin{quote}
+    TBD: introduce \emph{test driver}
+\end{quote}
+
 If we have evidence that does not fit into the user documentation, but
@@
     For user documentation,
-    Haddock does not have markups to describe evidence.
+    Haddock does not have markup to describe evidence.
 \item
@@
 we use Doctest for user documentation and Hspec for code documentation.
-The following subsections explain two new markups of Haddock
-and evidence combinators for Hspec.
+The following subsections explain two extensions to Haddock's markup.
+And combinators for expectations, that can be used with Hspec, are
+introduced.
 
@@
 
-\noindent First, we introduce new Haddock markup for examples.
+\noindent First, we introduce new Haddock markup for interactive examples.
 In this paper, we use {\tt >>>} which is the same as Python's one.
@@
 \subsubsection{Markup for properties}
-\label{sec:markups}
+\label{sec:markup}
 
@@
 
-\noindent Of course, we allow quantified properties by using a
-lambda expression. The property example in Section~\ref{sec:doctest-haddock}
-can be written as follows:
+\noindent And it is possible to quantify properties by using a
+lambda expression. The property from Section~\ref{sec:doctest-haddock}
+can be written as:
 
@@
 
-\noindent Note that the $\hookleftarrow$ symbol denotes that two lines are actually written in a single line.
-If a default type is not suitable, expression type-signatures can be specified for each parameter:
+\noindent Note that the $\hookleftarrow$ symbol denotes line
+continuation.
+If a default type is not suitable, expression type-signatures can be
+specified for each parameter:
 
@@
 
-We propose several combinators which can be used to express expectations
-about the outcome of examples.  These are built on top of HUnit.
+We propose several combinators which can be used to set expectations
+about the outcome of code examples.  These are built on top of HUnit:
 
@@
 \noindent A common expectation is, that a value should be equal to an other
-value.
+value:
 
@@
 
-\noindent Examples are often used to provide evidence for {\tt IO}
-actions.
+\noindent Code examples are often used to provide evidence for {\tt
+IO} actions, e.g.:
 
@@
 
-\noindent So we propose a shortcut for this common use case.
+\noindent So we propose a shortcut for this common use case:
 
@@
 servers two purposes.  It selects the type of the expected exception,
-and constraints the value.
+and constraints the value.  Here is an example:
 
@@
 
-\noindent To make it more convenient, to expect any exception of a
+\noindent To make it more convenient to expect any exception of a
 certain type, we provide type-restricted versions of {\tt (const
-True)} for all standard exceptions.
+True)} for all standard exceptions, e.g.:
 
@@
 
-We make creative use GHC's mechanism for custom preprocessors for
-that.  With our solution, the developer creates test driver, that only
+We make creative use GHC's support for custom preprocessors for
+that.  With our solution, the developer creates a test driver, that only
 contains a single line:
@@
 
-This tells GHC to run a preprocessor, \verb|hspec-discover| on the
+\noindent This tells GHC to run a preprocessor, \verb|hspec-discover|, on the
 source file.  \verb|hspec-discover| finds all specs for a project, and
@@
 
-This approach relies on the following conventions
+This approach relies on the following conventions:
 \begin{enumerate}
@@
     \item
-        Each spec file has to provide a top-level binding \verb|spec|,
+        Each spec file has to export a top-level binding \verb|spec|,
         of type \verb|Spec|.
@@
 
-It would be possible to lift some of those requirements.  But we think
-they are reasonable conventions, anyway.
+\noindent It would be possible to lift some of those requirements.
+But we think they are reasonable conventions, anyway.
 
@@
 
-Haddock version 2.7.2 or earlier provided markup syntax for
+\begin{quote}
+    TBD: This section needs more work, to get the wording right.
+\end{quote}
+
+Haddock version 2.7.2 and earlier provided markup syntax for
 code blocks, but did not provide any markup syntax for evidence.
-As we described, we implemented markup support for examples ({\tt >>>}) in 2010.
+As we described, we implemented markup support for interactive examples ({\tt >>>}) in 2010.
 It is available in Haddock version 2.8 or later.
@@
 will be available with a future release of Haddock.
-Both examples and properties in Haddock comments are pretty-printed when
-generating documentation, (e.g. as HTML).
+Both interactive examples and properties in Haddock comments are
+include into generated documentation.
 
@@
 
-To extract evidence from Haddock markup,
-Doctest used Haddock API in the past.
-Since Each release of Haddock supports only the latest version of GHC,
+To extract evidence from Haddock comments,
+Doctest used the Haddock API in the past.
+Since each release of Haddock supports only the latest version of GHC,
 it is difficult to support old versions of GHC with Haddock API.
@@
 The scope is cleared with \verb|:reload| before a new
-Example group is started.
+example group is started.
 
@@
 
-\noindent Here is the example in Section~\ref{sec:markups} again:
+\noindent Here is the example from Section~\ref{sec:markup}:
 
@@
 \noindent All combinators introduced in \ref{sec:code-doc} have been
-implemented in the {\tt hspec-shouldbe} package\footnote{%
-http://hackage.haskell.org/package/hspec-shouldbe};
+implemented.  They are available form Hackage\footnote{%
+http://hackage.haskell.org/package/hspec-shouldbe},
 and they will be incorporated into a future
@@
 
-\newpage % FIXME
 \subsection{hspec-discover}
@@
 As discussed in \ref{sec:solution-discovery}, we use a preprocessor to
-automatically generate test drivers.  GHC invokes preprocessors with
-three arguments.
+automatically generate test drivers for Hspec.  GHC invokes the
+preprocessor with three arguments:
 
@@
     \item[{\tt src}]
-        -- the original source file\\A preprocessor usually does not
+        -- path to the original source file\\
+        A preprocessor usually does not
         look at the content of this file.  It is only provided, so
@@
     \item[{\tt cur}]
-        -- the file that holds the input for the preprocessor
+        -- path to a file that holds the input for the preprocessor
     \item[{\tt dst}]
@@
 processes it, and writes the result to {\tt dst}.  If something goes
-wrong, it uses {\tt src} in error messages.
+wrong, it uses {\tt src} in error messages to refer to the source
+file.
 
@@
 directory component of {\tt src}, to find all files that are in the
-same directory as {\tt src}, or any subdirectories.  It than generates
+same directory as the source file, or in any subdirectories.  It then generates
 a test driver from all files that end in {\tt Spec.hs}.
@@
 \begin{quote}
+    sol: TBD. @kazu, the descriptions make no sense, can you try to
+    make them more precise?
+\end{quote}
+
+\begin{quote}
 \small
@@
 
-\noindent Now we run test suites until
-all test failures disappear.
+\noindent Now we run tests, and improve our implementation, until all
+tests pass.
 
@@
     \emph{should emphasize ``automatically discover tests''}
-\end{quote}
-
-\begin{quote}
-\emph{This is optional, so we can skip it, if necessary.}
 \end{quote}

### 2012-05-31 paper.tex
Do not yet reference text that may never get written
Extended conclusion, and other things

@@
 
-\noindent Cabals test feature can be used to automatically run tests~\cite{cabal}.  A cabal
+\noindent Cabal's test feature can be used to automatically run tests~\cite{cabal}.  A cabal
 file can contain multiple test suites, so it is possible to define test suites
-for Doctest and Hspec in a single cabal file.
+for Doctest and Hspec in the same cabal file.
 
@@
 The {\tt testpack} package provides a function to check
-exception, similar to {\tt shouldThrow} described in Section~\ref{sec:hspec-comb}.
+exception, similar to {\tt shouldThrow} described in Section~\ref{todo}.
 
@@
 prominently {\tt ErrorCall}) have no {\tt Eq} instance.
-It places an unnecessary {\tt Show} constraint on the
+And it places an unnecessary {\tt Show} constraint on the
 result of the {\tt IO} action, further limiting it's applicability.
-It only allows to expect specific values; expecting
+It only allows to expect specific values, expecting
 any exception of a certain type is not possible.
@@
 \begin{itemize}
-\item We discussed how to carry out BDD in Haskell.
-\item We proposed to use Doctest for user documentation
-and Hspec for code documentation.
-\item We enhanced Haddock to support examples and properties,
-and implemented Doctest integrated with QuickCheck.
-\item We also supplied expectation combinators to Hspec.
-\item With these two tools, specification can also be used for
+\item
+    We discussed how to apply BDD to Haskell.
+\item
+    We proposed to use Doctest for user documentation and Hspec for
+    code documentation.
+\item
+    We added markup support for properties and examples to Haddock.
+    And we implemented Doctest, to verify examples and test properties
+    from Haddock comments.
+\item
+    We extend Hspec with a set of combinators for expectations.
+\item
+    With these two tools, specification can also be used for
     documentation and automatic testing.
-\item test discovery
-\item This enhances the value of test code and would motivate
-Haskell programmers to write test code.
+\item
+    We describe a mechanism, that can automatically discover all specs
+    of a project.
+\item
+    We think that this enhances the value of test code, and we hope
+    that it will motivate Haskell programmers to write more and better
+    tests.
 \end{itemize}

### 2012-05-31 paper.tex
Wording

@@
 
-Evidence for users should be documented with Haddock markup syntax.
-For this purpose, we propose the followings:
+Evidence that is useful as user documentation, should be included in
+Haddock comments.  For this purpose, we propose the following:
 
 \begin{itemize}
-\item Define new Haddock markup for examples.
-\item Define new Haddock markup for properties.
+\item Add markup syntax for interactive examples to Haddock
+\item Add markup syntax for properties to Haddock
 \item Bring Doctest from the Python community to the Haskell

### 2012-05-31 paper.tex
Remove redundant comma
Wording

@@
 
-We agree that type signatures are good abstract documentation~\cite{free}, but
+We agree that a function's type can reveal something about a
+function's behavior~\cite{free}, but
 there are many functions whose behavior cannot be guessed from their names and
@@
 
-Evidence that is useful as user documentation, should be included in
+Evidence that is useful as user documentation should be included in
 Haddock comments.  For this purpose, we propose the following:

### 2012-05-31 paper.tex
Remove unneeded section
We cover this elsewhere now.

@@
 
-\subsection{Automatic discovery of specs}
-
-\begin{quote}
-    \emph{should emphasize ``automatically discover tests''}
-\end{quote}
-
-\noindent We think the following are good conventions for organizing specs:
-\begin{itemize}
-    \item create a separate spec file for each source file
-    \item append {\tt Spec.hs} to the name of spec files
-\end{itemize}
-
-\noindent If we have a source tree like so:
-
-\begin{quote}
-\small
-\begin{verbatim}
-src/Foo.hs
-src/Foo/Bar.hs
-src/Baz.hs
-\end{verbatim}
-\end{quote}
-
-\noindent We would end up with the following spec files.
-
-\begin{quote}
-\small
-\begin{verbatim}
-test/FooSpec.hs
-test/Foo/BarSpec.hs
-test/BazSpec.hs
-\end{verbatim}
-\end{quote}
-
-\begin{itemize}
-\item We still want to combine those specs into a single spec.
-\item By convention we put the combined spec into test/Spec.hs
-\end{itemize}
-
-\begin{itemize}
-\item We use a preprocessor (the {\tt hspec-discover} package\footnote{%
-http://hackage.haskell.org/package/hspec-discover}) to automatically generate this boilerplate.
-\end{itemize}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-05-31 paper.tex
Add top-level spec to Hspec example
Enhance "Integrating multiple test suites"

@@
 
-\begin{quote}
-    TBD: introduce \emph{test driver}
-\end{quote}
-
 If we have evidence that does not fit into the user documentation, but
@@
 import Test.Hspec.Monadic
+
 main :: IO ()
-main = spec $ do
+main = hspec spec
+
+spec :: Spec
+spec = do
   describe "reverse" $ do
@@
 \label{sec:problem-discovery}
+
+\begin{quote}
+    TBD: introduce \emph{test driver}
+\end{quote}
 
@@
 
-A driver for Hspec looks like so:
+As discussed in \ref{sec:solution-discovery}, we can use {\tt
+hspec-discover} to automatically generate a test driver for Hspec:
 
@@
 -- file Spec.hs
-import Test.Hspec.Monadic
-main = hspec spec
-\end{verbatim}
-\end{quote}
-
-\noindent And a corresponding Cabal test suite section like so:
+{-# OPTIONS_GHC -F -pgmF hspec-discover #-}
+\end{verbatim}
+\end{quote}
+
+\noindent A corresponding Cabal test suite section looks like this:
 

### 2012-05-31 paper.tex
Add comment for @kazu
Fix grammar issues

@@
 
-Haskell~\cite{haskell} has a very strong type system, which ensures that most type errors are
+Haskell~\cite{haskell} has a strong type system, which ensures that most type errors are
 detected at compile time rather than runtime.  Thanks to this, we assume that
@@
 For example, it does not ensure that division by zero does not occur.
-Another example is laws of type classes.
-For instance, the {\tt Monoid} class has three Monoid laws
-that its instance \emph{should} satisfy
-but
-it is the programmers responsibility to ensure that
+An other example is, that instances of {\tt Monoid} \emph{should} be
+monoids in the mathematical sense, but the type system does not
+enforce this.  It is the programmers responsibility to ensure that
 instance definitions \emph{behave} as \emph{expected}.
@@
 
-\noindent Note that the {\tt placeholders} library provides better keywords such as {\tt notImplemented} and {\tt todo}, which tells source locations when executed.
+\noindent Note that the {\tt placeholders} library provides better
+keywords such as {\tt notImplemented} and {\tt todo}, which tells
+source locations when executed.
+\begin{quote}
+    sol: @kazu Would it be fine with you to remove the reference to
+    placeholders?
+
+    \begin{itemize}
+        \item Do we need it to make our point?
+        \item It requires template-haskell and additional import, and
+            is thus not very practical.
+    \end{itemize}
+\end{quote}
+
 Next we write user documentation in Haddock markup:

### 2012-05-31 paper.tex
Fix grammar and semantic

@@
 enforce this.  It is the programmers responsibility to ensure that
-instance definitions \emph{behave} as \emph{expected}.
-
-Haskell programmers, however, are apt to rely on the type system too
-much and tend not to write test code.  For example, Haskell
-Platform 2011.4.0.0 contains 23 packages, not counting the 19 core
-libraries that come with GHC.  Only 8 of them have a test suite, 5
-provide example code, and 10 have neither tests nor examples.
-A lack of tests is not only a quality and maintainability issue.  It
-also is a lack of documentation.  Tests document how code is supposed
-to be used, and tests document the behavior of code.
+instance definitions \emph{behave} as \emph{expected}.  Hence
+automated testing is useful.
+
+However, some widely used Haskell libraries do not have automated
+tests.  For example, the Haskell Platform 2011.4.0.0 contains 23
+packages, not counting the 19 core libraries that come with GHC.  Only
+8 of them have a test suite, 5 provide example code, and 10 have
+neither tests nor examples.  A lack of tests is not only a quality and
+maintainability issue.  It also is a lack of documentation.  Tests
+document how code is supposed to be used, and tests document the
+behavior of code.
 

### 2012-05-31 paper.tex
Remove paragraph about "theorems for free"
automatic -> automated + add a TBD

@@
 Our method utilizes specification that can also be used for
-documentation and automatic testing. We hope that this three-in-one
+documentation and automated testing. We hope that this three-in-one
 quality motivates Haskellers to write more test code.
@@
 
-We agree that a function's type can reveal something about a
-function's behavior~\cite{free}, but
-there are many functions whose behavior cannot be guessed from their names and
-type signatures.  One example is \verb|nub :: [a] -> [a]| from {\tt Data.List}.
-
 In other programming language communities, behavior driven development
@@
 In BDD, specifications can be used for
-automatic testing.
+automated testing.
 BDD is meant to be an enjoyable experience for the developer.
@@
 \noindent Our BDD procedure
-can be used for design, documentation and automatic testing.
+can be used for design, documentation and automated testing.
 We hope that this kill-three-birds-with-one-stone aspect
@@
 
+\begin{quote}
+    TBD
+
 \begin{itemize}
@@
     suites.  Where should we put this?
+
+\item automatically sounds strange, better use automated somehow
 \end{itemize}
+\end{quote}
 

### 2012-05-31 paper.tex
Wording
Grammar

@@
 makes tests very easy to read.  And like TDD it puts emphasis on
-test automation, to take work load from the shoulders of
+test automation, to take workload off the shoulders of
 developers.
@@
 document and test Haskell code.
-Both, Doctest and Hspec are missing some features.  So we extend them
-as needed.  In particular, our contributions are:
+Both, Doctest and Hspec were missing some features.  So we extended
+them as needed.  In particular, our contributions are:
 

### 2012-05-31 paper.tex
Simplify a footnote

@@
 Hspec was originally implemented by Trystan Spangler, and is now
-maintained and actively/further developed by one of the authors.}.
+maintained by one of the authors.}.
 

### 2012-05-31 paper.tex
Add an answer for @kazu

@@
 \begin{quote}
-    \emph{kazu: Monoid is already described in Introduction. Why is this repeated? And ``code documentation'' is aslo already dicussed in Introduction. I think we can use ``code documentation'' here.}
+    kazu: Monoid is already described in Introduction. Why is this
+    repeated? And ``code documentation'' is aslo already dicussed in
+    Introduction. I think we can use ``code documentation'' here.
+
+    sol: We reuse Monoid here, but we make a different point.  So this
+    still adds something, and I think we need it.
 \end{quote}

### 2012-05-31 paper.tex
Wording

@@
 
-We will show how to use Doctest and Hspec in combination, to design,
-document and test Haskell code.
-Both, Doctest and Hspec were missing some features.  So we extended
-them as needed.  In particular, our contributions are:
+We show how to use Doctest and Hspec in combination, to design,
+document and test Haskell code.  Both, Doctest and Hspec are missing
+some features.  So we extend them as needed.  In particular, our
+contributions are:
 
@@
 \item
-    We added support for QuickCheck properties to Doctest and extended
+    We added support for QuickCheck properties to Doctest, and extended
     Haddock's markup to support properties
@@
 \item
-    We show best current practice of BDD in Haskell, using Doctest and
-    Hspec (Section \ref{ref:Best-current-practice})
+    We show best current practices of BDD in Haskell, using Doctest
+    and Hspec (Section \ref{ref:Best-current-practice})
 
@@
 
-\noindent Our BDD procedure
-can be used for design, documentation and automated testing.
-We hope that this kill-three-birds-with-one-stone aspect
-motivates Haskellers to write test code.
+\noindent Our approach can be used for design, documentation and
+automated testing.  We hope that this motivates Haskellers to write
+more and better tests.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Best current practice}
+\section{Best current practices}
 \label{ref:Best-current-practice}
 
-This section show best current practice of
+This section show best current practices of
 BDD in Haskell.

### 2012-05-31 paper.tex
Add mkl as author, commented
adding supplyment phrase.

@@
            {sol@typeful.net}
+%\authorinfo{Markus KLINIK}
+%           {Elektrobit Automotive GmbH}
+%           {markus.klinik@elektrobit.com}
 \authorinfo{Kazuhiko YAMAMOTO}
@@
 \section{Introduction}
+\label{sec:introduction}
 
@@
 
+As we discussed in Section~\ref{sec:introduction},
 We think that some evidence is useful as user documentation.
@@
 instance is a monoid in the mathematical sense is self-evident.
-
-\begin{quote}
-    kazu: Monoid is already described in Introduction. Why is this
-    repeated? And ``code documentation'' is aslo already dicussed in
-    Introduction. I think we can use ``code documentation'' here.
-
-    sol: We reuse Monoid here, but we make a different point.  So this
-    still adds something, and I think we need it.
-\end{quote}
 

### 2012-05-31 paper.tex
adding reference.

@@
     extended Haddock's markup to support examples, back in 2010
+    (Section \ref{sec:haddock} and \ref{sec:doctest})
 \item
@@
 \subsection{Haddock}
+\label{sec:haddock}
 

### 2012-05-31 paper.tex
let Emacs's hilighting work.

@@
 
-\begin{quote}\small\begin{verbatim}
+\begin{quote}
+\small
+\begin{verbatim}
 [1, 2, 3].reverse.should eq([3, 2, 1])
-\end{verbatim}\end{quote}
+\end{verbatim}
+\end{quote}
 
@@
 
-\begin{quote}\small\begin{verbatim}
+\begin{quote}
+\small
+\begin{verbatim}
 expect {launch_missiles}.to raise_error
-\end{verbatim}\end{quote}
+\end{verbatim}
+\end{quote}
 
@@
 
-\begin{quote}\small\begin{verbatim}
+\begin{quote}
+\small
+\begin{verbatim}
 type Expectation = Assertion
-\end{verbatim}\end{quote}
+\end{verbatim}
+\end{quote}
 
@@
 
-\begin{quote}\small\begin{verbatim}
+\begin{quote}
+\small
+\begin{verbatim}
 x `shouldBe` 23
-\end{verbatim}\end{quote}
+\end{verbatim}
+\end{quote}
 
@@
 
-\begin{quote}\small\begin{verbatim}
+\begin{quote}
+\small
+\begin{verbatim}
 shouldBe :: (Show a, Eq a) => a -> a -> Expectation
-\end{verbatim}\end{quote}
+\end{verbatim}
+\end{quote}
 
@@
 
-\begin{quote}\small\begin{verbatim}
+\begin{quote}
+\small
+\begin{verbatim}
 x `shouldSatisfy` (< 23)
-\end{verbatim}\end{quote}
+\end{verbatim}
+\end{quote}
 
@@
 
-\begin{quote}\small\begin{verbatim}
+\begin{quote}
+\small
+\begin{verbatim}
 shouldSatisfy :: Show a => a -> (a -> Bool)
   -> Expectation
-\end{verbatim}\end{quote}
+\end{verbatim}
+\end{quote}
 
@@
 
-\begin{quote}\small\begin{verbatim}
+\begin{quote}
+\small
+\begin{verbatim}
 launchMissiles >>=
   (`shouldBe` Left "not implemented")
-\end{verbatim}\end{quote}
+\end{verbatim}
+\end{quote}
 
@@
 
-\begin{quote}\small\begin{verbatim}
+\begin{quote}
+\small
+\begin{verbatim}
 launchMissiles `shouldReturn` Left "not implemented"
-\end{verbatim}\end{quote}
+\end{verbatim}
+\end{quote}
 
@@
 
-\begin{quote}\small\begin{verbatim}
+\begin{quote}
+\small
+\begin{verbatim}
 shouldReturn :: (Show a, Eq a) => IO a -> a
   -> Expectation
-\end{verbatim}\end{quote}
+\end{verbatim}
+\end{quote}
 
@@
 
-\begin{quote}\small\begin{verbatim}
+\begin{quote}
+\small
+\begin{verbatim}
 shouldThrow :: Exception e => IO a -> (e -> Bool)
   -> Expectation
-\end{verbatim}\end{quote}
+\end{verbatim}
+\end{quote}
 
@@
 
-\begin{quote}\small\begin{verbatim}
+\begin{quote}
+\small
+\begin{verbatim}
 launchMissiles `shouldThrow` (== ExitFailure 1)
-\end{verbatim}\end{quote}
+\end{verbatim}
+\end{quote}
 
@@
 
-\begin{quote}\small\begin{verbatim}
+\begin{quote}
+\small
+\begin{verbatim}
 launchMissiles `shouldThrow` isPermissionError
-\end{verbatim}\end{quote}
+\end{verbatim}
+\end{quote}
 
@@
 
-\begin{quote}\small\begin{verbatim}
+\begin{quote}
+\small
+\begin{verbatim}
 anyArithException :: ArithException -> Bool
 anyArithException = const True
-\end{verbatim}\end{quote}
+\end{verbatim}
+\end{quote}
 
@@
 
-\begin{quote}\small\begin{verbatim}
+\begin{quote}
+\small
+\begin{verbatim}
 evaluate (1 `div` 0) `shouldThrow` anyArithException
-\end{verbatim}\end{quote}
+\end{verbatim}
+\end{quote}
 
@@
 
-\begin{quote}\small\begin{verbatim}
+\begin{quote}
+\small
+\begin{verbatim}
 sort xs == (sort . sort) (xs :: [Int]))
-\end{verbatim}\end{quote}
+\end{verbatim}
+\end{quote}
 
@@
 
-\begin{quote}\small\begin{verbatim}
+\begin{quote}
+\small
+\begin{verbatim}
 \xs -> sort xs == (sort . sort) (xs :: [Int])
-\end{verbatim}\end{quote}
+\end{verbatim}
+\end{quote}
 

### 2012-05-31 paper.tex
Section~\ref{} style.

@@
     extended Haddock's markup to support examples, back in 2010
-    (Section \ref{sec:haddock} and \ref{sec:doctest})
+    (Section~\ref{sec:haddock} and Section~\ref{sec:doctest})
 \item
@@
     Haddock's markup to support properties
-    (Section
-    \ref{sec:user-doc} and Section \ref{sec:doctest})
+    (Section~\ref{sec:user-doc} and Section~\ref{sec:doctest})
 
@@
     We introduce a set of combinators to express expectations about
-    code examples for use with Hspec (Section \ref{sec:code-doc} and
-    Section \ref{sec:hspec-comb})
-
-\item
-    We improve CI support for Doctest and Hspec (Section
-    \ref{sec:TBD})
+    code examples for use with Hspec
+    (Section~\ref{sec:code-doc} and Section~\ref{sec:hspec-comb})
+
+\item
+    We improve CI support for Doctest and Hspec
+    (Section~\ref{sec:TBD})
 
@@
     We show best current practices of BDD in Haskell, using Doctest
-    and Hspec (Section \ref{ref:Best-current-practice})
+    and Hspec (Section~\ref{ref:Best-current-practice})
 
@@
 
-As we explained in Section \ref{sec:problem},
+As we explained in Section~\ref{sec:problem},
 we use Doctest for user documentation and Hspec for code documentation.
@@
 boilerplate code, similar to what is shown in
-\ref{sec:problem-discovery}.
+Section~\ref{sec:problem-discovery}.
 
@@
 
-\noindent All combinators introduced in \ref{sec:code-doc} have been
+\noindent All combinators introduced in Section~\ref{sec:code-doc} have been
 implemented.  They are available form Hackage\footnote{%
@@
 
-As discussed in \ref{sec:solution-discovery}, we use a preprocessor to
+As discussed in Section~\ref{sec:solution-discovery}, we use a preprocessor to
 automatically generate test drivers for Hspec.  GHC invokes the
@@
 
-As discussed in \ref{sec:solution-discovery}, we can use {\tt
-hspec-discover} to automatically generate a test driver for Hspec:
+As discussed in Section~\ref{sec:solution-discovery}, we can use
+{\tt hspec-discover} to automatically generate a test driver for Hspec:
 

### 2012-05-31 paper.tex
more \ref fix.

@@
     We show best current practices of BDD in Haskell, using Doctest
-    and Hspec (Section~\ref{ref:Best-current-practice})
+    and Hspec (Section~\ref{sec:best-current-practice})
 
@@
 \subsubsection{Expecting exceptions}
+\label{sec:expacting-exceptions}
 
@@
 \section{Best current practices}
-\label{ref:Best-current-practice}
+\label{sec:best-current-practice}
 
@@
 The {\tt testpack} package provides a function to check
-exception, similar to {\tt shouldThrow} described in Section~\ref{todo}.
+exception, similar to {\tt shouldThrow} described in Section~\ref{sec:expacting-exceptions}.
 

### 2012-05-31 paper.tex
adding "~" for \cite{}.
adding TBD placeholder.
adding references.

@@
     We improve CI support for Doctest and Hspec
-    (Section~\ref{sec:TBD})
+    (Section~\ref{sec:solution-discovery} and Section~\ref{sec:implementation-discovery})
 
@@
 \item[property]
-    Property means a QuickCheck property as defined in
-    \cite{quickcheck}.
+    Property means a QuickCheck property as defined in~\cite{quickcheck}.
 
@@
     BDD terminology.
+\item TBD for CI
 \end{itemize}
@@
 \label{sec:solution-discovery}
+
 We want to automatically discover all specs of a project, so that we

### 2012-05-31 paper.tex
removing TBD.
removing CI.

@@
 \item
-    We improve CI support for Doctest and Hspec
+    We improve automatic discovery support for Doctest and Hspec
     (Section~\ref{sec:solution-discovery} and Section~\ref{sec:implementation-discovery})
-
-    TBD:
-    \emph{This relates to hspec-discover and Doctest's Cabal support.}
 
@@
     BDD terminology.
-\item TBD for CI
+\item
+    Automatic discovery for specs is missing (TBD)
 \end{itemize}

### 2012-06-01 paper.tex
removing placeholder package.

@@
 
-\noindent Note that the {\tt placeholders} library provides better
-keywords such as {\tt notImplemented} and {\tt todo}, which tells
-source locations when executed.
-\begin{quote}
-    sol: @kazu Would it be fine with you to remove the reference to
-    placeholders?
-
-    \begin{itemize}
-        \item Do we need it to make our point?
-        \item It requires template-haskell and additional import, and
-            is thus not very practical.
-    \end{itemize}
-\end{quote}
-
-Next we write user documentation in Haddock markup:
+\noindent Next we write user documentation in Haddock markup:
 

### 2012-06-01 paper.tex
writting up Conclusion.

@@
 
-TBD:
-
-\begin{itemize}
-\item
-    We discussed how to apply BDD to Haskell.
-\item
-    We proposed to use Doctest for user documentation and Hspec for
-    code documentation.
-\item
-    We added markup support for properties and examples to Haddock.
-    And we implemented Doctest, to verify examples and test properties
-    from Haddock comments.
-\item
-    We extend Hspec with a set of combinators for expectations.
-\item
-    With these two tools, specification can also be used for
-    documentation and automatic testing.
-\item
-    We describe a mechanism, that can automatically discover all specs
-    of a project.
-\item
-    We think that this enhances the value of test code, and we hope
-    that it will motivate Haskell programmers to write more and better
-    tests.
-\end{itemize}
+This paper discussed how to apply behavior driven development to
+Haskell.  We proposed to use Doctest for user documentation and Hspec
+for code documentation.  For user documentation, two markups for
+properties and examples are added to Haddock.  And we implemented
+Doctest to verify examples and test properties from Haddock comments.
+For code documentation, a set of combinators for expectations is
+extended for Hspec.  With these two tools, specification can also be
+used for documentation and automatic testing.  We also described a
+mechanism that can automatically discover all specs of a project.  We
+think that our approach enhances the value of test code, and we hope
+that three-in-one aspect will motivate Haskell programmers to write
+more and better
 

### 2012-06-01 paper.tex
brush up Conclusion.

@@
 Haskell.  We proposed to use Doctest for user documentation and Hspec
-for code documentation.  For user documentation, two markups for
-properties and examples are added to Haddock.  And we implemented
+for code documentation. We added two markups for
+properties and examples to Haddock and implemented
 Doctest to verify examples and test properties from Haddock comments.
-For code documentation, a set of combinators for expectations is
-extended for Hspec.  With these two tools, specification can also be
+We also extended a set of combinators for expectations for Hspec.
+With these two tools, specification can also be
 used for documentation and automatic testing.  We also described a
-mechanism that can automatically discover all specs of a project.  We
-think that our approach enhances the value of test code, and we hope
+mechanis that can automatically discover all specs of a project.  We
+think that our approach enhances the value of test code, and hope
 that three-in-one aspect will motivate Haskell programmers to write
-more and better
+more and better test code.
 

### 2012-06-01 example/Step1.hs
Add an example for "Best current practices"

@@ -1,0 +1,11 @@
+module Fib where
+
+
+
+
+
+
+
+-- | Calculate Fibonacci numbers.
+fib :: Int -> Integer
+fib = undefined

### 2012-06-01 example/Step2.hs

@@ -1,0 +1,14 @@
+module Fib where
+
+
+
+
+
+
+
+-- | Calculate Fibonacci numbers.
+--
+-- >>> fib 10
+-- 55
+fib :: Int -> Integer
+fib = const 55

### 2012-06-01 example/Step3.hs

@@ -1,0 +1,29 @@
+module Fib where
+import Control.Applicative
+
+
+
+import Test.Hspec.ShouldBe
+import Test.QuickCheck
+
+-- | Calculate Fibonacci numbers.
+--
+-- >>> fib 10
+-- 55
+fib :: Int -> Integer
+fib 0 = 0
+fib 1 = 1
+fib n = fib (n - 1) + fib (n - 2)
+
+-- small non-negative numbers
+newtype Small = Small Int
+  deriving Show
+
+instance Arbitrary Small where
+  arbitrary = Small . (`mod` 10) <$> arbitrary
+
+main = hspec $ do
+  describe "fib" $ do
+    it "calculates arbitrary Fibonacci numbers" $ do
+      property $ \(Small n) ->
+        fib n == fib (n + 2) - fib (n + 1)

### 2012-06-01 example/Step4.hs

@@ -1,0 +1,32 @@
+module Fib where
+import Control.Applicative
+import Control.Exception
+import Data.List
+import System.Timeout
+import Test.Hspec.ShouldBe
+import Test.QuickCheck
+
+-- | Calculate Fibonacci numbers.
+--
+-- >>> fib 10
+-- 55
+fib :: Int -> Integer
+fib n = xs !! n
+  where
+    xs = 0 : 1 : zipWith (+) xs (tail xs)
+
+-- small non-negative numbers
+newtype Small = Small Int
+  deriving Show
+
+instance Arbitrary Small where
+  arbitrary = Small . (`mod` 1000) <$> arbitrary
+
+main = hspec $ do
+  describe "fib" $ do
+    it "calculates arbitrary Fibonacci numbers" $ do
+      property $ \(Small n) ->
+        fib n == fib (n + 2) - fib (n + 1)
+
+    it "is efficient" $ do
+      timeout 10000 (evaluate $ fib 32) `shouldReturn` Just 2178309

### 2012-06-01 example/Step5.hs

@@ -1,0 +1,35 @@
+module Fib where
+import Control.Applicative
+import Control.Exception
+import Data.List
+import System.Timeout
+import Test.Hspec.ShouldBe
+import Test.QuickCheck
+
+-- | Calculate Fibonacci numbers.
+--
+-- >>> fib 10
+-- 55
+fib :: Int -> Integer
+fib n = xs !! n
+  where
+    xs = 0 : 1 : zipWith (+) xs (tail xs)
+
+-- small non-negative numbers
+newtype Small = Small Int
+  deriving Show
+
+instance Arbitrary Small where
+  arbitrary = Small . (`mod` 1000) <$> arbitrary
+
+main = hspec $ do
+  describe "fib" $ do
+    it "calculates arbitrary Fibonacci numbers" $ do
+      property $ \(Small n) ->
+        fib n == fib (n + 2) - fib (n + 1)
+
+    it "is efficient" $ do
+      timeout 10000 (evaluate $ fib 32) `shouldReturn` Just 2178309
+
+    it "throws an exception on negative input" $ do
+      evaluate (fib (-10)) `shouldThrow` anyException

### 2012-06-01 example/Step4.hs
Extend example for "best current practices"

@@ -11,9 +11,10 @@ import Test.QuickCheck
 -- >>> fib 10
 -- 55
 fib :: Int -> Integer
-fib n = xs !! n
-  where
-    xs = 0 : 1 : zipWith (+) xs (tail xs)
+fib n = fibs !! n
+
+
+fibs = 0 : 1 : zipWith (+) fibs (tail fibs)
 
 -- small non-negative numbers
 newtype Small = Small Int

### 2012-06-01 example/Step5.hs

@@ -11,9 +11,10 @@ import Test.QuickCheck
 -- >>> fib 10
 -- 55
 fib :: Int -> Integer
-fib n = xs !! n
-  where
-    xs = 0 : 1 : zipWith (+) xs (tail xs)
+fib n = fibs !! n
+
+
+fibs = 0 : 1 : zipWith (+) fibs (tail fibs)
 
 -- small non-negative numbers
 newtype Small = Small Int
@@ -31,5 +32,5 @@ main = hspec $ do
     it "is efficient" $ do
       timeout 10000 (evaluate $ fib 32) `shouldReturn` Just 2178309
 
-    it "throws an exception on negative input" $ do
-      evaluate (fib (-10)) `shouldThrow` anyException
+    it "throws ErrorCall on negative input" $ do
+      evaluate (fib (-10)) `shouldThrow` anyErrorCall

### 2012-06-01 example/Step6.hs

@@ -1,8 +1,8 @@
 module Fib where
 import Control.Applicative
-
-
-
+import Control.Exception
+import Data.List
+import System.Timeout
 import Test.Hspec.ShouldBe
 import Test.QuickCheck
 
@@ -11,20 +11,26 @@ import Test.QuickCheck
 -- >>> fib 10
 -- 55
 fib :: Int -> Integer
-fib 0 = 0
-fib 1 = 1
-fib n = fib (n - 1) + fib (n - 2)
+fib n | n < 0     = 0
+      | otherwise = fibs !! n
 
+fibs = 0 : 1 : zipWith (+) fibs (tail fibs)
 
 -- small non-negative numbers
 newtype Small = Small Int
   deriving Show
 
 instance Arbitrary Small where
-  arbitrary = Small . (`mod` 10) <$> arbitrary
+  arbitrary = Small . (`mod` 1000) <$> arbitrary
 
 main = hspec $ do
   describe "fib" $ do
     it "calculates arbitrary Fibonacci numbers" $ do
       property $ \(Small n) ->
         fib n == fib (n + 2) - fib (n + 1)
+
+    it "is efficient" $ do
+      timeout 10000 (evaluate $ fib 32) `shouldReturn` Just 2178309
+
+    it "returns 0 on negative input" $ do
+      fib (-10) `shouldBe` 0

### 2012-06-01 example/Step7.hs

@@ -1,8 +1,8 @@
 module Fib where
 import Control.Applicative
-
-
-
+import Control.Exception
+import Data.List
+import System.Timeout
 import Test.Hspec.ShouldBe
 import Test.QuickCheck
 
@@ -11,20 +11,27 @@ import Test.QuickCheck
 -- >>> fib 10
 -- 55
 fib :: Int -> Integer
-fib 0 = 0
-fib 1 = 1
-fib n = fib (n - 1) + fib (n - 2)
+fib n | n < 0     = 0
+      | otherwise = fibs !! n
 
+fibs = 0 : 1 : zipWith (+) fibs (tail fibs)
 
 -- small non-negative numbers
 newtype Small = Small Int
   deriving Show
 
 instance Arbitrary Small where
-  arbitrary = Small . (`mod` 10) <$> arbitrary
+  arbitrary = Small . (`mod` 1000) <$> arbitrary
 
 main = hspec $ do
   describe "fib" $ do
     it "calculates arbitrary Fibonacci numbers" $ do
       property $ \(Small n) ->
         fib n == fib (n + 2) - fib (n + 1)
+
+    it "is efficient" $ do
+      timeout 10000 (evaluate $ fib 32) `shouldReturn` Just 2178309
+
+    it "returns 0 on negative input" $ do
+      property $ \n -> n < 0 ==>
+        fib n == 0

### 2012-06-01 paper.tex
"Integrating multiple test suites" -> "Cabal integration"
More on cabal integration

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Integrating multiple test suites}
+\subsection{Cabal integration}
 
@@
 
-As discussed in Section~\ref{sec:solution-discovery}, we can use
-{\tt hspec-discover} to automatically generate a test driver for Hspec:
+As discussed in Section~\ref{sec:solution-discovery},
+{\tt hspec-discover} can be used to automatically generate a test
+driver for Hspec:
 
@@
 
-\noindent A driver for Doctest looks like so:
+\noindent Doctest provides both, an executable and a library.  The library
+exposes a function \verb|doctest| of type:
+
+\begin{quote}
+\small
+\begin{verbatim}
+doctest :: [String] -> IO ()
+\end{verbatim}
+\end{quote}
+
+\noindent Doctest's own `main` is simply:
+\begin{quote}
+\small
+\begin{verbatim}
+main = getArgs >>= doctest
+\end{verbatim}
+\end{quote}
+
+\noindent Consequently, it is possible to create a custom executable
+for a project, by passing all command-line arguments that are required
+for that project to `doctest`.  A simple example looks like this:
 
@@
 import Test.DocTest
-main = doctest ["src/Main.hs"]
-\end{verbatim}
-\end{quote}
-
-\noindent And a corresponding Cabal test suite section like so:
+main = doctest ["--optghc=isrc", "src/Main.hs"]
+\end{verbatim}
+\end{quote}
+
+\noindent And a corresponding Cabal test suite section like this:
 

### 2012-06-01 paper.tex
Wording

@@
 
-The {\tt testpack} package provides a function to check
-exception, similar to {\tt shouldThrow} described in Section~\ref{sec:expacting-exceptions}.
+The {\tt testpack} package provides a combinator to expect exceptions,
+similar to {\tt shouldThrow} from
+Section~\ref{sec:expacting-exceptions}:
 
@@
 result of the {\tt IO} action, further limiting it's applicability.
-It only allows to expect specific values, expecting
-any exception of a certain type is not possible.
+It only allows to expect specific exception values, expecting
+arbitrary exceptions of a certain type is not possible.
 

### 2012-06-01 paper.tex
Semantic fix
Wording

@@
 SmallCheck~\cite{smallcheck} is a tool for
-automatic exhaustive testing at a certain level.
+automatic exhaustive testing up to a certain depth.
 Both Doctest and Hspec do not have support for SmallCheck
 at this moment.
-Hspec's mechanism for evidence is extensible.
-If necessary, it would be easy to support SmallCheck.
+Hspec's mechanism for evidence is extensible, and it would be easy to
+add support for SmallCheck.
 

### 2012-06-01 paper.tex
Remove comment
Comment a comment

@@
 
-\begin{quote}
-    sol: We e.g. say ill-behaved Monoid instance.  So is
-    ``behavior'' really a ``bad'' word?  But I agree that some people in
-    the community might associated it with something imperative.
-
-    kazu: ``behavior'' sounded strange at least to me at the first time.
-    We alreay supported this word with typeclassopedia. So, I think
-    this paragraph is completed. Should we replace ``is associated'' with
-    ``might be associated''?
-
-    sol: @kazu, this was primarily meant for @mkl as reference.
-
-    One might also argue that ``should'' is not very mathematical, but
-    it is common language when we talk about mathematics.  E.g. Brent
-    Yorgey uses both ``should'' and ``behave'' a lot in the
-    Typeclassopedia (e.g. ``\ldots {\tt Alternative} should satisfy
-    the monoid laws\ldots'')
-\end{quote}
+%\begin{quote}
+%    sol: We e.g. say ill-behaved Monoid instance.  So is
+%    ``behavior'' really a ``bad'' word?  But I agree that some people in
+%    the community might associated it with something imperative.
+%
+%    kazu: ``behavior'' sounded strange at least to me at the first time.
+%    We alreay supported this word with typeclassopedia. So, I think
+%    this paragraph is completed. Should we replace ``is associated'' with
+%    ``might be associated''?
+%
+%    sol: @kazu, this was primarily meant for @mkl as reference.
+%
+%    One might also argue that ``should'' is not very mathematical, but
+%    it is common language when we talk about mathematics.  E.g. Brent
+%    Yorgey uses both ``should'' and ``behave'' a lot in the
+%    Typeclassopedia (e.g. ``\ldots {\tt Alternative} should satisfy
+%    the monoid laws\ldots'')
+%\end{quote}
 
@@
 
-\begin{quote}
-    sol: I think you can have several files with tests, but you have
-    to combine the test suites manually.
-\end{quote}
-
 The {\tt testpack} package provides a combinator to expect exceptions,

### 2012-06-01 paper.tex
Update comment

@@
 \begin{quote}
-    \emph{sol: Can we remove what follows?}
-
-    \emph{kazu: I want to keep them for readability.}
-
-    \emph{sol: For me it really depends on how often we use something.
-    If we only use it once or twice, I'd try make it clear from the
-    context what is meant.}
+    \emph TBD: Check, terminology for consistency, and remove
+    unnecessary terminology (if any).  This has to be done as one of
+    the last steps.
 \end{quote}

### 2012-06-01 paper.tex
Wording
Best current practices: mkl's initial version

@@
 
+% These are ignored, unless 'preprint' option specified.
 \titlebanner{A draft paper to be submitted to Haskell Symposium 2012}
-\preprintfooter{TBD: short description of paper}   % 'preprint' option specified.
+\preprintfooter{TBD: short description of paper}
 
@@
 
+\authorinfo{Kazuhiko YAMAMOTO}
+           {IIJ Innovation Institute Inc.}
+           {kazu@iij.ad.jp}
 \authorinfo{Simon HENGEL}
@@
            {sol@typeful.net}
-%\authorinfo{Markus KLINIK}
-%           {Elektrobit Automotive GmbH}
-%           {markus.klinik@elektrobit.com}
-\authorinfo{Kazuhiko YAMAMOTO}
-           {IIJ Innovation Institute Inc.}
-           {kazu@iij.ad.jp}
+\authorinfo{Markus KLINIK}
+           {Elektrobit Automotive GmbH}
+           {markus.klinik@elektrobit.com}
 
@@
 
-The word ``behavior'' is associated with imperative actions.  Since
-pure functions of Haskell are static, this word might be improper for
-Haskell.  However, this word is already used in the Haskell
-community~\cite{typeclassopedia}.  So, we use ``behavior'' instead of
-defining terminology in this paper.
+\noindent The word ``behavior'' might be associated with ``imperative
+action''.  Since pure functions of Haskell are static, this word might
+be improper for Haskell.  However, this word is already used in the
+Haskell community~\cite{typeclassopedia}.  So, we use ``behavior''
+instead of defining terminology in this paper.
 
@@
 
-This section show best current practices of
-BDD in Haskell.
-
-\subsection{Procedure of BDD}
-
-First, we write a signature and leave the definition of the function undefined.
-
-\begin{quote}
-\small
-\begin{verbatim}
-reverse :: [a] -> [a]
-reverse = undefined
-\end{verbatim}
-\end{quote}
-
-\noindent Next we write user documentation in Haddock markup:
-
-\begin{quote}
-\small
-\begin{alltt}
--- |
--- 'reverse' @xs@ returns the elements of @xs@
--- in reverse order. @xs@ must be finite.
+This section shows how behaviour driven development works in practice,
+by means of the fibonacci function.  Of course this is a very simple
+example, and the reader is encouraged to imagine a more sophisticated
+library, where design decisions might not be as obvious as here.
+
+We present a step-by-step refinement of a single file {\tt Fib.hs}
+that contains both implementation and specs.  Each line of code is
+only given once when introduced, subsequent code snippets are
+incremental updates.  The reader is asked to substitute code parts as
+appropriate.
+
+\subsection{Step 1}
+
+We start with the function's documentation, which gives a usage
+example, together with the simplest possible implementation that makes
+the example work.
+
+\begin{quote}
+\small
+\begin{verbatim}
+module Fib where
+-- | Calculate Fibonacci numbers.
 --
--- >>> reverse [1,2,3]
--- [3,2,1]
+-- >>> fib 10
+-- 55
+fib :: Int -> Integer
+fib = const 55
+\end{verbatim}
+\end{quote}
+
+This first step might appear useless, but as of now we already have
+made some choices with regard to the API of our library.  By giving a
+usage example, we were forced to think about which arguments our
+function takes.  We can now run Doctest to verify that the example
+works.
+
+\begin{quote}
+\small
+\begin{verbatim}
+$ doctest Fib.hs
+There is one test, with one single interaction.
+Examples: 1  Tried: 1  Errors: 0  Failures: 0
+\end{verbatim}
+\end{quote}
+
+\subsection{Step 2}
+
+We add a specification that forces our function to actually calculate
+fibonacci numbers.
+
+\begin{quote}
+\small
+\begin{verbatim}
+import Control.Applicative
+import Test.Hspec.ShouldBe
+import Test.QuickCheck
+
+-- small non-negative numbers
+newtype Small = Small Int
+  deriving Show
+
+instance Arbitrary Small where
+  arbitrary = Small . (`mod` 10) <$> arbitrary
+
+main = hspec $ do
+  describe "fib" $ do
+    it "calculates arbitrary Fibonacci numbers" $ do
+      property $ \(Small n) ->
+        fib n == fib (n + 2) - fib (n + 1)
+\end{verbatim}
+\end{quote}
+
+\noindent Of course, this test now fails when we run it,
+
+\begin{quote}
+\small
+\begin{verbatim}
+$ runhaskell Fib.hs
+fib
+ - calculates arbitrary Fibonacci numbers FAILED [1]
+\end{verbatim}
+\end{quote}
+
+\noindent so we add a working implementation.  Both Doctest and Hspec
+are happy now.
+
+\begin{quote}
+\small
+\begin{verbatim}
+fib 0 = 0
+fib 1 = 1
+fib n = fib (n - 1) + fib (n - 2)
+\end{verbatim}
+\end{quote}
+
+\subsection{Step 3}
+
+As the seasoned Haskeller might have already noticed, our
+implementation is quite inefficient, as can be proved by this
+specification:
+
+\begin{quote}
+\small
+\begin{verbatim}
+import Control.Exception
+import System.Timeout
+
+it "is efficient" $ do
+  timeout 10000
+      (evaluate $ fib 32) `shouldReturn` Just 2178309
+\end{verbatim}
+\end{quote}
+
+\noindent When we run it, hspec tells us
+
+\begin{quote}
+\small
+\begin{verbatim}
+$ runhaskell Fib.hs
+fib
+ - calculates arbitrary Fibonacci numbers
+ - is efficient FAILED [1]
+\end{verbatim}
+\end{quote}
+
+\noindent and we respond with an efficient implementation.
+
+\begin{quote}
+\small
+\begin{verbatim}
+import Data.List
+
+fib :: Int -> Integer
+fib n = fibs !! n
+fibs = 0 : 1 : zipWith (+) fibs (tail fibs)
+\end{verbatim}
+\end{quote}
+
+\subsection{Step 4}
+
+So far, so good.  We have an implementaion of the fibonacci function
+that swiftly calculates correct values for all positive integers.
+But what about negative ones? Our implementation uses the list index
+operator {\tt (!!)} which is a partial function that throws an
+exeption for negative indexes.  Let's specify that behaviour.
+
+\begin{quote}
+\small
+\begin{verbatim}
+it "throws ErrorCall on negative input" $ do
+  evaluate (fib (-10)) `shouldThrow` anyErrorCall
+\end{verbatim}
+\end{quote}
+
+\noindent Throwing exceptions from pure code is considered bad
+practice, so we guard the {\tt (!!)} operator against negative
+indexes.
+
+\begin{quote}
+\small
+\begin{verbatim}
+fib n | n < 0     = 0
+      | otherwise = fibs !! n
+\end{verbatim}
+\end{quote}
+
+\noindent Now our specification fails, so we adapt it accordingly.
+
+\begin{quote}
+\small
+\begin{verbatim}
+it "returns 0 on negative input" $ do
+  fib (-10) `shouldBe` 0
+\end{verbatim}
+\end{quote}
+
+\noindent In fact, it is the case that fib returns 0 for all negative
+numbers.
+
+\begin{quote}
+\small
+\begin{verbatim}
+it "returns 0 on negative input" $ do
+  property $ \n -> n < 0 ==>
+    fib n == 0
+\end{verbatim}
+\end{quote}
+
+\noindent This concludes our example coding session.  We have seen how
+to utilize behavior driven development to incrementally build a small
+library.  The combinators {\tt shouldBe}, {\tt shouldReturn} and {\tt
+shouldThrow} can be used to specify behavior of functions in an
+intuitive, easy to read manner.  The final version of our library
+looks like this.
+
+\begin{quote}
+\small
+\begin{verbatim}
+module Fib where
+import Control.Applicative
+import Control.Exception
+import Data.List
+import System.Timeout
+import Test.Hspec.ShouldBe
+import Test.QuickCheck
+
+-- | Calculate Fibonacci numbers.
 --
--- prop> reverse [] == reverse []
--- prop> reverse (xs ++ ys) == reverse ys ++ \(\hookleftarrow\)
-                               reverse (xs::[Int])
--- prop> reverse (reverse xs) == (xs::[Int])
-
-reverse :: [a] -> [a]
-reverse = undefined
-\end{alltt}
-\end{quote}
-
-\noindent We should execute Doctest to ensure no syntax errors.
-Also, we write code documentation with Hspec:
-
-\begin{quote}
-    sol: TBD. @kazu, the descriptions make no sense, can you try to
-    make them more precise?
-\end{quote}
-
-\begin{quote}
-\small
-\begin{verbatim}
-describe "reverse" $ do
-  it "returns the first element in the last" $
-    property $ \xs -> not (null xs) ==>
-      head (reverse xs) == last (xs :: [Int])
-
-  it "returns the last element in the first" $
-    property $ \xs -> not (null xs) ==>
-      last (reverse xs) == head (xs :: [Int])
-\end{verbatim}
-\end{quote}
-
-\noindent Also, we should execute this Hspec module
-to ensure no syntax errors.
-Then we implement the function.
-
-\begin{quote}
-\small
-\begin{verbatim}
-reverse :: [a] -> [a]
-reverse = foldl (flip (:)) []
-\end{verbatim}
-\end{quote}
-
-\noindent Now we run tests, and improve our implementation, until all
-tests pass.
+-- >>> fib 10
+-- 55
+fib :: Int -> Integer
+fib n | n < 0     = 0
+      | otherwise = fibs !! n
+
+fibs = 0 : 1 : zipWith (+) fibs (tail fibs)
+
+-- small non-negative numbers
+newtype Small = Small Int
+  deriving Show
+
+instance Arbitrary Small where
+  arbitrary = Small . (`mod` 1000) <$> arbitrary
+
+main = hspec $ do
+  describe "fib" $ do
+    it "calculates arbitrary Fibonacci numbers" $ do
+      property $ \(Small n) ->
+        fib n == fib (n + 2) - fib (n + 1)
+
+    it "is efficient" $ do
+      timeout 10000 (evaluate $ fib 32)
+          `shouldReturn` Just 2178309
+
+    it "returns 0 on negative input" $ do
+      property $ \n -> n < 0 ==>
+        fib n == 0
+\end{verbatim}
+\end{quote}
 
@@
 
-\noindent This can be used to document and automatically test
-arbitrary Monoid instances.
+\noindent This can be used to document and test arbitrary Monoid
+instances.
 Here is an example how to use it for lists:
@@
 \begin{itemize}
-    \item \emph{optional}: Bridging the gap between HUnit and QuickCheck
     \item
@@
     \item
-        Are more high-lever approaches to BDD, akin to Ruby's
-        Cucumber, applicable to Haskell? (Observation: Cucumber's
-        approach is tightly tied to OO)
 \end{itemize}
+
+\begin{quote}
+    TBD: add ``The Cucumber Book'' to the bibliography as {\tt
+    cucumber-book}.
+\end{quote}
+
+Cucumber~\cite{cucumber-book} is a higher-level approach to BDD for Ruby, and
+it is typically used in combination with RSpec.
+It focuses on user stories, to assist the developer in delivering
+value to the stakeholder.
+There is nothing comparable to Cucumber for Haskell, yet.  And we think
+it would be interesting to explore, whether it is possible to apply
+Cucumber's approach to BDD to Haskell.
 
@@
 
-The {\tt test-framework} family is a TDD tool,
-which currently integrates HUnit, QuickCheck, SmallCheck, and Doctest.
-Automatic test discovery based on prefixes of function name
-and is implemented using Template Haskell~\cite{template-haskell}.
-All examples and properties must be written in a single
-source file.
+The {\tt test-framework} family is a TDD tool, which currently
+integrates HUnit, QuickCheck, SmallCheck, and Doctest.
+There is support for automatic test discovery that relies on naming
+conventions for test functions\footnote{%
+http://hackage.haskell.org/package/test-framework-th}.  It is
+implemented with Template Haskell~\cite{template-haskell}, and it only
+works within a single source file.  Automatic discovery across
+multiple source files is not possible with a mechanism that is based
+on Template Haskell, because Template Haskell can not generate import
+statements\footnote{%
+See http://hackage.haskell.org/trac/ghc/ticket/1475 }.
 
@@
 \section{Conclusion}
+
+\begin{quote}
+    TBD: Add "best current practice"
+\end{quote}
 
@@
 
-We are grateful to David Waern, Haddock's maintainer for reviewing and
+We are grateful to David Waern, Haddock's maintainer, for reviewing and
 accepting our patches to Haddock.

### 2012-06-01 paper.tex
Discuss Cucumber

@@
     \item
-        Are more high-lever approaches to BDD, akin to Ruby's
-        Cucumber, applicable to Haskell? (Observation: Cucumber's
-        approach is tightly tied to OO)
 \end{itemize}
+
+\begin{quote}
+    TBD: add ``The Cucumber Book'' to the bibliography as {\tt
+    cucumber-book}.
+\end{quote}
+
+Cucumber~\cite{cucumber-book} is a higher-level approach to BDD for Ruby, and
+it is typically used in combination with RSpec.
+It focuses on user stories, to assist the developer in delivering
+value to the stakeholder.
+There is nothing comparable to Cucumber for Haskell yet.  And we think
+it would be interesting to explore, whether it is possible to apply
+Cucumber's approach to BDD to Haskell.
+
+Are more high-lever approaches to BDD, akin to Ruby's Cucumber,
+applicable to Haskell? (Observation: Cucumber's approach is tightly
+tied to OO)
 

### 2012-06-01 paper.tex
Add a LaTeX comment about how to use the class
Add a comma
Remove leftover
Add mkl as author (author order is reverse alphabetic order)

@@
 value to the stakeholder.
-There is nothing comparable to Cucumber for Haskell yet.  And we think
+There is nothing comparable to Cucumber for Haskell, yet.  And we think
 it would be interesting to explore, whether it is possible to apply
 Cucumber's approach to BDD to Haskell.
-
-Are more high-lever approaches to BDD, akin to Ruby's Cucumber,
-applicable to Haskell? (Observation: Cucumber's approach is tightly
-tied to OO)
 

### 2012-06-01 paper.tex
Add URL for testpack

@@
 
-This section show best current practices of
-BDD in Haskell.
-
-\subsection{Procedure of BDD}
-
-First, we write a signature and leave the definition of the function undefined.
-
-\begin{quote}
-\small
-\begin{verbatim}
-reverse :: [a] -> [a]
-reverse = undefined
-\end{verbatim}
-\end{quote}
-
-\noindent Next we write user documentation in Haddock markup:
-
-\begin{quote}
-\small
-\begin{alltt}
--- |
--- 'reverse' @xs@ returns the elements of @xs@
--- in reverse order. @xs@ must be finite.
+This section shows how behaviour driven development works in practice,
+by means of the fibonacci function.  Of course this is a very simple
+example, but the method is also applicable to more sophisticated
+situations where design decisions might not be as obvious as here.
+
+We present a step-by-step refinement of a single file {\tt Fib.hs}
+that contains both implementation and specs.  Each line of code is
+only given once when introduced, subsequent code fragments are
+incremental updates.  The reader is asked to substitute code parts as
+appropriate.
+
+\subsection{Step 1}
+
+We start with the function's documentation, which gives a usage
+example, together with the simplest possible implementation that makes
+the example work.
+
+\begin{quote}
+\small
+\begin{verbatim}
+module Fib where
+-- | Calculate Fibonacci numbers.
 --
--- >>> reverse [1,2,3]
--- [3,2,1]
+-- >>> fib 10
+-- 55
+fib :: Int -> Integer
+fib = const 55
+\end{verbatim}
+\end{quote}
+
+This first step might appear useless, but as of now we already have
+made some choices with regard to the API of our library.  By giving a
+usage example, we were forced to think about which arguments our
+function takes.  We can now run Doctest to verify that the example
+works.
+
+\begin{quote}
+\small
+\begin{verbatim}
+$ doctest Fib.hs
+There is one test, with one single interaction.
+Examples: 1  Tried: 1  Errors: 0  Failures: 0
+\end{verbatim}
+\end{quote}
+
+\subsection{Step 2}
+
+We add a specification that forces our function to actually calculate
+fibonacci numbers.
+
+\begin{quote}
+\small
+\begin{verbatim}
+import Control.Applicative
+import Test.Hspec.ShouldBe
+import Test.QuickCheck
+
+-- small non-negative numbers
+newtype Small = Small Int
+  deriving Show
+
+instance Arbitrary Small where
+  arbitrary = Small . (`mod` 10) <$> arbitrary
+
+main = hspec $ do
+  describe "fib" $ do
+    it "calculates arbitrary Fibonacci numbers" $ do
+      property $ \(Small n) ->
+        fib n == fib (n + 2) - fib (n + 1)
+\end{verbatim}
+\end{quote}
+
+\noindent Of course, this test now fails when we run it,
+
+\begin{quote}
+\small
+\begin{verbatim}
+$ runhaskell Fib.hs
+fib
+ - calculates arbitrary Fibonacci numbers FAILED [1]
+\end{verbatim}
+\end{quote}
+
+\noindent so we add a working implementation.  Both Doctest and Hspec
+are happy now.
+
+\begin{quote}
+\small
+\begin{verbatim}
+fib 0 = 0
+fib 1 = 1
+fib n = fib (n - 1) + fib (n - 2)
+\end{verbatim}
+\end{quote}
+
+\subsection{Step 3}
+
+As the seasoned Haskeller might have already noticed, our
+implementation is quite inefficient, as can be proved by this
+specification:
+
+\begin{quote}
+\small
+\begin{verbatim}
+import Control.Exception
+import System.Timeout
+
+it "is efficient" $ do
+  timeout 10000
+      (evaluate $ fib 32) `shouldReturn` Just 2178309
+\end{verbatim}
+\end{quote}
+
+\noindent When we run it, hspec tells us
+
+\begin{quote}
+\small
+\begin{verbatim}
+$ runhaskell Fib.hs
+fib
+ - calculates arbitrary Fibonacci numbers
+ - is efficient FAILED [1]
+\end{verbatim}
+\end{quote}
+
+\noindent and we respond with an efficient implementation.
+
+\begin{quote}
+\small
+\begin{verbatim}
+import Data.List
+
+fib :: Int -> Integer
+fib n = fibs !! n
+fibs = 0 : 1 : zipWith (+) fibs (tail fibs)
+\end{verbatim}
+\end{quote}
+
+\subsection{Step 4}
+
+So far, so good.  We have an implementaion of the fibonacci function
+that swiftly calculates correct values for all positive integers.
+But what about negative ones? Our implementation uses the list index
+operator {\tt (!!)} which is a partial function that throws an
+exeption for negative indexes.  Let's specify that behaviour.
+
+\begin{quote}
+\small
+\begin{verbatim}
+it "throws ErrorCall on negative input" $ do
+  evaluate (fib (-10)) `shouldThrow` anyErrorCall
+\end{verbatim}
+\end{quote}
+
+\noindent Throwing exceptions from pure code is considered bad
+practice, so we guard the {\tt (!!)} operator against negative
+indexes.
+
+\begin{quote}
+\small
+\begin{verbatim}
+fib n | n < 0     = 0
+      | otherwise = fibs !! n
+\end{verbatim}
+\end{quote}
+
+\noindent Now our specification fails, so we adapt it accordingly.
+
+\begin{quote}
+\small
+\begin{verbatim}
+it "returns 0 on negative input" $ do
+  fib (-10) `shouldBe` 0
+\end{verbatim}
+\end{quote}
+
+\noindent In fact, it is the case that fib returns 0 for all negative
+numbers.
+
+\begin{quote}
+\small
+\begin{verbatim}
+it "returns 0 on negative input" $ do
+  property $ \n -> n < 0 ==>
+    fib n == 0
+\end{verbatim}
+\end{quote}
+
+\noindent This concludes our example coding session.  We have seen how
+to utilize behavior driven development to incrementally build a small
+library.  The combinators {\tt shouldBe}, {\tt shouldReturn} and {\tt
+shouldThrow} can be used to specify behavior of functions in an
+intuitive, easy to read manner.  The final version of our library
+looks like this.
+
+\begin{quote}
+\small
+\begin{verbatim}
+module Fib where
+import Control.Applicative
+import Control.Exception
+import Data.List
+import System.Timeout
+import Test.Hspec.ShouldBe
+import Test.QuickCheck
+
+-- | Calculate Fibonacci numbers.
 --
--- prop> reverse [] == reverse []
--- prop> reverse (xs ++ ys) == reverse ys ++ \(\hookleftarrow\)
-                               reverse (xs::[Int])
--- prop> reverse (reverse xs) == (xs::[Int])
-
-reverse :: [a] -> [a]
-reverse = undefined
-\end{alltt}
-\end{quote}
-
-\noindent We should execute Doctest to ensure no syntax errors.
-Also, we write code documentation with Hspec:
-
-\begin{quote}
-    sol: TBD. @kazu, the descriptions make no sense, can you try to
-    make them more precise?
-\end{quote}
-
-\begin{quote}
-\small
-\begin{verbatim}
-describe "reverse" $ do
-  it "returns the first element in the last" $
-    property $ \xs -> not (null xs) ==>
-      head (reverse xs) == last (xs :: [Int])
-
-  it "returns the last element in the first" $
-    property $ \xs -> not (null xs) ==>
-      last (reverse xs) == head (xs :: [Int])
-\end{verbatim}
-\end{quote}
-
-\noindent Also, we should execute this Hspec module
-to ensure no syntax errors.
-Then we implement the function.
-
-\begin{quote}
-\small
-\begin{verbatim}
-reverse :: [a] -> [a]
-reverse = foldl (flip (:)) []
-\end{verbatim}
-\end{quote}
-
-\noindent Now we run tests, and improve our implementation, until all
-tests pass.
+-- >>> fib 10
+-- 55
+fib :: Int -> Integer
+fib n | n < 0     = 0
+      | otherwise = fibs !! n
+
+fibs = 0 : 1 : zipWith (+) fibs (tail fibs)
+
+-- small non-negative numbers
+newtype Small = Small Int
+  deriving Show
+
+instance Arbitrary Small where
+  arbitrary = Small . (`mod` 1000) <$> arbitrary
+
+main = hspec $ do
+  describe "fib" $ do
+    it "calculates arbitrary Fibonacci numbers" $ do
+      property $ \(Small n) ->
+        fib n == fib (n + 2) - fib (n + 1)
+
+    it "is efficient" $ do
+      timeout 10000 (evaluate $ fib 32)
+          `shouldReturn` Just 2178309
+
+    it "returns 0 on negative input" $ do
+      property $ \n -> n < 0 ==>
+        fib n == 0
+\end{verbatim}
+\end{quote}
 
@@
 
-The {\tt testpack} package provides a combinator to expect exceptions,
+The {\tt testpack}\footnote{%
+http://hackage.haskell.org/package/testpack}
+package provides a combinator to expect exceptions,
 similar to {\tt shouldThrow} from

### 2012-06-01 paper.tex
BCP: wording

@@
 
-The {\tt testpack}\footnote{%
-http://hackage.haskell.org/package/testpack}
-package provides a combinator to expect exceptions,
+The {\tt testpack} package provides a combinator to expect exceptions,
 similar to {\tt shouldThrow} from

### 2012-06-01 paper.tex
BCP: Step 1, 2: wording

--- 2012-06-01 paper.tex
@@
 
-This first step might appear useless, but as of now we already have
-made some choices with regard to the API of our library.  By giving a
-usage example, we were forced to think about which arguments our
-function takes.  We can now run Doctest to verify that the example
-works.
+\noindent This first step might appear useless, but as of now we
+already have made some choices with regard to the API of our library.
+By giving a usage example, we were forced to think about which
+arguments our function takes.  We can now run Doctest to verify that
+the example works.
 
@@
 
-We add a specification that forces our function to actually calculate
-fibonacci numbers.
+In the next step, we add a specification that forces our function to
+actually calculate fibonacci numbers.
 
@@
 
-\noindent Of course, this test now fails when we run it,
+\noindent Of course, this spec fails when we run it,
 
@@
 \noindent so we add a working implementation.  Both Doctest and Hspec
-are happy now.
+report success after this.
 

### 2012-06-01 paper.tex
BCP: bad practice -> poor design
BCP: fibonacci -> Fibonacci

@@
 This section shows how behaviour driven development works in practice,
-by means of the fibonacci function.  Of course this is a very simple
+by means of the Fibonacci function.  Of course this is a very simple
 example, but the method is also applicable to more sophisticated
@@
 In the next step, we add a specification that forces our function to
-actually calculate fibonacci numbers.
+actually calculate Fibonacci numbers.
 
@@
 
-So far, so good.  We have an implementaion of the fibonacci function
+So far, so good.  We have an implementaion of the Fibonacci function
 that swiftly calculates correct values for all positive integers.
@@
 
-\noindent Throwing exceptions from pure code is considered bad
-practice, so we guard the {\tt (!!)} operator against negative
+\noindent Throwing exceptions from pure code is considered poor
+design, so we guard the {\tt (!!)} operator against negative
 indexes.

### 2012-06-01 paper.tex
BCP: Step 4: spec first, then implementation

@@
 that swiftly calculates correct values for all positive integers.
-But what about negative ones? Our implementation uses the list index
-operator {\tt (!!)} which is a partial function that throws an
-exeption for negative indexes.  Let's specify that behaviour.
+But what about negative ones? Let's assume that we want our
+implementation to throw an exception in this case. We would
+specify it like this.
 
@@
 
-\noindent Throwing exceptions from pure code is considered poor
-design, so we guard the {\tt (!!)} operator against negative
-indexes.
+\noindent Our implementation already fulfils this spec, so we are
+fine. Some time later however, the one person in our team that knows
+more about Haskell than we do, convinces us that throwing exceptions
+from pure code is considered poor design. Hence, we change the
+specification so that zero should be the result for negative inputs.
+
+\begin{quote}
+\small
+\begin{verbatim}
+it "returns 0 on negative input" $ do
+  fib (-10) `shouldBe` 0
+\end{verbatim}
+\end{quote}
+
+\noindent Now our specification fails, so we adapt it accordingly.
 
@@
       | otherwise = fibs !! n
-\end{verbatim}
-\end{quote}
-
-\noindent Now our specification fails, so we adapt it accordingly.
-
-\begin{quote}
-\small
-\begin{verbatim}
-it "returns 0 on negative input" $ do
-  fib (-10) `shouldBe` 0
 \end{verbatim}
@@
 to utilize behavior driven development to incrementally build a small
-library.  The combinators {\tt shouldBe}, {\tt shouldReturn} and {\tt
-shouldThrow} can be used to specify behavior of functions in an
-intuitive, easy to read manner.  The final version of our library
-looks like this.
+library.  We always extend or adapt the specification and watch it
+fail first.  Then, we fix the implementation.  The combinators {\tt
+shouldBe}, {\tt shouldReturn} and {\tt shouldThrow} can be used to
+specify behavior of functions in an intuitive, easy to read manner.
+The final version of our library looks like this.
 

### 2012-06-01 paper.tex
BCP: adapt the implementation, not the spec
BCP: remove import lists and final version

@@
 \begin{verbatim}
-import Control.Applicative
-import Test.Hspec.ShouldBe
-import Test.QuickCheck
-
 -- small non-negative numbers
@@
 \begin{verbatim}
-import Control.Exception
-import System.Timeout
-
 it "is efficient" $ do
@@
 \begin{verbatim}
-import Data.List
-
 fib :: Int -> Integer
@@
 
-\noindent Now our specification fails, so we adapt it accordingly.
+\noindent Now our specification fails, so we adapt the implementation
+accordingly.
 
@@
 specify behavior of functions in an intuitive, easy to read manner.
-The final version of our library looks like this.
-
-\begin{quote}
-\small
-\begin{verbatim}
-module Fib where
-import Control.Applicative
-import Control.Exception
-import Data.List
-import System.Timeout
-import Test.Hspec.ShouldBe
-import Test.QuickCheck
-
--- | Calculate Fibonacci numbers.
---
--- >>> fib 10
--- 55
-fib :: Int -> Integer
-fib n | n < 0     = 0
-      | otherwise = fibs !! n
-
-fibs = 0 : 1 : zipWith (+) fibs (tail fibs)
-
--- small non-negative numbers
-newtype Small = Small Int
-  deriving Show
-
-instance Arbitrary Small where
-  arbitrary = Small . (`mod` 1000) <$> arbitrary
-
-main = hspec $ do
-  describe "fib" $ do
-    it "calculates arbitrary Fibonacci numbers" $ do
-      property $ \(Small n) ->
-        fib n == fib (n + 2) - fib (n + 1)
-
-    it "is efficient" $ do
-      timeout 10000 (evaluate $ fib 32)
-          `shouldReturn` Just 2178309
-
-    it "returns 0 on negative input" $ do
-      property $ \n -> n < 0 ==>
-        fib n == 0
-\end{verbatim}
-\end{quote}
 

### 2012-06-01 paper.tex
Wording

@@
 
-We present a step-by-step refinement of a single file {\tt Fib.hs}
+We present a step-by-step refinement of a library
 that contains both implementation and specs.  Each line of code is
@@
 to utilize behavior driven development to incrementally build a small
-library.  We always extend or adapt the specification and watch it
-fail first.  Then, we fix the implementation.  The combinators {\tt
+library.  We always extend or adapt the specification first, watch it
+fail,  and then fix the implementation.  The combinators {\tt
 shouldBe}, {\tt shouldReturn} and {\tt shouldThrow} can be used to
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Reusable spec}
+\subsection{Reusable pieces of specification}
 

### 2012-06-01 paper.tex
BCP: wording in conclusion
BCP: we refine a library, not a file
Use \subsubsections for steps in BCP
BCP: wording

@@
 
+\subsection{Applying BDD to Haskell: An Example}
 This section shows how behaviour driven development works in practice,
@@
 
-\subsection{Step 1}
+\subsubsection{Step 1}
 
@@
 
-\noindent This first step might appear useless, but as of now we
+\noindent This first step might appear excessively trivial, but as of now we
 already have made some choices with regard to the API of our library.
@@
 
-\subsection{Step 2}
+\subsubsection{Step 2}
 
@@
 
-\noindent Of course, this spec fails when we run it,
-
-\begin{quote}
-\small
-\begin{verbatim}
-$ runhaskell Fib.hs
+\noindent Of course this spec fails when we run it,
+
+\begin{quote}
+\small
+\begin{verbatim}
+$ runhaskell FibSpec.hs
 fib
@@
 \noindent so we add a working implementation.  Both Doctest and Hspec
-report success after this.
+report success after that.
 
@@
 
-\subsection{Step 3}
+\subsubsection{Step 3}
 
 As the seasoned Haskeller might have already noticed, our
-implementation is quite inefficient, as can be proved by this
+implementation is quite inefficient, as made evident by this
 specification:
@@
 \begin{verbatim}
-$ runhaskell Fib.hs
+$ runhaskell FibSpec.hs
 fib
@@
 
-\subsection{Step 4}
+\subsubsection{Step 4}
 

### 2012-06-01 paper.tex
Add TBD
Move "reusable specs" to BCP

@@
 
+\begin{quote}
+    TBD: Section -> section
+\end{quote}
+
 %%%% Describe the problem.
@@
 
+\begin{quote}
+    TBD: in this section we:
+    \begin{itemize}
+        \item We show how to do BDD in Haskell
+        \item How to reuse pieces of specification
+        \item How to use Doctest and Hspec with Cabal's test feature
+    \end{itemize}
+\end{quote}
+
 \subsection{Applying BDD to Haskell: An Example}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Cabal integration}
-
-\begin{quote}
-    TBD
-
-\begin{itemize}
-
-\item It is important that testing is automated, including running of test
-    suites.  Where should we put this?
-
-\item automatically sounds strange, better use automated somehow
-\end{itemize}
-\end{quote}
-
-\noindent Cabal's test feature can be used to automatically run tests~\cite{cabal}.  A cabal
-file can contain multiple test suites, so it is possible to define test suites
-for Doctest and Hspec in the same cabal file.
-
-As discussed in Section~\ref{sec:solution-discovery},
-{\tt hspec-discover} can be used to automatically generate a test
-driver for Hspec:
-
-\begin{quote}
-\small
-\begin{verbatim}
--- file Spec.hs
-{-# OPTIONS_GHC -F -pgmF hspec-discover #-}
-\end{verbatim}
-\end{quote}
-
-\noindent A corresponding Cabal test suite section looks like this:
-
-\begin{quote}
-\small
-\begin{verbatim}
-test-suite spec
-  type:          exitcode-stdio-1.0
-  main-is:       Spec.hs
-  build-depends: base, hspec
-\end{verbatim}
-\end{quote}
-
-\noindent Doctest provides both, an executable and a library.  The library
-exposes a function \verb|doctest| of type:
-
-\begin{quote}
-\small
-\begin{verbatim}
-doctest :: [String] -> IO ()
-\end{verbatim}
-\end{quote}
-
-\noindent Doctest's own `main` is simply:
-\begin{quote}
-\small
-\begin{verbatim}
-main = getArgs >>= doctest
-\end{verbatim}
-\end{quote}
-
-\noindent Consequently, it is possible to create a custom executable
-for a project, by passing all command-line arguments that are required
-for that project to `doctest`.  A simple example looks like this:
-
-\begin{quote}
-\small
-\begin{verbatim}
--- file doctests.hs
-import Test.DocTest
-main = doctest ["--optghc=isrc", "src/Main.hs"]
-\end{verbatim}
-\end{quote}
-
-\noindent And a corresponding Cabal test suite section like this:
-
-\begin{quote}
-\small
-\begin{verbatim}
-test-suite doctests
-  type:          exitcode-stdio-1.0
-  main-is:       doctests.hs
-  build-depends: base, doctest
-\end{verbatim}
-\end{quote}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Open problems and future work}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Reusable pieces of specification}
+\label{sec:reusable-specs}
 
@@
 prepare reusable spec to cover variety of type classes.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Cabal integration}
+
+\begin{quote}
+    TBD
+
+\begin{itemize}
+
+\item It is important that testing is automated, including running of test
+    suites.  Where should we put this?
+
+\item automatically sounds strange, better use automated somehow
+\end{itemize}
+\end{quote}
+
+\noindent Cabal's test feature can be used to automatically run tests~\cite{cabal}.  A cabal
+file can contain multiple test suites, so it is possible to define test suites
+for Doctest and Hspec in the same cabal file.
+
+As discussed in Section~\ref{sec:solution-discovery},
+{\tt hspec-discover} can be used to automatically generate a test
+driver for Hspec:
+
+\begin{quote}
+\small
+\begin{verbatim}
+-- file Spec.hs
+{-# OPTIONS_GHC -F -pgmF hspec-discover #-}
+\end{verbatim}
+\end{quote}
+
+\noindent A corresponding Cabal test suite section looks like this:
+
+\begin{quote}
+\small
+\begin{verbatim}
+test-suite spec
+  type:          exitcode-stdio-1.0
+  main-is:       Spec.hs
+  build-depends: base, hspec
+\end{verbatim}
+\end{quote}
+
+\noindent Doctest provides both, an executable and a library.  The library
+exposes a function \verb|doctest| of type:
+
+\begin{quote}
+\small
+\begin{verbatim}
+doctest :: [String] -> IO ()
+\end{verbatim}
+\end{quote}
+
+\noindent Doctest's own `main` is simply:
+\begin{quote}
+\small
+\begin{verbatim}
+main = getArgs >>= doctest
+\end{verbatim}
+\end{quote}
+
+\noindent Consequently, it is possible to create a custom executable
+for a project, by passing all command-line arguments that are required
+for that project to `doctest`.  A simple example looks like this:
+
+\begin{quote}
+\small
+\begin{verbatim}
+-- file doctests.hs
+import Test.DocTest
+main = doctest ["--optghc=isrc", "src/Main.hs"]
+\end{verbatim}
+\end{quote}
+
+\noindent And a corresponding Cabal test suite section like this:
+
+\begin{quote}
+\small
+\begin{verbatim}
+test-suite doctests
+  type:          exitcode-stdio-1.0
+  main-is:       doctests.hs
+  build-depends: base, doctest
+\end{verbatim}
+\end{quote}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Open problems and future work}
+
+\begin{itemize}
+    \item
+        In section~\ref{sec:reusable-specs}, we showed how to reuse
+        pieces of specification.
+    \item
+        It would be possible and useful to write a library that
+        provides reusable pieces of specification, for many of
+        Haskell's standard classes.
+        (similar in purpose to
+        {\tt
+        checkers}\footnote{http://hackage.haskell.org/package/checkers})
+\end{itemize}
 

### 2012-06-01 paper.tex
BCP: Step 4, conclusion: wording

@@
 from pure code is considered poor design. Hence, we change the
-specification so that zero should be the result for negative inputs.
+specification so that negative inputs should have zero as result.
 
@@
 
-\noindent In fact, it is the case that fib returns 0 for all negative
-numbers.
+\noindent In fact, should be the case that fib returns 0 for all
+negative numbers.
 
@@
 
-\noindent This concludes our example coding session.  We have seen how
-to utilize behavior driven development to incrementally build a small
-library.  We always extend or adapt the specification first, watch it
-fail,  and then fix the implementation.  The combinators {\tt
-shouldBe}, {\tt shouldReturn} and {\tt shouldThrow} can be used to
-specify behavior of functions in an intuitive, easy to read manner.
+\noindent Again, our implementation already fulfils that spec.
+
+This concludes our example coding session.  We have seen how to utilize
+behavior driven development to incrementally build a small library.
+We always extend or adapt the specification first, watch it fail,  and
+then tighten the implementation.  The combinators {\tt shouldBe}, {\tt
+shouldReturn} and {\tt shouldThrow} can be used to specify behavior of
+functions in an intuitive, easy to read manner.
 

### 2012-06-01 paper.tex
BCP: should be the case -> it should be the case
State that hspec generates a report

@@
 A spec can be run with the \texttt{hspec} function.  When a spec is
-run, all evidence is tested.  Behavior with evidence that does not
-hold is reported to the developer.
+run, all evidence is tested, and a report is generated.
+Behavior with evidence that does not hold is marked in the generated
+report as {\tt FAILED}.
 
@@
 
-\noindent In fact, should be the case that fib returns 0 for all
+\noindent In fact, it should be the case that fib returns 0 for all
 negative numbers.

### 2012-06-01 paper.tex
Work on "Open problems"

@@
 
+\begin{quote}
+    TBD: A library for reusable pieces of specification would be
+    useful.
 \begin{itemize}
@@
 \end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Source locations for failing examples}
-
-TBD:
-
-We do not get source locations for failing tests. Doctest solves this.
-Something like JHC's
-\verb|SRCLOC_ANNOTATE|
-pragma\footnote{%
-http://repetae.net/computer/jhc/jhc.shtml} could help
-for Hspec/QuickCheck/HUnit.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{TBD}
-
-\begin{itemize}
-    \item
-        Feedback for failing QuickCheck properties is not always as
-        useful as it could be -> combinators that use QuickCheck's
-        Result type could help
-    \item
-\end{itemize}
+\end{quote}
+
+When Hspec finds evidence that does not hold, it reports the
+corresponding behavior to the developer.  It would be useful to
+include the source file and location where the failing evidence is
+defined, so that the developer can find it immediately.
+Extending GHC with a mechanism similar to JHC's
+\verb|SRCLOC_ANNOTATE| pragma\footnote{%
+http://repetae.net/computer/jhc/jhc.shtml} would allow to solve this.
+
+\begin{quote}
+    sol: should we just remove the next paragraph?
+\end{quote}
+
+Feedback for failing QuickCheck properties is not always as useful as
+it could be.
+Combinators, similar to {\tt shouldBe}, that construct QuickCheck {\tt
+Result}s could improve the situation.
 

### 2012-06-01 paper.tex
Update comment
Update comments

@@
 \begin{quote}
-    TBD: Section -> section
+    TBD: ``Section'' to ``section''
 \end{quote}
@@
     TBD: introduce \emph{test driver}
+
+    sol: not sure if it is necessary, @mkl, @kazu?
 \end{quote}
@@
 
+    sol: I think it is not critical.  I'll try to resolve this
+    tomorrow, but I think we can always remove it.
+
 \begin{itemize}
@@
 \begin{quote}
-    TBD: A library for reusable pieces of specification would be
-    useful.
+    TBD: state that ``A library for reusable pieces of specification would be
+    useful.''
 \begin{itemize}

### 2012-06-02 paper.tex
make Haddock markup valid.

@@
 \begin{quote}
-    sol: @kazu should we make the Haddock examples in this section
-    valid Haddock comments?
-\end{quote}
-
-\begin{quote}
 \small
@@
 \begin{verbatim}
->>> writeFile "tmpfile" "Hello"
->>> readFile "tmpfile"
-"Hello"
+-- |
+-- >>> writeFile "tmpfile" "Hello"
+-- >>> readFile "tmpfile"
+-- "Hello"
 \end{verbatim}
@@
 \begin{verbatim}
->>> head []
-*** Exception: Prelude.head: empty list
+-- |
+-- >>> head []
+-- *** Exception: Prelude.head: empty list
 \end{verbatim}
@@
 \begin{verbatim}
-prop> Data.Map.null empty == True
+-- |
+-- prop> Data.Map.null empty == True
 \end{verbatim}
@@
 \begin{alltt}
-prop> \verb|\|x -> null (takeExtension x) == \(\hookleftarrow\)
+-- |
+-- prop> \verb|\|x -> null (takeExtension x) == \(\hookleftarrow\)
             not (hasExtension x)
@@
 \begin{verbatim}
-prop> \xs -> sort xs == (sort . sort) (xs :: [Int])
+-- |
+-- prop> \xs -> sort xs == (sort . sort) (xs :: [Int])
 \end{verbatim}
@@
 \begin{verbatim}
-prop> sort xs == (sort . sort) (xs :: [Int])
+-- |
+-- prop> sort xs == (sort . sort) (xs :: [Int])
 \end{verbatim}

### 2012-06-02 paper.tex
removing comments.
indenting for readablity.
adding a comment.

@@
     TBD: ``Section'' to ``section''
+
+kazu: ``Section X.Y'' is perfect English, i.e. a proper noun.
+      ``The following three sections'' is also perfect, i.e. a ordinal noun.
 \end{quote}
@@
 
-\begin{quote}
-\begin{itemize}
-    \item
-        We already added markup support for interactive examples to
-        Haddock, back in 2010; and Doctest can verify them.
-    \item
-        But it is not possible to use properties in Haddock comments
-        and automatically test them.
-
-    \emph{kazu: I guess that it's enough to describe this in the Implementation Section.}
-\end{itemize}
-\end{quote}
-
 \noindent First, we introduce new Haddock markup for interactive examples.

### 2012-06-02 paper.tex
adding a comment.
Add a reply
adding cucumber book.

@@
 
+\begin{quote}
+kazu: I guess ``Doctest integrated with QuickCheck'' is better.
+\end{quote}
+
 \end{abstract}
@@
 
-\begin{quote}
-    TBD: add ``The Cucumber Book'' to the bibliography as {\tt
-    cucumber-book}.
-\end{quote}
-
-Cucumber~\cite{cucumber-book} is a higher-level approach to BDD for Ruby, and
+Cucumber~\cite{cucumber} is a higher-level approach to BDD for Ruby, and
 it is typically used in combination with RSpec.

### 2012-06-02 paper.tex
revert mis-operation.
adopting a comment.

@@
 behavior driven development with two tools:
-QuickCheck integrated in Doctest for user documentation and Hspec
-with new mnemonic combinators for code documentation.
+Doctest integrated with QuickCheck for user documentation and
+Hspec with new mnemonic combinators for code documentation.
 Our method utilizes specification that can also be used for
@@
 %% Say what follows from your solution.
-
-\begin{quote}
-kazu: I guess ``Doctest integrated with QuickCheck'' is better.
-\end{quote}
 

### 2012-06-02 paper.tex
Remove comment
Terminology.

@@
 
-\begin{quote}
-    TBD: ``Section'' to ``section''
-
-kazu: ``Section X.Y'' is perfect English, i.e. a proper noun.
-      ``The following three sections'' is also perfect, i.e. a ordinal noun.
-\end{quote}
-
 %%%% Describe the problem.
@@
 \item[property]
-    Property means a QuickCheck property as defined in~\cite{quickcheck}.
+    A property is something that is testable with
+    QuickCheck~\cite{quickcheck}.
 
@@
     \item
-        In section~\ref{sec:reusable-specs}, we showed how to reuse
+        In Section~\ref{sec:reusable-specs}, we showed how to reuse
         pieces of specification.

### 2012-06-02 paper.tex
Terminology
Wording

@@
     Expectations are used to define the expected outcome of a code
-    example.  They serve the same purpose as \emph{assertions} in xUnit
-    testing frameworks, but use a different terminology.
+    example.  They serve the same purpose as \emph{assertions} in
+    traditional unit testing frameworks, but use a different
+    terminology.
 
@@
 
-Haddock has markup support for code blocks, and they are widely used
+Haddock has markup support for code blocks, and code blocks are widely used
 to include properties and examples into documentation\footnote{

### 2012-06-02 paper.tex
Grammar
Wording

@@
 {\tt bytestring} or
-{\tt text}}.
+{\tt text}.}.
 Here is an example from the {\tt filepath} package.
@@
 behavior and some evidence (a code example or a property) into a spec.
-Code examples can be arbitrary HUnit assertions.  Properties must be of type {\tt
-Property}; QuickCheck's \texttt{property} function can be used to turn
-arbitrary properties into a \texttt{Property}.
+Code examples can be arbitrary HUnit test cases\footnote{%
+Wrapping the test case with HUnit's {\tt TestCase} constructor is not
+necessary.}.  Properties must be of type {\tt
+Property}.  QuickCheck's \texttt{property} function can be used to turn
+an arbitrary properties into values of type \texttt{Property}.
 

### 2012-06-02 paper.tex
Wording

@@
 \item[expectation]
-    Expectations are used to define the expected outcome of a code
+    Expectations are used to express the expected outcome of a code
     example.  They serve the same purpose as \emph{assertions} in
@@
 A spec can be run with the \texttt{hspec} function.  When a spec is
-run, all evidence is tested, and a report is generated.
-Behavior with evidence that does not hold is marked in the generated
-report as {\tt FAILED}.
+run, all evidence is tested, and a report is generated.  Behavior with
+evidence that does not hold is marked as {\tt FAILED} in the generated
+report.
 

### 2012-06-02 paper.tex
Add a TBD
Grammar
Update a comment

@@
 \label{sec:problem}
+\begin{itemize}
+    \item TBD: use \emph{code documentation} somewhere here.
+\end{itemize}
 
@@
 \begin{quote}
-    TBD: introduce \emph{test driver}
-
-    sol: not sure if it is necessary, @mkl, @kazu?
+    TBD:
+
+    \begin{itemize}
+        \item Do we have to introduce \emph{test driver} here?
+        \item Better section titles for all the sections that discuss
+            this (problem/solution/implementation).
+    \end{itemize}
 \end{quote}
@@
 community and integrate it with QuickCheck.
-Examples and properties in Haddock comments is automatically
+Examples and properties in Haddock comments are automatically
 tested by the Doctest command.

### 2012-06-02 paper.tex
behaviour -> behavior

@@
 \subsection{Applying BDD to Haskell: An Example}
-This section shows how behaviour driven development works in practice,
+This section shows how behavior driven development works in practice,
 by means of the Fibonacci function.  Of course this is a very simple

### 2012-06-02 paper.tex
Add a comment
BCP: Spec and library are in separate files now

@@
 
+\begin{quote}
+\begin{itemize}
+    \item
+        We already added markup support for interactive examples to
+        Haddock, back in 2010; and Doctest can verify them.
+    \item
+        But it is not possible to use properties in Haddock comments
+        and automatically test them.
+
+    \emph{kazu: I guess that it's enough to describe this in the Implementation Section.}
+\end{itemize}
+\end{quote}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 We present a step-by-step refinement of a library
-that contains both implementation and specs.  Each line of code is
+that, and a corresponding spec.  Each line of code is
 only given once when introduced, subsequent code fragments are

### 2012-06-02 paper.tex
Fix a link

@@
 implemented.  They are available form Hackage\footnote{%
-http://hackage.haskell.org/package/hspec-shouldbe},
+http://hackage.haskell.org/package/test-shouldbe},
 and they will be incorporated into a future

### 2012-06-02 paper.tex
Strip some of doctest's output
Step 1 -> Start with a trivial example
Add a comment

@@
 
-\subsubsection{Step 1}
+\subsubsection{Start with a minimal implementation}
 
@@
 $ doctest Fib.hs
-There is one test, with one single interaction.
 Examples: 1  Tried: 1  Errors: 0  Failures: 0
@@
 actually calculate Fibonacci numbers.
+
+\begin{quote}
+    TBD: do we want to add something like this?
+
+    At this stage we do not care about invalid input or performance.
+    We are happy with an implementation that works for small
+    non-negative numbers.
+\end{quote}
 

### 2012-06-02 paper.tex
Alternative section titles
BCP: Adapt Arbitrary instance

@@
 
-\begin{quote}
-    TBD: This section needs more work, to get the wording right.
-\end{quote}
-
-Haddock version 2.7.2 and earlier provided markup syntax for
-code blocks, but did not provide any markup syntax for evidence.
-As we described, we implemented markup support for interactive examples ({\tt >>>}) in 2010.
+Haddock before version 2.8 had markup syntax for
+code blocks, but no markup syntax for evidence.
+As mentioned before, we implemented support for interactive examples ({\tt >>>}) in 2010.
 It is available in Haddock version 2.8 or later.
-We also extended Haddock to support markup for properties ({\tt prop>}) which
-will be available with a future release of Haddock.
+We also added support for markup of properties ({\tt prop>}) which
+will be available in a future release of Haddock.
 Both interactive examples and properties in Haddock comments are
-include into generated documentation.
+included in the generated documentation.
 
@@
 
-\subsubsection{Step 2}
+\subsubsection{Get it right}
 
@@
 
-\subsubsection{Step 3}
+\subsubsection{Get it fast}
 
@@
 fibs = 0 : 1 : zipWith (+) fibs (tail fibs)
+\end{verbatim}
+\end{quote}
+
+\noindent TBD: With a much more efficient implementation at hand, we
+can also adapt our definition of ``small non-negative number'', so
+that a wide range of input values is tested:
+
+\begin{quote}
+\small
+\begin{verbatim}
+instance Arbitrary Small where
+  arbitrary = Small . (`mod` 1000) <$> arbitrary
 \end{verbatim}

### 2012-06-02 paper.tex
Implementation: Haddock: wording
Implementation: Doctest: wording

@@
 To extract evidence from Haddock comments,
-Doctest used the Haddock API in the past.
-Since each release of Haddock supports only the latest version of GHC,
-it is difficult to support old versions of GHC with Haddock API.
-Doctest is now directly using GHC API for portability.
+Doctest initially used the Haddock API.
+Since each release of Haddock only supports the latest version of GHC,
+it was difficult to support old versions of GHC with Doctest.
+Doctest now uses the GHC API directly for portability.
 

### 2012-06-02 paper.tex
Implementation: Testing examples: wording

@@
 
-To verify examples, Doctest uses GHCi.
-Test is done by comparing result string in Haddock notation and
-result string from GHCi.
-Since verification based on string comparison, exceptions can be handled.
+Doctest uses GHCi to verify examples.  Verification is done by
+comparing the expected output given in the Haddock comment with the
+actual output of GHCi.  Since verification is based on string
+comparison, exceptions can be handled.
 
@@
 and share the same scope.
-The scope is cleared with \verb|:reload| before a new
-example group is started.
+Doctest clears the scope with \verb|:reload| before it tests a new
+example group.
 

### 2012-06-02 paper.tex
Work on reusable specs

@@
 Haddock before version 2.8 had markup syntax for
-code blocks, but no markup syntax for evidence.
-As mentioned before, we implemented support for interactive examples ({\tt >>>}) in 2010.
-It is available in Haddock version 2.8 or later.
+code blocks, but not for evidence.
+We implemented support for interactive examples ({\tt >>>}) in
+Haddock 2.8.
 We also added support for markup of properties ({\tt prop>}) which
-will be available in a future release of Haddock.
+will be available in a future release.
 Both interactive examples and properties in Haddock comments are
@@
 
-Reusable spec is an idea to scrap boilerplate of evidence.
-For instance, without reusable spec,
-each developer of an instance of a type class should
-express evidence for its laws.
-Since instances of a type class share API,
-we can avoid such boilerplate.
-The following is an example to implement
-evidence for Monoid laws:
+RSpec provides an explicit mechanism to facilitate reuse of pieces of
+specification.~\cite{rspec}
+In Haskell we get reusability for free, it does not require any
+explicit support from Hspec.
+
+\begin{quote}
+  TBD: not sure, if this adds anything
+
+This not only shows that functional programming can deliver what OO
+has always claimed, but it is also awfully useful.
+\end{quote}
+
+Here is a specification that documents that a list is a monoid:
+
+\begin{quote}
+\small
+\begin{verbatim}
+spec = do
+  describe "List as a Monoid" $ do
+    describe "mempty" $ do
+      it "is a left identity" $ property $ \x ->
+        mempty <> x == x
+
+      it "is a right identity" $ property $ \x ->
+        x <> mempty == x
+
+    describe "<>" $ do
+      it "is associative" $ property $ \x y z ->
+        (x <> y) <> z == x <> (y <> z)
+\end{verbatim}
+\end{quote}
+
+\noindent And it is possible to to make this specification reusable,
+by abstracting over the type:
 
@@
 \noindent This can be used to document and test arbitrary Monoid
-instances.
-Here is an example how to use it for lists:
-
-\begin{quote}
-\small
-\begin{verbatim}
-describe "List as a Monoid" $
+instances, the specification from above would become:
+
+\begin{quote}
+\small
+\begin{verbatim}
+spec = do
+  describe "List as a Monoid" $ do
     shouldSatisfyMonoidLaws (undefined :: [Int])
@@
 \end{quote}
-
-\noindent We have already implemented this experimentally.
-We should ensure the practicality of reusable spec and
-prepare reusable spec to cover variety of type classes.
 
@@
         checkers}\footnote{http://hackage.haskell.org/package/checkers})
+      \item copied here from Section~\ref{sec:reuseable-pecs}:
+        We should ensure the practicality of reusable spec and
+        prepare reusable spec to cover variety of type classes.
+
 \end{itemize}

### 2012-06-02 paper.tex
Commented code on QuickCheck -> /dev/null

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \subsection{QuickCheck}
-
-%% \emph{TBD What does this have to do with the topic of the paper?}
-%% \emph{Acknowledgment is enough.}
-
-%% It is enough to test non-quantified properties just one.
-%% However, QuickCheck 2.3 or earlier tests
-%% non-quantified properties 100 times.
-%% More precisely, the evaluation results are memorized and they are
-%% compared with {\tt True} 100 times.
-%% The current maintainer of QuickCheck modified so that
-%% non-quantified properties
-%% are tested just once in its source repository.
-%% This feature will be available in the future release.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Doctest}
@@
 \noindent All combinators introduced in Section~\ref{sec:code-doc} have been
-implemented.  They are available form Hackage\footnote{%
+implemented.  They are available from Hackage\footnote{%
 http://hackage.haskell.org/package/test-shouldbe},
-and they will be incorporated into a future
+and will be incorporated into a future
 release of Hspec.
@@
 
-A fully functional implementations is available from
+A fully functional implementation is available from
 Hackage\footnote{http://hackage.haskell.org/package/hspec-discover},
-and it will be incorporate into a future release of Hspec.
+and it will be incorporated into a future release of Hspec.
 

### 2012-06-02 paper.tex
Add a comment
typo: form -> from

--- 2012-06-02 paper.tex
@@
 \noindent All combinators introduced in Section~\ref{sec:code-doc} have been
-implemented.  They are available form Hackage\footnote{%
+implemented.  They are available from Hackage\footnote{%
 http://hackage.haskell.org/package/test-shouldbe},

### 2012-06-02 paper.tex
remove repeated 'they'

@@
 http://hackage.haskell.org/package/test-shouldbe},
-and they will be incorporated into a future
+and will be incorporated into a future
 release of Hspec.

### 2012-06-02 paper.tex
Update comment

@@
 \subsubsection{Testing examples}
+
+\begin{quote}
+    TBD: We read both, stdout + stderr from GHCi and compare that with the
+    expected result.  On Windows, only stdout works (IIRC, when an
+    example causes output to stderr on windows, doctest exits with an
+    error, see http://hackage.haskell.org/trac/ghc/ticket/5761).
+
+    sol: relevant?
+\end{quote}
 
@@
 \noindent This can be used to document and test arbitrary Monoid
-instances, the specification from above would become:
+instances.
+The specification from above becomes:
 
@@
       \item copied here from Section~\ref{sec:reuseable-pecs}:
+
+        Reusable spec is an idea to scrap boilerplate of evidence.
+        For instance, without reusable spec,
+        each developer of an instance of a type class should
+        express evidence for its laws.
+        Since instances of a type class share API,
+        we can avoid such boilerplate.
+        The following is an example to implement
+        evidence for Monoid laws:
+
         We should ensure the practicality of reusable spec and

### 2012-06-02 paper.tex
Add more comments

--- 2012-06-02 paper.tex
@@
 
-%% State the problem.
-%% Say why it's an interesting problem.
-%% Say what your solution achieves.
-%% Say what follows from your solution.
+\begin{itemize}
+    \item We show how to reuse pieces of specifications (and/or that
+        specifications are composable?).
+\end{itemize}
+
+\begin{itemize}
+    \item State the problem.
+    \item Say why it's an interesting problem.
+    \item Say what your solution achieves.
+    \item Say what follows from your solution.
+\end{itemize}
 
@@
     \item
+        We think a library of reuseable specs would be a good idea
+    \item
         It would be possible and useful to write a library that
@@
 \begin{quote}
-    TBD: Add "best current practice"
+    \begin{itemize}
+        \item Add "best current practice"
+        \item We showed how to reuse pieces of specifications
+    \end{itemize}
 \end{quote}

### 2012-06-02 paper.tex
Fix code example

@@
       it "is a left identity" $ property $ \x ->
-        mempty <> x == x
+        mempty <> x == (x :: [Int])
 
       it "is a right identity" $ property $ \x ->
-        x <> mempty == x
+        x <> mempty == (x :: [Int])
 
@@
       it "is associative" $ property $ \x y z ->
-        (x <> y) <> z == x <> (y <> z)
+        (x <> y) <> z == (x <> (y <> z) :: [Int])
 \end{verbatim}

### 2012-06-02 paper.tex
Specs -> Spec

@@
 shouldSatisfyMonoidLaws ::
-  (Eq a, Show a, Monoid a, Arbitrary a) => a -> Specs
+  (Eq a, Show a, Monoid a, Arbitrary a) => a -> Spec
 shouldSatisfyMonoidLaws t = do

### 2012-06-02 paper.tex
Semantics: positive -> non-negative
Wording

@@
 So far, so good.  We have an implementaion of the Fibonacci function
-that swiftly calculates correct values for all positive integers.
+that swiftly calculates correct values for all non-negative integers.
 But what about negative ones? Let's assume that we want our
@@
 RSpec provides an explicit mechanism to facilitate reuse of pieces of
-specification.~\cite{rspec}
+specification~\cite{rspec}.
 In Haskell we get reusability for free, it does not require any
 explicit support from Hspec.
+The following example illustrates this.
 
@@
 \end{quote}
-
 Here is a specification that documents that a list is a monoid:
@@
 
-\noindent And it is possible to to make this specification reusable,
-by abstracting over the type:
+\noindent It is possible to to make this specification reusable, by
+abstracting over the type:
 

### 2012-06-02 paper.tex
Fix typo
BCP: adapt code

@@
 \begin{verbatim}
-module Fib where
 -- | Calculate Fibonacci numbers.
@@
 
-main = hspec $ do
+main = hspec spec
+
+spec = do
   describe "fib" $ do
@@
 
-So far, so good.  We have an implementaion of the Fibonacci function
+So far, so good.  We have an implementation of the Fibonacci function
 that swiftly calculates correct values for all non-negative integers.

### 2012-06-02 paper.tex
Update a comment
BCP: break code into two pieces (we still need some "separator" text)

@@
 
-    At this stage we do not care about invalid input or performance.
+    At this stage we do not think about invalid input or performance.
     We are happy with an implementation that works for small
-    non-negative numbers.
+    non-negative integers.
 \end{quote}
@@
   arbitrary = Small . (`mod` 10) <$> arbitrary
-
+\end{verbatim}
+\end{quote}
+
+\noindent TBD: can we put some useful text here, so that the code
+example is divided into two pieces, and hence each piece is smaller?
+
+\begin{quote}
+\small
+\begin{verbatim}
 main = hspec spec

### 2012-06-02 example/Fib.hs
Add final code from BCP

@@ -1,0 +1,11 @@
+module Fib (fib) where
+
+-- | Calculate Fibonacci numbers.
+--
+-- >>> fib 10
+-- 55
+fib :: Int -> Integer
+fib n | n < 0     = 0
+      | otherwise = fibs !! n
+
+fibs = 0 : 1 : zipWith (+) fibs (tail fibs)

### 2012-06-02 example/FibSpec.hs

@@ -1,0 +1,32 @@
+module FibSpec (main, spec) where
+
+import Control.Applicative
+import Control.Exception
+import System.Timeout
+
+import Test.QuickCheck
+import Test.Hspec.ShouldBe
+
+import Fib
+
+-- small non-negative numbers
+newtype Small = Small Int
+  deriving Show
+
+instance Arbitrary Small where
+  arbitrary = Small . (`mod` 1000) <$> arbitrary
+
+main = hspec spec
+
+spec = do
+  describe "fib" $ do
+    it "calculates arbitrary Fibonacci numbers" $ do
+      property $ \(Small n) ->
+        fib n == fib (n + 2) - fib (n + 1)
+
+    it "is efficient" $ do
+      timeout 10000 (evaluate $ fib 32) `shouldReturn` Just 2178309
+
+    it "returns 0 on negative input" $ do
+      property $ \n -> n < 0 ==>
+        fib n == 0

### 2012-06-02 paper.tex
Step 4 -> Dealing with invalid input
BCP: Split last step into two pieces
Add final code for example from BCP to appendix

@@
 it "is efficient" $ do
-  timeout 10000
-      (evaluate $ fib 32) `shouldReturn` Just 2178309
+  timeout 10000 (evaluate $ fib 32)
+    `shouldReturn` Just 2178309
 \end{verbatim}
@@
 
-\subsubsection{Step 4}
+\subsubsection{Handling invalid input}
 
@@
 \noindent Our implementation already fulfils this spec, so we are
-fine. Some time later however, the one person in our team that knows
+fine.
+
+\subsubsection{Dealing with specification changes}
+
+Some time later however, the one person in our team that knows
 more about Haskell than we do, convinces us that throwing exceptions
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\appendix
+\section{TBD}
+
+\begin{quote}
+\small
+\begin{verbatim}
+-- file Fib.hs
+module Fib (fib) where
+
+-- | Calculate Fibonacci numbers.
+--
+-- >>> fib 10
+-- 55
+fib :: Int -> Integer
+fib n | n < 0     = 0
+      | otherwise = fibs !! n
+
+fibs = 0 : 1 : zipWith (+) fibs (tail fibs)
+
+-- file FibSpec.hs
+module FibSpec (main, spec) where
+
+import Control.Applicative
+import Control.Exception
+import System.Timeout
+
+import Test.QuickCheck
+import Test.Hspec.ShouldBe
+
+import Fib
+
+-- small non-negative numbers
+newtype Small = Small Int
+  deriving Show
+
+instance Arbitrary Small where
+  arbitrary = Small . (`mod` 1000) <$> arbitrary
+
+main = hspec spec
+
+spec = do
+  describe "fib" $ do
+    it "calculates arbitrary Fibonacci numbers" $ do
+      property $ \(Small n) ->
+        fib n == fib (n + 2) - fib (n + 1)
+
+    it "is efficient" $ do
+      timeout 10000 (evaluate $ fib 32)
+        `shouldReturn` Just 2178309
+
+    it "returns 0 on negative input" $ do
+      property $ \n -> n < 0 ==>
+        fib n == 0
+\end{verbatim}
+\end{quote}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \acks

### 2012-06-02 paper.tex
test-shouldbe

@@
 implemented.  They are available from Hackage\footnote{%
-http://hackage.haskell.org/package/test-shouldbe},
+http://hackage.haskell.org/package/hspec-shouldbe},
 and will be incorporated into a future

### 2012-06-02 paper.tex
wording.

@@
 \subsection{Applying BDD to Haskell: An Example}
-This section shows how behavior driven development works in practice,
+This subsection shows how behavior driven development works in practice,
 by means of the Fibonacci function.  Of course this is a very simple

### 2012-06-02 paper.tex
More conservative wording

@@
 
-We present a step-by-step refinement of a library
-that, and a corresponding spec.  Each line of code is
-only given once when introduced, subsequent code fragments are
-incremental updates.  The reader is asked to substitute code parts as
-appropriate.
-
-\subsubsection{Start with a minimal implementation}
+We present a step-by-step refinement of a library, and a corresponding
+spec.  Each line of code is only given once when introduced,
+subsequent code fragments are incremental updates.  The reader is
+asked to substitute code parts as appropriate.
+
+\subsubsection{Starting with a minimal implementation}
 
@@
 
-\subsubsection{Get it right}
-
-In the next step, we add a specification that forces our function to
-actually calculate Fibonacci numbers.
-
-\begin{quote}
-    TBD: do we want to add something like this?
-
-    At this stage we do not think about invalid input or performance.
-    We are happy with an implementation that works for small
-    non-negative integers.
-\end{quote}
-
-\begin{quote}
-\small
-\begin{verbatim}
--- small non-negative numbers
+\subsubsection{Getting it right}
+
+Next, we add a specification that forces our function to actually
+calculate Fibonacci numbers.
+At this stage we do not worry about invalid input or performance.  We
+are happy with an implementation that works for small non-negative
+integers.
+
+\begin{quote}
+\small
+\begin{verbatim}
 newtype Small = Small Int
@@
   arbitrary = Small . (`mod` 10) <$> arbitrary
-\end{verbatim}
-\end{quote}
-
-\noindent TBD: can we put some useful text here, so that the code
-example is divided into two pieces, and hence each piece is smaller?
-
-\begin{quote}
-\small
-\begin{verbatim}
+
 main = hspec spec
@@
 
-\noindent Of course this spec fails when we run it,
+\noindent This spec fails when we run it:
 
@@
 
-\noindent so we add a working implementation.  Both Doctest and Hspec
-report success after that.
+\noindent So we add a working implementation:
 
@@
 
-\subsubsection{Get it fast}
-
-As the seasoned Haskeller might have already noticed, our
-implementation is quite inefficient, as made evident by this
+\noindent Both Doctest and Hspec report success now.
+
+\subsubsection{Getting it fast}
+
+Our implementation is quite inefficient, as made evident by this
 specification:
@@
 
-\noindent When we run it, hspec tells us
+\noindent When we run it, hspec tells us:
 
@@
 
-\noindent and we respond with an efficient implementation.
+\noindent And we respond with an efficient implementation:
 
@@
 
-\noindent TBD: With a much more efficient implementation at hand, we
-can also adapt our definition of ``small non-negative number'', so
-that a wide range of input values is tested:
+\noindent With a much more efficient implementation at hand, we can
+now adapt our definition of small non-negative integers, so that
+a wider range of input values is tested:
 
@@
 
-So far, so good.  We have an implementation of the Fibonacci function
-that swiftly calculates correct values for all non-negative integers.
+We have an implementation of the Fibonacci function
+that calculates correct values for all non-negative integers.
 But what about negative ones? Let's assume that we want our
@@
 from pure code is considered poor design. Hence, we change the
-specification so that negative inputs should have zero as result.
+specification so that negative input values should result in zero.
 

### 2012-06-02 paper.tex
marking with TBD.
Comment paragraph about "behavior is imperative"

@@
 \authorinfo{Simon HENGEL}
-           {}
+           {TBD}
            {sol@typeful.net}
@@
 
+TBD:
+
 \begin{itemize}
@@
 
-\noindent The word ``behavior'' might be associated with ``imperative
-action''.  Since pure functions of Haskell are static, this word might
-be improper for Haskell.  However, this word is already used in the
-Haskell community~\cite{typeclassopedia}.  So, we use ``behavior''
-instead of defining terminology in this paper.
+%\noindent The word ``behavior'' might be associated with ``imperative
+%action''.  Since pure functions of Haskell are static, this word might
+%be improper for Haskell.  However, this word is already used in the
+%Haskell community~\cite{typeclassopedia}.  So, we use ``behavior''
+%instead of defining terminology in this paper.
 
@@
 \begin{quote}
+TBD:
 \begin{itemize}
@@
 \begin{quote}
+TBD:
     \begin{itemize}

### 2012-06-02 paper.tex
for submit-1.

--- 2012-06-02 paper.tex
@@
 \authorinfo{Simon HENGEL}
-           {}
+           {Freelance software developer}
            {sol@typeful.net}
@@
 
-\begin{itemize}
-    \item We show how to reuse pieces of specifications (and/or that
-        specifications are composable?).
-\end{itemize}
-
-\begin{itemize}
-    \item State the problem.
-    \item Say why it's an interesting problem.
-    \item Say what your solution achieves.
-    \item Say what follows from your solution.
-\end{itemize}
-
 \end{abstract}
@@
     evidence.
-
-\item[spec]
-    TBD: BDD terminology for test suite.
-\end{description}
-
-\begin{quote}
-    \emph TBD: Check, terminology for consistency, and remove
-    unnecessary terminology (if any).  This has to be done as one of
-    the last steps.
-\end{quote}
-
-\begin{description}
-\item[developer]
-    A programmer who develops a library.
-
-\item[user]
-    A programmer who uses a library.
-
-\item[user documentation]
-    Explanation of functions for users.
-\item[code documentation]
-    Explanation of functions for developers.
-
-\item[non-quantified property]
-    Property which does not take parameters.
-
-\item[quantified property]
-    Property which takes parameters.
 
@@
 \label{sec:problem}
-\begin{itemize}
-    \item TBD: use \emph{code documentation} somewhere here.
-\end{itemize}
 
@@
 \label{sec:problem-discovery}
-
-\begin{quote}
-    TBD:
-
-    \begin{itemize}
-        \item Do we have to introduce \emph{test driver} here?
-        \item Better section titles for all the sections that discuss
-            this (problem/solution/implementation).
-    \end{itemize}
-\end{quote}
 
@@
 \item
-    Automatic discovery for specs is missing (TBD)
+    Automatic discovery for specs is missing.
 \end{itemize}
@@
 
-\begin{quote}
-\begin{itemize}
-    \item
-        We already added markup support for interactive examples to
-        Haddock, back in 2010; and Doctest can verify them.
-    \item
-        But it is not possible to use properties in Haddock comments
-        and automatically test them.
-
-    \emph{kazu: I guess that it's enough to describe this in the Implementation Section.}
-\end{itemize}
-\end{quote}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsubsection{Testing examples}
-
-\begin{quote}
-    TBD: We read both, stdout + stderr from GHCi and compare that with the
-    expected result.  On Windows, only stdout works (IIRC, when an
-    example causes output to stderr on windows, doctest exits with an
-    error, see http://hackage.haskell.org/trac/ghc/ticket/5761).
-
-    sol: relevant?
-\end{quote}
 
@@
 
-\begin{quote}
-    TBD: The implementation of \verb|shouldThrow| may still be
-    interesting.  We can describe it, if we have time left.
-\end{quote}
-
-\begin{quote}
-\small
-\begin{verbatim}
-shouldThrow ::
-  Exception e => IO a -> Selector e -> Assertion
-\end{verbatim}
-\end{quote}
+%% \begin{quote}
+%%     TBD: The implementation of \verb|shouldThrow| may still be
+%%     interesting.  We can describe it, if we have time left.
+%% \end{quote}
+
+%% \begin{quote}
+%% \small
+%% \begin{verbatim}
+%% shouldThrow ::
+%%   Exception e => IO a -> Selector e -> Assertion
+%% \end{verbatim}
+%% \end{quote}
 
@@
 
-\begin{quote}
-    TBD: in this section we:
-    \begin{itemize}
-        \item We show how to do BDD in Haskell
-        \item How to reuse pieces of specification
-        \item How to use Doctest and Hspec with Cabal's test feature
-    \end{itemize}
-\end{quote}
-
 \subsection{Applying BDD to Haskell: An Example}
@@
 \begin{quote}
-  TBD: not sure, if this adds anything
-
-This not only shows that functional programming can deliver what OO
-has always claimed, but it is also awfully useful.
-\end{quote}
-Here is a specification that documents that a list is a monoid:
-
-\begin{quote}
 \small
@@
 
-\begin{quote}
-    TBD
-
-    sol: I think it is not critical.  I'll try to resolve this
-    tomorrow, but I think we can always remove it.
-
-\begin{itemize}
-
-\item It is important that testing is automated, including running of test
-    suites.  Where should we put this?
-
-\item automatically sounds strange, better use automated somehow
-\end{itemize}
-\end{quote}
-
-\noindent Cabal's test feature can be used to automatically run tests~\cite{cabal}.  A cabal
+Cabal's test feature can be used to automatically run tests~\cite{cabal}.  A cabal
 file can contain multiple test suites, so it is possible to define test suites
@@
 \section{Open problems and future work}
-
-\begin{quote}
-    TBD: state that ``A library for reusable pieces of specification would be
-    useful.''
-\begin{itemize}
-    \item
-        In Section~\ref{sec:reusable-specs}, we showed how to reuse
-        pieces of specification.
-    \item
-        We think a library of reuseable specs would be a good idea
-    \item
-        It would be possible and useful to write a library that
-        provides reusable pieces of specification, for many of
-        Haskell's standard classes.
-        (similar in purpose to
-        {\tt
-        checkers}\footnote{http://hackage.haskell.org/package/checkers})
-      \item copied here from Section~\ref{sec:reuseable-pecs}:
-
-        Reusable spec is an idea to scrap boilerplate of evidence.
-        For instance, without reusable spec,
-        each developer of an instance of a type class should
-        express evidence for its laws.
-        Since instances of a type class share API,
-        we can avoid such boilerplate.
-        The following is an example to implement
-        evidence for Monoid laws:
-
-        We should ensure the practicality of reusable spec and
-        prepare reusable spec to cover variety of type classes.
-
-\end{itemize}
-\end{quote}
 
@@
 http://repetae.net/computer/jhc/jhc.shtml} would allow to solve this.
-
-\begin{quote}
-    sol: should we just remove the next paragraph?
-\end{quote}
-
-Feedback for failing QuickCheck properties is not always as useful as
-it could be.
-Combinators, similar to {\tt shouldBe}, that construct QuickCheck {\tt
-Result}s could improve the situation.
 
@@
 \section{Conclusion}
-
-\begin{quote}
-    \begin{itemize}
-        \item Add "best current practice"
-        \item We showed how to reuse pieces of specifications
-    \end{itemize}
-\end{quote}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\appendix
-\section{TBD}
-
-\begin{quote}
-\small
-\begin{verbatim}
--- file Fib.hs
-module Fib (fib) where
-
--- | Calculate Fibonacci numbers.
---
--- >>> fib 10
--- 55
-fib :: Int -> Integer
-fib n | n < 0     = 0
-      | otherwise = fibs !! n
-
-fibs = 0 : 1 : zipWith (+) fibs (tail fibs)
-
--- file FibSpec.hs
-module FibSpec (main, spec) where
-
-import Control.Applicative
-import Control.Exception
-import System.Timeout
-
-import Test.QuickCheck
-import Test.Hspec.ShouldBe
-
-import Fib
-
--- small non-negative numbers
-newtype Small = Small Int
-  deriving Show
-
-instance Arbitrary Small where
-  arbitrary = Small . (`mod` 1000) <$> arbitrary
-
-main = hspec spec
-
-spec = do
-  describe "fib" $ do
-    it "calculates arbitrary Fibonacci numbers" $ do
-      property $ \(Small n) ->
-        fib n == fib (n + 2) - fib (n + 1)
-
-    it "is efficient" $ do
-      timeout 10000 (evaluate $ fib 32)
-        `shouldReturn` Just 2178309
-
-    it "returns 0 on negative input" $ do
-      property $ \n -> n < 0 ==>
-        fib n == 0
-\end{verbatim}
-\end{quote}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \acks

### 2012-06-02 paper.tex
Doctest is the solution, so we do not mention it in "Problems" anymore.
simplify

--- 2012-06-02 paper.tex
@@
 
-\subsection{Doctest and Haddock}
+\subsection{Haddock}
 \label{sec:doctest-haddock}
@@
 
-So, we need common syntax to describe examples and properties
-independently in Haddock. Doctest should automatically verify
-examples and test properties.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Hspec provides an EDSL for defining specs.  A spec organizes evidence
-in a tree structure (or a forest, to be accurate), and it is defined
+in a tree structure (or a forest, to be accurate), and is defined
 with \texttt{it} and \texttt{describe}.

### 2012-06-02 paper.tex
Wording + grammar

@@
 
-Hspec provides an EDSL for defining specs.  A spec organizes evidence
+Hspec provides an EDSL for defining specs.  A \emph{spec} organizes evidence
 in a tree structure (or a forest, to be accurate), and is defined
 with \texttt{it} and \texttt{describe}.
+
+Here is an example that specifies some behavior of the {\tt reverse}
+function.
 
@@
 Property}.  QuickCheck's \texttt{property} function can be used to turn
-an arbitrary properties into values of type \texttt{Property}.
+arbitrary properties into values of type \texttt{Property}.
 

### 2012-06-02 paper.tex
Wording + syntax

@@
 Wrapping the test case with HUnit's {\tt TestCase} constructor is not
-necessary.}.  Properties must be of type {\tt
-Property}.  QuickCheck's \texttt{property} function can be used to turn
-arbitrary properties into values of type \texttt{Property}.
+necessary.}.  If a property is used as evidence, it must be wrapped
+with QuickCheck's {\tt property} function.
 
@@
 run, all evidence is tested, and a report is generated.  Behavior with
-evidence that does not hold is marked as {\tt FAILED} in the generated
+evidence that does not hold is marked as {\tt FAILED} in the
 report.
@@
 % \subsubsection{Hspec does not provide a mechanism for expectations}
-BDD frameworks use \emph{should} or \emph{must} instead of
-\emph{assert} to express expectations.  Here is an example that uses
+BDD frameworks use ``should'' or ``must'' instead of
+``assert'' to express expectations.  Here is an example that uses
 RSpec.

### 2012-06-02 paper.tex
Wording + better explanations

@@
 % \subsubsection{Hspec provides no mechanism to expect exceptions}
-Moreover RSpec provides a mechanism to expect exceptions.
+Moreover RSpec provides a mechanism to expect exceptions.  It is used
+like this:
 
@@
 
-\noindent This is neither challenging nor interesting, so it should be
-automated.  But Hspec does not provide a solution for that.
+\noindent This is error prone, and neither challenging nor
+interesting.  So it should be automated.  But Hspec does not provide a
+solution for generating such \emph{test drivers}.
 
@@
 
-\subsection{hspec-discover}
+\subsection{Automatic discovery of specs}
 \label{sec:implementation-discovery}

### 2012-06-02 paper.tex
Wording of "Problem summary"

@@
 
-Here is a summary of the problem:
+Here is a summary of the problems:
 
@@
 \item
-    For user documentation,
-    Haddock does not have markup to describe evidence.
+    There is no way to automatically test examples and properties
+    contained in user documentation.
 \item
     For code documentation,
-    Hspec does not provide combinators to express expectations with
+    Hspec does not provide combinators to express expectations in
     BDD terminology.
 \item
-    Automatic discovery for specs is missing (TBD)
+    There is no facility that automatically finds and combines all
+    specs of a project.
 \end{itemize}

### 2012-06-02 paper.tex
Wording of "Solution"

@@
 The following subsections explain two extensions to Haddock's markup.
-And combinators for expectations, that can be used with Hspec, are
-introduced.
+Subsequently, combinators for expectations, that can be used with
+Hspec, are introduced.
 
@@
 \item Add markup syntax for properties to Haddock
-\item Bring Doctest from the Python community to the Haskell
-community and integrate it with QuickCheck.
-Examples and properties in Haddock comments are automatically
-tested by the Doctest command.
-Its implementation will be discussed in Section~\ref{sec:doctest}.
+\item
+    Implement a tool, similar to Python's Doctest, that can verify
+    examples and test properties in Haddock comments.
+    The implementation is discussed in Section~\ref{sec:doctest}.
 

### 2012-06-02 paper.tex
Polished "Markup for examples"

@@
 
-\noindent First, we introduce new Haddock markup for interactive examples.
-In this paper, we use {\tt >>>} which is the same as Python's one.
+We introduce new Haddock markup for interactive examples.  A line
+starting with a \emph{prompt} (``{\tt >>>}'') denotes an expression.
+Lines following an expression denote the result of that expression.
 Here is an example:
@@
 
-\noindent The first line starts with {\tt >>>} and an expression follows. The second line contains its result. We can write this by copy-and-pasting from an interpreter and modifying the prompt.
-It's worth explaining that we can specify operation sequences:
+\noindent The result ends at the next empty line, or another prompt.
+It is possible to specify example sequences:
 
@@
 
-\noindent Also, we can show exceptions.
+\noindent Result is defined by what GHCi prints to stdout and stderr
+when evaluating the expression.  This, e.g. also allows to document
+exceptions:
 

### 2012-06-02 paper.tex
Polish "Markup for properties"

@@
 
-\noindent To specify properties, we also propose new markup. In this paper, we
-use the {\tt prop>} prefix. We can write non-quantified properties as follows:
+\noindent To specify properties, we also introduce new markup. A line
+starting with ``{\tt prop>}'' indicates a property.
 
@@
 
-\noindent And it is possible to quantify properties by using a
+\noindent It is possible to quantify properties by using a
 lambda expression. The property from Section~\ref{sec:doctest-haddock}
@@
 
-\noindent Note that the $\hookleftarrow$ symbol denotes line
-continuation.
-If a default type is not suitable, expression type-signatures can be
-specified for each parameter:
+\noindent The $\hookleftarrow$ symbol denotes line continuation.
+
+GHCi's type defaulting~\cite{ghc-manual} applies to properties.  If a default type is
+not suitable explicit type-signatures can be used:
+
+\begin{quote}
+    TBD: @kazu, can you add this to the bib:
+    \url{http://www.haskell.org/ghc/docs/7.4.1/html/users_guide/interactive-evaluation.html}
+\end{quote}
 
@@
 
-\noindent Since the right hand side of {\tt ->} is verbose as documentation,
-lambda terms without the lambda prefix are also allowed:
+\noindent Since the lambda abstraction adds nothing to the
+documentation, it can be omitted:
 
@@
 \noindent We call such properties \emph{implicitly quantified}.  The
-given property is implicitly quantified over \verb|xs|.
+given property is implicitly quantified over {\tt xs}.
 

### 2012-06-02 paper.tex
an other -> another

@@
 
-\noindent A common expectation is, that a value should be equal to an other
+\noindent A common expectation is, that a value should be equal to another
 value:
@@
 
-\noindent This is just an other name for HUnit's \texttt{@?=}
+\noindent This is just another name for HUnit's \texttt{@?=}
 operator.  The type is:

### 2012-06-02 paper.tex
Polish "expecting exceptions"

@@
 
-\noindent Or we may expect that a value should satisfies some
+\noindent Or we may expect that a value should satisfy some
 predicate.
@@
 
-HUnit does not provide any versatile mechanism to state that an action
+HUnit does not provide a convenient mechanism to state that an action
 should throw an exception.  So we propose a combinator for that.
@@
 servers two purposes.  It selects the type of the expected exception,
-and constraints the value.  Here is an example:
+and constrains the value.  Here is an example:
 
@@
 
-\noindent And it can be used with predicates from {\tt
-System.IO.Error}, to require specific {\tt IOException}, e.g.  with
-{\tt isPermissionError}:
-
-\begin{quote}
-\small
-\begin{verbatim}
-launchMissiles `shouldThrow` isPermissionError
-\end{verbatim}
-\end{quote}
+%\noindent And it can be used with predicates from {\tt
+%System.IO.Error}, to require specific {\tt IOException}, e.g.  with
+%{\tt isPermissionError}:
+%
+%\begin{quote}
+%\small
+%\begin{verbatim}
+%launchMissiles `shouldThrow` isPermissionError
+%\end{verbatim}
+%\end{quote}
 
@@
 
-\noindent Which can be used like so:
+\noindent Which can be used like this:
 

### 2012-06-02 paper.tex
Remove comment
Simplify Implementation -> Haddock

@@
 \section{Implementation}
-
-\begin{quote}
-TBD:
-\begin{itemize}
-    \item
-        We already added markup support for interactive examples to
-        Haddock, back in 2010; and Doctest can verify them.
-    \item
-        But it is not possible to use properties in Haddock comments
-        and automatically test them.
-
-    \emph{kazu: I guess that it's enough to describe this in the Implementation Section.}
-\end{itemize}
-\end{quote}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Haddock before version 2.8 had markup syntax for
-code blocks, but not for evidence.
-We implemented support for interactive examples ({\tt >>>}) in
-Haddock 2.8.
-We also added support for markup of properties ({\tt prop>}) which
-will be available in a future release.
-Both interactive examples and properties in Haddock comments are
-included in the generated documentation.
+We added markup support for interactive examples to Haddock, and it is
+available since version 2.8.
+We also added markup support for properties, which will be included in
+a future release of Haddock.
 

### 2012-06-02 paper.tex
Polish implementation of Doctest

@@
 
-To extract evidence from Haddock comments,
-Doctest initially used the Haddock API.
-Since each release of Haddock only supports the latest version of GHC,
-it was difficult to support old versions of GHC with Doctest.
-Doctest now uses the GHC API directly for portability.
+Doctest uses the GHC API to extract properties and examples from
+Haddock comments, and uses GHCi to test them.
+Before testing, Doctest brings the module that contains the Haddock
+comment from which the evidence was extracted into scope.
 
@@
 
-\begin{quote}
-    TBD: We read both, stdout + stderr from GHCi and compare that with the
-    expected result.  On Windows, only stdout works (IIRC, when an
-    example causes output to stderr on windows, doctest exits with an
-    error, see http://hackage.haskell.org/trac/ghc/ticket/5761).
-
-    sol: relevant?
-\end{quote}
-
-Doctest uses GHCi to verify examples.  Verification is done by
+Verification of examples is done by
 comparing the expected output given in the Haddock comment with the
-actual output of GHCi.  Since verification is based on string
-comparison, exceptions can be handled.
+actual output of GHCi.
 
@@
 
-\noindent QuickCheck and GHCi are used to test properties.  Before we
-test a property, we bring the module that contains the Haddock comment
-from which we extracted the property into scope.
-
+QuickCheck is used to test properties.
 Implicitly quantified properties need to be closed before they can be
@@
     \item
-        Find all free variables of the property
-    \item
-        Try to lookup each free variable in the module context
+        Find all free variables of the property.
+    \item
+        Try to lookup each free variable in the module context.
     \item
         Close the property over all free variables, for which lookup
-        failed
+        failed.
 \end{enumerate}
@@
 
-The implementation is fully functional, and it will be included in the
+\noindent The implementation is fully functional, and it will be included in the
 next release of Doctest.

### 2012-06-02 paper.tex
remove TBD impl of wouldThrow
BCP: section description

@@
 
-\begin{quote}
-    TBD: The implementation of \verb|shouldThrow| may still be
-    interesting.  We can describe it, if we have time left.
-\end{quote}
-
-\begin{quote}
-\small
-\begin{verbatim}
-shouldThrow ::
-  Exception e => IO a -> Selector e -> Assertion
-\end{verbatim}
-\end{quote}
-
 \subsection{Automatic discovery of specs}
@@
 
-\begin{quote}
-    TBD: in this section we:
-    \begin{itemize}
-        \item We show how to do BDD in Haskell
-        \item How to reuse pieces of specification
-        \item How to use Doctest and Hspec with Cabal's test feature
-    \end{itemize}
-\end{quote}
+In this section we show how to do BDD in Haskell, how to reuse pieces
+of specification, and how to use Doctest and Hspec in combination with
+Cabal's test feature.
 

### 2012-06-02 paper.tex
BCP: use section instead of subsection

@@
 \subsection{Applying BDD to Haskell: An Example}
-This subsection shows how behavior driven development works in practice,
+This section shows how behavior driven development works in practice,
 by means of the Fibonacci function.  Of course this is a very simple

### 2012-06-02 paper.tex
BCP: be more conservative

@@
 
-Some time later however, the one person in our team that knows
-more about Haskell than we do, convinces us that throwing exceptions
-from pure code is considered poor design. Hence, we change the
-specification so that negative input values should result in zero.
+At some point, we realize that throwing an exception on invalid input
+is not suitable for our use-case.  Hence, we change the specification,
+so that negative input values should result in zero.
 

### 2012-06-02 paper.tex
BCP: remove 'in fact

@@
 spec.  Each line of code is only given once when introduced,
-subsequent code fragments are incremental updates.  The reader is
-asked to substitute code parts as appropriate.
+subsequent code fragments are incremental updates.
 
@@
 
-\noindent In fact, it should be the case that fib returns 0 for all
-negative numbers.
+\noindent It should be the case that fib returns 0 for all negative
+numbers, so we state that in the specification:
 
@@
 
-This concludes our example coding session.  We have seen how to utilize
+This concludes our example.  We have seen how to utilize
 behavior driven development to incrementally build a small library.

### 2012-06-02 paper.tex
Reusable spec: fix quotation and tt

@@
 The following example illustrates this.
-
-\begin{quote}
-  TBD: not sure, if this adds anything
-
-This not only shows that functional programming can deliver what OO
-has always claimed, but it is also awfully useful.
-\end{quote}
 Here is a specification that documents that a list is a monoid:
@@
 \noindent Doctest provides both, an executable and a library.  The library
-exposes a function \verb|doctest| of type:
+exposes a function ``{\tt doctest}'' of type:
 
@@
 
-\noindent Doctest's own `main` is simply:
+\noindent Doctest's own {\tt main} is simply:
 \begin{quote}
@@
 for a project, by passing all command-line arguments that are required
-for that project to `doctest`.  A simple example looks like this:
+for that project to {\tt doctest}.  A simple example looks like this:
 

### 2012-06-02 paper.tex
Open problems: review

@@
 
-\begin{quote}
-    TBD: state that ``A library for reusable pieces of specification would be
-    useful.''
-\begin{itemize}
-    \item
-        In Section~\ref{sec:reusable-specs}, we showed how to reuse
-        pieces of specification.
-    \item
-        We think a library of reuseable specs would be a good idea
-    \item
-        It would be possible and useful to write a library that
-        provides reusable pieces of specification, for many of
-        Haskell's standard classes.
-        (similar in purpose to
-        {\tt
-        checkers}\footnote{http://hackage.haskell.org/package/checkers})
-      \item copied here from Section~\ref{sec:reuseable-pecs}:
-
-        Reusable spec is an idea to scrap boilerplate of evidence.
-        For instance, without reusable spec,
-        each developer of an instance of a type class should
-        express evidence for its laws.
-        Since instances of a type class share API,
-        we can avoid such boilerplate.
-        The following is an example to implement
-        evidence for Monoid laws:
-
-        We should ensure the practicality of reusable spec and
-        prepare reusable spec to cover variety of type classes.
-
-\end{itemize}
-\end{quote}
-
-When Hspec finds evidence that does not hold, it reports the
-corresponding behavior to the developer.  It would be useful to
-include the source file and location where the failing evidence is
-defined, so that the developer can find it immediately.
-Extending GHC with a mechanism similar to JHC's
-\verb|SRCLOC_ANNOTATE| pragma\footnote{%
-http://repetae.net/computer/jhc/jhc.shtml} would allow to solve this.
-
-\begin{quote}
-    sol: should we just remove the next paragraph?
-\end{quote}
-
-Feedback for failing QuickCheck properties is not always as useful as
-it could be.
-Combinators, similar to {\tt shouldBe}, that construct QuickCheck {\tt
-Result}s could improve the situation.
+In Section~\ref{sec:reusable-specs}, we showed how to reuse pieces of
+specification.  It would be possible and useful to write a library
+that provides reusable pieces of specification, for many of Haskell's
+standard classes (similar in purpose to {\tt
+checkers}\footnote{http://hackage.haskell.org/package/checkers}).
+This would provide a convenient mechanism for developers to document
+and test that their instance definitions are well-behaved.
+
+When Hspec finds failing evidence, it reports the corresponding
+behavior to the developer.  It would be useful to include source
+locations for failing evidence, so that the programmer can find the
+corresponding spec immediately. Solving this with GHC requires
+Template Haskell.  Extending GHC with a mechanism
+similar to JHC's \verb|SRCLOC_ANNOTATE| pragma\footnote{%
+http://repetae.net/computer/jhc/jhc.shtml} would allow to solve this
+without Template Haskell.
 
@@
 it would be interesting to explore, whether it is possible to apply
-Cucumber's approach to BDD to Haskell.
+Cucumber's approach to Haskell.
 

### 2012-06-02 paper.tex
Related work: test-framework
We conclude

@@
 Hspec with new mnemonic combinators for code documentation.
-Our method utilizes specification that can also be used for
-documentation and automated testing. We hope that this three-in-one
-quality motivates Haskellers to write more test code.
+Our approach can be used for design, documentation and automated
+testing.  We hope that this motivates Haskell programmers to write
+more and better tests.
 
@@
 %%%% Describe the problem.
+\begin{itemize}
+    \item ref to appendix
+\end{itemize}
 
@@
 \noindent Our approach can be used for design, documentation and
-automated testing.  We hope that this motivates Haskellers to write
+automated testing.  We hope that this motivates Haskell programmers to write
 more and better tests.
@@
 \subsection{Applying BDD to Haskell: An Example}
+\label{sec:bcp-example}
 This section shows how behavior driven development works in practice,
@@
 
-The {\tt test-framework} family is a TDD tool, which currently
-integrates HUnit, QuickCheck, SmallCheck, and Doctest.
+{\tt test-framework} is a library that allows TDD in Haskell, and it
+has support for HUnit, QuickCheck and SmallCheck.
 There is support for automatic test discovery that relies on naming
@@
 
-\begin{quote}
-TBD:
-    \begin{itemize}
-        \item Add "best current practice"
-        \item We showed how to reuse pieces of specifications
-    \end{itemize}
-\end{quote}
-
-This paper discussed how to apply behavior driven development to
+We showed how to apply behavior driven development to
 Haskell.  We proposed to use Doctest for user documentation and Hspec
-for code documentation. We added two markups for
-properties and examples to Haddock and implemented
-Doctest to verify examples and test properties from Haddock comments.
-We also extended a set of combinators for expectations for Hspec.
-With these two tools, specification can also be
-used for documentation and automatic testing.  We also described a
-mechanis that can automatically discover all specs of a project.  We
-think that our approach enhances the value of test code, and hope
-that three-in-one aspect will motivate Haskell programmers to write
-more and better test code.
+for code documentation. We extended Haddock's markup with support for
+examples and properties.  And we implemented Doctest, which verifies
+examples and tests properties in Haddock comments.
+We added two missing features to Hspec: Combinators to express
+expectations, and a mechanism for automatic discovery of specs.
+And we showed how to reuse pieces of specification.
+Our approach can be used for design, documentation and automated
+testing.  We hope that this motivates Haskell programmers to write
+more and better tests.
 
@@
 \appendix
-\section{TBD}
+\section{The final code from Section~\ref{sec:bcp-example}}
+\label{sec:appendix-code}
 

### 2012-06-02 paper.tex
Apply consistency regarding how to format Python's doctest
Add missing comma

@@
 \item
-    Implement a tool, similar to Python's Doctest, that can verify
-    examples and test properties in Haddock comments.
+    Implement a tool, similar to Python's {\tt doctest} module, that
+    can verify examples and test properties in Haddock comments.
     The implementation is discussed in Section~\ref{sec:doctest}.
@@
 GHCi's type defaulting~\cite{ghc-manual} applies to properties.  If a default type is
-not suitable explicit type-signatures can be used:
+not suitable, explicit type-signatures can be used:
 

### 2012-06-02 paper.tex
Remove a comment
Make abstract more focused

@@
 \begin{abstract}
-
-When wading through the numerous libraries written by
-the Haskell community, one sometimes gets the impression that the type
-system is being too much relied on, at the expense of testing.
-While the type system cannot always guarantee semantic correctness,
-test code has more applications than getting the code correct.
+We observed that several widely used Haskell libraries do not have
+tests.
 Test code assists developers in maintaining correctness when extending
@@
 
-\begin{quote}
-    TBD
-
-    sol: I think it is not critical.  I'll try to resolve this
-    tomorrow, but I think we can always remove it.
-
-\begin{itemize}
-
-\item It is important that testing is automated, including running of test
-    suites.  Where should we put this?
-
-\item automatically sounds strange, better use automated somehow
-\end{itemize}
-\end{quote}
-
 \noindent Cabal's test feature can be used to automatically run tests~\cite{cabal}.  A cabal

### 2012-06-02 paper.tex
Even more focused Abstract

@@
 behavior driven development with two tools:
-Doctest integrated with QuickCheck for user documentation and
-Hspec with new mnemonic combinators for code documentation.
+Doctest for user documentation and
+Hspec for code documentation.
 Our approach can be used for design, documentation and automated

### 2012-06-02 paper.tex
Remove TBD
Terminology: explain /spec/

@@
 more and better tests.
-
-TBD:
-
-\begin{itemize}
-    \item We show how to reuse pieces of specifications (and/or that
-        specifications are composable?).
-\end{itemize}
-
-\begin{itemize}
-    \item State the problem.
-    \item Say why it's an interesting problem.
-    \item Say what your solution achieves.
-    \item Say what follows from your solution.
-\end{itemize}
-
 \end{abstract}
@@
 
-The definitions of words used in this paper is as follows:
+We use the following terminology:
 
@@
 \item[spec]
-    TBD: BDD terminology for test suite.
+    A spec is similar to a test suite in TDD.  But it is written with
+    specification terminology, and produces a specification document
+    when run.
 \end{description}

### 2012-06-02 paper.tex
Finalize terminology

@@
 For example, it does not ensure that division by zero does not occur.
-An other example is, that instances of {\tt Monoid} \emph{should} be
+Another example is, that instances of {\tt Monoid} should be
 monoids in the mathematical sense, but the type system does not
 enforce this.  It is the programmers responsibility to ensure that
-instance definitions \emph{behave} as \emph{expected}.  Hence
+instance definitions behave as expected.  Hence
 automated testing is useful.
@@
 document how code is supposed to be used, and tests document the
-behavior of code.
+\emph{behavior} of code.
 
@@
 Hspec is thus useful for \emph{code documentation}, but we observed that parts
-of a specification can be useful as user documentation. To that end,
+of a specification can be useful as \emph{user documentation}. To that end,
 Python's {\tt doctest} module~\cite{doctest} is of interest for us.
@@
 \item
-    We introduce a set of combinators to express expectations about
+    We introduce a set of combinators to express \emph{expectations} about
     code examples for use with Hspec
@@
     when run.
-\end{description}
-
-\begin{quote}
-    \emph TBD: Check, terminology for consistency, and remove
-    unnecessary terminology (if any).  This has to be done as one of
-    the last steps.
-\end{quote}
-
-\begin{description}
-\item[developer]
-    A programmer who develops a library.
-
-\item[user]
-    A programmer who uses a library.
 
 \item[user documentation]
-    Explanation of functions for users.
+    User documentation means the documentation for users of a library.
+
 \item[code documentation]
-    Explanation of functions for developers.
-
-\item[non-quantified property]
-    Property which does not take parameters.
-
-\item[quantified property]
-    Property which takes parameters.
+    Code documentation means a detailed description, or specification,
+    of a library.
 
@@
 \noindent Hspec does not yet provide a mechanism to express
-\emph{expectations} with that terminology.  The developer has to use
+expectations with that terminology.  The developer has to use
 HUnit assertions instead.
@@
 
-We are grateful to David Waern, Haddock's maintainer, for reviewing and
-accepting our patches to Haddock.
-We would like to thank
-Nick Smallbone for having improved QuickCheck so that
-non-quantified properties are tested just once.
+We are grateful to David Waern, Haddock's maintainer, for reviewing
+and accepting our patches to Haddock.
+%We would like to thank
+%Nick Smallbone for having improved QuickCheck so that
+%non-quantified properties are tested just once.
 

### 2012-06-02 paper.tex
Add ref to appendix

@@
 %%%% Describe the problem.
-\begin{itemize}
-    \item ref to appendix
-\end{itemize}
-
 Haskell~\cite{haskell} has a strong type system, which ensures that most type errors are
@@
 This section shows how behavior driven development works in practice,
-by means of the Fibonacci function.  Of course this is a very simple
+by means of the Fibonacci function.  This is a very simple
 example, but the method is also applicable to more sophisticated
@@
 spec.  Each line of code is only given once when introduced,
-subsequent code fragments are incremental updates.
+subsequent code fragments are incremental updates.  Appendix
+\ref{sec:appendix-code} contains the complete code of this section.
 

### 2012-06-02 paper.tex
Add preprintfooter
Remove affiliation
Remove a TBD
addint footnote instead of reference.

@@
 \titlebanner{A draft paper to be submitted to Haskell Symposium 2012}
-\preprintfooter{TBD: short description of paper}
+\preprintfooter{BDD in Haskell}
 
@@
 \authorinfo{Simon HENGEL}
-           {TBD}
+           {}
            {sol@typeful.net}
@@
 
-GHCi's type defaulting~\cite{ghc-manual} applies to properties.  If a default type is
+GHCi's type defaulting~\footnote{%
+http://www.haskell.org/ghc/docs/7.4.1/html/users\_guide/interactive-evaluation.html} applies to properties.  If a default type is
 not suitable, explicit type-signatures can be used:
-
-\begin{quote}
-    TBD: @kazu, can you add this to the bib:
-    \url{http://www.haskell.org/ghc/docs/7.4.1/html/users_guide/interactive-evaluation.html}
-\end{quote}
 

### 2012-06-02 paper.tex
Add section text to "Implementation"

@@
 \item[property]
-    A property is something that is testable with
+    A property is something testable with
     QuickCheck~\cite{quickcheck}.
@@
 \section{Implementation}
+
+In this section we describe implementation aspects of the proposed
+solution.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 examples and tests properties in Haddock comments.
-We added two missing features to Hspec: Combinators to express
+We added two missing features to Hspec: combinators to express
 expectations, and a mechanism for automatic discovery of specs.

### 2012-06-02 paper.tex
Shorten BCD, resolve a TBD

@@
 
-GHCi's type defaulting~\footnote{%
-http://www.haskell.org/ghc/docs/7.4.1/html/users\_guide/interactive-evaluation.html} applies to properties.  If a default type is
+GHCi's type defaulting\footnote{%
+\url{http://www.haskell.org/ghc/docs/7.4.1/html/users_guide/interactive-evaluation.html#extended-default-rules}
+} applies to properties.  If a default type is
 not suitable, explicit type-signatures can be used:
@@
 by means of the Fibonacci function.  This is a very simple
-example, but the method is also applicable to more sophisticated
+example, but the method is also applicable to more complicated
 situations where design decisions might not be as obvious as here.
@@
 \noindent This first step might appear excessively trivial, but as of now we
-already have made some choices with regard to the API of our library.
+already have made some choices with regard to the API of the library.
 By giving a usage example, we were forced to think about which
-arguments our function takes.  We can now run Doctest to verify that
+arguments the function takes.  We can now run Doctest to verify that
 the example works.
@@
 it "returns 0 on negative input" $ do
-  fib (-10) `shouldBe` 0
+  property $ \n -> n < 0 ==>
+    fib n == 0
 \end{verbatim}
@@
 
-\noindent It should be the case that fib returns 0 for all negative
-numbers, so we state that in the specification:
-
-\begin{quote}
-\small
-\begin{verbatim}
-it "returns 0 on negative input" $ do
-  property $ \n -> n < 0 ==>
-    fib n == 0
-\end{verbatim}
-\end{quote}
-
-\noindent Again, our implementation already fulfils that spec.
-
-This concludes our example.  We have seen how to utilize
+\noindent This concludes our example.  We have seen how to utilize
 behavior driven development to incrementally build a small library.

### 2012-06-03 paper.tex
Minor change to abstract
wording.
fixing overflow.
Polish contributions

@@
 \begin{abstract}
-We observed that several widely used Haskell libraries do not have
-tests.
-Test code assists developers in maintaining correctness when extending
-software, and shows users how an API is supposed to be used.
+We observe that several widely used Haskell libraries do not have
+tests.  Test code assists developers in maintaining correctness when
+extending software, and shows users how an API is supposed to be used.
 To ease the production and enhance the value of test code, we advocate
-behavior driven development with two tools:
-Doctest for user documentation and
-Hspec for code documentation.
-Our approach can be used for design, documentation and automated
-testing.  We hope that this motivates Haskell programmers to write
-more and better tests.
+behavior driven development with two tools: Doctest for user
+documentation and Hspec for code documentation.  Our approach can be
+used for design, documentation and automated testing.  We hope that
+this motivates Haskell programmers to write more and better tests.
 \end{abstract}
@@
     extended Haddock's markup to support examples, back in 2010
-    (Section~\ref{sec:haddock} and Section~\ref{sec:doctest})
+    (Section~\ref{sec:haddock} and Section~\ref{sec:doctest}).
 \item
@@
     Haddock's markup to support properties
-    (Section~\ref{sec:user-doc} and Section~\ref{sec:doctest})
+    (Section~\ref{sec:user-doc} and Section~\ref{sec:doctest}).
 
@@
     code examples for use with Hspec
-    (Section~\ref{sec:code-doc} and Section~\ref{sec:hspec-comb})
-
-\item
-    We improve automatic discovery support for Doctest and Hspec
-    (Section~\ref{sec:solution-discovery} and Section~\ref{sec:implementation-discovery})
+    (Section~\ref{sec:code-doc} and Section~\ref{sec:hspec-comb}).
+
+\item
+    We implemented automatic test discovery for Hspec
+    (Section~\ref{sec:solution-discovery} and
+    Section~\ref{sec:implementation-discovery}).
 
@@
     We show best current practices of BDD in Haskell, using Doctest
-    and Hspec (Section~\ref{sec:best-current-practice})
+    and Hspec (Section~\ref{sec:best-current-practice}).
 

### 2012-06-03 paper.tex
Minor wording change
State, that the solution section describes "automatic spec discovery"
Consistency

@@
 documentation.  The following subsections discuss problems that we
-have to solve before we can do this.
+have to solve before we can do so.
 
@@
 Hspec, are introduced.
+And a mechanism for automatic spec discovery is described.
 
@@
 \begin{itemize}
-\item Add markup syntax for interactive examples to Haddock
-\item Add markup syntax for properties to Haddock
+\item Add markup syntax for interactive examples to Haddock.
+\item Add markup syntax for properties to Haddock.
 \item
@@
 \noindent To specify properties, we also introduce new markup. A line
-starting with ``{\tt prop>}'' indicates a property.
+starting with ``{\tt prop>}'' indicates a property:
 

### 2012-06-19 paper.tex
Remove Show instance of Small
It is not required.
BCP: get rid of a usage of we
we use `we' all the time, so one time less won't hurt
Remove Show instance for Small in appendix
typo fix.

@@
 As we discussed in Section~\ref{sec:introduction},
-We think that some evidence is useful as user documentation.
+we think that some evidence is useful as user documentation.
 But there is also evidence, that is not
@@
 
-\noindent Where \verb|shouldReturn| has the following type:
+\noindent where \verb|shouldReturn| has the following type:
 
@@
 newtype Small = Small Int
-  deriving Show
 
@@
 
-\noindent This spec fails when we run it:
+\noindent This spec fails when run:
 

### 2012-10-20 paper.tex
Show full output for failing examples (as suggested by reviewer 1)

@@
 $ runhaskell FibSpec.hs
+
 fib
  - calculates arbitrary Fibonacci numbers FAILED [1]
+
+1) fib calculates arbitrary Fibonacci numbers FAILED
+*** Falsifiable (after 1 test):
+Small 1
+
+Finished in 0.0021 seconds
+
+1 example, 1 failure
 \end{verbatim}
@@
 $ runhaskell FibSpec.hs
+
 fib
@@
  - is efficient FAILED [1]
+
+1) fib is efficient FAILED
+expected: Just 2178309
+ but got: Nothing
+
+Finished in 0.0685 seconds
+
+2 examples, 1 failure
 \end{verbatim}

### 2012-10-20 paper.tex
Derive Show instance for Small
Update example code

@@
 newtype Small = Small Int
+  deriving Show
 
@@
 
+fibs :: [Integer]
 fibs = 0 : 1 : zipWith (+) fibs (tail fibs)
@@
 
+import Test.Hspec
 import Test.QuickCheck
-import Test.Hspec.ShouldBe
 
@@
 
--- small non-negative numbers
+-- small non-negative integers
 newtype Small = Small Int
+  deriving Show
 
@@
 
+main :: IO ()
 main = hspec spec
 
+spec :: Spec
 spec = do

### 2012-10-20 example/Fib.hs

@@ -8,4 +8,5 @@ fib :: Int -> Integer
 fib n | n < 0     = 0
       | otherwise = fibs !! n
 
+fibs :: [Integer]
 fibs = 0 : 1 : zipWith (+) fibs (tail fibs)

### 2012-10-20 example/FibSpec.hs

@@ -4,20 +4,22 @@ import Control.Applicative
 import Control.Exception
 import System.Timeout
 
+import Test.Hspec
 import Test.QuickCheck
-import Test.Hspec.ShouldBe
 
 import Fib
 
--- small non-negative numbers
+-- small non-negative integers
 newtype Small = Small Int
   deriving Show
 
 instance Arbitrary Small where
   arbitrary = Small . (`mod` 1000) <$> arbitrary
 
+main :: IO ()
 main = hspec spec
 
+spec :: Spec
 spec = do
   describe "fib" $ do
     it "calculates arbitrary Fibonacci numbers" $ do

### 2012-10-20 paper.tex
Getting it fast -> Making it fast (as suggested by reviewer 3)
Fix typo
Remove \titlebanner, etc.

@@
 
-\conferenceinfo{Haskell Symposium 2012}{13th September, Copenhagen.}
+%\conferenceinfo{Haskell Symposium 2012}{13th September, Copenhagen.}
 \copyrightyear{2012}
@@
 % These are ignored, unless 'preprint' option specified.
-\titlebanner{A draft paper to be submitted to Haskell Symposium 2012}
+%\titlebanner{A draft paper to be submitted to Haskell Symposium 2012}
 \preprintfooter{BDD in Haskell}
@@
 spec.  Each line of code is only given once when introduced,
-subsequent code fragments are incremental updates.  Appendix
+subsequent code fragments are incremental updates, import statements
+are omitted.  Appendix
 \ref{sec:appendix-code} contains the complete code of this section.
@@
 
-\subsubsection{Getting it fast}
+\subsubsection{Making it fast}
 
@@
 
-\noindent It is possible to to make this specification reusable, by
+\noindent It is possible to make this specification reusable, by
 abstracting over the type:

### 2012-10-20 paper.tex
Remove everything except "Best current practices"

@@
 
-\category{D.2.2}{Software Engineering}{Testing tools}
-
-\terms Languages, Design, Verification
-
-\keywords Doctest, Haddock, Hspec, QuickCheck
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-\label{sec:introduction}
-
-%%%% Describe the problem.
-Haskell~\cite{haskell} has a strong type system, which ensures that most type errors are
-detected at compile time rather than runtime.  Thanks to this, we assume that
-programs written in Haskell have less bugs than programs written in programming
-languages that do not provide such a strong type system.
-
-Unfortunately, the type system of Haskell does not always guarantee
-correct semantics.
-For example, it does not ensure that division by zero does not occur.
-Another example is, that instances of {\tt Monoid} should be
-monoids in the mathematical sense, but the type system does not
-enforce this.  It is the programmers responsibility to ensure that
-instance definitions behave as expected.  Hence
-automated testing is useful.
-
-However, some widely used Haskell libraries do not have automated
-tests.  For example, the Haskell Platform 2011.4.0.0 contains 23
-packages, not counting the 19 core libraries that come with GHC.  Only
-8 of them have a test suite, 5 provide example code, and 10 have
-neither tests nor examples.  A lack of tests is not only a quality and
-maintainability issue.  It also is a lack of documentation.  Tests
-document how code is supposed to be used, and tests document the
-\emph{behavior} of code.
-
-In other programming language communities, behavior driven development
-(BDD) is getting much attention, and is slowly replacing test driven
-development (TDD).  TDD is an iterative style of software development
-where the developer first writes a few test cases, then implements the
-software in the simplest possible way which makes the test cases pass.
-That could mean yielding a constant value at first.  Subsequently,
-more test cases are written to require more details and drive the
-implementation closer to the desired goal.  Rinse and repeat.
-
-BDD includes TDD, amongst other software development methodologies,
-but with a terminology and mindset shifted towards
-specification instead of testing.
-In BDD, specifications can be used for
-automated testing.
-BDD is meant to be an enjoyable experience for the developer.
-Compared to TDD it provides a much nicer syntax that
-makes tests very easy to read.  And like TDD it puts emphasis on
-test automation, to take workload off the shoulders of
-developers.
-BDD-style TDD has been pioneered by
-RSpec~\cite{rspec}.  In this paper, we explore how to apply the TDD
-aspect of BDD to Haskell.
-
-In BDD, all behavior should be documented, together with some testable
-\emph{evidence} for that behavior.  Most existing BDD tools only use
-\emph{code examples} as evidence for behavior.
-Hspec\footnote{%
-http://hackage.haskell.org/package/hspec},
-a BDD tool for Haskell inspired by RSpec, supports
-both code examples and \emph{properties} as evidence.
-It can verify examples and test properties, and it can create an
-informal specification document from that.
-
-Hspec is thus useful for \emph{code documentation}, but we observed that parts
-of a specification can be useful as \emph{user documentation}. To that end,
-Python's {\tt doctest} module~\cite{doctest} is of interest for us.
-It is not commonly described as a BDD tool, but we think it can be
-used for BDD.
-There is a Haskell version called Doctest\footnote{%
-http://hackage.haskell.org/package/doctest}, which is integrated with Haddock~\cite{haddock},
-the standard documentation mechanism for Haskell.
-It extracts \emph{interactive examples} from Haddock comments and
-verifies them automatically.
-
-We show how to use Doctest and Hspec in combination, to design,
-document and test Haskell code.  Both, Doctest and Hspec are missing
-some features.  So we extend them as needed.  In particular, our
-contributions are:
-
-\begin{itemize}
-\item
-    We originally implemented Haskell's version of Doctest, and
-    extended Haddock's markup to support examples, back in 2010
-    (Section~\ref{sec:haddock} and Section~\ref{sec:doctest}).
-\item
-    We added support for QuickCheck properties to Doctest, and extended
-    Haddock's markup to support properties
-    (Section~\ref{sec:user-doc} and Section~\ref{sec:doctest}).
-
-\item
-    We introduce a set of combinators to express \emph{expectations} about
-    code examples for use with Hspec
-    (Section~\ref{sec:code-doc} and Section~\ref{sec:hspec-comb}).
-
-\item
-    We implemented automatic test discovery for Hspec
-    (Section~\ref{sec:solution-discovery} and
-    Section~\ref{sec:implementation-discovery}).
-
-\item
-    We show best current practices of BDD in Haskell, using Doctest
-    and Hspec (Section~\ref{sec:best-current-practice}).
-
-\end{itemize}
-
-\noindent Our approach can be used for design, documentation and
-automated testing.  We hope that this motivates Haskell programmers to write
-more and better tests.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Terminology}
-
-We use the following terminology:
-
-\begin{description}
-
-\item[example]
-    An example shows how a function is used.  It is either a \emph{code
-    example} in the form of an HUnit test case; or an \emph{interactive
-    example}, given as a short GHCi session.
-
-\item[expectation]
-    Expectations are used to express the expected outcome of a code
-    example.  They serve the same purpose as \emph{assertions} in
-    traditional unit testing frameworks, but use a different
-    terminology.
-
-\item[property]
-    A property is something testable with
-    QuickCheck~\cite{quickcheck}.
-
-\item[behavior]
-    Behavior means an invariant of a function.  We use an informal description
-    to refer to a particular behavior, e.g.  "sorts a list" or "launches
-    missiles when run".
-
-\item[evidence]
-    Evidence provides some confidence, that a function exhibits a
-    certain behavior.  In Haskell we use examples and properties as
-    evidence.
-
-\item[spec]
-    A spec is similar to a test suite in TDD.  But it is written with
-    specification terminology, and produces a specification document
-    when run.
-
-\item[user documentation]
-    User documentation means the documentation for users of a library.
-
-\item[code documentation]
-    Code documentation means a detailed description, or specification,
-    of a library.
-
-\end{description}
-
-%\noindent The word ``behavior'' might be associated with ``imperative
-%action''.  Since pure functions of Haskell are static, this word might
-%be improper for Haskell.  However, this word is already used in the
-%Haskell community~\cite{typeclassopedia}.  So, we use ``behavior''
-%instead of defining terminology in this paper.
-
-%\begin{quote}
-%    sol: We e.g. say ill-behaved Monoid instance.  So is
-%    ``behavior'' really a ``bad'' word?  But I agree that some people in
-%    the community might associated it with something imperative.
-%
-%    kazu: ``behavior'' sounded strange at least to me at the first time.
-%    We alreay supported this word with typeclassopedia. So, I think
-%    this paragraph is completed. Should we replace ``is associated'' with
-%    ``might be associated''?
-%
-%    sol: @kazu, this was primarily meant for @mkl as reference.
-%
-%    One might also argue that ``should'' is not very mathematical, but
-%    it is common language when we talk about mathematics.  E.g. Brent
-%    Yorgey uses both ``should'' and ``behave'' a lot in the
-%    Typeclassopedia (e.g. ``\ldots {\tt Alternative} should satisfy
-%    the monoid laws\ldots'')
-%\end{quote}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Problem}
-\label{sec:problem}
-
-BDD tools can be used to document the behavior of a system in a way that
-is similar to a specification.  All facets of the system should be
-described with evidence, including every possible corner case.
-
-As we discussed in Section~\ref{sec:introduction},
-we think that some evidence is useful as user documentation.
-But there is also evidence, that is not
-useful as user documentation.  E.g. that a particular \texttt{Monoid}
-instance is a monoid in the mathematical sense is self-evident.
-
-We propose to use Doctest and Haddock for evidence that is useful as
-user documentation, and Hspec for evidence that is not useful as user
-documentation.  The following subsections discuss problems that we
-have to solve before we can do so.
-
-\subsection{Haddock}
-\label{sec:doctest-haddock}
-
-Haddock has markup support for code blocks, and code blocks are widely used
-to include properties and examples into documentation\footnote{
-See e.g. the documentation of
-{\tt containers},
-{\tt filepath},
-{\tt bytestring} or
-{\tt text}.}.
-Here is an example from the {\tt filepath} package.
-
-\begin{quote}
-\small
-\begin{verbatim}
--- | Does the given filename have an extension?
---
--- > null (takeExtension x) == not (hasExtension x)
-hasExtension :: FilePath -> Bool
-hasExtension = any isExtSeparator . takeFileName
-\end{verbatim}
-\end{quote}
-
-\noindent Though code blocks in Haddock comments are a useful feature,
-we still think that this particular usage has problems.  The code
-block contains a property.  And it would be useful to have a tool that
-automatically extracts and tests the property.
-But writing such a tool is hard, because it is not easy to decide
-whether the code block is a property, an example, or something else.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Hspec}
-
-If we have evidence that does not fit into the user documentation, but
-would rather be part of a design document, we use Hspec\footnote{%
-Hspec was originally implemented by Trystan Spangler, and is now
-maintained by one of the authors.}.
-
-Hspec provides an EDSL for defining specs.  A \emph{spec} organizes evidence
-in a tree structure (or a forest, to be accurate), and is defined
-with \texttt{it} and \texttt{describe}.
-
-Here is an example that specifies some behavior of the {\tt reverse}
-function.
-
-\begin{quote}
-\small
-\begin{verbatim}
-import Test.Hspec.Monadic
-
-main :: IO ()
-main = hspec spec
-
-spec :: Spec
-spec = do
-  describe "reverse" $ do
-    it "reverses a list" $
-      assertEqual "" [3, 2, 1] (reverse [1, 2, 3])
-
-    it "gives the original list, if applied twice" $
-      property $ \xs ->
-        (reverse . reverse) xs == (xs :: [Int])
-\end{verbatim}
-\end{quote}
-
-\noindent The {\tt it} function combines a text description for a
-behavior and some evidence (a code example or a property) into a spec.
-Code examples can be arbitrary HUnit test cases\footnote{%
-Wrapping the test case with HUnit's {\tt TestCase} constructor is not
-necessary.}.  If a property is used as evidence, it must be wrapped
-with QuickCheck's {\tt property} function.
-
-The \texttt{describe} function combines a list of specs into a larger spec.
-% It can be used to organize a spec as a tree.
-% Internally, a \texttt{Writer} is used to build up specs in a monadic way.
-A spec can be run with the \texttt{hspec} function.  When a spec is
-run, all evidence is tested, and a report is generated.  Behavior with
-evidence that does not hold is marked as {\tt FAILED} in the
-report.
-
-%%%%%%%%%%%%%%%%
-% \subsubsection{Hspec does not provide a mechanism for expectations}
-BDD frameworks use ``should'' or ``must'' instead of
-``assert'' to express expectations.  Here is an example that uses
-RSpec.
-
-\begin{quote}
-\small
-\begin{verbatim}
-[1, 2, 3].reverse.should eq([3, 2, 1])
-\end{verbatim}
-\end{quote}
-
-\noindent Hspec does not yet provide a mechanism to express
-expectations with that terminology.  The developer has to use
-HUnit assertions instead.
-
-%%%%%%%%%%%%%%%%
-% \subsubsection{Hspec provides no mechanism to expect exceptions}
-Moreover RSpec provides a mechanism to expect exceptions.  It is used
-like this:
-
-\begin{quote}
-\small
-\begin{verbatim}
-expect {launch_missiles}.to raise_error
-\end{verbatim}
-\end{quote}
-
-\noindent Neither Hspec nor HUnit provide a convenient equivalent.
-
-\subsection{Automatic discovery of specs}
-\label{sec:problem-discovery}
-
-It is a common convention to have one spec file for each source file.
-To run all specs for a given project together, it is necessary to
-combine all the specs from all spec files into a single spec.  For
-example, if we have three modules, \verb|Foo|, \verb|Foo.Bar| and
-\verb|Baz|, and corresponding specs in \verb|FooSpec|,
-\verb|Foo.BarSpec| and \verb|BazSpec|, we are required to write the
-following boilerplate:
-
-\begin{quote}
-\small
-\begin{verbatim}
-import Test.Hspec.Monadic
-
-import qualified FooSpec
-import qualified Foo.BarSpec
-import qualified BazSpec
-
-main :: IO ()
-main = hspec $ do
-  describe "Foo"     FooSpec.spec
-  describe "Foo.Bar" Foo.BarSpec.spec
-  describe "Baz"     BazSpec.spec
-\end{verbatim}
-\end{quote}
-
-\noindent This is error prone, and neither challenging nor
-interesting.  So it should be automated.  But Hspec does not provide a
-solution for generating such \emph{test drivers}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Problem summary}
-
-Here is a summary of the problems:
-
-\begin{itemize}
-\item
-    There is no way to automatically test examples and properties
-    contained in user documentation.
-\item
-    For code documentation,
-    Hspec does not provide combinators to express expectations in
-    BDD terminology.
-\item
-    There is no facility that automatically finds and combines all
-    specs of a project.
-\end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Solution}
-
-As we explained in Section~\ref{sec:problem},
-we use Doctest for user documentation and Hspec for code documentation.
-The following subsections explain two extensions to Haddock's markup.
-Subsequently, combinators for expectations, that can be used with
-Hspec, are introduced.
-And a mechanism for automatic spec discovery is described.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{User documentation}
-\label{sec:user-doc}
-
-Evidence that is useful as user documentation should be included in
-Haddock comments.  For this purpose, we propose the following:
-
-\begin{itemize}
-\item Add markup syntax for interactive examples to Haddock.
-\item Add markup syntax for properties to Haddock.
-\item
-    Implement a tool, similar to Python's {\tt doctest} module, that
-    can verify examples and test properties in Haddock comments.
-    The implementation is discussed in Section~\ref{sec:doctest}.
-
-\end{itemize}
-
-%%%%%%%%%%%%%%%%
-\subsubsection{Markup for examples}
-
-We introduce new Haddock markup for interactive examples.  A line
-starting with a \emph{prompt} (``{\tt >>>}'') denotes an expression.
-Lines following an expression denote the result of that expression.
-Here is an example:
-
-\begin{quote}
-\small
-\begin{verbatim}
--- |
--- >>> length []
--- 0
-\end{verbatim}
-\end{quote}
-
-\noindent The result ends at the next empty line, or another prompt.
-It is possible to specify example sequences:
-
-\begin{quote}
-\small
-\begin{verbatim}
--- |
--- >>> writeFile "tmpfile" "Hello"
--- >>> readFile "tmpfile"
--- "Hello"
-\end{verbatim}
-\end{quote}
-
-\noindent Result is defined by what GHCi prints to stdout and stderr
-when evaluating the expression.  This, e.g. also allows to document
-exceptions:
-
-\begin{quote}
-\small
-\begin{verbatim}
--- |
--- >>> head []
--- *** Exception: Prelude.head: empty list
-\end{verbatim}
-\end{quote}
-
-%%%%%%%%%%%%%%%%
-\subsubsection{Markup for properties}
-\label{sec:markup}
-
-\noindent To specify properties, we also introduce new markup. A line
-starting with ``{\tt prop>}'' indicates a property:
-
-\begin{quote}
-\small
-\begin{verbatim}
--- |
--- prop> Data.Map.null empty == True
-\end{verbatim}
-\end{quote}
-
-\noindent It is possible to quantify properties by using a
-lambda expression. The property from Section~\ref{sec:doctest-haddock}
-can be written as:
-
-\begin{quote}
-\small
-\begin{alltt}
--- |
--- prop> \verb|\|x -> null (takeExtension x) == \(\hookleftarrow\)
-               not (hasExtension x)
-\end{alltt}
-\end{quote}
-
-\noindent The $\hookleftarrow$ symbol denotes line continuation.
-
-GHCi's type defaulting\footnote{%
-\url{http://www.haskell.org/ghc/docs/7.4.1/html/users_guide/interactive-evaluation.html#extended-default-rules}
-} applies to properties.  If a default type is
-not suitable, explicit type-signatures can be used:
-
-\begin{quote}
-\small
-\begin{verbatim}
--- |
--- prop> \xs -> sort xs == (sort . sort) (xs :: [Int])
-\end{verbatim}
-\end{quote}
-
-\noindent Since the lambda abstraction adds nothing to the
-documentation, it can be omitted:
-
-\begin{quote}
-\small
-\begin{verbatim}
--- |
--- prop> sort xs == (sort . sort) (xs :: [Int])
-\end{verbatim}
-\end{quote}
-
-\noindent We call such properties \emph{implicitly quantified}.  The
-given property is implicitly quantified over {\tt xs}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Code documentation}
-\label{sec:code-doc}
-
-We propose the following extensions for Hspec to
-describe evidence for developers.
-
-%%%%%%%%%%%%%%%%
-\subsubsection{Setting expectations}
-
-We propose several combinators which can be used to set expectations
-about the outcome of code examples.  These are built on top of HUnit:
-
-\begin{quote}
-\small
-\begin{verbatim}
-type Expectation = Assertion
-\end{verbatim}
-\end{quote}
-
-\noindent A common expectation is, that a value should be equal to another
-value:
-
-\begin{quote}
-\small
-\begin{verbatim}
-x `shouldBe` 23
-\end{verbatim}
-\end{quote}
-
-\noindent This is just another name for HUnit's \texttt{@?=}
-operator.  The type is:
-
-\begin{quote}
-\small
-\begin{verbatim}
-shouldBe :: (Show a, Eq a) => a -> a -> Expectation
-\end{verbatim}
-\end{quote}
-
-\noindent Or we may expect that a value should satisfy some
-predicate.
-
-\begin{quote}
-\small
-\begin{verbatim}
-x `shouldSatisfy` (< 23)
-\end{verbatim}
-\end{quote}
-
-\noindent The value is required to be in the \texttt{Show} class, so that a
-useful error message can be given when the predicate is not satisfied.
-
-\begin{quote}
-\small
-\begin{verbatim}
-shouldSatisfy :: Show a => a -> (a -> Bool)
-  -> Expectation
-\end{verbatim}
-\end{quote}
-
-\noindent Code examples are often used to provide evidence for {\tt
-IO} actions, e.g.:
-
-\begin{quote}
-\small
-\begin{verbatim}
-launchMissiles >>=
-  (`shouldBe` Left "not implemented")
-\end{verbatim}
-\end{quote}
-
-\noindent So we propose a shortcut for this common use case:
-
-\begin{quote}
-\small
-\begin{verbatim}
-launchMissiles `shouldReturn` Left "not implemented"
-\end{verbatim}
-\end{quote}
-
-\noindent where \verb|shouldReturn| has the following type:
-
-\begin{quote}
-\small
-\begin{verbatim}
-shouldReturn :: (Show a, Eq a) => IO a -> a
-  -> Expectation
-\end{verbatim}
-\end{quote}
-
-%%%%%%%%%%%%%%%%
-\subsubsection{Expecting exceptions}
-\label{sec:expacting-exceptions}
-
-HUnit does not provide a convenient mechanism to state that an action
-should throw an exception.  So we propose a combinator for that.
-
-\begin{quote}
-\small
-\begin{verbatim}
-shouldThrow :: Exception e => IO a -> (e -> Bool)
-  -> Expectation
-\end{verbatim}
-\end{quote}
-
-\noindent It takes an {\tt IO} action, and a predicate.  The predicate
-servers two purposes.  It selects the type of the expected exception,
-and constrains the value.  Here is an example:
-
-\begin{quote}
-\small
-\begin{verbatim}
-launchMissiles `shouldThrow` (== ExitFailure 1)
-\end{verbatim}
-\end{quote}
-
-%\noindent And it can be used with predicates from {\tt
-%System.IO.Error}, to require specific {\tt IOException}, e.g.  with
-%{\tt isPermissionError}:
-%
-%\begin{quote}
-%\small
-%\begin{verbatim}
-%launchMissiles `shouldThrow` isPermissionError
-%\end{verbatim}
-%\end{quote}
-
-\noindent To make it more convenient to expect any exception of a
-certain type, we provide type-restricted versions of {\tt (const
-True)} for all standard exceptions, e.g.:
-
-\begin{quote}
-\small
-\begin{verbatim}
-anyArithException :: ArithException -> Bool
-anyArithException = const True
-\end{verbatim}
-\end{quote}
-
-\noindent Which can be used like this:
-
-\begin{quote}
-\small
-\begin{verbatim}
-evaluate (1 `div` 0) `shouldThrow` anyArithException
-\end{verbatim}
-\end{quote}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Automatic discovery of specs}
-\label{sec:solution-discovery}
-
-We want to automatically discover all specs of a project, so that we
-can run them all together.  This requires to automatically generate
-boilerplate code, similar to what is shown in
-Section~\ref{sec:problem-discovery}.
-
-We make creative use GHC's support for custom preprocessors for
-that.  With our solution, the developer creates a test driver, that only
-contains a single line:
-
-\begin{quote}
-\small
-\begin{verbatim}
-{-# OPTIONS_GHC -F -pgmF hspec-discover #-}
-\end{verbatim}
-\end{quote}
-
-\noindent This tells GHC to run a preprocessor, \verb|hspec-discover|, on the
-source file.  \verb|hspec-discover| finds all specs for a project, and
-generates a suitable test driver.
-
-This approach relies on the following conventions:
-\begin{enumerate}
-    \item
-        Spec files have to be placed into the same directory as the
-        test driver, or into a subdirectory.
-    \item
-        The name of a spec file has to end in \verb|Spec.hs|, and the
-        module name has to match the file name.
-    \item
-        Each spec file has to export a top-level binding \verb|spec|,
-        of type \verb|Spec|.
-\end{enumerate}
-
-\noindent It would be possible to lift some of those requirements.
-But we think they are reasonable conventions, anyway.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Implementation}
-
-In this section we describe implementation aspects of the proposed
-solution.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Haddock}
-\label{sec:haddock}
-
-We added markup support for interactive examples to Haddock, and it is
-available since version 2.8.
-We also added markup support for properties, which will be included in
-a future release of Haddock.
-
-\subsection{Doctest}
-\label{sec:doctest}
-
-Doctest uses the GHC API to extract properties and examples from
-Haddock comments, and uses GHCi to test them.
-Before testing, Doctest brings the module that contains the Haddock
-comment from which the evidence was extracted into scope.
-
-\subsubsection{Testing examples}
-
-Verification of examples is done by
-comparing the expected output given in the Haddock comment with the
-actual output of GHCi.
-
-Doctest reuses the same GHCi session for all examples.
-Examples from the same Haddock comment are grouped together,
-and share the same scope.
-Doctest clears the scope with \verb|:reload| before it tests a new
-example group.
-
-\subsubsection{Testing properties}
-
-QuickCheck is used to test properties.
-Implicitly quantified properties need to be closed before they can be
-tested.  This requires the following steps.
-\begin{enumerate}
-    \item
-        Find all free variables of the property.
-    \item
-        Try to lookup each free variable in the module context.
-    \item
-        Close the property over all free variables, for which lookup
-        failed.
-\end{enumerate}
-
-\noindent Here is the example from Section~\ref{sec:markup}:
-
-\begin{quote}
-\small
-\begin{verbatim}
-sort xs == (sort . sort) (xs :: [Int]))
-\end{verbatim}
-\end{quote}
-
-\noindent Both \verb|xs| and \verb|sort| are free variables, but
-\verb|sort| is bound by the module context.  So we need to close the
-term over \verb|xs|.
-
-\begin{quote}
-\small
-\begin{verbatim}
-\xs -> sort xs == (sort . sort) (xs :: [Int])
-\end{verbatim}
-\end{quote}
-
-\noindent The implementation is fully functional, and it will be included in the
-next release of Doctest.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Hspec combinators for expectations}
-\label{sec:hspec-comb}
-
-\noindent All combinators introduced in Section~\ref{sec:code-doc} have been
-implemented.  They are available from Hackage\footnote{%
-http://hackage.haskell.org/package/hspec-shouldbe},
-and will be incorporated into a future
-release of Hspec.
-The implementation is straightforward, so we will not further discuss
-it.
-
-\subsection{Automatic discovery of specs}
-\label{sec:implementation-discovery}
-
-As discussed in Section~\ref{sec:solution-discovery}, we use a preprocessor to
-automatically generate test drivers for Hspec.  GHC invokes the
-preprocessor with three arguments:
-
-\begin{description}
-    \item[{\tt src}]
-        -- path to the original source file\\
-        A preprocessor usually does not
-        look at the content of this file.  It is only provided, so
-        that it can be used in error messages.
-    \item[{\tt cur}]
-        -- path to a file that holds the input for the preprocessor
-    \item[{\tt dst}]
-        -- a path, where the preprocessor should write it's result to
-\end{description}
-
-\noindent A typical preprocessor reads the content of {\tt cur},
-processes it, and writes the result to {\tt dst}.  If something goes
-wrong, it uses {\tt src} in error messages to refer to the source
-file.
-
-\verb|hspec-discover| ignores {\tt cur} altogether.  It uses the
-directory component of {\tt src}, to find all files that are in the
-same directory as the source file, or in any subdirectories.  It then generates
-a test driver from all files that end in {\tt Spec.hs}.
-
-A fully functional implementation is available from
-Hackage\footnote{http://hackage.haskell.org/package/hspec-discover},
-and it will be incorporated into a future release of Hspec.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Best current practices}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Open problems and future work}
-
-In Section~\ref{sec:reusable-specs}, we showed how to reuse pieces of
-specification.  It would be possible and useful to write a library
-that provides reusable pieces of specification, for many of Haskell's
-standard classes (similar in purpose to {\tt
-checkers}\footnote{http://hackage.haskell.org/package/checkers}).
-This would provide a convenient mechanism for developers to document
-and test that their instance definitions are well-behaved.
-
-When Hspec finds failing evidence, it reports the corresponding
-behavior to the developer.  It would be useful to include source
-locations for failing evidence, so that the programmer can find the
-corresponding spec immediately. Solving this with GHC requires
-Template Haskell.  Extending GHC with a mechanism
-similar to JHC's \verb|SRCLOC_ANNOTATE| pragma\footnote{%
-http://repetae.net/computer/jhc/jhc.shtml} would allow to solve this
-without Template Haskell.
-
-Cucumber~\cite{cucumber} is a higher-level approach to BDD for Ruby, and
-it is typically used in combination with RSpec.
-It focuses on user stories, to assist the developer in delivering
-value to the stakeholder.
-There is nothing comparable to Cucumber for Haskell, yet.  And we think
-it would be interesting to explore, whether it is possible to apply
-Cucumber's approach to Haskell.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Related work}
-
-SmallCheck~\cite{smallcheck} is a tool for
-automatic exhaustive testing up to a certain depth.
-Both Doctest and Hspec do not have support for SmallCheck
-at this moment.
-Hspec's mechanism for evidence is extensible, and it would be easy to
-add support for SmallCheck.
-
-{\tt test-framework} is a library that allows TDD in Haskell, and it
-has support for HUnit, QuickCheck and SmallCheck.
-There is support for automatic test discovery that relies on naming
-conventions for test functions\footnote{%
-http://hackage.haskell.org/package/test-framework-th}.  It is
-implemented with Template Haskell~\cite{template-haskell}, and it only
-works within a single source file.  Automatic discovery across
-multiple source files is not possible with a mechanism that is based
-on Template Haskell, because Template Haskell can not generate import
-statements\footnote{%
-See http://hackage.haskell.org/trac/ghc/ticket/1475 }.
-
-The {\tt testpack}\footnote{%
-http://hackage.haskell.org/package/testpack}
-package provides a combinator to expect exceptions,
-similar to {\tt shouldThrow} from
-Section~\ref{sec:expacting-exceptions}:
-
-\begin{quote}
-\small
-\begin{verbatim}
-assertRaises :: (Show a, Exception e, Show e, Eq e) =>
-  String -> e -> IO a -> IO ()
-\end{verbatim}
-\end{quote}
-
-\noindent But it has several shortcomings:
-It only works for exceptions with an {\tt Eq} instance,
-but several standard exceptions (most
-prominently {\tt ErrorCall}) have no {\tt Eq} instance.
-And it places an unnecessary {\tt Show} constraint on the
-result of the {\tt IO} action, further limiting it's applicability.
-It only allows to expect specific exception values, expecting
-arbitrary exceptions of a certain type is not possible.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusion}
-
-We showed how to apply behavior driven development to
-Haskell.  We proposed to use Doctest for user documentation and Hspec
-for code documentation. We extended Haddock's markup with support for
-examples and properties.  And we implemented Doctest, which verifies
-examples and tests properties in Haddock comments.
-We added two missing features to Hspec: combinators to express
-expectations, and a mechanism for automatic discovery of specs.
-And we showed how to reuse pieces of specification.
-Our approach can be used for design, documentation and automated
-testing.  We hope that this motivates Haskell programmers to write
-more and better tests.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \appendix
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\acks
-
-We are grateful to David Waern, Haddock's maintainer, for reviewing
-and accepting our patches to Haddock.
-%We would like to thank
-%Nick Smallbone for having improved QuickCheck so that
-%non-quantified properties are tested just once.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \bibliographystyle{abbrvnat}

### 2012-10-20 paper.tex
Shorten abstract

@@
 \begin{abstract}
-We observe that several widely used Haskell libraries do not have
-tests.  Test code assists developers in maintaining correctness when
-extending software, and shows users how an API is supposed to be used.
-To ease the production and enhance the value of test code, we advocate
-behavior driven development with two tools: Doctest for user
-documentation and Hspec for code documentation.  Our approach can be
-used for design, documentation and automated testing.  We hope that
-this motivates Haskell programmers to write more and better tests.
+Several widely used Haskell libraries do not have tests.  Test code
+assists developers in maintaining correctness when extending software,
+and shows users how an API is supposed to be used.
+
+We show how to do behavior-driven development in Haskell with two
+tools: Doctest and Hspec.  Our approach can be used for design,
+documentation and automated testing.  We hope that this motivates
+Haskell programmers to write more and better tests.
 \end{abstract}

### 2012-10-20 paper.tex
behavior driven -> behavior-driven

@@
 %\titlebanner{A draft paper to be submitted to Haskell Symposium 2012}
-\preprintfooter{BDD in Haskell}
+\preprintfooter{Behavior-Driven Development in Haskell, a tutorial
+introduction}
 
-\title{Behavior Driven Development in Haskell}
-%\subtitle{Subtitle Text, if any}
+\title{Behavior-Driven Development in Haskell}
+\subtitle{a tutorial introduction}
 
@@
 \label{sec:bcp-example}
-This section shows how behavior driven development works in practice,
+This section shows how behavior-driven development works in practice,
 by means of the Fibonacci function.  This is a very simple
@@
 \noindent This concludes our example.  We have seen how to utilize
-behavior driven development to incrementally build a small library.
+behavior-driven development to incrementally build a small library.
 We always extend or adapt the specification first, watch it fail,  and

### 2012-10-20 paper.tex
State explicitly that import statements are omitted

--- 2012-10-20 paper.tex
@@
 
-%\conferenceinfo{Haskell Symposium 2012}{13th September, Copenhagen.}
+\conferenceinfo{Haskell Symposium 2012}{13th September, Copenhagen.}
 \copyrightyear{2012}
@@
 % These are ignored, unless 'preprint' option specified.
-%\titlebanner{A draft paper to be submitted to Haskell Symposium 2012}
+\titlebanner{A draft paper to be submitted to Haskell Symposium 2012}
 \preprintfooter{BDD in Haskell}
@@
 
+\category{D.2.2}{Software Engineering}{Testing tools}
+
+\terms Languages, Design, Verification
+
+\keywords Doctest, Haddock, Hspec, QuickCheck
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Introduction}
+\label{sec:introduction}
+
+%%%% Describe the problem.
+Haskell~\cite{haskell} has a strong type system, which ensures that most type errors are
+detected at compile time rather than runtime.  Thanks to this, we assume that
+programs written in Haskell have less bugs than programs written in programming
+languages that do not provide such a strong type system.
+
+Unfortunately, the type system of Haskell does not always guarantee
+correct semantics.
+For example, it does not ensure that division by zero does not occur.
+Another example is, that instances of {\tt Monoid} should be
+monoids in the mathematical sense, but the type system does not
+enforce this.  It is the programmers responsibility to ensure that
+instance definitions behave as expected.  Hence
+automated testing is useful.
+
+However, some widely used Haskell libraries do not have automated
+tests.  For example, the Haskell Platform 2011.4.0.0 contains 23
+packages, not counting the 19 core libraries that come with GHC.  Only
+8 of them have a test suite, 5 provide example code, and 10 have
+neither tests nor examples.  A lack of tests is not only a quality and
+maintainability issue.  It also is a lack of documentation.  Tests
+document how code is supposed to be used, and tests document the
+\emph{behavior} of code.
+
+In other programming language communities, behavior driven development
+(BDD) is getting much attention, and is slowly replacing test driven
+development (TDD).  TDD is an iterative style of software development
+where the developer first writes a few test cases, then implements the
+software in the simplest possible way which makes the test cases pass.
+That could mean yielding a constant value at first.  Subsequently,
+more test cases are written to require more details and drive the
+implementation closer to the desired goal.  Rinse and repeat.
+
+BDD includes TDD, amongst other software development methodologies,
+but with a terminology and mindset shifted towards
+specification instead of testing.
+In BDD, specifications can be used for
+automated testing.
+BDD is meant to be an enjoyable experience for the developer.
+Compared to TDD it provides a much nicer syntax that
+makes tests very easy to read.  And like TDD it puts emphasis on
+test automation, to take workload off the shoulders of
+developers.
+BDD-style TDD has been pioneered by
+RSpec~\cite{rspec}.  In this paper, we explore how to apply the TDD
+aspect of BDD to Haskell.
+
+In BDD, all behavior should be documented, together with some testable
+\emph{evidence} for that behavior.  Most existing BDD tools only use
+\emph{code examples} as evidence for behavior.
+Hspec\footnote{%
+http://hackage.haskell.org/package/hspec},
+a BDD tool for Haskell inspired by RSpec, supports
+both code examples and \emph{properties} as evidence.
+It can verify examples and test properties, and it can create an
+informal specification document from that.
+
+Hspec is thus useful for \emph{code documentation}, but we observed that parts
+of a specification can be useful as \emph{user documentation}. To that end,
+Python's {\tt doctest} module~\cite{doctest} is of interest for us.
+It is not commonly described as a BDD tool, but we think it can be
+used for BDD.
+There is a Haskell version called Doctest\footnote{%
+http://hackage.haskell.org/package/doctest}, which is integrated with Haddock~\cite{haddock},
+the standard documentation mechanism for Haskell.
+It extracts \emph{interactive examples} from Haddock comments and
+verifies them automatically.
+
+We show how to use Doctest and Hspec in combination, to design,
+document and test Haskell code.  Both, Doctest and Hspec are missing
+some features.  So we extend them as needed.  In particular, our
+contributions are:
+
+\begin{itemize}
+\item
+    We originally implemented Haskell's version of Doctest, and
+    extended Haddock's markup to support examples, back in 2010
+    (Section~\ref{sec:haddock} and Section~\ref{sec:doctest}).
+\item
+    We added support for QuickCheck properties to Doctest, and extended
+    Haddock's markup to support properties
+    (Section~\ref{sec:user-doc} and Section~\ref{sec:doctest}).
+
+\item
+    We introduce a set of combinators to express \emph{expectations} about
+    code examples for use with Hspec
+    (Section~\ref{sec:code-doc} and Section~\ref{sec:hspec-comb}).
+
+\item
+    We implemented automatic test discovery for Hspec
+    (Section~\ref{sec:solution-discovery} and
+    Section~\ref{sec:implementation-discovery}).
+
+\item
+    We show best current practices of BDD in Haskell, using Doctest
+    and Hspec (Section~\ref{sec:best-current-practice}).
+
+\end{itemize}
+
+\noindent Our approach can be used for design, documentation and
+automated testing.  We hope that this motivates Haskell programmers to write
+more and better tests.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Terminology}
+
+We use the following terminology:
+
+\begin{description}
+
+\item[example]
+    An example shows how a function is used.  It is either a \emph{code
+    example} in the form of an HUnit test case; or an \emph{interactive
+    example}, given as a short GHCi session.
+
+\item[expectation]
+    Expectations are used to express the expected outcome of a code
+    example.  They serve the same purpose as \emph{assertions} in
+    traditional unit testing frameworks, but use a different
+    terminology.
+
+\item[property]
+    A property is something testable with
+    QuickCheck~\cite{quickcheck}.
+
+\item[behavior]
+    Behavior means an invariant of a function.  We use an informal description
+    to refer to a particular behavior, e.g.  "sorts a list" or "launches
+    missiles when run".
+
+\item[evidence]
+    Evidence provides some confidence, that a function exhibits a
+    certain behavior.  In Haskell we use examples and properties as
+    evidence.
+
+\item[spec]
+    A spec is similar to a test suite in TDD.  But it is written with
+    specification terminology, and produces a specification document
+    when run.
+
+\item[user documentation]
+    User documentation means the documentation for users of a library.
+
+\item[code documentation]
+    Code documentation means a detailed description, or specification,
+    of a library.
+
+\end{description}
+
+%\noindent The word ``behavior'' might be associated with ``imperative
+%action''.  Since pure functions of Haskell are static, this word might
+%be improper for Haskell.  However, this word is already used in the
+%Haskell community~\cite{typeclassopedia}.  So, we use ``behavior''
+%instead of defining terminology in this paper.
+
+%\begin{quote}
+%    sol: We e.g. say ill-behaved Monoid instance.  So is
+%    ``behavior'' really a ``bad'' word?  But I agree that some people in
+%    the community might associated it with something imperative.
+%
+%    kazu: ``behavior'' sounded strange at least to me at the first time.
+%    We alreay supported this word with typeclassopedia. So, I think
+%    this paragraph is completed. Should we replace ``is associated'' with
+%    ``might be associated''?
+%
+%    sol: @kazu, this was primarily meant for @mkl as reference.
+%
+%    One might also argue that ``should'' is not very mathematical, but
+%    it is common language when we talk about mathematics.  E.g. Brent
+%    Yorgey uses both ``should'' and ``behave'' a lot in the
+%    Typeclassopedia (e.g. ``\ldots {\tt Alternative} should satisfy
+%    the monoid laws\ldots'')
+%\end{quote}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Problem}
+\label{sec:problem}
+
+BDD tools can be used to document the behavior of a system in a way that
+is similar to a specification.  All facets of the system should be
+described with evidence, including every possible corner case.
+
+As we discussed in Section~\ref{sec:introduction},
+we think that some evidence is useful as user documentation.
+But there is also evidence, that is not
+useful as user documentation.  E.g. that a particular \texttt{Monoid}
+instance is a monoid in the mathematical sense is self-evident.
+
+We propose to use Doctest and Haddock for evidence that is useful as
+user documentation, and Hspec for evidence that is not useful as user
+documentation.  The following subsections discuss problems that we
+have to solve before we can do so.
+
+\subsection{Haddock}
+\label{sec:doctest-haddock}
+
+Haddock has markup support for code blocks, and code blocks are widely used
+to include properties and examples into documentation\footnote{
+See e.g. the documentation of
+{\tt containers},
+{\tt filepath},
+{\tt bytestring} or
+{\tt text}.}.
+Here is an example from the {\tt filepath} package.
+
+\begin{quote}
+\small
+\begin{verbatim}
+-- | Does the given filename have an extension?
+--
+-- > null (takeExtension x) == not (hasExtension x)
+hasExtension :: FilePath -> Bool
+hasExtension = any isExtSeparator . takeFileName
+\end{verbatim}
+\end{quote}
+
+\noindent Though code blocks in Haddock comments are a useful feature,
+we still think that this particular usage has problems.  The code
+block contains a property.  And it would be useful to have a tool that
+automatically extracts and tests the property.
+But writing such a tool is hard, because it is not easy to decide
+whether the code block is a property, an example, or something else.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Hspec}
+
+If we have evidence that does not fit into the user documentation, but
+would rather be part of a design document, we use Hspec\footnote{%
+Hspec was originally implemented by Trystan Spangler, and is now
+maintained by one of the authors.}.
+
+Hspec provides an EDSL for defining specs.  A \emph{spec} organizes evidence
+in a tree structure (or a forest, to be accurate), and is defined
+with \texttt{it} and \texttt{describe}.
+
+Here is an example that specifies some behavior of the {\tt reverse}
+function.
+
+\begin{quote}
+\small
+\begin{verbatim}
+import Test.Hspec.Monadic
+
+main :: IO ()
+main = hspec spec
+
+spec :: Spec
+spec = do
+  describe "reverse" $ do
+    it "reverses a list" $
+      assertEqual "" [3, 2, 1] (reverse [1, 2, 3])
+
+    it "gives the original list, if applied twice" $
+      property $ \xs ->
+        (reverse . reverse) xs == (xs :: [Int])
+\end{verbatim}
+\end{quote}
+
+\noindent The {\tt it} function combines a text description for a
+behavior and some evidence (a code example or a property) into a spec.
+Code examples can be arbitrary HUnit test cases\footnote{%
+Wrapping the test case with HUnit's {\tt TestCase} constructor is not
+necessary.}.  If a property is used as evidence, it must be wrapped
+with QuickCheck's {\tt property} function.
+
+The \texttt{describe} function combines a list of specs into a larger spec.
+% It can be used to organize a spec as a tree.
+% Internally, a \texttt{Writer} is used to build up specs in a monadic way.
+A spec can be run with the \texttt{hspec} function.  When a spec is
+run, all evidence is tested, and a report is generated.  Behavior with
+evidence that does not hold is marked as {\tt FAILED} in the
+report.
+
+%%%%%%%%%%%%%%%%
+% \subsubsection{Hspec does not provide a mechanism for expectations}
+BDD frameworks use ``should'' or ``must'' instead of
+``assert'' to express expectations.  Here is an example that uses
+RSpec.
+
+\begin{quote}
+\small
+\begin{verbatim}
+[1, 2, 3].reverse.should eq([3, 2, 1])
+\end{verbatim}
+\end{quote}
+
+\noindent Hspec does not yet provide a mechanism to express
+expectations with that terminology.  The developer has to use
+HUnit assertions instead.
+
+%%%%%%%%%%%%%%%%
+% \subsubsection{Hspec provides no mechanism to expect exceptions}
+Moreover RSpec provides a mechanism to expect exceptions.  It is used
+like this:
+
+\begin{quote}
+\small
+\begin{verbatim}
+expect {launch_missiles}.to raise_error
+\end{verbatim}
+\end{quote}
+
+\noindent Neither Hspec nor HUnit provide a convenient equivalent.
+
+\subsection{Automatic discovery of specs}
+\label{sec:problem-discovery}
+
+It is a common convention to have one spec file for each source file.
+To run all specs for a given project together, it is necessary to
+combine all the specs from all spec files into a single spec.  For
+example, if we have three modules, \verb|Foo|, \verb|Foo.Bar| and
+\verb|Baz|, and corresponding specs in \verb|FooSpec|,
+\verb|Foo.BarSpec| and \verb|BazSpec|, we are required to write the
+following boilerplate:
+
+\begin{quote}
+\small
+\begin{verbatim}
+import Test.Hspec.Monadic
+
+import qualified FooSpec
+import qualified Foo.BarSpec
+import qualified BazSpec
+
+main :: IO ()
+main = hspec $ do
+  describe "Foo"     FooSpec.spec
+  describe "Foo.Bar" Foo.BarSpec.spec
+  describe "Baz"     BazSpec.spec
+\end{verbatim}
+\end{quote}
+
+\noindent This is error prone, and neither challenging nor
+interesting.  So it should be automated.  But Hspec does not provide a
+solution for generating such \emph{test drivers}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Problem summary}
+
+Here is a summary of the problems:
+
+\begin{itemize}
+\item
+    There is no way to automatically test examples and properties
+    contained in user documentation.
+\item
+    For code documentation,
+    Hspec does not provide combinators to express expectations in
+    BDD terminology.
+\item
+    There is no facility that automatically finds and combines all
+    specs of a project.
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Solution}
+
+As we explained in Section~\ref{sec:problem},
+we use Doctest for user documentation and Hspec for code documentation.
+The following subsections explain two extensions to Haddock's markup.
+Subsequently, combinators for expectations, that can be used with
+Hspec, are introduced.
+And a mechanism for automatic spec discovery is described.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{User documentation}
+\label{sec:user-doc}
+
+Evidence that is useful as user documentation should be included in
+Haddock comments.  For this purpose, we propose the following:
+
+\begin{itemize}
+\item Add markup syntax for interactive examples to Haddock.
+\item Add markup syntax for properties to Haddock.
+\item
+    Implement a tool, similar to Python's {\tt doctest} module, that
+    can verify examples and test properties in Haddock comments.
+    The implementation is discussed in Section~\ref{sec:doctest}.
+
+\end{itemize}
+
+%%%%%%%%%%%%%%%%
+\subsubsection{Markup for examples}
+
+We introduce new Haddock markup for interactive examples.  A line
+starting with a \emph{prompt} (``{\tt >>>}'') denotes an expression.
+Lines following an expression denote the result of that expression.
+Here is an example:
+
+\begin{quote}
+\small
+\begin{verbatim}
+-- |
+-- >>> length []
+-- 0
+\end{verbatim}
+\end{quote}
+
+\noindent The result ends at the next empty line, or another prompt.
+It is possible to specify example sequences:
+
+\begin{quote}
+\small
+\begin{verbatim}
+-- |
+-- >>> writeFile "tmpfile" "Hello"
+-- >>> readFile "tmpfile"
+-- "Hello"
+\end{verbatim}
+\end{quote}
+
+\noindent Result is defined by what GHCi prints to stdout and stderr
+when evaluating the expression.  This, e.g. also allows to document
+exceptions:
+
+\begin{quote}
+\small
+\begin{verbatim}
+-- |
+-- >>> head []
+-- *** Exception: Prelude.head: empty list
+\end{verbatim}
+\end{quote}
+
+%%%%%%%%%%%%%%%%
+\subsubsection{Markup for properties}
+\label{sec:markup}
+
+\noindent To specify properties, we also introduce new markup. A line
+starting with ``{\tt prop>}'' indicates a property:
+
+\begin{quote}
+\small
+\begin{verbatim}
+-- |
+-- prop> Data.Map.null empty == True
+\end{verbatim}
+\end{quote}
+
+\noindent It is possible to quantify properties by using a
+lambda expression. The property from Section~\ref{sec:doctest-haddock}
+can be written as:
+
+\begin{quote}
+\small
+\begin{alltt}
+-- |
+-- prop> \verb|\|x -> null (takeExtension x) == \(\hookleftarrow\)
+               not (hasExtension x)
+\end{alltt}
+\end{quote}
+
+\noindent The $\hookleftarrow$ symbol denotes line continuation.
+
+GHCi's type defaulting\footnote{%
+\url{http://www.haskell.org/ghc/docs/7.4.1/html/users_guide/interactive-evaluation.html#extended-default-rules}
+} applies to properties.  If a default type is
+not suitable, explicit type-signatures can be used:
+
+\begin{quote}
+\small
+\begin{verbatim}
+-- |
+-- prop> \xs -> sort xs == (sort . sort) (xs :: [Int])
+\end{verbatim}
+\end{quote}
+
+\noindent Since the lambda abstraction adds nothing to the
+documentation, it can be omitted:
+
+\begin{quote}
+\small
+\begin{verbatim}
+-- |
+-- prop> sort xs == (sort . sort) (xs :: [Int])
+\end{verbatim}
+\end{quote}
+
+\noindent We call such properties \emph{implicitly quantified}.  The
+given property is implicitly quantified over {\tt xs}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Code documentation}
+\label{sec:code-doc}
+
+We propose the following extensions for Hspec to
+describe evidence for developers.
+
+%%%%%%%%%%%%%%%%
+\subsubsection{Setting expectations}
+
+We propose several combinators which can be used to set expectations
+about the outcome of code examples.  These are built on top of HUnit:
+
+\begin{quote}
+\small
+\begin{verbatim}
+type Expectation = Assertion
+\end{verbatim}
+\end{quote}
+
+\noindent A common expectation is, that a value should be equal to another
+value:
+
+\begin{quote}
+\small
+\begin{verbatim}
+x `shouldBe` 23
+\end{verbatim}
+\end{quote}
+
+\noindent This is just another name for HUnit's \texttt{@?=}
+operator.  The type is:
+
+\begin{quote}
+\small
+\begin{verbatim}
+shouldBe :: (Show a, Eq a) => a -> a -> Expectation
+\end{verbatim}
+\end{quote}
+
+\noindent Or we may expect that a value should satisfy some
+predicate.
+
+\begin{quote}
+\small
+\begin{verbatim}
+x `shouldSatisfy` (< 23)
+\end{verbatim}
+\end{quote}
+
+\noindent The value is required to be in the \texttt{Show} class, so that a
+useful error message can be given when the predicate is not satisfied.
+
+\begin{quote}
+\small
+\begin{verbatim}
+shouldSatisfy :: Show a => a -> (a -> Bool)
+  -> Expectation
+\end{verbatim}
+\end{quote}
+
+\noindent Code examples are often used to provide evidence for {\tt
+IO} actions, e.g.:
+
+\begin{quote}
+\small
+\begin{verbatim}
+launchMissiles >>=
+  (`shouldBe` Left "not implemented")
+\end{verbatim}
+\end{quote}
+
+\noindent So we propose a shortcut for this common use case:
+
+\begin{quote}
+\small
+\begin{verbatim}
+launchMissiles `shouldReturn` Left "not implemented"
+\end{verbatim}
+\end{quote}
+
+\noindent where \verb|shouldReturn| has the following type:
+
+\begin{quote}
+\small
+\begin{verbatim}
+shouldReturn :: (Show a, Eq a) => IO a -> a
+  -> Expectation
+\end{verbatim}
+\end{quote}
+
+%%%%%%%%%%%%%%%%
+\subsubsection{Expecting exceptions}
+\label{sec:expacting-exceptions}
+
+HUnit does not provide a convenient mechanism to state that an action
+should throw an exception.  So we propose a combinator for that.
+
+\begin{quote}
+\small
+\begin{verbatim}
+shouldThrow :: Exception e => IO a -> (e -> Bool)
+  -> Expectation
+\end{verbatim}
+\end{quote}
+
+\noindent It takes an {\tt IO} action, and a predicate.  The predicate
+servers two purposes.  It selects the type of the expected exception,
+and constrains the value.  Here is an example:
+
+\begin{quote}
+\small
+\begin{verbatim}
+launchMissiles `shouldThrow` (== ExitFailure 1)
+\end{verbatim}
+\end{quote}
+
+%\noindent And it can be used with predicates from {\tt
+%System.IO.Error}, to require specific {\tt IOException}, e.g.  with
+%{\tt isPermissionError}:
+%
+%\begin{quote}
+%\small
+%\begin{verbatim}
+%launchMissiles `shouldThrow` isPermissionError
+%\end{verbatim}
+%\end{quote}
+
+\noindent To make it more convenient to expect any exception of a
+certain type, we provide type-restricted versions of {\tt (const
+True)} for all standard exceptions, e.g.:
+
+\begin{quote}
+\small
+\begin{verbatim}
+anyArithException :: ArithException -> Bool
+anyArithException = const True
+\end{verbatim}
+\end{quote}
+
+\noindent Which can be used like this:
+
+\begin{quote}
+\small
+\begin{verbatim}
+evaluate (1 `div` 0) `shouldThrow` anyArithException
+\end{verbatim}
+\end{quote}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Automatic discovery of specs}
+\label{sec:solution-discovery}
+
+We want to automatically discover all specs of a project, so that we
+can run them all together.  This requires to automatically generate
+boilerplate code, similar to what is shown in
+Section~\ref{sec:problem-discovery}.
+
+We make creative use GHC's support for custom preprocessors for
+that.  With our solution, the developer creates a test driver, that only
+contains a single line:
+
+\begin{quote}
+\small
+\begin{verbatim}
+{-# OPTIONS_GHC -F -pgmF hspec-discover #-}
+\end{verbatim}
+\end{quote}
+
+\noindent This tells GHC to run a preprocessor, \verb|hspec-discover|, on the
+source file.  \verb|hspec-discover| finds all specs for a project, and
+generates a suitable test driver.
+
+This approach relies on the following conventions:
+\begin{enumerate}
+    \item
+        Spec files have to be placed into the same directory as the
+        test driver, or into a subdirectory.
+    \item
+        The name of a spec file has to end in \verb|Spec.hs|, and the
+        module name has to match the file name.
+    \item
+        Each spec file has to export a top-level binding \verb|spec|,
+        of type \verb|Spec|.
+\end{enumerate}
+
+\noindent It would be possible to lift some of those requirements.
+But we think they are reasonable conventions, anyway.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Implementation}
+
+In this section we describe implementation aspects of the proposed
+solution.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Haddock}
+\label{sec:haddock}
+
+We added markup support for interactive examples to Haddock, and it is
+available since version 2.8.
+We also added markup support for properties, which will be included in
+a future release of Haddock.
+
+\subsection{Doctest}
+\label{sec:doctest}
+
+Doctest uses the GHC API to extract properties and examples from
+Haddock comments, and uses GHCi to test them.
+Before testing, Doctest brings the module that contains the Haddock
+comment from which the evidence was extracted into scope.
+
+\subsubsection{Testing examples}
+
+Verification of examples is done by
+comparing the expected output given in the Haddock comment with the
+actual output of GHCi.
+
+Doctest reuses the same GHCi session for all examples.
+Examples from the same Haddock comment are grouped together,
+and share the same scope.
+Doctest clears the scope with \verb|:reload| before it tests a new
+example group.
+
+\subsubsection{Testing properties}
+
+QuickCheck is used to test properties.
+Implicitly quantified properties need to be closed before they can be
+tested.  This requires the following steps.
+\begin{enumerate}
+    \item
+        Find all free variables of the property.
+    \item
+        Try to lookup each free variable in the module context.
+    \item
+        Close the property over all free variables, for which lookup
+        failed.
+\end{enumerate}
+
+\noindent Here is the example from Section~\ref{sec:markup}:
+
+\begin{quote}
+\small
+\begin{verbatim}
+sort xs == (sort . sort) (xs :: [Int]))
+\end{verbatim}
+\end{quote}
+
+\noindent Both \verb|xs| and \verb|sort| are free variables, but
+\verb|sort| is bound by the module context.  So we need to close the
+term over \verb|xs|.
+
+\begin{quote}
+\small
+\begin{verbatim}
+\xs -> sort xs == (sort . sort) (xs :: [Int])
+\end{verbatim}
+\end{quote}
+
+\noindent The implementation is fully functional, and it will be included in the
+next release of Doctest.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Hspec combinators for expectations}
+\label{sec:hspec-comb}
+
+\noindent All combinators introduced in Section~\ref{sec:code-doc} have been
+implemented.  They are available from Hackage\footnote{%
+http://hackage.haskell.org/package/hspec-shouldbe},
+and will be incorporated into a future
+release of Hspec.
+The implementation is straightforward, so we will not further discuss
+it.
+
+\subsection{Automatic discovery of specs}
+\label{sec:implementation-discovery}
+
+As discussed in Section~\ref{sec:solution-discovery}, we use a preprocessor to
+automatically generate test drivers for Hspec.  GHC invokes the
+preprocessor with three arguments:
+
+\begin{description}
+    \item[{\tt src}]
+        -- path to the original source file\\
+        A preprocessor usually does not
+        look at the content of this file.  It is only provided, so
+        that it can be used in error messages.
+    \item[{\tt cur}]
+        -- path to a file that holds the input for the preprocessor
+    \item[{\tt dst}]
+        -- a path, where the preprocessor should write it's result to
+\end{description}
+
+\noindent A typical preprocessor reads the content of {\tt cur},
+processes it, and writes the result to {\tt dst}.  If something goes
+wrong, it uses {\tt src} in error messages to refer to the source
+file.
+
+\verb|hspec-discover| ignores {\tt cur} altogether.  It uses the
+directory component of {\tt src}, to find all files that are in the
+same directory as the source file, or in any subdirectories.  It then generates
+a test driver from all files that end in {\tt Spec.hs}.
+
+A fully functional implementation is available from
+Hackage\footnote{http://hackage.haskell.org/package/hspec-discover},
+and it will be incorporated into a future release of Hspec.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Best current practices}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Open problems and future work}
+
+In Section~\ref{sec:reusable-specs}, we showed how to reuse pieces of
+specification.  It would be possible and useful to write a library
+that provides reusable pieces of specification, for many of Haskell's
+standard classes (similar in purpose to {\tt
+checkers}\footnote{http://hackage.haskell.org/package/checkers}).
+This would provide a convenient mechanism for developers to document
+and test that their instance definitions are well-behaved.
+
+When Hspec finds failing evidence, it reports the corresponding
+behavior to the developer.  It would be useful to include source
+locations for failing evidence, so that the programmer can find the
+corresponding spec immediately. Solving this with GHC requires
+Template Haskell.  Extending GHC with a mechanism
+similar to JHC's \verb|SRCLOC_ANNOTATE| pragma\footnote{%
+http://repetae.net/computer/jhc/jhc.shtml} would allow to solve this
+without Template Haskell.
+
+Cucumber~\cite{cucumber} is a higher-level approach to BDD for Ruby, and
+it is typically used in combination with RSpec.
+It focuses on user stories, to assist the developer in delivering
+value to the stakeholder.
+There is nothing comparable to Cucumber for Haskell, yet.  And we think
+it would be interesting to explore, whether it is possible to apply
+Cucumber's approach to Haskell.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Related work}
+
+SmallCheck~\cite{smallcheck} is a tool for
+automatic exhaustive testing up to a certain depth.
+Both Doctest and Hspec do not have support for SmallCheck
+at this moment.
+Hspec's mechanism for evidence is extensible, and it would be easy to
+add support for SmallCheck.
+
+{\tt test-framework} is a library that allows TDD in Haskell, and it
+has support for HUnit, QuickCheck and SmallCheck.
+There is support for automatic test discovery that relies on naming
+conventions for test functions\footnote{%
+http://hackage.haskell.org/package/test-framework-th}.  It is
+implemented with Template Haskell~\cite{template-haskell}, and it only
+works within a single source file.  Automatic discovery across
+multiple source files is not possible with a mechanism that is based
+on Template Haskell, because Template Haskell can not generate import
+statements\footnote{%
+See http://hackage.haskell.org/trac/ghc/ticket/1475 }.
+
+The {\tt testpack}\footnote{%
+http://hackage.haskell.org/package/testpack}
+package provides a combinator to expect exceptions,
+similar to {\tt shouldThrow} from
+Section~\ref{sec:expacting-exceptions}:
+
+\begin{quote}
+\small
+\begin{verbatim}
+assertRaises :: (Show a, Exception e, Show e, Eq e) =>
+  String -> e -> IO a -> IO ()
+\end{verbatim}
+\end{quote}
+
+\noindent But it has several shortcomings:
+It only works for exceptions with an {\tt Eq} instance,
+but several standard exceptions (most
+prominently {\tt ErrorCall}) have no {\tt Eq} instance.
+And it places an unnecessary {\tt Show} constraint on the
+result of the {\tt IO} action, further limiting it's applicability.
+It only allows to expect specific exception values, expecting
+arbitrary exceptions of a certain type is not possible.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusion}
+
+We showed how to apply behavior driven development to
+Haskell.  We proposed to use Doctest for user documentation and Hspec
+for code documentation. We extended Haddock's markup with support for
+examples and properties.  And we implemented Doctest, which verifies
+examples and tests properties in Haddock comments.
+We added two missing features to Hspec: combinators to express
+expectations, and a mechanism for automatic discovery of specs.
+And we showed how to reuse pieces of specification.
+Our approach can be used for design, documentation and automated
+testing.  We hope that this motivates Haskell programmers to write
+more and better tests.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \appendix
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\acks
+
+We are grateful to David Waern, Haddock's maintainer, for reviewing
+and accepting our patches to Haddock.
+%We would like to thank
+%Nick Smallbone for having improved QuickCheck so that
+%non-quantified properties are tested just once.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \bibliographystyle{abbrvnat}

### 2012-10-20 paper.tex
Emphasis first use of BDD

--- 2012-10-20 paper.tex
@@
 
-In this section we show how to do BDD in Haskell, how to reuse pieces
+In this section we show how to do \emph{behavior-drive development
+(BDD)} in Haskell, how to reuse pieces
 of specification, and how to use Doctest and Hspec in combination with
